### 2018-01-30 Radiocarbon2018/abstract/RC-18HGISAbst.Rmd
add Radiocarbon abstract

---
title: "Data acquisition and analysis at NOSAMS"
author:
  - name: Brett E. Longworth
    email: blongworth@whoi.edu
    affiliation: Woods Hole Oceanographic Institution 
    footnote: Corresponding Author
  - name: Alan R. Gagnon
    affiliation: Woods Hole Oceanographic Institution 
  - name: Kalina Godispodinova
    affiliation: Woods Hole Oceanographic Institution 
  - name: Joshua Hlavenka
    affiliation: Woods Hole Oceanographic Institution 
  - name: Ann P. McNichol
    affiliation: Woods Hole Oceanographic Institution 
  - name: Mark L. Roberts 
    affiliation: Woods Hole Oceanographic Institution 
address:
  - code: Woods Hole Oceanographic Institution 
    address: WHOI/NOSAMS, Woods Hole, MA, 02543
abstract: |

Here we present progress on development of a hybrid sputter gas ion source and associated gas inlets.

We are focussing on two inlet technologies (compare and contrast) modified
hardware from a vg stable isotope mass spectrometer and an open split.

progress

application

goals

keywords: ion source, gas ion source, inlet, accelerator mass
spectrometry, radiocarbon output: rticles::elsevier_article ---

### 2018-04-03 Radiocarbon2018/abstract/RC-18HGISAbst.Rmd
Final draft

@@
 ---
-title: "Data acquisition and analysis at NOSAMS"
+title: "Development of a Hybrid Gas Ion Source and Gas Inlets at NOSAMS"
 author:
@@
     affiliation: Woods Hole Oceanographic Institution 
-  - name: Kalina Godispodinova
+  - name: Joshua R. Burton
     affiliation: Woods Hole Oceanographic Institution 
-  - name: Joshua Hlavenka
+  - name: Kalina D. Godispodinova
+    affiliation: Woods Hole Oceanographic Institution 
+  - name: Joshua D. Hlavenka
     affiliation: Woods Hole Oceanographic Institution 
@@
 
-Here we present progress on development of a hybrid sputter gas ion source and associated gas inlets.
+  Here we present progress on development of gas inlet systems for a hybrid sputter gas ion source (HGIS).
+Our HGIS is a lightly modified gas-capable 134-sample MCSNICS source from NEC.
+The ion optics and cesium delivery have been upgraded following Roberts et al.
+[-@Roberts2010].
+Gas delivery to the source is via a short glass capillary which crosses the source HV potential.
 
-We are focussing on two inlet technologies (compare and contrast) modified
-hardware from a vg stable isotope mass spectrometer and an open split.
+  We are developing two inlet technologies, one delivering gas to the source at low pressure using modified gas handling hardware from a VG stable isotope mass spectrometer and the other sampling gas at atmospheric pressure using an open split.
+The gas handling system designed for use with SIRMS is nearly ideal for our purposes, providing low volume, low pressure delivery of gas, and control of mixing ratios and delivery pressure using variable volumes.
+We will use the open split with large samples, focussing first on carbonate samples, hydrolysed and delivered using the gasbench system developed for use with the CFAMS microwave plasma GIS [@Roberts2013].
 
-progress
-
-application
-
-goals
-
-keywords: ion source, gas ion source, inlet, accelerator mass
-spectrometry, radiocarbon output: rticles::elsevier_article ---
+keywords: ion source, gas ion source, inlet, accelerator mass spectrometry, radiocarbon 
+bibliography: RC-18HGISAbst.bib
+csl: nuclear-inst-and-methods-in-physics-research-b.csl
+output: rticles::elsevier_article 
+---

### 2020-12-21 hgis_summary.Rmd
Add summary document

---
title: "USAMS Hybrid Source Summary"
author: "Brett Longworth"
date: "12/20/2020"
output: html_document
---

This document summarizes the operating parameters, development, and questions
for the hybrid gas ion source on USAMS.

# Source parameters

Best currents for the source are around 10uA 12C- or roughly 1/10 of graphite
currents. This is in line with what other hybrid sources produce.  The source
typically produces best current at around 10uLs-1 CO2. Literature values give
1.5uLs-1 as optimal, but we've never seen that, and pressures, capillaries and
flows in at least one of these papers doesn't make sense.  Precision on modern
targets has been as good as 1% (SD of multiple measurements of a target) with
directly introduced bottle gas. Typical precision in tests of inlet systems has
been around 3%. Measured value of tank modern gas agrees with the accepted
value within error when normalizing to solid graphite standards.  The system
seems to have a constant blank contribution from the target and source of about
100nA 12C- or 1% of a sample producing 10uA 12C-. Helium does not seem to add
significantly to this blank. Since the blank current is constant, the blank
contribution is inversely related to the sample current, ie. if the sample
current is 5uA, 2% of that will be from the blank and so on.

# Open split inlet

The open split interface supplies a helium-co2 mixture to a split where a
capillary delivers ~10uL/min to the source and the rest is vented. The current
split starts showing "breakthrough" of lab air at around 150uLs-1 flow.
Currently using 3m of 50um capillary with a bulkhead connector at the source
cage.  Precision of measurements of modern CO2 has been 3% at best, 10% at
worst, with better precision most recently. This is much less precise than the
expected 1%.  Accuracy is reasonable within the large errors on gas
measurements. Uncorrected measurements of dead CO2 range from 1% to 5%
depending on sample current.

## Questions

Q: Background from the source/target is relatively stable at roughly 100nA or
1% of maximum CO2 current. Is the relative blank contribution proportional to
CO2 current?

A: Based on limited data, it's looking like yes, this follows a constant
contamination model.

Q: If the relative blank contribution is proportional to sample current, can we
correct for a varying sample current over a run (due to variable dillution,
pressure, etc.)?

A: If the sample current varies widely, this will likely require taking data in
shorter segments and blank correcting each segment using the mean current for
that segment.

Q: Why is the precision not as expected? Can we replicate the
literature/initial test precision?

A: Unknown. We can rerun a test with the best performing conditions (bottle gas
through capillary to source) to eliminate as many variables as possible.

## Development

2020-12-18 - Continue testing lower open split flows with leak checker, run
dead and live gas at lower flows. Spoke with Josh B. about double needle
setups.

2020-12-11 - Open split flow reduction. Test flow controller for helium, adjust
needle valves to allow lower CO2 flows. Measure blanks of target, helium, CO2
and dilluted CO2.

2020-12-04 - Measure blank contributions of target/source, helium, CO2, and
dilluted CO2. Install flow controller for helium.

2020-12-03 - Prepare and load new HGIS test wheel.

2020-11-17 - Diagnose issue with low currents, measure diluted and pure live
and dead gas.

2020-11-03 - Troubleshoot low currents, measure dilluted and pure live and dead
gas.

2020-10-15 - Troubleshoot low currents. Lots of tune and target position
testing.

2020-10-13 - Prepare and load new HGIS test wheel.

2020-10-06 - Dillution tests of live and dead gas. Test stability of currents
over time.

2020-09-29 - Starting up HGIS. Load sample wheel. Reterminate capillaries at
cage wall and open split.

### 2021-01-19 hgis_summary.Rmd
Update summary.

@@
 1.5uLs-1 as optimal, but we've never seen that, and pressures, capillaries and
-flows in at least one of these papers doesn't make sense.  Precision on modern
-targets has been as good as 1% (SD of multiple measurements of a target) with
-directly introduced bottle gas. Typical precision in tests of inlet systems has
-been around 3%. Measured value of tank modern gas agrees with the accepted
-value within error when normalizing to solid graphite standards.  The system
-seems to have a constant blank contribution from the target and source of about
-100nA 12C- or 1% of a sample producing 10uA 12C-. Helium does not seem to add
-significantly to this blank. Since the blank current is constant, the blank
-contribution is inversely related to the sample current, ie. if the sample
-current is 5uA, 2% of that will be from the blank and so on.
+flows in at least one of these papers doesn't make sense.
+
+Initially, precision on modern targets was thought to be as good as 1% (SD of
+multiple measurements of a modern target) with directly introduced bottle gas,
+which is in agreement with literature values. Revisiting these data show
+precision in the range of 2-4%. Typical precision in tests of inlet systems has
+been in the range of 2-4%.
+
+Measured value of tank modern gas agrees with the accepted value within error
+when normalizing to solid graphite standards.  The system seems to have a
+constant blank contribution from the target and source of about 100nA 12C- or 1%
+of a sample producing 10uA 12C-. Helium does not seem to add significantly to
+this blank. Since the blank current is constant, the blank contribution is
+inversely related to the sample current, ie. if the sample current is 5uA, 2% of
+that will be from the blank and so on.
+
+The above measurements are at optimum sample current of 6-10uA 12C-. Precision
+and measured ratio both decrease as sample current drops below the optimum
+range. The ratio effects may be related to the "constant current" blank
+discussed above.
 
@@
 capillary delivers ~10uL/min to the source and the rest is vented. The current
-split starts showing "breakthrough" of lab air at around 150uLs-1 flow.
+split starts showing "breakthrough" of lab air at around 150uL/min flow.
 Currently using 3m of 50um capillary with a bulkhead connector at the source
-cage.  Precision of measurements of modern CO2 has been 3% at best, 10% at
+cage.  Precision of measurements of modern CO2 has been 2% at best, 10% at
 worst, with better precision most recently. This is much less precise than the
@@
 depending on sample current.
+
+## Helium displacement from vials
+
+First tests of displacing pure CO2 from a vial into the open split using helium
+are promising. Helium flowing at 100uLm-1 produces current from 7mL of CO2 for
+over 2 hours. Currents do decrease with time due to dilution, but by supplying
+higher than optimal CO2 flow to the source initially, stable currents can be
+maintained for some time.
 
@@
 Q: If the relative blank contribution is proportional to sample current, can we
-correct for a varying sample current over a run (due to variable dillution,
+correct for a varying sample current over a run (due to variable dilution,
 pressure, etc.)?
@@
 shorter segments and blank correcting each segment using the mean current for
-that segment.
+that segment. We may be able to mitigate by setting capillary length such that
+CO2 flow starts above optimal and drops to optimal by the end of the
+acquisition.
 
@@
 
-A: Unknown. We can rerun a test with the best performing conditions (bottle gas
-through capillary to source) to eliminate as many variables as possible.
+A: Unknown. Tests with the best performing conditions (bottle gas through
+capillary to source) give 2% precision, and revisiting early test data shows
+precision of 2%. Precision is driven by variability between measurements of a
+sample. Investigating correlations here may help. We should also look at
+variability between identical samples to get a sense of reproducibility.
 
 ## Development
+2021-01-15 - Pure CO2 and displacement tests using vial and double needle.
+Tested precision with vial setup by flowing CO2 at 200kPa (high flow, roughly 400ul/min) into open split.
+Followed this by switching to helium flow to vial at 100kPa (200ul/min).
+Repeated test with a second vial.
+Built second dual needle setup for use in vial filling (can also use Gilson).
+
+2021-01-08 - Precision tests with direct CO2-source connection.
+Tested source pressure and current vs CO2 regulator pressure.
+Tested flow vs pressure with leak checker.
+Several capillary swaps.
+Acquired data at several flow rates for precision testing.
+
+2021-01-07 - Set up for precision test using 7m x 50um capillary from CO2/he via Valco valve to cage wall.
+
+2020-12-31 - Add CO2/He valve to vial test setup. VICI 4-way valve connected
+with 1/16" ss tubing from CO2 and He regulators allows switching between gasses
+delivered to a glass capillary leading to outer needle of dual needle. Leak
+checking and capillary adjustments for good flow range at reasonable pressures.
+
+2020-12-24 - Build dual needle setup. Supply displacement helium through outer
+1/16" tubing, sample mixture through inner needle. Use VICI Luer adapter on top
+of needle as open split with inserted capillary to source.
 

### 2021-01-24 hgis_summary.Rmd
Update summary.

@@
 currents. This is in line with what other hybrid sources produce.  The source
-typically produces best current at around 10uLs-1 CO2. Literature values give
+typically produces best current at around 5uLs-1 CO2. Literature values give
 1.5uLs-1 as optimal, but we've never seen that, and pressures, capillaries and
@@
 
-Initially, precision on modern targets was thought to be as good as 1% (SD of
-multiple measurements of a modern target) with directly introduced bottle gas,
-which is in agreement with literature values. Revisiting these data show
-precision in the range of 2-4%. Typical precision in tests of inlet systems has
-been in the range of 2-4%.
+Precision taken as the standard error of the mean of repeated runs of a sample is around 1%.
+Precision is typically a little worse than predicted by counting statistics.
+More work is needed to look at whether this reflects variability between samples.
 
-Measured value of tank modern gas agrees with the accepted value within error
-when normalizing to solid graphite standards.  The system seems to have a
-constant blank contribution from the target and source of about 100nA 12C- or 1%
-of a sample producing 10uA 12C-. Helium does not seem to add significantly to
-this blank. Since the blank current is constant, the blank contribution is
-inversely related to the sample current, ie. if the sample current is 5uA, 2% of
-that will be from the blank and so on.
+Measured value of tank modern gas normalized to solid graphite standards ranges from agreeing with solid OX-I to 7% lower than expected for low current samples.
+Normalizing to tank CO2 is probably the best method to account for variability in hybrid source performance.
+
+The system seems to have a constant blank contribution from the target and
+source of about 100nA 12C- or 1% of a sample producing 10uA 12C-. Helium does
+not seem to add significantly to this blank. Since the blank current is
+constant, the blank contribution is inversely related to the sample current, ie.
+if the sample current is 5uA, 2% of that will be from the blank and so on.
 
@@
 range. The ratio effects may be related to the "constant current" blank
-discussed above.
+discussed above. More work is needed to see whether this is just a blank effect or whether there are other factors at low current.
 
@@
 capillary delivers ~10uL/min to the source and the rest is vented. The current
-split starts showing "breakthrough" of lab air at around 150uL/min flow.
-Currently using 3m of 50um capillary with a bulkhead connector at the source
-cage.  Precision of measurements of modern CO2 has been 2% at best, 10% at
-worst, with better precision most recently. This is much less precise than the
-expected 1%.  Accuracy is reasonable within the large errors on gas
-measurements. Uncorrected measurements of dead CO2 range from 1% to 5%
-depending on sample current.
+split starts showing "breakthrough" of lab air at around 150uL/min flow. A new open split with lower volume should have improved this, but testing is needed.
+
+Currently, 250cm of 50um capillary defines flow rate from the open split into
+the source, with a bulkhead connector at the source cage.  Precision of
+measurements of modern CO2 has been better than 1% (std error of mean of
+measurements of a target) for high current samples, with precision dropping off
+as current decreases below 6uA or so. Accuracy is reasonable within the large
+errors on gas measurements. Uncorrected measurements of dead CO2 range from 1%
+to 5% of modern depending on sample current.
 
@@
 maintained for some time.
+Targets have shown "pulses" of higher current.
+Not sure why this is, but I suspect it's either buildup of graphite or other available carbon on the target, or target switching to the "high state" as sometimes happens with solid targets.
 
@@
 
-Q: Why is the precision not as expected? Can we replicate the
-literature/initial test precision?
+Q: Does standard error of the mean as a sample error agree with between target variability?
 
-A: Unknown. Tests with the best performing conditions (bottle gas through
-capillary to source) give 2% precision, and revisiting early test data shows
-precision of 2%. Precision is driven by variability between measurements of a
-sample. Investigating correlations here may help. We should also look at
-variability between identical samples to get a sense of reproducibility.
+A: Unknown. We should look at variability between identical samples to get a sense of reproducibility. Some data are available, but no controlled replicates have been run.
 
 ## Development
+
+2021-01-22 - Ran pure CO2 at different pressures to look at effect of current on precision and ratio. Repeated test for live and dead tank CO2.
+
 2021-01-15 - Pure CO2 and displacement tests using vial and double needle.

### 2021-01-24 hgis_summary.Rmd
Fix typos.

@@
 needle valves to allow lower CO2 flows. Measure blanks of target, helium, CO2
-and dilluted CO2.
+and diluted CO2.
 
 2020-12-04 - Measure blank contributions of target/source, helium, CO2, and
-dilluted CO2. Install flow controller for helium.
+diluted CO2. Install flow controller for helium.
 
@@
 
-2020-11-03 - Troubleshoot low currents, measure dilluted and pure live and dead
+2020-11-03 - Troubleshoot low currents, measure diluted and pure live and dead
 gas.
@@
 
-2020-10-06 - Dillution tests of live and dead gas. Test stability of currents
+2020-10-06 - Dilution tests of live and dead gas. Test stability of currents
 over time.

### 2021-01-25 hgis_summary.Rmd
Update summary.

@@
 The open split interface supplies a helium-co2 mixture to a split where a
-capillary delivers ~10uL/min to the source and the rest is vented. The current
+capillary delivers gas to the source and the rest is vented. The current
 split starts showing "breakthrough" of lab air at around 150uL/min flow. A new open split with lower volume should have improved this, but testing is needed.
@@
 maintained for some time.
+Setting flow to the source to 10uL/min provides excess CO2 to source until the helium dilution ratio increases to > 1:1.
+Another approach may be to backfill vials with helium and run at a higher dilution ratio and flow to the source.
+This should produce a smaller change in He:CO2 over the course of a run.
+MICADAS runs at 95% He:CO2, which would mean producing < 0.5mL CO2 in a 7mL vial and backfilling with helium before displacing to the source.
+Some intermediate dilution may be appropriate.
+
 Targets have shown "pulses" of higher current.
@@
 A: Unknown. We should look at variability between identical samples to get a sense of reproducibility. Some data are available, but no controlled replicates have been run.
+
+## To do
+
+* Repeated measurements under identical conditions to determine reproducibility
+* Test breakthrough with new open split
+* Establish best helium flow for displacement
+* Fine tune capillary length for best flow to source over pCO2 range of full size sample
+* How to handle < 1 Atm in vials. Backfill?
+* Measure test samples prepped on Gilson (7mL CO2, Atm pressure)
+* Define blank, determine whether current-balance correction is needed.
+* Refine normalization and error propagation for data reduction.
 

### 2021-02-01 hgis_summary.Rmd
Update notes and summary with most recent data.

@@
 flows in at least one of these papers doesn't make sense.
+
+Source currents are variable by day and by target.
+"Good" currents for a given day vary between 5-10uA, and it's not yet clear what causes this variability.
+Alignment, source condition, and settings of the ionizer and cesium are all factors.
+Current also varies by target for a given set of gas and source conditions.
+Some targets are clearly "bad", producing currents much lower than other targets for a given set of conditions.
 
@@
 to 5% of modern depending on sample current.
+
+Measurement of a small number of replicates shows that between sample reproducibility is around 1%.
+Samples displaced by helium have slightly higher variability, but more tests are needed to better assess reproducibility.
 
@@
 
-A: Unknown. We should look at variability between identical samples to get a sense of reproducibility. Some data are available, but no controlled replicates have been run.
+A: Initial data from identical samples run under similar conditions show standard deviation of around 1-1.5% for modern samples.
 
@@
 
-* Repeated measurements under identical conditions to determine reproducibility
+* More repeated measurements under identical conditions to determine reproducibility
 * Test breakthrough with new open split
@@
 ## Development
+
+2021-01-29 - Ran a series of vial tests with modern and dead CO2 to assess reproducibility.
 

### 2021-02-22 hgis_summary.Rmd
Use include to use summary in README and summary doc.

--- 2021-01-24 hgis_summary.Rmd
@@
 author: "Brett Longworth"
-date: "12/20/2020"
+date: "`r Sys.Date()`"
 output: html_document
@@
 
-# Source parameters
-
-Best currents for the source are around 10uA 12C- or roughly 1/10 of graphite
-currents. This is in line with what other hybrid sources produce.  The source
-typically produces best current at around 5uLs-1 CO2. Literature values give
-1.5uLs-1 as optimal, but we've never seen that, and pressures, capillaries and
-flows in at least one of these papers doesn't make sense.
-
-Precision taken as the standard error of the mean of repeated runs of a sample is around 1%.
-Precision is typically a little worse than predicted by counting statistics.
-More work is needed to look at whether this reflects variability between samples.
-
-Measured value of tank modern gas normalized to solid graphite standards ranges from agreeing with solid OX-I to 7% lower than expected for low current samples.
-Normalizing to tank CO2 is probably the best method to account for variability in hybrid source performance.
-
-The system seems to have a constant blank contribution from the target and
-source of about 100nA 12C- or 1% of a sample producing 10uA 12C-. Helium does
-not seem to add significantly to this blank. Since the blank current is
-constant, the blank contribution is inversely related to the sample current, ie.
-if the sample current is 5uA, 2% of that will be from the blank and so on.
-
-The above measurements are at optimum sample current of 6-10uA 12C-. Precision
-and measured ratio both decrease as sample current drops below the optimum
-range. The ratio effects may be related to the "constant current" blank
-discussed above. More work is needed to see whether this is just a blank effect or whether there are other factors at low current.
-
-# Open split inlet
-
-The open split interface supplies a helium-co2 mixture to a split where a
-capillary delivers ~10uL/min to the source and the rest is vented. The current
-split starts showing "breakthrough" of lab air at around 150uL/min flow. A new open split with lower volume should have improved this, but testing is needed.
-
-Currently, 250cm of 50um capillary defines flow rate from the open split into
-the source, with a bulkhead connector at the source cage.  Precision of
-measurements of modern CO2 has been better than 1% (std error of mean of
-measurements of a target) for high current samples, with precision dropping off
-as current decreases below 6uA or so. Accuracy is reasonable within the large
-errors on gas measurements. Uncorrected measurements of dead CO2 range from 1%
-to 5% of modern depending on sample current.
-
-## Helium displacement from vials
-
-First tests of displacing pure CO2 from a vial into the open split using helium
-are promising. Helium flowing at 100uLm-1 produces current from 7mL of CO2 for
-over 2 hours. Currents do decrease with time due to dilution, but by supplying
-higher than optimal CO2 flow to the source initially, stable currents can be
-maintained for some time.
-Targets have shown "pulses" of higher current.
-Not sure why this is, but I suspect it's either buildup of graphite or other available carbon on the target, or target switching to the "high state" as sometimes happens with solid targets.
-
-## Questions
-
-Q: Background from the source/target is relatively stable at roughly 100nA or
-1% of maximum CO2 current. Is the relative blank contribution proportional to
-CO2 current?
-
-A: Based on limited data, it's looking like yes, this follows a constant
-contamination model.
-
-Q: If the relative blank contribution is proportional to sample current, can we
-correct for a varying sample current over a run (due to variable dilution,
-pressure, etc.)?
-
-A: If the sample current varies widely, this will likely require taking data in
-shorter segments and blank correcting each segment using the mean current for
-that segment. We may be able to mitigate by setting capillary length such that
-CO2 flow starts above optimal and drops to optimal by the end of the
-acquisition.
-
-Q: Does standard error of the mean as a sample error agree with between target variability?
-
-A: Unknown. We should look at variability between identical samples to get a sense of reproducibility. Some data are available, but no controlled replicates have been run.
-
-## Development
-
-2021-01-22 - Ran pure CO2 at different pressures to look at effect of current on precision and ratio. Repeated test for live and dead tank CO2.
-
-2021-01-15 - Pure CO2 and displacement tests using vial and double needle.
-Tested precision with vial setup by flowing CO2 at 200kPa (high flow, roughly 400ul/min) into open split.
-Followed this by switching to helium flow to vial at 100kPa (200ul/min).
-Repeated test with a second vial.
-Built second dual needle setup for use in vial filling (can also use Gilson).
-
-2021-01-08 - Precision tests with direct CO2-source connection.
-Tested source pressure and current vs CO2 regulator pressure.
-Tested flow vs pressure with leak checker.
-Several capillary swaps.
-Acquired data at several flow rates for precision testing.
-
-2021-01-07 - Set up for precision test using 7m x 50um capillary from CO2/he via Valco valve to cage wall.
-
-2020-12-31 - Add CO2/He valve to vial test setup. VICI 4-way valve connected
-with 1/16" ss tubing from CO2 and He regulators allows switching between gasses
-delivered to a glass capillary leading to outer needle of dual needle. Leak
-checking and capillary adjustments for good flow range at reasonable pressures.
-
-2020-12-24 - Build dual needle setup. Supply displacement helium through outer
-1/16" tubing, sample mixture through inner needle. Use VICI Luer adapter on top
-of needle as open split with inserted capillary to source.
-
-2020-12-18 - Continue testing lower open split flows with leak checker, run
-dead and live gas at lower flows. Spoke with Josh B. about double needle
-setups.
-
-2020-12-11 - Open split flow reduction. Test flow controller for helium, adjust
-needle valves to allow lower CO2 flows. Measure blanks of target, helium, CO2
-and diluted CO2.
-
-2020-12-04 - Measure blank contributions of target/source, helium, CO2, and
-diluted CO2. Install flow controller for helium.
-
-2020-12-03 - Prepare and load new HGIS test wheel.
-
-2020-11-17 - Diagnose issue with low currents, measure diluted and pure live
-and dead gas.
-
-2020-11-03 - Troubleshoot low currents, measure diluted and pure live and dead
-gas.
-
-2020-10-15 - Troubleshoot low currents. Lots of tune and target position
-testing.
-
-2020-10-13 - Prepare and load new HGIS test wheel.
-
-2020-10-06 - Dilution tests of live and dead gas. Test stability of currents
-over time.
-
-2020-09-29 - Starting up HGIS. Load sample wheel. Reterminate capillaries at
-cage wall and open split.
+```{r child='summary.Rmd'}
+```

### 2021-04-05 doc/SNEAP2021/SNEAP2021_Abstract.Rmd
SNEAP abstract first commit.

---
title: "Progress with the NOSAMS Hybrid Gas Ion Source"
author: "Brett Longworth"
date: "3/31/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

A hybrid sputter ion source produces carbon ions from CO2 gas for measurement of 14/12C via accelerator mass spectrometry. CO2 is introduced onto a titanium surface within a cathode sputtered by cesium ions, dissociating the CO2 and ionizing some of the carbon ions for extraction as in a solid sputter source. At NOSAMS we have been working to find ideal conditions for this process and develop interfaces for producing CO2 and introducing it to the source under these conditions. Early development with pure CO2 from gas cylinders gave way to introduction of CO2 diluted with helium carrier in a closed system, and most recent development introduces a CO2/He mixture to the source via an open split. Current development focuses on producing CO2 from carbonates (CaCO3) which is delivered to the open split and ion source using a flow of helium carrier gas. We will discuss efficiency of the system, issues with background carbon current and precision and accuracy of the measurement.

hybrid source introduction
direct injection of CO2
closed source development with irms inlet
open split development to introduce CO2 from vials
advantages and disadvantages of closed and open methods

### 2021-04-11 doc/SNEAP2021/SNEAP2021_Abstract.Rmd
edit for clarity and length

@@
 title: "Progress with the NOSAMS Hybrid Gas Ion Source"
-author: "Brett Longworth"
-date: "3/31/2021"
-output: html_document
+author:
+- Brett Longworth
+- Josh Burton
+- Josh Hlavenka
+- Mark Roberts
+date: "`r format(Sys.time(), '%B %d, %Y')`"
+output:
+  pdf_document: default
+  html_document: default
+  word_document: default
 ---
@@
 
-A hybrid sputter ion source produces carbon ions from CO2 gas for measurement of 14/12C via accelerator mass spectrometry.
-CO2 is introduced onto a titanium surface within a cathode sputtered by cesium ions, dissociating the CO2 and ionizing some of the carbon ions for extraction as in a solid sputter source.
-At NOSAMS we have been working to find ideal conditions for this process and develop interfaces for producing CO2 and introducing it to the source under these conditions.
-Early development with pure CO2 from gas cylinders gave way to introduction of CO2 diluted with helium carrier in a closed system, and most recent development introduces a CO2/He mixture to the source via an open split.
-Current development focuses on producing CO2 from carbonates (CaCO3) which is delivered to the open split and ion source using a flow of helium carrier gas.
-We will discuss efficiency of the system, issues with background carbon current and precision and accuracy of the measurement.
-
-hybrid source introduction
-direct injection of CO2
-closed source development with irms inlet
-open split development to introduce CO2 from vials
-advantages and disadvantages of closed and open methods
+A hybrid gas-sputter ion source produces carbon ions from CO~2~ gas for measurement of ^14/12^C via accelerator mass spectrometry.
+CO~2~ is introduced onto a titanium surface within a cathode sputtered by cesium ions, dissociating CO~2~ and ionizing some of the carbon for extraction as in a solid sputter source.
+We are working to optimize ionization and develop interfaces that produce CO~2~ from samples and introduce it to the source.
+Early tests with CO~2~ standard gas gave way to introduction of CO~2~ diluted with helium carrier in a closed system, and most recent work introduces a CO~2~/He mixture to the source via an open split.
+Current development focuses on producing CO~2~ from carbonates (CaCO~3~) which is delivered to the open split and ion source using a flow of helium carrier gas.
+We will discuss efficiency, background carbon current, precision, and accuracy of the method.

### 2021-04-19 doc/SNEAP2021/SNEAP2021_Abstract.Rmd
Final version

@@
 - Mark Roberts
-date: "`r format(Sys.time(), '%B %d, %Y')`"
 output:
@@
 
-A hybrid gas-sputter ion source produces carbon ions from CO~2~ gas for measurement of ^14/12^C via accelerator mass spectrometry.
-CO~2~ is introduced onto a titanium surface within a cathode sputtered by cesium ions, dissociating CO~2~ and ionizing some of the carbon for extraction as in a solid sputter source.
+### National Ocean Sciences AMS Facility, Woods Hole Oceanographic Institution, Woods Hole, MA, USA
+
+We are using a hybrid gas-sputter ion source to produce carbon ions from CO~2~ gas for measurement of ^14/12^C via accelerator mass spectrometry.
+In our source, CO~2~ is introduced onto a titanium surface within a cathode sputtered by cesium ions, dissociating CO~2~ and ionizing some of the carbon for extraction as in a solid sputter source.
 We are working to optimize ionization and develop interfaces that produce CO~2~ from samples and introduce it to the source.

### 2021-04-28 doc/SNEAP2021/SNEAP2021_figures.Rmd
figures notebook

---
title: "SNEAP2021 Figures"
output:
  html_document:
    df_print: paged
---

Figures for SNEAP talk

# Setup

```{r}
library(tidyverse)
library(here)
library(amstools)
library(HybridGIS)
```

# Current vs flow

Good figure in first tests. Combine with data from dual bellows?

```{r, message = FALSE}

# factors for flow calc
r <- 2.5E-5
u <- 1.49E-5
x <- 8.2

data <- get_hgis_data(here("data/USAMS040317R.txt")) %>% 
  filter(Pos == 13)

# Assign Run numbers
counter <- 0
Run <- c(1:190)
for (i in 1:nrow(data)) {
  if (data$Meas[i] == 1) {
    counter <- counter + 1
  }
  Run[i] <- counter
} 
data$Run <- Run

# Load pressure data
pres <- read.csv(here("data/USAMS040317Pressure.csv"))
pres$Run <- as.numeric(row.names(pres))

# Join to run data
data <- inner_join(data, pres) %>%
  mutate(Source = Source * 1E-7, # I think the source pressures were to the -7, but unsure...
         flow = flowcalc(Bottle, r, u, x),
         eff = hgis_eff(flow, he12C, cs = 1)) # cs 1 to estimate LE current

# summarize
d.s <- data %>% 
  group_by(Run, Bottle, Source, flow) %>% 
  summarize(le12C = mean(le12C),
            he12C = mean(he12C),
            counts = sum(CntTotGT),
            interr = 1/sqrt(counts),
            C1412he = mean(X14.12he),
            C1412he_sd = sd(X14.12he),
            ferr = interr * C1412he,
            cor1412he = mean(cor1412he),
            eff = mean(eff))

# plot
filter(d.s, flow > 0) %>%
  ggplot(aes(flow, le12C)) +
  geom_smooth() +
  geom_point() +
  labs(x = expression(CO[2]~flow~rate~~mu~L %.% min^{-1}),
       y = bquote('12C-  ' *mu~ 'A')) +
  scale_y_reverse() +
  ylim(0,-10) +
  xlim(0,15.5)
```

# Efficiency vs flow

# Dual bellows

Breakseal results for OX-I, C-1, OX-II

# Open split

## Dillution curve

From helium_displacement.Rmd

Maybe also show long dillution runs plotted by CO2 flow?

Yes. Calculate flow at time using concCO2

## Reproducibility of standards

Data from 1-08, 1-22, 1-29

## Blanks - blank values

compare dead gas, vial dead gas, C1, and TIRI-F

## Blanks - estimation of constant current contaminant

combine data from 12-04, 12-11, 12-18,

## Results for carbonates

Combine all runs 2-05, 3-05, 4-09, 4-16

show as agreement with consensus vs consensus value with sample names

Need normalized values, BC values, Errors, and rec_num or consensus value

```{r}
df <- map_dfr(list.files(here("data_analysed"), full.names = TRUE), read_csv)
```

```{r}
ggplot(df, aes(fm_consensus, mean - fm_consensus)) +
  geom_hline(yintercept = 0) +
  geom_errorbar(aes(ymin = diff - merr, ymax = diff + merr)) +
  geom_point() +
  ggtitle("Difference in LBC-corrected vs consensus")
```

```{r}
ggplot(df, aes(fm_consensus, diff)) +
  geom_hline(yintercept = 0) +
  geom_errorbar(aes(ymin = diff - merr, ymax = diff + merr)) +
  geom_point() +
  ggtitle("Difference in LBC-corrected vs consensus")
```

```{r}
ggplot(df, aes(fm_consensus, sigma)) +
  geom_hline(yintercept = 0) +
  geom_point() +
  ggtitle("Sigma of LBC-corrected vs consensus")
```

### 2021-04-29 doc/SNEAP2021/SNEAP2021_figures.Rmd
dual bellows dillution current vs flow figure. clean up secondary standards figure

@@
 library(HybridGIS)
+library(patchwork)
 ```
@@
 
+Data from dual bellows tests
+
+```{r}
+read.csv(here("data/14Nov2017LPtest.csv"), skip = 4) %>%
+  mutate(flow = flowcalc(supply/10, 2E-5, 1.49E-5, 1.58)) %>% 
+  ggplot(aes(flow, current)) +
+  geom_smooth() + 
+  geom_point() +
+  labs(x = "Flow (ul/min)", y = "Source current (uA)")
+```
+
 # Efficiency vs flow
@@
 Yes. Calculate flow at time using concCO2
+
+```{r}
+data <- get_hgis_data(here("data/USAMS040121R.txt"), as.Date("2021-04-02")) %>% 
+  filter(Pos == 5) %>% 
+  mutate(time = cum_acqtime/60,
+         co2flow = concCO2(time, r = 244) * 30 #30ul/min
+         )
+```
+
+Sample current vs time
+
+```{r}
+# flow vs time subplot
+flow_time <- data.frame(x = 0:125) %>% 
+  ggplot(aes(x)) +
+  stat_function(fun=function(x) concCO2(x, r = 244) * 30, aes(color = "CO2")) +
+  stat_function(fun=function(x) 30 - (concCO2(x, r = 244) * 30), aes(color = "Helium")) +
+  scale_color_manual("Gas", values = c("green", "blue")) +
+  labs(title = "Gas flows to source",
+       subtitle = "7mL vial, 244uL/min helium, 30ul/min to source",
+       y = "Gas flow (uL/min)") +
+  theme(axis.title.x = element_blank(),
+        axis.text.x = element_blank())
+
+# current vs time subplot
+cur_time <- ggplot(data, aes(time, he12C)) +
+  geom_smooth(span = .4, se=FALSE, color = "green") +
+  geom_point() +
+  xlim(0, 125) +
+  labs(title = "Ion current",
+       x = "Time (min)",
+       y = "12C current (uA)")
+
+flow_time / cur_time
+```
+
+### Expected CO2 ratio/flow
+
+Plot volume of CO2 in sample vial during displacement by helium.
+It looks like optimal currents are maintained until about 2000s, which corresponds to about 2mL of CO2.
+Stable ratios until about 1.25 mL CO2.
+Can we back calculate flow to source if current falls off below ~2ul/min CO2?
+
+Add flow at time to data and plot current vs flow. 30 ul/m is estimate of capillary flow based on past results.
+
+```{r}
+ggplot(data, aes(co2flow, he12C)) +
+  geom_smooth(span = .3, se=FALSE) +
+  geom_point() +
+  labs(title = "Vial dillution current",
+       subtitle = "250uL/min displacement, 30ul/min delivery",
+       x = "CO2 flow (ul/m)",
+       y = "12C current (uA)")
+```
 
@@
 
-```{r}
-df <- map_dfr(list.files(here("data_analysed"), full.names = TRUE), read_csv)
+```{r, message = FALSE}
+df <- map_dfr(list.files(here("data_analysed"), full.names = TRUE), read_csv) %>% 
+  mutate(fm_normdiff = normFm(mean, fm_consensus),
+         fm_sigma = sigma(mean, fm_consensus, merr),
+         fmbc_normdiff = normFm(fm_lbc, fm_consensus),
+         fmbc_sigma = sigma(fm_lbc, fm_consensus, merr))
 ```
@@
 ```{r}
-ggplot(df, aes(fm_consensus, mean - fm_consensus)) +
+df %>% 
+  filter(fm_normdiff < 100) %>% 
+ggplot(aes(fm_consensus, fm_normdiff)) +
   geom_hline(yintercept = 0) +
-  geom_errorbar(aes(ymin = diff - merr, ymax = diff + merr)) +
+  geom_errorbar(aes(ymin = fm_normdiff - merr, ymax = fm_normdiff + merr)) +
   geom_point() +
-  ggtitle("Difference in LBC-corrected vs consensus")
+  ggtitle("Difference in measured Fm vs. consensus")
 ```
 
+```{r}
+df %>% 
+  filter(fm_normdiff < 1) %>% 
+ggplot(aes(fm_consensus, fm_normdiff)) +
+  geom_hline(yintercept = 0) +
+  geom_errorbar(aes(ymin = fm_normdiff - merr, ymax = fm_normdiff + merr)) +
+  geom_point() +
+  ggtitle("Difference in measured Fm vs. consensus")
+```
 ```{r}

### 2021-04-30 doc/SNEAP2021/SNEAP2021_figures.Rmd
Redo secondaries figure. convert plots to svg

@@
 df <- map_dfr(list.files(here("data_analysed"), full.names = TRUE), read_csv) %>% 
-  mutate(fm_normdiff = normFm(mean, fm_consensus),
+  filter(!is.na(rec_num)) %>% 
+  mutate(fm_diff = mean - fm_consensus,
+         fm_normdiff = normFm(mean, fm_consensus),
          fm_sigma = sigma(mean, fm_consensus, merr),
+         fmbc_diff = fm_lbc - fm_consensus,
          fmbc_normdiff = normFm(fm_lbc, fm_consensus),
@@
 ```
-```{r}
-ggplot(df, aes(fm_consensus, diff)) +
-  geom_hline(yintercept = 0) +
-  geom_errorbar(aes(ymin = diff - merr, ymax = diff + merr)) +
-  geom_point() +
-  ggtitle("Difference in LBC-corrected vs consensus")
-```
 
@@
 ```
+
+### difference from consensus
+
+```{r}
+secs <- df %>% 
+  filter(fm_diff < .1) %>% 
+ggplot(aes(fm_consensus, fm_diff, color = Cur)) +
+  geom_hline(yintercept = 0) +
+  geom_smooth(method = "lm", se = FALSE) +
+  geom_pointrange(aes(ymin = fm_diff - merr, ymax = fm_diff + merr)) +
+  geom_point() +
+  lims(x = c(0, 1.1),
+       y = c(-.15, .15)) +
+  ggtitle("Difference from consensus",
+          subtitle = "Uncorrected") +
+  theme(axis.title.x = element_blank(),
+        axis.text.x = element_blank())
+```
+
+```{r}
+secs_lbc <- df %>% 
+  filter(fm_diff < .1) %>% 
+  mutate(mean_diff = mean(fmbc_diff)) %>% 
+ggplot(aes(fm_consensus, fmbc_diff, color = Cur)) +
+  geom_hline(yintercept = 0) +
+  geom_smooth(method = "lm", se = FALSE) +
+  #geom_hline(yintercept = mean_diff) +
+  geom_pointrange(aes(ymin = fmbc_diff - merr, ymax = fmbc_diff + merr)) +
+  geom_point() +
+  lims(x = c(0, 1.1),
+       y = c(-.15, .15)) +
+  labs(subtitle = "LBC-corrected")
+```
+
+```{r}
+secs / secs_lbc
+```

### 2021-05-01 doc/SNEAP2021/SNEAP2021_figures.Rmd
rewrite inlet and blank slides. Add inlet diagram. Add current dependent blank figures

--- 2021-04-29 doc/SNEAP2021/SNEAP2021_figures.Rmd
@@
 
-combine data from 12-04, 12-11, 12-18,
+```{r}
+data<- get_hgis_data(here("data/USAMS101320R.txt"), as.Date("2020-11-17")) %>% 
+  mutate(Num = ifelse(Pos %in% c(2, 4), "S", ifelse(Num == "S", "U", Num)))
+stdrat <- mean(data$cor1412he[data$Num == "S" & !data$outlier])
+data1013 <- data %>% mutate(normFm = norm_gas(cor1412he, stdrat))
+
+data <- get_hgis_data(here("data/USAMS120320R.txt"), as.Date("2020-12-04")) %>% 
+  mutate(Num = ifelse(Pos %in% c(2, 4), "S", ifelse(Num == "S", "U", Num)))
+stdrat <- mean(data$cor1412he[data$Num == "S" & !data$outlier])
+data1204 <- data %>% mutate(normFm = norm_gas(cor1412he, stdrat))
+
+data1211 <- get_hgis_data(here("data/USAMS120320R.txt")) %>% 
+  filter(Pos %in% 1:4 | as.Date(ts) == "2020-12-11")
+
+data <- get_hgis_data(here("data/USAMS120320R.txt"), as.Date("2020-12-18")) %>% 
+  filter(Pos != 0) %>% 
+  mutate(Num = ifelse(Pos %in% c(2, 4), "S", ifelse(Num == "S", "U", Num)))
+stdrat <- mean(data$cor1412he[data$Num == "S" & !data$outlier])
+data1218 <- data %>% mutate(normFm = norm_gas(cor1412he, stdrat))
+
+data0108 <- get_hgis_data(here("data/USAMS120320R.txt"), as.Date("2021-01-08"))
+
+# no solid standards
+# df <- get_hgis_data(here("data/USAMS120320R.txt"), as.Date("2021-01-15")) 
+# stdrat <- mean(df$cor1412he[df$pos_name == "8 - LiveGasOS" & !df$outlier])
+# data0115 <- mutate(df, normFm = norm_gas(cor1412he, stdrat))
+
+data <- get_hgis_data(here("data/USAMS020521R.txt")) %>% 
+  mutate(Num = ifelse(Pos %in% c(2, 4), "S", ifelse(Num == "S", "U", Num)))
+stdrat <- mean(data$cor1412he[data$Num == "S" & !data$outlier])
+data205 <- data %>% mutate(normFm = norm_gas(cor1412he, stdrat))
+
+data <- get_hgis_data(here("data/USAMS030421R.txt"), as.Date("2021-03-05")) %>% 
+  mutate(Num = ifelse(Pos %in% c(2, 4), "S", ifelse(Num == "S", "U", Num)))
+stdrat <- mean(data$cor1412he[data$Num == "S" & !data$outlier])
+data304 <- data %>% mutate(normFm = norm_gas(cor1412he, stdrat))
+
+data <- get_hgis_data(here("data/USAMS040121R.txt"), as.Date("2021-04-09")) %>% 
+  mutate(Num = ifelse(Pos %in% c(22, 24), "S", ifelse(Num == "S", "U", Num)))
+stdrat <- mean(data$cor1412he[data$Num == "S" & !data$outlier])
+data409 <- data %>% mutate(normFm = norm_gas(cor1412he, stdrat))
+
+data <- get_hgis_data(here("data/USAMS041521R.txt"), as.Date("2021-04-16"))  %>% 
+  mutate(Num = ifelse(Pos %in% c(22, 24), "S", ifelse(Num == "S", "U", Num)))
+stdrat <- mean(data$cor1412he[data$Num == "S" & !data$outlier])
+data415 <- data %>% mutate(normFm = norm_gas(cor1412he, stdrat))
+
+std <- getStdTable()
+
+data <- rbind(data1013, data1204, data1211, data1218, data205, data304, data409, data415) %>% 
+  mutate(rec_num = case_when(str_detect(Sample.Name, "LiveGas") ~ 101730,
+                             str_starts(Sample.Name, "C-1") ~ 83028,
+                             str_starts(Sample.Name, "C1") ~ 83028,
+                             str_starts(Sample.Name, "TIRI-F") ~ 2138,
+                             str_starts(Sample.Name, "TIRI-I") ~ 17185,
+                             str_starts(Sample.Name, "C-2") ~ 1082,
+                             str_starts(Sample.Name, "C2") ~ 1082,
+                             str_starts(Sample.Name, "NOSAMS") ~ 38809,
+                             str_detect(Sample.Name, "DeadGas") ~ 72446)) %>% 
+ left_join(select(std, rec_num, fm_consensus), by = "rec_num") %>% 
+ mutate(fm_consensus = case_when(rec_num == 101730 ~ 1.0398,
+                                 rec_num == 72446 ~ 0.0013,
+                                 TRUE ~ fm_consensus))
+
+xl <- 8
+
+livefm_cur <- data %>% 
+  filter(!(he12C < 2.5 & normFm > 1.05) & fm_consensus > 0.9) %>% 
+  ggplot(aes(he12C, normFm)) + 
+  geom_point() +
+  xlim(0, xl) +
+  ylim(.8, 1.2) +
+  theme(axis.title.x = element_blank(),
+        axis.text.x = element_blank())
+
+deadfm_cur <- data %>% 
+  filter(fm_consensus < 0.1 & normFm < 0.3 & Sample.Name != "DilDeadGas6") %>% 
+  ggplot(aes(he12C, normFm)) + 
+  geom_point() +
+  xlim(0, xl) +
+  ylim(0, .4)
+
+livefm_cur / deadfm_cur
+```
+
+### Current dependent model
+
+```{r}
+ds <- sum_hgis(data) %>% 
+  filter((str_detect(Sample.Name, "LiveGas") & Cur > 1) |
+         (str_detect(Sample.Name, "DeadGas") & Cur > 1) |
+          str_detect(Sample.Name, "Blank")) %>% 
+  mutate(Gas = case_when(str_detect(Sample.Name, "LiveGas") ~ "LiveGas",
+                         str_detect(Sample.Name, "DeadGas") ~ "DeadGas",
+                         TRUE ~ "Blank"),
+         Cur_inv = 1/Cur)
+
+# Fit linear model
+fits <- ds %>% 
+  ungroup() %>% 
+  nest(data = -Gas) %>% 
+  mutate(fit = map(data, ~lm(mean ~ Cur_inv, data = .x)),
+         tidied = map(fit, tidy)) %>% 
+  unnest(tidied)
+
+blankfit <- fits %>% 
+  select(Gas, term, estimate) %>% 
+  pivot_wider(names_from = c(Gas, term), values_from = estimate) %>% 
+  mutate(inv_m_blank = -(`DeadGas_(Intercept)` - `LiveGas_(Intercept)`)/(DeadGas_Cur_inv - LiveGas_Cur_inv),
+         Fm_blank = `DeadGas_(Intercept)` + DeadGas_Cur_inv * inv_m_blank,
+         m_blank = 1/inv_m_blank)
+
+# Modeled Fm and mass of blank
+blankfit %>% 
+  select(Fm_blank, m_blank)
+
+# Plot of data and model with measured blanks
+ggplot(ds, aes(Cur_inv, mean, color = Gas)) +
+  geom_abline(slope = blankfit$LiveGas_Cur_inv, 
+              intercept = blankfit$`LiveGas_(Intercept)`) +
+  geom_abline(slope = blankfit$DeadGas_Cur_inv, 
+              intercept = blankfit$`DeadGas_(Intercept)`) +
+  geom_pointrange(aes(ymin = mean - merr, ymax = mean + merr), size = .3) + 
+  xlim(0, 10) +
+  #ylim(0, 1.2)
+  ggtitle("Fm of live and dead CO2 vs. inverse 12C current",
+          subtitle = "Fit lines should cross at Fm and current of blank")
+```
 
@@
 df <- map_dfr(list.files(here("data_analysed"), full.names = TRUE), read_csv) %>% 
-  mutate(fm_normdiff = normFm(mean, fm_consensus),
+  filter(!is.na(rec_num)) %>% 
+  mutate(fm_diff = mean - fm_consensus,
          fm_sigma = sigma(mean, fm_consensus, merr),
-         fmbc_normdiff = normFm(fm_lbc, fm_consensus),
+         fmbc_diff = fm_lbc - fm_consensus,
          fmbc_sigma = sigma(fm_lbc, fm_consensus, merr))
@@
 
-```{r}
-df %>% 
-  filter(fm_normdiff < 100) %>% 
-ggplot(aes(fm_consensus, fm_normdiff)) +
+### difference from consensus
+
+```{r, fig.height=5, dev='svg'}
+
+dfc <- df %>% 
+  filter(Cur > 2)
+
+ps <- 0.3
+dw <- 0.03
+yl <- .1
+
+secs <- ggplot(dfc, aes(fm_consensus, fm_diff, color = Cur)) +
   geom_hline(yintercept = 0) +
-  geom_errorbar(aes(ymin = fm_normdiff - merr, ymax = fm_normdiff + merr)) +
-  geom_point() +
-  ggtitle("Difference in measured Fm vs. consensus")
+  geom_smooth(method = "lm", se = FALSE) +
+  geom_pointrange(aes(ymin = fm_diff - merr, ymax = fm_diff + merr), 
+                  position = position_dodge2(width = dw),
+                  size = ps,) +
+  #geom_point() +
+  lims(x = c(-0.05, 1.1),
+       y = c(-yl, yl)) +
+  labs(title = "Difference from consensus",
+       subtitle = "Uncorrected",
+       y = "Fm difference") +
+  theme(axis.title.x = element_blank(),
+        axis.text.x = element_blank(),
+        legend.direction = "horizontal",
+        legend.position = c(.88, 1),
+        legend.justification = "bottom")
+
+secs_lbc <- ggplot(dfc, aes(fm_consensus, fmbc_diff, color = Cur)) +
+  geom_hline(yintercept = 0) +
+  geom_smooth(method = "lm", se = FALSE) +
+  #geom_hline(yintercept = mean_diff) +
+  geom_pointrange(aes(ymin = fmbc_diff - merr, ymax = fmbc_diff + merr), 
+                  position = position_dodge2(width = dw),
+                  size = ps) +
+  #geom_point() +
+  lims(x = c(-0.05, 1.1),
+       y = c(-yl, yl)) +
+  labs(subtitle = "LBC-corrected",
+       x = "Fm consensus",
+       y = "Fm difference") +
+  theme(legend.position = "None")
+
+secs / secs_lbc
 ```
-
-```{r}
-df %>% 
-  filter(fm_normdiff < 1) %>% 
-ggplot(aes(fm_consensus, fm_normdiff)) +
-  geom_hline(yintercept = 0) +
-  geom_errorbar(aes(ymin = fm_normdiff - merr, ymax = fm_normdiff + merr)) +
-  geom_point() +
-  ggtitle("Difference in measured Fm vs. consensus")
-```
-```{r}
-ggplot(df, aes(fm_consensus, diff)) +
-  geom_hline(yintercept = 0) +
-  geom_errorbar(aes(ymin = diff - merr, ymax = diff + merr)) +
-  geom_point() +
-  ggtitle("Difference in LBC-corrected vs consensus")
-```
+### same for sigma
 
@@
 ```
+
+### Blanks - current independent
+
+```{r}
+dfc
+```

### 2021-05-01 doc/SNEAP2021/SNEAP2021_figures.Rmd
Add dual bellows results. Add open split blanks fig.

@@
 Breakseal results for OX-I, C-1, OX-II
+
+```{r}
+data <- get_hgis_data(here("data/USAMS052318RHIGS.txt")) %>% 
+                  mutate(Num = ifelse(Pos == 15, "S", "U"))
+stdrat <- mean(data$cor1412he[data$Num == "S" & data$outlier == FALSE])
+data_523 <- mutate(data, normFm = norm_gas(cor1412he, stdrat))
+
+data <- get_hgis_data(here("data/USAMS053018R.txt"), as.Date("2018-05-30")) %>% 
+           mutate(Pos = ifelse(Sample.Name == "OX-II", 6, Pos),
+                  Num = ifelse(Pos == 5, "S", "U"))
+stdrat <- mean(data$cor1412he[data$Num == "S"])
+data_530 <- mutate(data, normFm = norm_gas(cor1412he, stdrat))
+
+data <- get_hgis_data(here("data/USAMS060618R.txt")) %>% 
+  filter(Pos != 1,                           #remove spurious pos1 runs
+         !(Pos == 3 & X13.12he > 0.011)) %>%     #remove solid OX-I fliers
+                  mutate(Num = ifelse(Pos == 5, "S", "U"))
+stdrat <- mean(data$cor1412he[data$Num == "S" & data$outlier == FALSE])
+data_606 <- mutate(data, normFm = norm_gas(cor1412he, stdrat))
+
+dbd <- rbind(data_523, data_530, data_606)
+
+std <- getStdTable()
+
+dbds <- sum_hgis(dbd) %>% 
+  mutate(rec_num = case_when(str_detect(Sample.Name, "Live") ~ 101730,
+                             Sample.Name == "OX-I" ~ 34148,
+                             Sample.Name == "OX-II" ~ 34149,
+                             str_starts(Sample.Name, "C1") ~ 83028)) %>% 
+ left_join(select(std, rec_num, fm_consensus), by = "rec_num") %>% 
+ mutate(fm_consensus = case_when(rec_num == 101730 ~ 1.0398,
+                                 rec_num == 72446 ~ 0.0013,
+                                 TRUE ~ fm_consensus),
+        fm_diff = mean - fm_consensus)
+
+dbds %>% 
+  filter(Cur < 10) %>% 
+  ggplot(aes(fm_consensus, fm_diff)) +
+  geom_hline(yintercept = 0) +
+  geom_pointrange(aes(ymin = fm_diff - merr, ymax = fm_diff + merr),
+                 position = position_dodge2(width = .1),
+                 size = .5) +
+  labs(title = "Closed inlet")
+```
 
@@
 ```{r}
-dfc
+df %>% filter(fm_consensus < 0.1 &
+              Cur > 2) %>% 
+  ggplot(aes(Cur, mean)) +
+  geom_errorbar(aes(ymin = mean - merr, ymax = mean + merr)) +
+  geom_errorbarh(aes(xmin = Cur - Cur.sd, xmax = Cur + Cur.sd)) +
+  geom_point(aes(color = wheel), size = 3) +
+  labs(title = "Blanks",
+       x = "Current (uA)",
+       y = "Fraction modern") +
+  theme(legend.position = c(0.85, 0.8),
+        legend.background = element_rect(fill = "white", color = "black")) 
 ```

### 2021-05-02 doc/SNEAP2021/SNEAP2021_HGIS_presentation/SNEAP2021_HGIS_presentation.Rmd
Move presentation and required files to separate directory

---
title: "Progress with the NOSAMS"
subtitle: "Hybrid Gas Ion Source"
author: "Brett Longworth"
institute: "NOSAMS/WHOI"
date: "2021/05/03"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.showtext = TRUE)

# Libraries
library(tidyverse)
library(xaringanthemer)
library(here)
library(HybridGIS)
library(patchwork)
library(amstools)
library(gt)
library(knitr)
library(broom)

# Colors
dark_blue <- "#041e42"
light_blue <- "#00a9e0"
medium_blue <- "0069b1"
aqua_blue <- "#00b7bd"
cool_gray_11 <- "#53565a"
cool_gray_4 <- "#bbbcbc"
background <- "#e6e7e8"

style_mono_light(
  # colors
  base_color = dark_blue,
  # primary_color = background,
  # secondary_color = dark_blue,

  # fonts
  header_font_google = google_font("Montserrat", "800"),
  text_font_google = google_font("Red Hat Display", "500")
)
style_extra_css(
  css = list(
    ".title-slide" = list(
      "background-image" = "url(images/whoi_seal_title_bkgnd.jpg)",
      "background-size" = "cover"
    )
  ),
  append = TRUE,
  heading = "Extra CSS"
)

```

# Why run gas instead of graphite?

.pull-left[
### Benefits

- Skip a time-consuming reduction to graphite
- Eliminate blank associated with graphitization

### Challenges

- Ultimate precision is lower
- Lower currents/counts (time to precision)
- variability in blank (current dependent)
- variability in sample performance
]

.pull-right[

<br>
<br>
```{r method-table}
methods <- tribble(~" ", ~"Graphite", ~"CFAMS ECR GIS", ~"Hybrid GIS",
                   "Continous flow", "No", "Yes", "No",
                   "Measurement time", "30 min", "5 min", "15-30 min",
                   "Precision", "0.3%", "1%", "1%",
                   "Background", "0.1 pmc", "1 pmc", "1 pmc")
methods %>% 
  kable(format = "html",
        caption = "Comparison of AMS methods")

```
]

???

Performance = currents counts stability efficiency time

---

# How do you sputter a gas?

### $$CO_2+e^- \longrightarrow C^-+2O^-$$

<br>
.pull-left[
* Inject CO<sub>2</sub> gas onto Ti target
* Sputter with Cs<sup>+</sup>
* Dissociate molecules
* Produce C<sup>-</sup> ions (and O<sup>-</sup>, CH<sup>-</sup>, OH<sup>-</sup>, CO<sup>-</sup>...)
* Extract C<sup>-</sup> beam
]

.pull-right[
![Salazar, 2013](images/salazar_hgis_target_flow.png)
.right[Salazar, 2013]
]

???

introduction onto Ti surface in target recess

Predicted chemistry/physics of decomposition and ionization

---

# The NEC HGIS

.pull-left[
* Introduce gas to back of cathode
* Rest is gas handling
<br>
<br>

```{r hgis-arm, echo=FALSE, fig.align='center', out.width="80%"}
knitr::include_graphics("images/hgis_arm_source.jpeg")
```
]

.pull-right[
.center[
<img src="images/HGIS_arm_drawing.svg" width="60%" />
]
]

???

Figure/photo of NEC bellows system

---

# Targets

.pull-left[
Potential for:

* Lower cost
* Better performance
]

.pull-right[
![HGIS Target](images/target_cutaway.png)
NEC hybrid target cutaway
]

---

# First tests

.pull-left[
### Conditions
* Modern and dead CO<sub>2</sub>
* 8m x 50 μm capillary
* 0-4 atm supply pressure

### Results
* ~10μA <sup>12</sup>C<sup>-</sup>
* <1% precision (SE of repeat measurements)
* 1% blank
* 1-3% total efficiency
]

.pull-right[
```{r cur-flow, fig.height=6, dev='svg'}
read.csv(here("data/14Nov2017LPtest.csv"), skip = 4) %>%
  mutate(flow = flowcalc(supply/10, 2E-5, 1.49E-5, 1.58)) %>% 
  ggplot(aes(flow, current)) +
  geom_smooth(se = FALSE, span = 0.7, color = dark_blue) + 
  geom_point() +
  labs(title = "Current vs gas flow",
          x = bquote('CO'[2]~'Flow (μl/min)'), 
          y = "12C current (μA)") +
  theme_classic() +
  theme_xaringan(title_font_size = 20,
                 text_font_size = 16,
                 background_color = "#ffffff")
```
]

---

# Inlet/interface systems

.pull-left[
Adapter between CO<sub>2</sub> producers and source.

Goal: constant flow of CO<sub>2</sub> to source at ideal rate.

Increase flow by
* increasing pCO<sub>2</sub>
* reducing volume

Gas mixtures - partial pressure
* adding another gas increases pressure and total flow
* reduces pCO<sub>2</sub>
* CO<sub>2</sub> remains constant? No- viscosity
]

.pull-right[

<br>
### $$PV=nRT$$

![](images/inlet_flow_diag.svg)

]

???

Pressure of CO<sub>2</sub> in inlet controls flow in capillary

P = flow = current

* increase partial pressure of CO<sub>2</sub> by dilution?
Gas inlet controls flow of CO<sub>2</sub> to source by varying pressure, volume, and gas composition.

Dilution with helium

Makes moving small amounts of gas easier by increasing pressure

Source doesn't care about helium as long as it gets its CO<sub>2</sub>

Introduce steady, reproducible current as key to data quality
Constant flow of CO<sub>2</sub> to source produces constant currents, which reduces ratio variability.

---

.pull-left[
# Closed inlet

### Flow control
* Change volume
* Add CO<sub>2</sub>
* Add Helium

### For a given sample size:
* High inlet volume = low inlet pressure (NEC inlet)
* Low inlet volume = high inlet pressure (Ionplus inlet)

<br>
Smallest samples require minimizing inlet volume and pressure while maintaining good flow control
]

.pull-right.center[
<br><br><br>
<img src="images/dual_bellows.png" width="80%" />
]

???

* Mechanically more complex
* Extra sample transfer time
    * but can be done during burn-in
* Data/performance simpler

---

# NOSAMS Closed inlet

.pull-left[
#### IRMS dual bellows

* VG Prism dual bellows
* LabVIEW interface
* 2 x bellows
* Cryotraps
* Manifold capable

![IRMS inlet diagram](images/HGIS_dual_bellows_drawing.svg)
]

.pull-right[
```{r dual-bellows-results, fig.height=6, fig.width=7.5, dev='svg'}
data <- get_hgis_data(here("data/USAMS052318RHIGS.txt")) %>% 
                  mutate(Num = ifelse(Pos == 15, "S", "U"))
stdrat <- mean(data$cor1412he[data$Num == "S" & data$outlier == FALSE])
data_523 <- mutate(data, normFm = norm_gas(cor1412he, stdrat))

data <- get_hgis_data(here("data/USAMS053018R.txt"), as.Date("2018-05-30")) %>% 
           mutate(Pos = ifelse(Sample.Name == "OX-II", 6, Pos),
                  Num = ifelse(Pos == 5, "S", "U"))
stdrat <- mean(data$cor1412he[data$Num == "S"])
data_530 <- mutate(data, normFm = norm_gas(cor1412he, stdrat))

data <- get_hgis_data(here("data/USAMS060618R.txt")) %>% 
  filter(Pos != 1,                           #remove spurious pos1 runs
         !(Pos == 3 & X13.12he > 0.011)) %>%     #remove solid OX-I fliers
                  mutate(Num = ifelse(Pos == 5, "S", "U"))
stdrat <- mean(data$cor1412he[data$Num == "S" & data$outlier == FALSE])
data_606 <- mutate(data, normFm = norm_gas(cor1412he, stdrat))

dbd <- rbind(data_523, data_530, data_606)

std <- getStdTable()

dbds <- sum_hgis(dbd) %>% 
  mutate(rec_num = case_when(str_detect(Sample.Name, "Live") ~ 101730,
                             Sample.Name == "OX-I" ~ 34148,
                             Sample.Name == "OX-II" ~ 34149,
                             str_starts(Sample.Name, "C1") ~ 83028)) %>% 
 left_join(select(std, rec_num, fm_consensus), by = "rec_num") %>% 
 mutate(fm_consensus = case_when(rec_num == 101730 ~ 1.0398,
                                 rec_num == 72446 ~ 0.0013,
                                 TRUE ~ fm_consensus),
        fm_diff = mean - fm_consensus)

dbds %>% 
  filter(Cur < 10) %>% 
  ggplot(aes(fm_consensus, fm_diff)) +
  geom_hline(yintercept = 0) +
  geom_pointrange(aes(ymin = fm_diff - merr, ymax = fm_diff + merr),
                 position = position_dodge2(width = .1),
                 size = .5) +
  labs(title = "Closed inlet",
       x = "Expected Fm",
       y = "Fm difference") +
  theme_classic() +
  theme_xaringan(title_font_size = 20,
                 text_font_size = 16,
                 background_color = "#ffffff")
```
]

---

.pull-left[
# Open split

P and V are held constant, so only mixture controls pCO<sub>2</sub>

Use helium to displace CO<sub>2</sub> from vials into split.

Capillary to source controls gas flow

Open inlets maintain constant atmospheric pressure, producing constant<sup>1</sup> flow to source

![open split diagram](images/HGIS_open_split_drawing.svg)

.footnote[ [1] flow varies by composition due to differing viscosity]
]

.pull-right.center[
<br><br><br>
<img src="images/gas_control.jpg" width="75%" />
]

???

* Mechanically/operationally simple
* Harder to control current variability
    * during run
    * if sample size differs

---

# Carbonate method

.pull-left[
#### Sample preparation

* Evacuate vial
* Add acid to carbonate
* Fills vial with CO<sub>2</sub> at 1 atm

#### Measurement

* Pre-sputter target
* Displace sample 
* Measurement
]

.pull-right[
.center[![gilson autosampler](images/gilson.png)]
]

???

table: comparison of carbonate methods
figure: DAG graph of process

---

.pull-left[
# Currents and flows

Helium displacement of CO<sub>2</sub> from vial. Monitor currents and ratios as pCO<sub>2</sub> decreases.

* No current suppresion at high CO<sub>2</sub>
* Current stable until 40 min
* Corresponds to 2mL CO<sub>2</sub>

Balance flow rates to maximize stable currents and minimize sample size
]

.pull-right[
```{r cur-time-os, fig.height=8, dev='svg'}
data <- get_hgis_data(here("data/USAMS040121R.txt"), as.Date("2021-04-02")) %>% 
  filter(Pos == 5) %>% 
  mutate(time = cum_acqtime/60,
         co2flow = concCO2(time, r = 244) * 30) #30ul/min

# flow vs time subplot
flow_time <- data.frame(x = 0:125) %>% 
  ggplot(aes(x)) +
  stat_function(fun=function(x) concCO2(x, r = 244) * 30, aes(color = "CO2")) +
  stat_function(fun=function(x) 30 - (concCO2(x, r = 244) * 30), aes(color = "Helium")) +
  scale_color_manual("Gas", values = c("green", "blue")) +
  labs(title = "Gas flows to source",
       subtitle = "7mL vial, 244μL/min helium, 30μL/min to source",
       y = "Gas flow (μL/min)") +
  theme_classic() +
  theme_xaringan(title_font_size = 20,
                 text_font_size = 18,
                 background_color = "#ffffff") +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        legend.position = c(0.87, 0.65),
        legend.background = element_rect(fill = "white", color = "black"))

# current vs time subplot
cur_time <- ggplot(data, aes(time, he12C)) +
  geom_smooth(span = .4, se=FALSE, color = "green") +
  geom_point() +
  xlim(0, 125) +
  labs(title = "Ion current",
       x = "Time (min)",
       y = "12C current (μA)") +
  theme_classic() +
  theme_xaringan(title_font_size = 20,
                 text_font_size = 18,
                 background_color = "#ffffff")

flow_time / cur_time
```
]

---

# Currents and flows

.pull-left[
* Note similarity to first tests
* Flat top means stable currents/ratios for long enough to measure
* Sample size can be reduced to 2mL
* Further reduction with increased flow to source
]

.pull-right[
```{r cur-flow-os, fig.height=6, dev='svg'}
ggplot(data, aes(co2flow, he12C)) +
  geom_smooth(span = .3, se=FALSE, color = medium_blue) +
  geom_point() +
  labs(title = "Vial dillution current",
       subtitle = "250μL/min displacement, 30μl/min delivery",
       x = bquote('CO'[2]~'Flow (μl/min)'), 
       y = "12C current (μA)") +
  theme_classic() +
  theme_xaringan(title_font_size = 20,
                 text_font_size = 16,
                 background_color = "#ffffff")
```
]

???

compare to first tests

Note that our flat-topped current curve allows variable flow of the vial system. MICADAS has a more pronounced peak so this may not work as well. (Check papers for peak shape)

Also note current dependent ratio (blank) as a complicating factor

Lead into blank on next slide

---

```{r cur-blank-data, fig.height = 4, dev='svg'}
data<- get_hgis_data(here("data/USAMS101320R.txt"), as.Date("2020-11-17")) %>% 
  mutate(Num = ifelse(Pos %in% c(2, 4), "S", ifelse(Num == "S", "U", Num)))
stdrat <- mean(data$cor1412he[data$Num == "S" & !data$outlier])
data1013 <- data %>% mutate(normFm = norm_gas(cor1412he, stdrat))

data <- get_hgis_data(here("data/USAMS120320R.txt"), as.Date("2020-12-04")) %>% 
  mutate(Num = ifelse(Pos %in% c(2, 4), "S", ifelse(Num == "S", "U", Num)))
stdrat <- mean(data$cor1412he[data$Num == "S" & !data$outlier])
data1204 <- data %>% mutate(normFm = norm_gas(cor1412he, stdrat))

data1211 <- get_hgis_data(here("data/USAMS120320R.txt")) %>% 
  filter(Pos %in% 1:4 | as.Date(ts) == "2020-12-11")

data <- get_hgis_data(here("data/USAMS120320R.txt"), as.Date("2020-12-18")) %>% 
  filter(Pos != 0) %>% 
  mutate(Num = ifelse(Pos %in% c(2, 4), "S", ifelse(Num == "S", "U", Num)))
stdrat <- mean(data$cor1412he[data$Num == "S" & !data$outlier])
data1218 <- data %>% mutate(normFm = norm_gas(cor1412he, stdrat))

data0108 <- get_hgis_data(here("data/USAMS120320R.txt"), as.Date("2021-01-08"))

data <- get_hgis_data(here("data/USAMS020521R.txt")) %>% 
  mutate(Num = ifelse(Pos %in% c(2, 4), "S", ifelse(Num == "S", "U", Num)))
stdrat <- mean(data$cor1412he[data$Num == "S" & !data$outlier])
data205 <- data %>% mutate(normFm = norm_gas(cor1412he, stdrat))

data <- get_hgis_data(here("data/USAMS030421R.txt"), as.Date("2021-03-05")) %>% 
  mutate(Num = ifelse(Pos %in% c(2, 4), "S", ifelse(Num == "S", "U", Num)))
stdrat <- mean(data$cor1412he[data$Num == "S" & !data$outlier])
data304 <- data %>% mutate(normFm = norm_gas(cor1412he, stdrat))

data <- get_hgis_data(here("data/USAMS040121R.txt"), as.Date("2021-04-09")) %>% 
  mutate(Num = ifelse(Pos %in% c(22, 24), "S", ifelse(Num == "S", "U", Num)))
stdrat <- mean(data$cor1412he[data$Num == "S" & !data$outlier])
data409 <- data %>% mutate(normFm = norm_gas(cor1412he, stdrat))

data <- get_hgis_data(here("data/USAMS041521R.txt"), as.Date("2021-04-16"))  %>% 
  mutate(Num = ifelse(Pos %in% c(22, 24), "S", ifelse(Num == "S", "U", Num)))
stdrat <- mean(data$cor1412he[data$Num == "S" & !data$outlier])
data415 <- data %>% mutate(normFm = norm_gas(cor1412he, stdrat))

std <- getStdTable()

data <- rbind(data1013, data1204, data1211, data1218, data205, data304, data409, data415) %>% 
  mutate(rec_num = case_when(str_detect(Sample.Name, "LiveGas") ~ 101730,
                             str_starts(Sample.Name, "C-1") ~ 83028,
                             str_starts(Sample.Name, "C1") ~ 83028,
                             str_starts(Sample.Name, "TIRI-F") ~ 2138,
                             str_starts(Sample.Name, "TIRI-I") ~ 17185,
                             str_starts(Sample.Name, "C-2") ~ 1082,
                             str_starts(Sample.Name, "C2") ~ 1082,
                             str_starts(Sample.Name, "NOSAMS") ~ 38809,
                             str_detect(Sample.Name, "DeadGas") ~ 72446)) %>% 
 left_join(select(std, rec_num, fm_consensus), by = "rec_num") %>% 
 mutate(fm_consensus = case_when(rec_num == 101730 ~ 1.0398,
                                 rec_num == 72446 ~ 0.0013,
                                 TRUE ~ fm_consensus))

# Summarize by target
ds <- sum_hgis(data) %>% 
  filter((str_detect(Sample.Name, "LiveGas") & Cur > 1) |
         (str_detect(Sample.Name, "DeadGas") & Cur > 1) |
          str_detect(Sample.Name, "Blank")) %>% 
  mutate(Gas = case_when(str_detect(Sample.Name, "LiveGas") ~ "LiveGas",
                         str_detect(Sample.Name, "DeadGas") ~ "DeadGas",
                         TRUE ~ "Blank"),
         Cur_inv = 1/Cur)

# Fit linear model
fits <- ds %>% 
  ungroup() %>% 
  nest(data = -Gas) %>% 
  mutate(fit = map(data, ~lm(mean ~ Cur_inv, data = .x)),
         tidied = map(fit, tidy)) %>% 
  unnest(tidied)

blankfit <- fits %>% 
  select(Gas, term, estimate) %>% 
  pivot_wider(names_from = c(Gas, term), values_from = estimate) %>% 
  mutate(inv_m_blank = -(`DeadGas_(Intercept)` - `LiveGas_(Intercept)`)/(DeadGas_Cur_inv - LiveGas_Cur_inv),
         Fm_blank = `DeadGas_(Intercept)` + DeadGas_Cur_inv * inv_m_blank,
         m_blank = 1/inv_m_blank)

# Modeled Fm and mass of blank
blank_val <- blankfit %>% 
  select(Fm_blank, m_blank)
```

.pull-left[

# Current dependent blank

$$I_{sample} + I_{blank} = I_{meas}$$
$$Fm_m=Fm_bI_b+Fm_sI_s$$

$$Fm_s = Fm_m + \frac{I_b(Fm_m-Fm_b)}{I_m-I_b}$$

0.1μA blank current with 10μA total means 1% blank

Modeled blank:
* Fm: `r format(blank_val$Fm_blank, digits = 3)`
* Current: `r format(blank_val$m_blank, digits = 3)` μA

Measured blank:
* Fm: `r ds %>% filter(Gas == "Blank") %>% pull(mean) %>% mean() %>% format(digits = 3)`
* Current: `r ds %>% filter(Gas == "Blank") %>% pull(Cur) %>% mean() %>% format(digits = 3)` μA

]

.pull-right[

```{r ratio-cur, fig.height = 4, dev='svg'}
xl <- 8

livefm_cur <- data %>% 
  filter(!(he12C < 2.5 & normFm > 1.05) & fm_consensus > 0.9) %>% 
  ggplot(aes(he12C, normFm)) + 
  geom_point() +
  xlim(0, xl) +
  ylim(.8, 1.2) +
  ggtitle("Ratio vs current") +
  theme_classic() +
  theme_xaringan(title_font_size = 20,
                 text_font_size = 18,
                 background_color = "#ffffff") +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_blank())

deadfm_cur <- data %>% 
  filter(fm_consensus < 0.1 & normFm < 0.3 & Sample.Name != "DilDeadGas6") %>% 
  ggplot(aes(he12C, normFm)) + 
  geom_point() +
  xlim(0, xl) +
  ylim(0, .4) +
  xlab("12C Current (μA)") +
  theme_classic() +
  theme_xaringan(title_font_size = 20,
                 text_font_size = 18,
                 background_color = "#ffffff")

(livefm_cur / deadfm_cur) +
  ylab("Fraction modern") +
  theme(axis.title.y = element_text(hjust=-1)) 
```

```{r blank-model, fig.height = 4, dev='svg'}
# Plot of data and model with measured blanks
ggplot(ds, aes(Cur_inv, mean, color = Gas)) +
  geom_abline(slope = blankfit$LiveGas_Cur_inv, 
              intercept = blankfit$`LiveGas_(Intercept)`) +
  geom_abline(slope = blankfit$DeadGas_Cur_inv, 
              intercept = blankfit$`DeadGas_(Intercept)`) +
  geom_pointrange(aes(ymin = mean - merr, ymax = mean + merr), size = .3) + 
  xlim(0, 10) +
  labs(title = "Modelled blank",
       subtitle = "Fits cross at Fm and current of blank",
       x = "Inverse 12C Current (μA-1)",
       y = "Fraction modern") +
  theme_classic() +
  theme_xaringan(title_font_size = 20,
                 text_font_size = 16,
                 background_color = "#ffffff") +
  theme(legend.position = "none",
        legend.background = element_rect(fill = "white", color = "black")) 
```
]

???

Blank = Ion current not from sample gas

Targets alone and targets + helium (fm and current)

...but if current is kept constant, current drops out of correction
---

```{r carb-data}
carb_data <- map_dfr(list.files(here("data_analysed"), full.names = TRUE), read_csv) %>% 
  filter(!is.na(rec_num),
         Cur > 2) %>% 
  mutate(fm_diff = mean - fm_consensus,
         fm_sigma = sigma(mean, fm_consensus, merr),
         fmbc_diff = fm_lbc - fm_consensus,
         fmbc_sigma = sigma(fm_lbc, fm_consensus, merr))
```

# Current independent blank

.pull-left[
If standards, blanks, and unknowns have same current:

linear fit used to correct

$$Fm_s = Fm_m - Fm_{blk}\frac{Fm_{std} - Fm_m}{Fm_{std}}$$

#### Mean Fm of blanks:
`r carb_data %>% filter(fm_consensus < 0.1 & Cur > 2) %>% pull(mean) %>% mean() %>% format(digits = 3)`

]

.pull-right[

```{r blanks, fig.height=5, dev='svg'}
carb_data %>% filter(fm_consensus < 0.1 &
              Cur > 2) %>% 
  ggplot(aes(Cur, mean)) +
  geom_errorbar(aes(ymin = mean - merr, ymax = mean + merr)) +
  geom_errorbarh(aes(xmin = Cur - Cur.sd, xmax = Cur + Cur.sd)) +
  geom_point(aes(color = wheel), size = 3) +
  labs(title = "Blanks",
       x = "12C Current (μA)",
       y = "Fraction modern") +
  theme_classic() +
  theme_xaringan(title_font_size = 20,
                 text_font_size = 16,
                 background_color = "#ffffff") +
  theme(legend.position = c(0.82, 0.76),
        legend.background = element_rect(fill = "white", color = "black")) 
```

]

???

## Current independent correction

Current/target dependent?
Constant blanks and currents allow single mass independent correction

3 ways to control blank
* Reduce blank - better cleaning, presputtering
* Hold standard, blank, unknown current constant and use current independent correction (linear correction based on dead sample Fm)   
* Determine Fm and current of blank and use current dependent correction (current balance model)
Idea: use data from dillution runs of live and dead gas to create curves

It's relatively easy to produce runs with terrible current, but usually not good for blank as caused by other issues.

---

.left-column[
# Secondary standards

* Currents > 2μA
* Note current dependence

### Data summary
* mean difference: `r format(mean(carb_data$fmbc_diff), digits = 1)`
* sd of difference: `r format(sd(carb_data$fmbc_diff), digits = 2)`

]

.right-column[
.right[
```{r secondaries, fig.height=6, fig.width=7.5, dev='svg'}

ps <- 0.3
dw <- 0.03
yl <- .1

secs <- ggplot(carb_data, aes(fm_consensus, fm_diff, color = Cur)) +
  geom_hline(yintercept = 0) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_pointrange(aes(ymin = fm_diff - merr, ymax = fm_diff + merr), 
                  position = position_dodge2(width = dw),
                  size = ps,) +
  lims(x = c(-0.05, 1.1),
       y = c(-yl, yl)) +
  labs(title = "Difference from expected Fm",
       subtitle = "Uncorrected",
       y = "Fm difference") +
  theme_classic() +
  theme_xaringan(title_font_size = 20,
                 text_font_size = 16,
                 background_color = "#ffffff") +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        legend.direction = "horizontal",
        legend.position = c(.86, 1),
        legend.justification = "bottom")

secs_lbc <- ggplot(carb_data, aes(fm_consensus, fmbc_diff, color = Cur)) +
  geom_hline(yintercept = 0) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_pointrange(aes(ymin = fmbc_diff - merr, ymax = fmbc_diff + merr), 
                  position = position_dodge2(width = dw),
                  size = ps) +
  lims(x = c(-0.05, 1.1),
       y = c(-yl, yl)) +
  labs(subtitle = "Blank corrected",
       x = "Fm expected",
       y = "Fm difference") +
  theme_classic() +
  theme_xaringan(title_font_size = 20,
                 text_font_size = 16,
                 background_color = "#ffffff") +
  theme(legend.position = "None")

secs / secs_lbc
```
]
]

???

precision, accuracy
issues with blank and normalization

---

# Sample size reduction

Open split defines pressure, therefore

smaller samples = higher dilution = higher flow

plot: sample size vs. required flow

pressure control to maintain constant CO<sub>2</sub> to source

figure: setup with pressure control

---

# MICADAS

.pull-left[
### Coming soon(ish)...

Source differences
* one target in source at a time - lower blank
* gas supplied via conical joint in side of target

Inlet differences
* MICADAS uses closed inlet
* CO<sub>2</sub> from carbonates transferred onto resin trap in stream of helium
* Sample introduced by adding helium and adjusting flow using variable volume
]

.pull-right[![](images/MICADAS_HGIS_interface.jpeg)]

???

More pronounced current vs. flow curve makes measurement more difficult at variable CO<sub>2</sub> flow.
Salazar and Szidat are developing an open split interface with EA - squish peak with dead volume

photo: micadas source

---

# Summary

NOSAMS has developed a successful method for analysing carbonates with an open-split interface delivering gas to a NEC MCSNICS gas-accepting Cesium sputter ion source.

Current goals:
* finish characterization of the method
* Comparison to graphite with test samples
* Improve accuracy
* Better understand method blank
* Reduce sample size

---

# Thank You!

Mark Roberts, Josh Burton, Josh Hlavenka, Al Gagnon, Kalina Gospodinova, Mark Kurz, and the staff of NOSAMS

Karl von Reden, Ann McNichol

Cam McIntyre, SUERC

Ted Ognibene, CAMS

<br>
<br>

<table style="width:100%">
<tr>
<td>
<img src="images/NOSAMS_logo_transparent.png">
</td>
<td>
<img src="images/WHOI_PrimaryLogo_DarkBlueType_RGB.png">
</td>
<td>
<img src="images/NSF_4-Color_bitmap_Logo.png">
</td>
</tr>
</table>

### 2021-05-02 doc/SNEAP2021/SNEAP2021_HGIS_presentation/SNEAP2021_HGIS_presentation.Rmd
Another round of edits.

@@
 ![Salazar, 2013](images/salazar_hgis_target_flow.png)
-.right[Salazar, 2013]
+.right[Gas density in HGIS Cathode
+
+(Salazar, 2013)]
 ]
@@
 .pull-left[
+
+* Pneumatic arm brings gas to source
 * Introduce gas to back of cathode
-* Rest is gas handling
+* The rest is gas handling
 <br>
@@
 
-```{r hgis-arm, echo=FALSE, fig.align='center', out.width="80%"}
-knitr::include_graphics("images/hgis_arm_source.jpeg")
-```
+.center[<img src="images/hgis_arm_source.jpeg" width="80%" />]
+.right[MCSNICS source with gas arm]
 ]
@@
 ![HGIS Target](images/target_cutaway.png)
-NEC hybrid target cutaway
+.right[NEC hybrid target cutaway]
 ]
@@
 ### Results
-* ~10μA <sup>12</sup>C<sup>-</sup>
+* ~8 μA <sup>12</sup>C<sup>-</sup>
 * <1% precision (SE of repeat measurements)
@@
   geom_point() +
-  labs(title = "Current vs gas flow",
+  labs(title = "Ion current and gas flow",
           x = bquote('CO'[2]~'Flow (μl/min)'), 
@@
 
-Increase flow by
-* increasing pCO<sub>2</sub>
-* reducing volume
+Increase CO<sub>2</sub> flow by increasing pCO<sub>2</sub>:
+* Add CO<sub>2</sub> 
+* Reduce volume
 
-Gas mixtures - partial pressure
-* adding another gas increases pressure and total flow
-* reduces pCO<sub>2</sub>
-* CO<sub>2</sub> remains constant? No- viscosity
+Gas mixtures - helium dilution:
+* Increase pressure and total flow
+* Reduce pCO<sub>2</sub>
+* CO<sub>2</sub> flow remains constant? No - viscosity
 ]
@@
 <img src="images/dual_bellows.png" width="80%" />
+
+.right[Dual-bellows closed inlet]
 ]
@@
 
-* VG Prism dual bellows
-* LabVIEW interface
-* 2 x bellows
-* Cryotraps
+* VG Prism inlet
+* Bellows for volume control
+* Cryotraps to move and isolate CO<sub>2</sub>
+* LabVIEW control
 * Manifold capable
@@
 ![IRMS inlet diagram](images/HGIS_dual_bellows_drawing.svg)
+.right[Closed inlet schematic]
 ]
@@
 
-P and V are held constant, so only mixture controls pCO<sub>2</sub>
+* Pressure is held constant at 1 atm
+* Gas mixture controls pCO<sub>2</sub>
+* Helium displaces CO<sub>2</sub> from vials into split
+* Capillary to source controls gas flow
 
-Use helium to displace CO<sub>2</sub> from vials into split.
-
-Capillary to source controls gas flow
-
-Open inlets maintain constant atmospheric pressure, producing constant<sup>1</sup> flow to source
+Open inlets maintain constant atmospheric pressure, producing constant<sup>1</sup> flow to source.
 
 ![open split diagram](images/HGIS_open_split_drawing.svg)
-
-.footnote[ [1] flow varies by composition due to differing viscosity]
 ]
@@
 .pull-right.center[
-<br><br><br>
+<br><br>
 <img src="images/gas_control.jpg" width="75%" />
+.right[Open inlet setup]
 ]
+
+.footnote[[1] flow varies by composition due to differing viscosity]
 
@@
 
+* Weigh carbonate into vial
 * Evacuate vial
 * Add acid to carbonate
-* Fills vial with CO<sub>2</sub> at 1 atm
+* Reaction fills vial with CO<sub>2</sub> at 1 atm
 
@@
 * Pre-sputter target
-* Displace sample 
-* Measurement
+* Displace sample from vial to split with He 
+* Measure ratios for 15-30 min.
 ]
@@
 .center[![gilson autosampler](images/gilson.png)]
+.right[Automated gas handling system]
 ]
@@
 
-* No current suppresion at high CO<sub>2</sub>
+* No current suppression at high pCO<sub>2</sub>
 * Current stable until 40 min
-* Corresponds to 2mL CO<sub>2</sub>
-
-Balance flow rates to maximize stable currents and minimize sample size
+* 40 min ≈ 2 mL CO<sub>2</sub>
 ]
@@
 .pull-right[
-```{r cur-time-os, fig.height=8, dev='svg'}
+<br>
+<br>
+```{r cur-time-os, fig.height=7, dev='svg'}
 data <- get_hgis_data(here("data/USAMS040121R.txt"), as.Date("2021-04-02")) %>% 
@@
 
+.pull-left[
 # Currents and flows
 
-.pull-left[
 * Note similarity to first tests
 * Flat top means stable currents/ratios for long enough to measure
-* Sample size can be reduced to 2mL
+* Sample size can be reduced to 2 mL
 * Further reduction with increased flow to source
+
+### Balance flow rates to maximize stable currents and minimize sample size
 ]
@@
 .pull-right[
+<br>
+<br>
+<br>
 ```{r cur-flow-os, fig.height=6, dev='svg'}
@@
 
-0.1μA blank current with 10μA total means 1% blank
+0.1 μA blank current with 10 μA total = 1% blank
 
@@
 .pull-left[
-If standards, blanks, and unknowns have same current:
+If **standards**, **blanks**, and **unknowns** have the **same** blank and current:
 
-linear fit used to correct
+### Linear fit used to correct:
 
@@
 
-#### Mean Fm of blanks:
-`r carb_data %>% filter(fm_consensus < 0.1 & Cur > 2) %>% pull(mean) %>% mean() %>% format(digits = 3)`
+### Mean Fm of blanks: `r carb_data %>% filter(fm_consensus < 0.1 & Cur > 2) %>% pull(mean) %>% mean() %>% format(digits = 3)`
 
@@
 
-* Currents > 2μA
+* Currents > 2 μA
 * Note current dependence
@@
 
-# Sample size reduction
-
-Open split defines pressure, therefore
-
-smaller samples = higher dilution = higher flow
-
-plot: sample size vs. required flow
-
-pressure control to maintain constant CO<sub>2</sub> to source
-
-figure: setup with pressure control
-
----
-
 # MICADAS
@@
 
-Source differences
+#### Source differences
 * one target in source at a time - lower blank
@@
 
-Inlet differences
+#### Inlet differences
 * MICADAS uses closed inlet
@@
 
-.pull-right[![](images/MICADAS_HGIS_interface.jpeg)]
+.pull-right[
+![](images/MICADAS_HGIS_interface.jpeg)
+.right[MICADAS Gas Interface]
+]
 
@@
 
-NOSAMS has developed a successful method for analysing carbonates with an open-split interface delivering gas to a NEC MCSNICS gas-accepting Cesium sputter ion source.
+NOSAMS has developed a successful method for analyzing radiocarbon in carbonates with an open-split interface delivering CO<sub>2</sub> gas to a NEC MCSNICS gas-accepting cesium sputter ion source.
 
 Current goals:
-* finish characterization of the method
-* Comparison to graphite with test samples
+* Finish characterization of the method
+* Compare to graphite with test samples
 * Improve accuracy
@@
 * Reduce sample size
+* Automation of data acquisition and reduction
 
@@
 </table>
+
+---
+
+# Sample size reduction
+
+Open split defines pressure, therefore
+
+smaller samples = higher dilution = higher flow
+
+plot: sample size vs. required flow
+
+pressure control to maintain constant CO<sub>2</sub> to source
+
+figure: setup with pressure control

### 2021-05-02 doc/SNEAP2021/SNEAP2021_HGIS_presentation/SNEAP2021_HGIS_presentation.Rmd
theming plots

@@
   base_color = dark_blue,
+  white_color = lighten_color("#d9d9d6", .5),
   # primary_color = background,
@@
   ggplot(aes(flow, current)) +
-  geom_smooth(se = FALSE, span = 0.7, color = dark_blue) + 
-  geom_point() +
+  geom_smooth(se = FALSE, span = 0.7, color = "#00b7bd") + 
+  geom_point(size = 4, color = "#0069b1") +
   labs(title = "Ion current and gas flow",

### 2021-05-02 doc/SNEAP2021/SNEAP2021_HGIS_presentation/SNEAP2021_HGIS_presentation.Rmd
Update plot colors

@@
                  position = position_dodge2(width = .1),
-                 size = .5) +
+                 size = .6,
+                 color = "#0069b1") +
   labs(title = "Closed inlet",
@@
   ggplot(aes(x)) +
-  stat_function(fun=function(x) concCO2(x, r = 244) * 30, aes(color = "CO2")) +
-  stat_function(fun=function(x) 30 - (concCO2(x, r = 244) * 30), aes(color = "Helium")) +
-  scale_color_manual("Gas", values = c("green", "blue")) +
+  stat_function(fun=function(x) concCO2(x, r = 244) * 30, aes(color = "CO2"), size = 1.5) +
+  stat_function(fun=function(x) 30 - (concCO2(x, r = 244) * 30), aes(color = "Helium"), size = 1.5) +
+  scale_color_manual("Gas", values = c("#00b7bd", "#b7bf10")) +
   labs(title = "Gas flows to source",
@@
 cur_time <- ggplot(data, aes(time, he12C)) +
-  geom_smooth(span = .4, se=FALSE, color = "green") +
-  geom_point() +
+  geom_smooth(span = .4, se=FALSE, color = "#00b7bd") +
+  geom_point(size = 3, color = "#0069b1") +
   xlim(0, 125) +
@@
 ggplot(data, aes(co2flow, he12C)) +
-  geom_smooth(span = .3, se=FALSE, color = medium_blue) +
-  geom_point() +
+  geom_smooth(span = .3, se=FALSE, color = "#00b7bd") +
+  geom_point(size = 3, color = "#0069b1") +
   labs(title = "Vial dillution current",
@@
   ggplot(aes(he12C, normFm)) + 
-  geom_point() +
+  geom_point(size = 2, color = "#0069b1") +
   xlim(0, xl) +
@@
   ggplot(aes(he12C, normFm)) + 
-  geom_point() +
+  geom_point(size = 2, color = "#00a9e0") +
   xlim(0, xl) +
@@
   geom_abline(slope = blankfit$LiveGas_Cur_inv, 
-              intercept = blankfit$`LiveGas_(Intercept)`) +
+              intercept = blankfit$`LiveGas_(Intercept)`,
+              color = "#0069b1") +
   geom_abline(slope = blankfit$DeadGas_Cur_inv, 
-              intercept = blankfit$`DeadGas_(Intercept)`) +
+              intercept = blankfit$`DeadGas_(Intercept)`,
+              color = "#00a9e0") +
   geom_pointrange(aes(ymin = mean - merr, ymax = mean + merr), size = .3) + 
+  scale_color_manual("Gas", values = c("#b7bf10", "#00a9e0", "#0069b1")) +
   xlim(0, 10) +
@@
         legend.background = element_rect(fill = "white", color = "black"))
+
+factor(ds$Gas)
 ```
@@
   geom_point(aes(color = wheel), size = 3) +
+  scale_color_manual("wheel", values = c("#b7bf10", "#00a9e0", "#0069b1")) +
   labs(title = "Blanks",

### 2021-05-02 doc/SNEAP2021/SNEAP2021_HGIS_presentation/SNEAP2021_HGIS_presentation.Rmd
Minor fixes

@@
 - Skip a time-consuming reduction to graphite
+- Rapid iteration for science/development
 - Eliminate blank associated with graphitization
@@
 ### Conditions
+* Pure CO<sub>2</sub> to source
 * Modern and dead CO<sub>2</sub>
-* 8m x 50 μm capillary
+* 8 m x 50 μm capillary
 * 0-4 atm supply pressure
@@
 * CO<sub>2</sub> flow remains constant? No - viscosity
+* He is 25% more viscous than CO<sub>2</sub>
 ]
@@
 
-<br>
 ### $$PV=nRT$$
@@
 
+### $$Q=\frac{\Delta P \pi R^4}{8\mu L}$$
 ]
@@
 P = flow = current
+
+1/viscosity = flow
 
@@
 * Measure ratios for 15-30 min.
+
 ]
@@
 .center[![gilson autosampler](images/gilson.png)]
-.right[Automated gas handling system]
+.right[Automated gas handling system (Roberts, 2013)
+]
 ]
@@
 * Flat top means stable currents/ratios for long enough to measure
-* Sample size can be reduced to 2 mL
+* Sample size can be reduced to 2 mL (~1 mg C)
 * Further reduction with increased flow to source
@@
   geom_point(size = 3, color = "#0069b1") +
-  labs(title = "Vial dillution current",
+  labs(title = "Vial dilution current",
        subtitle = "250μL/min displacement, 30μl/min delivery",
@@
 $$I_{sample} + I_{blank} = I_{meas}$$
-$$Fm_m=Fm_bI_b+Fm_sI_s$$
+$$Fm_mI_m=Fm_bI_b+Fm_sI_s$$
 
@@
   xlim(0, 10) +
-  labs(title = "Modelled blank",
+  labs(title = "Modeled blank",
        subtitle = "Fits cross at Fm and current of blank",
@@
         legend.background = element_rect(fill = "white", color = "black")) 
-
-factor(ds$Gas)
 ```
@@
 
-Karl von Reden, Ann McNichol
+Karl von Reden
 
@@
 </table>
-
----
-
-# Sample size reduction
-
-Open split defines pressure, therefore
-
-smaller samples = higher dilution = higher flow
-
-plot: sample size vs. required flow
-
-pressure control to maintain constant CO<sub>2</sub> to source
-
-figure: setup with pressure control

### 2021-05-02 doc/SNEAP2021/SNEAP2021_HGIS_presentation/SNEAP2021_HGIS_presentation.Rmd
Reorder first slides. Minor edits.

@@
 ---
-title: "Progress with the NOSAMS"
-subtitle: "Hybrid Gas Ion Source"
+title: "Progress with the NOSAMS<br>Hybrid Gas Ion Source"
 author: "Brett Longworth"
@@
 
+# How do you sputter a gas?
+
+### $$CO_2+e^- \longrightarrow C^-+2O^-$$
+
+.pull-left[
+* Inject CO<sub>2</sub> gas onto Ti target
+* Sputter with Cs<sup>+</sup>
+* Dissociate molecules
+* Produce C<sup>-</sup> ions (and O<sup>-</sup>, CH<sup>-</sup>, OH<sup>-</sup>, CO<sup>-</sup>...)
+* Extract C<sup>-</sup> beam
+]
+
+.pull-right[
+![Salazar, 2013](images/salazar_hgis_target_flow.png)
+.right[Gas density in HGIS Cathode
+
+(Salazar, 2013)]
+]
+
+???
+
+introduction onto Ti surface in target recess
+
+Predicted chemistry/physics of decomposition and ionization
+
+Plasma in target recess? Blue glow.
+
+---
+
 # Why run gas instead of graphite?
@@
 - Lower currents/counts (time to precision)
-- variability in blank (current dependent)
-- variability in sample performance
+- Variability in blank (current dependent)
+- Variability in sample performance
 ]
@@
 ???
-
-Performance = currents counts stability efficiency time
-
----
-
-# How do you sputter a gas?
-
-### $$CO_2+e^- \longrightarrow C^-+2O^-$$
-
-<br>
-.pull-left[
-* Inject CO<sub>2</sub> gas onto Ti target
-* Sputter with Cs<sup>+</sup>
-* Dissociate molecules
-* Produce C<sup>-</sup> ions (and O<sup>-</sup>, CH<sup>-</sup>, OH<sup>-</sup>, CO<sup>-</sup>...)
-* Extract C<sup>-</sup> beam
-]
-
-.pull-right[
-![Salazar, 2013](images/salazar_hgis_target_flow.png)
-.right[Gas density in HGIS Cathode
-
-(Salazar, 2013)]
-]
-
-???
-
-introduction onto Ti surface in target recess
-
-Predicted chemistry/physics of decomposition and ionization
 
@@
 
-Figure/photo of NEC bellows system
-
 ---
@@
 .pull-left[
-Potential for:
+### Goal:
 
+* Maximize gas density and residence time where sputtering and ionization are happening.
+
+### Potential for:
+
+* Better performance
 * Lower cost
-* Better performance
 ]
@@
 
+???
+
+CAMS and ETH have both done a lot of target geometry research.
+
+In the end, compromise between simplicity and function.
+
+Maybe still gains to be made. 
 ---
@@
 ]
+
+???
+
+Bottle gas allowed testing full range of flows
+
+Precision not accuracy: modern gas normalized to itself
 
@@
 
-* increase partial pressure of CO<sub>2</sub> by dilution?
 Gas inlet controls flow of CO<sub>2</sub> to source by varying pressure, volume, and gas composition.
 
-Dilution with helium
+increase partial pressure of CO<sub>2</sub> by dilution?
 
-Makes moving small amounts of gas easier by increasing pressure
+Dilution Makes moving small amounts of gas easier by increasing pressure
 
@@
 Introduce steady, reproducible current as key to data quality
+
 Constant flow of CO<sub>2</sub> to source produces constant currents, which reduces ratio variability.
@@
 * Add CO<sub>2</sub>
-* Add Helium
+* Add helium
 
@@
 
-<br>
-Smallest samples require minimizing inlet volume and pressure while maintaining good flow control
 ]
@@
 
+Smallest samples require minimizing inlet volume and pressure while maintaining good flow control
 ---
 
-# NOSAMS Closed inlet
+# NOSAMS closed inlet
 
@@
 * Note similarity to first tests
-* Flat top means stable currents/ratios for long enough to measure
+* Flat top = stable currents/ratios 
 * Sample size can be reduced to 2 mL (~1 mg C)
@@
 
-### Balance flow rates to maximize stable currents and minimize sample size
+#### Balance flow rates to maximize stable currents and minimize sample size
 ]
@@
 .pull-right[
+<br>
 <br>
@@
 
-Modeled blank:
+.pull-left[
+#### Modeled blank:
 * Fm: `r format(blank_val$Fm_blank, digits = 3)`
 * Current: `r format(blank_val$m_blank, digits = 3)` μA
+]
 
-Measured blank:
+.pull-right[
+#### Measured blank:
 * Fm: `r ds %>% filter(Gas == "Blank") %>% pull(mean) %>% mean() %>% format(digits = 3)`
 * Current: `r ds %>% filter(Gas == "Blank") %>% pull(Cur) %>% mean() %>% format(digits = 3)` μA
-
+]
 ]
@@
 
-### Mean Fm of blanks: `r carb_data %>% filter(fm_consensus < 0.1 & Cur > 2) %>% pull(mean) %>% mean() %>% format(digits = 3)`
+#### Mean Fm of blanks: `r carb_data %>% filter(fm_consensus < 0.1 & Cur > 2) %>% pull(mean) %>% mean() %>% format(digits = 3)`
 
@@
 ???
+
+We'll see what fit looks like on next slide.
 
@@
        subtitle = "Uncorrected",
+       colour = "Cur (μA)",
        y = "Fm difference") +
   theme_classic() +
-  theme_xaringan(title_font_size = 20,
+  theme_xaringan(title_font_size = 18,
                  text_font_size = 16,
@@
         legend.direction = "horizontal",
-        legend.position = c(.86, 1),
+        legend.position = c(.84, 1),
         legend.justification = "bottom")
@@
   theme_classic() +
-  theme_xaringan(title_font_size = 20,
+  theme_xaringan(title_font_size = 18,
                  text_font_size = 16,
@@
 #### Source differences
-* one target in source at a time - lower blank
-* gas supplied via conical joint in side of target
+* One target in source at a time - lower blank
+* Gas supplied via conical joint in side of target
 
@@
 
-Current goals:
+###Goals:
 * Finish characterization of the method
@@
 * Reduce sample size
-* Automation of data acquisition and reduction
+* Automate data acquisition and reduction
 

### 2021-05-02 doc/SNEAP2021/SNEAP2021_HGIS_presentation/SNEAP2021_HGIS_presentation.Rmd
Tweak figure legend

@@
         legend.direction = "horizontal",
-        legend.position = c(.84, 1),
+        legend.position = c(.80, 1),
         legend.justification = "bottom")

### 2021-05-03 doc/SNEAP2021/SNEAP2021_HGIS_presentation/SNEAP2021_HGIS_presentation.Rmd
Minor edits.

@@
 - Skip a time-consuming reduction to graphite
-- Rapid iteration for science/development
+- Rapid iteration for science and development
 - Eliminate blank associated with graphitization
@@
 - Lower currents/counts (time to precision)
-- Variability in blank (current dependent)
-- Variability in sample performance
+- Variable blank (current dependent)
+- Reproducible sample performance
 ]
@@
 ### Data summary
-* mean difference: `r format(mean(carb_data$fmbc_diff), digits = 1)`
-* sd of difference: `r format(sd(carb_data$fmbc_diff), digits = 2)`
+* Mean difference: `r format(mean(carb_data$fmbc_diff), digits = 1)`
+* Standard deviation of difference: `r format(sd(carb_data$fmbc_diff), digits = 2)`
 
@@
 
-###Goals:
+### Goals:
 * Finish characterization of the method

### 2021-05-03 doc/SNEAP2021/SNEAP2021_HGIS_presentation/SNEAP2021_HGIS_presentation.Rmd
Minor edits

@@
 .pull-left[
-### Goal:
+### Goal
 
-* Maximize gas density and residence time where sputtering and ionization are happening.
+* Maximize gas density and residence time where sputtering and ionization are happening
 
-### Potential for:
+### Potential for
 
@@
 
-Increase CO<sub>2</sub> flow by increasing pCO<sub>2</sub>:
+#### Increase CO<sub>2</sub> flow by increasing pCO<sub>2</sub>
 * Add CO<sub>2</sub> 
@@
 
-Gas mixtures - helium dilution:
+#### Gas mixtures - helium dilution
 * Increase pressure and total flow
@@
 
-### Flow control
+#### Flow control
 * Change volume
@@
 
-### For a given sample size:
+#### For a given sample size
 * High inlet volume = low inlet pressure (NEC inlet)
@@
 
-Helium displacement of CO<sub>2</sub> from vial. Monitor currents and ratios as pCO<sub>2</sub> decreases.
+Helium displacement of CO<sub>2</sub> from vial.
+
+Monitor currents and ratios as pCO<sub>2</sub> decreases.
 
@@
 .pull-left[
-#### Modeled blank:
+#### Modeled blank
 * Fm: `r format(blank_val$Fm_blank, digits = 3)`
@@
 .pull-right[
-#### Measured blank:
+#### Measured blank
 * Fm: `r ds %>% filter(Gas == "Blank") %>% pull(mean) %>% mean() %>% format(digits = 3)`
@@
 .pull-left[
-If **standards**, **blanks**, and **unknowns** have the **same** blank and current:
+If **standards**, **blanks**, and **unknowns** have the **same** blank and current, current drops out.
 
-### Linear fit used to correct:
+#### Linear fit used to correct:
 
@@
 * Currents > 2 μA
-* Note current dependence
+* Current dependence
+* Underestimation of error
 
@@
 
-NOSAMS has developed a successful method for analyzing radiocarbon in carbonates with an open-split interface delivering CO<sub>2</sub> gas to a NEC MCSNICS gas-accepting cesium sputter ion source.
+**NOSAMS has developed a successful method for analyzing radiocarbon in carbonates with an open-split interface delivering CO<sub>2</sub> gas to a NEC MCSNICS gas-accepting cesium sputter ion source.**
 

### 2021-05-03 doc/SNEAP2021/SNEAP2021_HGIS_presentation/SNEAP2021_HGIS_presentation.Rmd
Fix image, minor edits.

@@
 title: "Progress with the NOSAMS<br>Hybrid Gas Ion Source"
-author: "Brett Longworth"
+author: "Brett Longworth, Josh Burton,<br>Josh Hlavenka, and Mark Roberts"
 institute: "NOSAMS/WHOI"
-date: "2021/05/03"
+date: "May 3, 2021"
 output:
@@
 
-#### Flow control
+### Flow control
 * Change volume
@@
 
-#### For a given sample size
+### For a given sample size
 * High inlet volume = low inlet pressure (NEC inlet)

### 2021-05-21 doc/SNEAP2021/SNEAP2021_figures.Rmd
Lots of small formatting fixes
add changes from final presentation. Improve sigma plot.

@@
 library(patchwork)
+library(broom)
 ```
@@
 
-```{r, message = FALSE}
-df <- map_dfr(list.files(here("data_analysed"), full.names = TRUE), read_csv) %>% 
-  filter(!is.na(rec_num)) %>% 
+```{r carb-data}
+carb_data <- map_dfr(list.files(here("data_analysed"), full.names = TRUE), read_csv) %>% 
+  filter(!is.na(rec_num),
+         Cur > 2) %>% 
   mutate(fm_diff = mean - fm_consensus,
@@
 
-### difference from consensus
+### Current independent blank
 
-```{r, fig.height=5, dev='svg'}
+#### Linear fit used to correct:
 
-dfc <- df %>% 
-  filter(Cur > 2)
+$$Fm_s = Fm_m - Fm_{blk}\frac{Fm_{std} - Fm_m}{Fm_{std}}$$
+
+Mean Fm of blanks: `r carb_data %>% filter(fm_consensus < 0.1 & Cur > 2) %>% pull(mean) %>% mean() %>% format(digits = 3)`
+
+```{r blanks, fig.height=5, dev='svg'}
+carb_data %>% filter(fm_consensus < 0.1 &
+              Cur > 2) %>% 
+  ggplot(aes(Cur, mean)) +
+  geom_errorbar(aes(ymin = mean - merr, ymax = mean + merr)) +
+  geom_errorbarh(aes(xmin = Cur - Cur.sd, xmax = Cur + Cur.sd)) +
+  geom_point(aes(color = wheel), size = 3) +
+  scale_color_manual("wheel", values = c("#b7bf10", "#00a9e0", "#0069b1")) +
+  labs(title = "Blanks",
+       x = "12C Current (μA)",
+       y = "Fraction modern") +
+  theme_classic() +
+  theme(legend.position = c(0.82, 0.76),
+        legend.background = element_rect(fill = "white", color = "black")) 
+```
+
+### Secondary standards
+
+* Currents > 2 μA
+* Current dependence
+* Underestimation of error
+
+#### Data summary
+* Mean difference: `r format(mean(carb_data$fmbc_diff), digits = 1)`
+* Standard deviation of difference: `r format(sd(carb_data$fmbc_diff), digits = 2)`
+
+```{r secondaries, fig.height=6, fig.width=7.5, dev='svg'}
 
@@
 
-secs <- ggplot(dfc, aes(fm_consensus, fm_diff, color = Cur)) +
+secs <- ggplot(carb_data, aes(fm_consensus, fm_diff, color = Cur)) +
   geom_hline(yintercept = 0) +
@@
                   size = ps,) +
-  #geom_point() +
   lims(x = c(-0.05, 1.1),
        y = c(-yl, yl)) +
-  labs(title = "Difference from consensus",
+  labs(title = "Difference from expected Fm",
        subtitle = "Uncorrected",
+       colour = "Cur (μA)",
        y = "Fm difference") +
+  theme_classic() +
   theme(axis.title.x = element_blank(),
@@
         legend.direction = "horizontal",
-        legend.position = c(.88, 1),
+        legend.position = c(.80, 1),
         legend.justification = "bottom")
 
-secs_lbc <- ggplot(dfc, aes(fm_consensus, fmbc_diff, color = Cur)) +
+secs_lbc <- ggplot(carb_data, aes(fm_consensus, fmbc_diff, color = Cur)) +
   geom_hline(yintercept = 0) +
   geom_smooth(method = "lm", se = FALSE) +
-  #geom_hline(yintercept = mean_diff) +
   geom_pointrange(aes(ymin = fmbc_diff - merr, ymax = fmbc_diff + merr), 
@@
                   size = ps) +
-  #geom_point() +
   lims(x = c(-0.05, 1.1),
        y = c(-yl, yl)) +
-  labs(subtitle = "LBC-corrected",
-       x = "Fm consensus",
+  labs(subtitle = "Blank corrected",
+       x = "Fm expected",
        y = "Fm difference") +
+  theme_classic() +
   theme(legend.position = "None")
@@
 ```
+
 ### same for sigma
 
+```{r secsigma, fig.height=6, fig.width=7.5, dev='svg'}
+
+ps <- 0.3
+dw <- 0.03
+yl <- .1
+
+secs <- ggplot(carb_data, aes(fm_consensus, fm_sigma, color = Cur)) +
+  geom_hline(yintercept = 0) +
+  geom_smooth(method = "lm", se = FALSE) +
+  geom_point(position = position_dodge2(width = dw)) +
+  #lims(x = c(-0.05, 1.1),
+  #     y = c(-yl, yl)) +
+  labs(title = "Sigma vs. expected Fm",
+       subtitle = "Uncorrected",
+       colour = "Cur (μA)",
+       y = "Sigma") +
+  theme_classic() +
+  theme(axis.title.x = element_blank(),
+        axis.text.x = element_blank(),
+        legend.direction = "horizontal",
+        legend.position = c(.80, 1),
+        legend.justification = "bottom")
+
+secs_lbc <- ggplot(carb_data, aes(fm_consensus, fmbc_sigma, color = Cur)) +
+  geom_hline(yintercept = 0) +
+  geom_smooth(method = "lm", se = FALSE) +
+  geom_point(position = position_dodge2(width = dw)) +
+  #lims(x = c(-0.05, 1.1),
+  #     y = c(-yl, yl)) +
+  labs(subtitle = "Blank corrected",
+       x = "Fm expected",
+       y = "Sigma") +
+  theme_classic() +
+  theme(legend.position = "None")
+
+secs / secs_lbc
+```
+
 ```{r}
-ggplot(df, aes(fm_consensus, sigma)) +
+ggplot(carb_data, aes(fm_consensus, sigma)) +
   geom_hline(yintercept = 0) +
@@
 
-### Blanks - current independent
+### Data summary
 
-```{r}
-df %>% filter(fm_consensus < 0.1 &
-              Cur > 2) %>% 
-  ggplot(aes(Cur, mean)) +
-  geom_errorbar(aes(ymin = mean - merr, ymax = mean + merr)) +
-  geom_errorbarh(aes(xmin = Cur - Cur.sd, xmax = Cur + Cur.sd)) +
-  geom_point(aes(color = wheel), size = 3) +
-  labs(title = "Blanks",
-       x = "Current (uA)",
-       y = "Fraction modern") +
-  theme(legend.position = c(0.85, 0.8),
-        legend.background = element_rect(fill = "white", color = "black")) 
-```
+* Mean difference: `r format(mean(carb_data$fmbc_diff), digits = 1)`
+* Standard deviation of difference: `r format(sd(carb_data$fmbc_diff), digits = 2)`
+* Number of samples: `r length(carb_data$fmbc_diff)`

### 2021-06-06 doc/hgis_data_reduction.Rmd
Describe data reduction.

---
title: "Data reduction for HGIS"
output: html_notebook
---

This document describes the procedure and functions used for reducing, normalizing, and blank correcting NOSAMS hybrid gas ion source data.

Data reduction proceeds as follows:

### Read raw file

Raw files are acquisition data from the AMS system, in the same format as graphite data. A typical HGIS data acquisition for a target consists of 10 sequential runs of 90s each, giving 15 min. of acquisition time per sample.

### Apply 13C correction to each run of a target

14/12 ratios of each run are corrected for fractionation using 13/12 ratio measured on the AMS system. Livetime correction is done during acquisition on USAMS. The correction is done as follows:

$$ {}^{14/12}C_{corr} = \frac{{}^{14/12}C_{meas}}{{}^{13/12}C_{meas}^2} $$

### Flag outliers by interquartile distance

Outliers are defined as points 1.5 * interquartile range above the 3rd quartile or below the 1st quartile, and are dropped from the analysis.

### Determine per-target means, internal and external errors

The per-target 13C-corrected raw ratio is taken as the unweighted mean of runs of the target. The internal or counting error of a target is defined as:

$$ \sigma_{int} = \frac{1}{\sqrt{\sum{counts}}} $$

The external or statistical error is defined as:

$$ \sigma_{ext} = \sqrt{\frac{}} $$

### Normalize targets using mean of standards and propagate errors

Unknowns are normalized using the mean measured value of gas standards using this equation:

$$ R_{norm} = \frac{R_{sample}}{R_{std}}*1.0398$$

Normalization errors are propagated using this equation:

$$ \sigma_{norm} = \sqrt{\frac{\sigma_{sample}^2}{R_{sample}^2} + \frac{\sigma_{std}^2}{R_{std}^2}} $$

### Apply a large blank correction using the mean of blanks and propagate errors.

The mean of blanks is used to apply a large blank correction to the unknowns. The error in the blanks is taken as the larger of the standard deviation of blanks or half of the mean blank Fm. The correction is done as follows:

$$ R_{cor} = R_{norm} - \frac{R_{blank}(R_{std} - R_{norm})}{R_{std}} $$

The error in the blank is propagated using this equation:

$$ \sigma_{cor} = \sqrt{\sigma_{norm}^2 * \frac{1 + R_{blank}}{R_{std}}^2 + \sigma_{blank}^2 * \frac{R_{norm}-R_{std}}{R_{std}}^2} $$

### 2021-06-09 doc/SNEAP2021/SNEAP2021_figures.Rmd
Add summary table

@@
 library(broom)
+library(gt)
 ```
@@
 ```{r carb-data}
-carb_data <- map_dfr(list.files(here("data_analysed"), full.names = TRUE), read_csv) %>% 
+carb_data <- map_dfr(list.files(here("data_analysed"), pattern = "USAMS", full.names = TRUE), read_csv) %>% 
   filter(!is.na(rec_num),
@@
 * Number of samples: `r length(carb_data$fmbc_diff)`
+
+### Data table
+
+```{r}
+carb_data_sum <- carb_data %>% 
+  select(-dil_factor, -sd, -exterr, -interr, -diff, -normFm, -sigma) %>% 
+  rename(norm_ratio = mean,
+         sig_norm_ratio = merr)
+write_csv(carb_data_sum, here("data_analysed/carb_data_summary.csv"))
+```
+Carb data summary by type
+
+```{r}
+carb_data_sum %>% 
+  mutate(Name = case_when(rec_num == 101730 ~ "LiveGas",
+                          rec_num == 83028 ~ "C-1",
+                          rec_num == 2138 ~ "TIRI-F",
+                          rec_num == 17185 ~ "TIRI-I",
+                          rec_num == 1082 ~ "C-2",
+                          rec_num == 38809 ~ "NOSAMS2",
+                          rec_num == 72446 ~ "DeadGas")) %>% 
+  select(rec_num, Name, norm_ratio, fm_lbc, sig_norm_ratio) %>% 
+  group_by(Name) %>% 
+  summarize(across(c(norm_ratio, fm_lbc, sig_norm_ratio), c(mean = mean, sd = sd)), N = n()) %>%
+  arrange(norm_ratio_mean) %>% 
+  gt() %>% 
+  fmt_number(columns = 2:7,
+             decimals = 4)
+```

### 2021-06-10 doc/hgis_data_reduction.Rmd
Fix external error.
knit to html
Fix blank error propagation eqn

@@
 title: "Data reduction for HGIS"
-output: html_notebook
+output:
+  html_document:
+    df_print: paged
 ---
@@
 
-$$ \sigma_{ext} = \sqrt{\frac{}} $$
+$$ \sigma_{ext} = \frac{SE(R_{cor})}{mean(R_{cor})}$$
 
@@
 
-$$ \sigma_{cor} = \sqrt{\sigma_{norm}^2 * \frac{1 + R_{blank}}{R_{std}}^2 + \sigma_{blank}^2 * \frac{R_{norm}-R_{std}}{R_{std}}^2} $$
+$$ \sigma_{cor} = \sqrt{\sigma_{norm}^2 * \left(\frac{1 + R_{blank}}{R_{std}}\right)^2 + \sigma_{blank}^2 * \left(\frac{R_{norm}-R_{std}}{R_{std}}\right)^2} $$

### 2021-06-10 inst/rmd/HGIS_pendleton_2021-05-28.Rmd
Move Rmd's to inst/Rmd

---
title: "HGIS 2021-05-28 Carbonates"
author: "Brett Longworth"
date: "6/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

```{r}
library(tidyverse)
library(here)
library(openxlsx)
library(amstools)
library(HybridGIS)

options(digits = 4)
options(scipen = 3)
```

Data reduction and analysis for carbonate samples run on May 28, 2021. Includes C-1, C-2, and test shell carbonates from Simon Pendleton.

```{r}

#Raw AMS data results file path
file <- here("data/USAMS052621R.txt")

# Date samples run (needed to separate other experiments with this wheel/targets)
date <- as.Date("2021-05-28")

# Positions of normalizing standards
standards <- c(5, 8, 13, 17)

# Positions of blanks for blank correction
blanks <- c(6, 9, 11)
```

- Using standards in positions `r standards` for normalization.
- Using blanks in positions `r blanks` for blank correction.
- Outliers are defined as points 1.5 * interquartile range above the 3rd quartile or below the 1st quartile, and are dropped from the analysis.

Data reduction proceeds as follows:

1. Read raw file
1. Apply 13C correction to each run of a target
1. Flag outliers by interquartile distance
1. Determine per-target means, internal and external errors
1. Normalize targets using mean of standards and propagate errors
1. Apply a large blank correction using the mean of blanks and propagate errors.

## raw results

```{r}
df <- get_hgis_data(file, date, standards)
write_csv(df, here("data_analysed/HGIS_2021-05-28_raw.csv"))
write.xlsx(df, here("data_analysed/HGIS_2021-05-28_raw.xlsx"))
```

#### Currents and ratios vs time.

```{r}
df %>% 
  filter(!(Pos %in% 1:4)) %>% 
  plot_hgis_time(normFm, ce*normFm) +
  ggtitle("Sample ratio vs. time")
```

```{r}
df %>% 
  filter(!(Pos %in% 1:4)) %>% 
  plot_hgis_time(he12C) +
  ggtitle("Sample current vs time")
```

```{r}
df_sum <- sum_hgis_targets(df) %>% 
  norm_hgis() %>% 
  blank_cor_hgis(blanks = c(6, 9, 11))

write_csv(df_sum, here("data_analysed/HGIS_2021-05-28_results.csv"))
write.xlsx(df_sum, here("data_analysed/HGIS_2021-05-28_results.xlsx"))
```

## Results

```{r}
df_sum %>% 
  filter(as.numeric(Pos) > 4) %>% 
  mutate(sig_normFm = max_err * normFm) %>% 
  select(Pos, Sample.Name, he12C, n_runs, fm_corr, sig_fm_corr) %>% 
  arrange(Sample.Name) %>% 
  knitr::kable()
  # gt::gt()
```

```{r}
df_sum %>% 
  filter(as.numeric(Pos) > 4) %>% 
plot_hgis_summary()
```

## Agreement of replicates

Compare results for samples with more than one replicate.

```{r}
df_sum %>% 
  filter(as.numeric(Pos) > 4) %>% 
  compare_replicates() %>% 
  knitr::kable()
```

## Agreement with consensus

```{r}
cons <- df_sum %>% 
  filter(as.numeric(Pos) > 4,
         !is.na(fm_consensus)) %>% 
  select(Sample.Name, fm_consensus, fm_corr, sig_fm_corr) %>% 
  mutate(Fm_diff = fm_corr - fm_consensus,
         sigma = amstools::sigma(fm_corr, fm_consensus, sig_fm_corr)) %>% 
  arrange(Sample.Name)

knitr::kable(cons)
```

### 2021-06-10 inst/rmd/HGIS_small_samples.Rmd
Move Rmd's to inst/Rmd

---
title: "HGIS small(er) samples"
author: "Brett Longworth"
date: "6/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE}
# Load libraries and functions
library(tidyverse)
library(amstools)
library(here)
library(HybridGIS)
```

## 2021-06-04 Small(er) sample tests

Tested 5, 10, and 15 mg of C-2 carbonate in vials backfilled with Helium.

### Method

1. Weigh sample into vial
2. Evacuate and acidify as normal with gilson
3. Backfill to 1 ATM with He-filled syringe 2h after acidification.

Allowed vacuum in vial to pull in He from syringe, then added 1 additional mL He. Should calculate volume CO2 in vial and measure He added to check that things make sense.

Used same capillaries as usual, but decreased He displacement flow to 140 ul/min. Source capillary flow assumed to be around 30ul/min.

### Samples

```{r}
carb_data <- tribble(~vial, ~name, ~rec_num, ~mass,
                     1, "C-2_1", 1082, 6.03,
                     2, "C-2_2", 1082, 11.10,
                     3, "C-2_3", 1082, 15.52
                     )

write_csv(carb_data, here("data/carb_data_2021-06-04.csv"))
carb_data
```

### Data

```{r}
df <- get_hgis_data(here("data/USAMS052621R.txt"), as.Date("2021-06-04")) #, standards)
```

#### Currents and ratios vs time.

```{r}
df %>% 
  filter(!(Pos %in% 1:4)) %>% 
  plot_hgis_time(normFm, ce*normFm) +
  ggtitle("Sample ratio vs. time")
```

```{r}
df %>% 
  filter(!(Pos %in% 1:4)) %>% 
  plot_hgis_time(he12C) +
  ggtitle("Sample current vs time")
```

### Results

Note: no blanks run, so need to use average blank if LBC needed.

```{r}
df_sum <- sum_hgis_targets(df) %>% 
  norm_hgis()
```

```{r}
df_sum %>% 
  filter(as.numeric(Pos) > 4) %>% 
  mutate(sig_normFm = max_err * normFm) %>% 
  select(Pos, Sample.Name, he12C, n_runs, norm_ratio, sig_norm_ratio) %>% 
  arrange(Sample.Name)
```

```{r}
df_sum %>% 
  filter(as.numeric(Pos) > 4) %>% 
  ggplot(aes(Sample.Name, norm_ratio)) +
    geom_pointrange(aes(ymin = norm_ratio - sig_norm_ratio, ymax = norm_ratio + sig_norm_ratio),
                    size = .5) + 
    theme(axis.text.x = element_text(angle = 45, hjust=1)) +
      labs(x = NULL,
           y = "Fraction Modern")
```

## Agreement of replicates

Compare results for samples with more than one replicate.

```{r}
df_sum %>% 
  filter(as.numeric(Pos) > 4) %>% 
  compare_replicates()
```

## Agreement with consensus

```{r}
df_sum %>% 
  filter(as.numeric(Pos) > 4,
         !is.na(fm_consensus)) %>% 
  select(Sample.Name, fm_consensus, norm_ratio, sig_norm_ratio) %>% 
  mutate(Fm_diff = norm_ratio - fm_consensus,
         sigma = amstools::sigma(norm_ratio, fm_consensus, sig_norm_ratio)) %>% 
  arrange(Sample.Name)
```

### 2021-06-10 inst/rmd/RC2018Figures.Rmd
Move Rmd's to inst/Rmd

---
title: "Figures for Radiocarbon 2018 Poster"
output:
  pdf_document: default
  html_notebook: default
  html_document: default
---

```{r libraries, warning=FALSE, message=FALSE}
library(amstools)
library(tidyverse)
library(ggplot2)
```

## Calculated flow

To model laminar flow in a capillary, Poiseuille's equation is used.

$$
Q = \frac{\Delta P \pi R^4}{8 \mu \Delta x}
$$

Where $\Delta P$ is change in pressure (Pa ($\frac{kg}{m \cdot s^2}$)), $R$ is
radius of pipe (m), $\mu$ is dynamic viscosity ( $P \cdot s$ ($\frac{kg}{s
\cdot m}$)), and $\Delta x$ is the length of the pipe (m). Q is in
$\frac{m^3}{s}$

The following functions calculate flow using this equation and convert base
units to more convenient flow in $\frac{\mu L}{m}$ and pressure in kPa.

```{r}
# Poiseuille's; convert delta pressure in Pa to flow in m^3/s
prestoflow <- function(dp, r, u, l) {
  # r = radius of capillary
  # l = length of capillary
  # u = viscosity kg/m
  flow <- dp * pi * r^4 / 8 / u / l
  flow
}

# convert gauge kPa to uL/min
flowcalc <- function(pres, ...) {
  dppa <- pres * 1E3 #convert kPa to Pa
  flowcms <- prestoflow(dppa, ...)
  flowuls <- flowcms * 1E6 * 1E3
  flowulm <- flowuls * 60
  flowulm
}
```

```{r}
# Efficiency functions
# convert L CO2 to atoms C
lCO2toatC <- function (vol) {
  molC <- vol * 1/22.4 # 22.4 L of gas per mole at stp
  atC <- molC * 6.022E23
  atC
}

# convert current in A to atoms C/sec
CcurtoatC <- function(cur, cs){
  at12C <- cur * 6.25E18 # electrons per coulomb
  at12C <- at12C / cs # divide by charge state
  at12C
}

# calculate efficiency from ml/m CO2 and uA C12
hgis_eff <- function(ulm, uA, cs = 3) {

  # if flow is zero or less, efficiency doesn't make sense
  ulm <- ifelse(ulm <= 0, NA, ulm)

  vol <- ulm * 1E-6 # conv to L
  cur <- uA * 1E-6 # conv to A
  atCin <- lCO2toatC(vol) / 60 # gives atoms/s
  atC12out <- CcurtoatC(cur, cs)
  atCout <- atC12out / .99 # 99% of C is 12C, we need total C
  atC12out / atCin
}
```

## USAMS032817

```{r}
data <- readResfile("/mnt/shared/USAMS/Results/2017 Results/USAMS032817R.txt") %>%
  mungeResfile() 
```

### Currents

For the first target, we turned up the gas supply pressure while measuring
source current at the high energy FC. These currents are roughly equivalent to
le12C.

* TODO: Blank current subtraction

Load and massage the data...

```{r}
cur <- read.delim("../Data/1sttestcurrents.csv") %>%
  mutate(flow = flowcalc(regp, r, u, x),
         eff = hgis_eff(flow, he12c, cs = 1)) %>% # cs 1 to estimate LE current
  filter(t > 5) 
```

```{r}
ggplot(cur, aes(regp, he12c)) + geom_line() + geom_point()
ggplot(cur, aes(flow, he12c)) + geom_line() + geom_point()
```

Currents were very stable after an initial burn in of ~10m. Initial current
started at ~12uA without gas on target and decayed (roughly) exponentially to a
minimum around 1.2uA.

```{r}
data %>% filter(Pos %in% c(11,12)) %>%
       	ggplot(aes(ts, he12C, color = Pos)) +
       	geom_line() + geom_point() + ylim(c(0,8))
```

### Efficiency

Using efficiency functions defined above. Efficiency drops with increasing flow
as currents go up.

```{r, warning = FALSE}
ggplot(cur, aes(flow, eff)) + geom_line() + geom_point()
```

### Ratios

Ratios for live and dead CO2 were recorded in 180 blocks of 10s each. Solid AA
and OX-I were run with 10 blocks of 180s.

#### Summary table

```{r}
data.s <- data %>% group_by(Pos) %>% summarize(c1412.m = mean(X14.12he),
                                         c1412.s = sd(X14.12he),
                                         c1412.sdr = c1412.s/c1412.m,
                                         c1412.se = se(X14.12he),
                                         c1412.ser = c1412.se/c1412.m,
                                         counts = sum(CntTotGT),
                                         interr = 1/sqrt(counts))
knitr::kable(data.s, digits = 4)
```

#### Plots

Boxplots of modern and dead samples. Ratios were essentially stable over time.
Positions < 10 are solid samples run for comparison.

Modern samples.

```{r}
data %>% filter(X14.12he > .4) %>%
       	ggplot(aes(Pos, X14.12he, color = Num)) + 
        geom_boxplot()
```

Dead gas and alpha aesar.

```{r}
data %>% filter(X14.12he < .4) %>% ggplot(aes(Pos, X14.12he, color = Num)) + 
        geom_boxplot()
```

```{r, message = FALSE}
data2 <- readResfile("/mnt/shared/USAMS/Results/2017 Results/USAMS040317R.txt") %>%
  filter(Pos == 13) %>%
  mungeResfile()

# Assign Run numbers
counter <- 0
Run <- c(1:190)
for (i in 1:nrow(data2)) {
  if (data2$Meas[i] == 1) {
    counter <- counter + 1
  }
  Run[i] <- counter
} 
data2$Run <- Run

# Load pressure data
pres <- read.csv("../Data/USAMS040317Pressure.csv")
pres$Run <- as.numeric(row.names(pres))

# Join to run data
data2 <- inner_join(data2, pres) %>%
  mutate(Source = Source * 1E-7, # I think the source pressures were to the -7, but unsure...
         flow = flowcalc(Bottle, r, u, x),
         eff = hgis_eff(flow, he12C, cs = 1)) # cs 1 to estimate LE current

# summarize
d2.s <- data2 %>% group_by(Run, Bottle, Source, flow) %>% summarize(le12C = mean(le12C),
                                              he12C = mean(he12C),
                                              counts = sum(CntTotGT),
                                              interr = 1/sqrt(counts),
                                              C1412he = mean(X14.12he),
                                              ferr = interr * C1412he,
                                              cor1412he = mean(cor1412he),
                                              eff = mean(eff))
```

# Current vs Flow

```{r, message = FALSE, warning = FALSE}
ggplot(d2.s, aes(Bottle, he12C)) +
  geom_smooth() +
  geom_point() 
filter(d2.s, flow > 0) %>%
  ggplot(aes(flow, le12C)) +
  geom_smooth() +
  geom_point() +
  labs(x = "CO2 flow rate (uL/min)", y = "12C- current (uA)") +
  scale_y_reverse() +
  ylim(0,-10) +
  xlim(0,15.5)
```

# Efficiency vs Flow

```{r, message = FALSE, warning = FALSE}
ggplot(d2.s, aes(flow, eff*100)) + 
  geom_smooth() +
  geom_point() +
  labs(x = "CO2 flow rate (uL/min)", y = "Measurement efficiency (%)") +
  xlim(0,15.5)
```

# Modern Performance

# Blank

### 2021-06-10 inst/rmd/helium_displacement.Rmd
Move Rmd's to inst/Rmd

---
title: "Helium displacement"
output:
  html_document:
    df_print: paged
---

Calculate concentration of CO2 in a vial/effluent when introducing helium as a displacement gas.

```{r, message=FALSE}
library(tidyverse)
library(HybridGIS)
```

I think this can be solved with ODE:

Change in volume of helium over time...

$$ C^\prime(t) = r - \frac{r}{V}C(t) $$

Add some math...

$$ C(t) = V(1-e^{-\frac{r}{V}t}) $$

The fraction of CO2 in the effluent should be:

$$ C(t) = e^{-\frac{r}{V}t} $$

Put this into functions. Main function moved to HGISFunctions.R and documented there.

Function definition:

```{r}
concCO2
```

Helper function to calculate fraction/flow of helium:

```{r}
concHe <- function(t, ...) {
  1 - concCO2(t)
}
```

Plotted as fraction of CO2 and helium

```{r}
data.frame(x = 0:400) %>% 
  ggplot(aes(x)) +
  stat_function(fun=concCO2, aes(color = "CO2")) +
  stat_function(fun=concHe, aes(color = "Helium")) +
  scale_color_manual("Gas", values = c("green", "blue")) +
  labs(title = "Concentration of CO2 and Helium in vial over time",
       subtitle = "7mL vial, 100uL/min helium",
       x = "Time (min)",
       y = "Concentration (fractional)")
```

OK, what's the rate of flow of CO2 over time, in the effluent, and in a capillary taking Q/t ul/m?

If the vial is 7mL, flowing at 20ul/m and sampling 10ul/m with the open split, CO2 ul/m looks like:

```{r}
data.frame(x = 0:400) %>% 
  ggplot(aes(x)) +
  stat_function(fun=function(x) concCO2(x, r = 20, flow = 10), aes(color = "CO2")) +
  stat_function(fun=function(x) 10 - concCO2(x, r = 20, flow = 10), aes(color = "Helium")) +
  scale_color_manual("Gas", values = c("green", "blue")) +
  labs(title = "Flow of CO2 to source over time",
       subtitle = "7mL vial, 20uL/min helium, 10uL/min to source",
       x = "Time (min)",
       y = "Flow (uL/min)")
```

Takes about 250 min to drop from 10 to 5 uL/min CO2.

## High dillution

Model case of small amount of CO2 in He in vial. Requires higher capillary flow to source to maintain ~5uL/min CO2 to source in helium. This also requires higher displacement flow to maintain split ratio.

Flows are 200uL/min of displacement helium, and 100uL/min delivered to source. Aiming for 5uL/min CO2 to source.

Full plot starting at 100% CO2 in 7mL vial. Working on adding initial CO2 concentration in vial to `concCO2()`, but for now, use different start time to same effect.

```{r}
data.frame(x = 0:400) %>% 
  ggplot(aes(x)) +
  stat_function(fun=function(x) concCO2(x, r = 200, flow = 100), aes(color = "CO2")) +
  stat_function(fun=function(x) 100 - concCO2(x, r = 200, flow = 100), aes(color = "Helium")) +
  scale_color_manual("Gas", values = c("green", "blue")) +
  labs(title = "Flow of CO2 to source over time",
       subtitle = "7mL vial, 0.2mL/min helium, 100uL/min to source",
       x = "Time (min)",
       y = "Flow (uL/min)")
```

Zooming in on flows around 5uL/min CO2 to source in 100uL/min He...

```{r}
data.frame(x = 75:150) %>% 
  ggplot(aes(x)) +
  stat_function(fun=function(x) concCO2(x, r = 200, flow = 100), aes(color = "CO2")) +
  labs(title = "Flow of CO2 to source over time",
       subtitle = "7mL vial, 0.2mL/min helium, 100uL/min to source",
       x = "Time (min)",
       y = "Flow (uL/min)") +
  scale_color_manual("Gas", values = c("green", "blue")) +
  xlim(50, 450) +
  ylim(0, 10)
```

Takes about 50 min to drop from 10 to 5 uL/min CO2.

Dilution results in higher rate of change of CO2 entering source.

### 2021-06-10 inst/rmd/opensplittests.Rmd
Move Rmd's to inst/Rmd

---
title: "Open Split Tests"
output:
  html_document:
    df_print: paged
    toc: true
  pdf_document: default
editor_options: 
  markdown: 
    wrap: 72
---

```{r, message=FALSE, warning=FALSE}
# Load libraries and functions
library(tidyverse)
library(amstools)
library(here)
library(HybridGIS)
library(broom)
library(patchwork)

theme_set(theme_bw())
options(digits = 3)
```

Development notes for using the HGIS with an open split. Intention is to
use with gasbench for carbonates and/or large co2 samples.

# Capillary parameters

Calculate ideal capillary to provide \~7.5uL/min at atmospheric
pressure.

## Short capillary at 1 atm

40um capillary

```{r}
r <- 2E-5
u <- 1.49E-5
x <- 3.5
flowcalc(100, r, u, x)
```

```{r}
r <- 2E-5
u <- 1.49E-5
x <- 2
flowcalc(100, r, u, x)
```

50um capillary

```{r}
r <- 2.5E-5
u <- 1.49E-5
x <- 3
flowcalc(100, r, u, x)
```

```{r}
r <- 2.5E-5
u <- 1.49E-5
x <- 7
flowcalc(100, r, u, x)
```

# First Tests

## 2018-04-10 - First tests

Tried a bunch of capillary combinations. Other than initial pressure
burst when opening or closing HGIS valve, couldn't see any source
pressure changes or beam current. We believe the source end of the
stainless capillary was clogged by cesium residue.

Cleaned capillary end at next source cleaning, first with kimwipe and
water, then with isopropanol. Lots of black crud. Confirmed open passage
by blowing N2 through capillary into cup of isopropanol to observe
bubbles.

## 2018-05-23 - First tests

Tried 3m of 50um glass capillary. Insulated from source to cage with
teflon tubing as outer jacket. No problems with sparks.

Low currents, even when cutting capillary down to 2m. Either steel
capillary is partially clogged or tip is not making good seal with
target.

## 2018-05-31 - He Dilution

Open split set up with helium dilution. Using 100um glass capillary.
\~50cm source to bulkhead with teflon jacket, 285cm bulkhead to open
split.

### Flows

```{r}
flowcalc(100, r = 2.5E-5, u = 1.49E-5, l = 3.35)
```

This is way too high. I may have been using a 50um capillary for the
tests.

### Pressures

| Condition    | Torr  |
|--------------|-------|
| base w/cryo  | 1E-7  |
| base w/turbo | 4E-7  |
| capillary    | 36E-7 |

: Source pressures

### Dilution and current

Flow and current tests with varying capillary and dilution. Flows in
mL/min, pressure in Torr, current in uA. Replumbed helium to panel open
split with needle valve partway through to better control flow. Live CO2
set to 10mL/min.

```{r}
osdil <- read_csv(here("data/2018-05-31_dillution.csv")) %>% 
  mutate(ratio = helium/co2)
osdil
```

```{r}
osdil %>% 
  filter(caplen == 322) %>% 
  ggplot(aes(ratio, current)) +
    geom_line() + 
    ggtitle("Current vs dillution ratio for 10 and 18 mL/m CO2") +
    facet_grid(co2 ~ .)
```

### Test Acquisitions

Ran 4 test Acquisitions with open split: pure CO2, pure dead CO2,
diluted live, diluted dead. Data are 14/12 ratio with a 13C correction
applied. NormFm is values normalized to the ratio of the solid OX-I.
There are also dual inlet samples on this wheel, filtered out by date.
Positions are incorrect in file because wheel was advanced manually.
Correct positions in notebook and fixed below. Added correct dilution
factors below.

```{r}
data <- get_hgis_data(here("data/USAMS053018R.txt"), 
                      as.Date(c("2018-05-30", "2018-05-31"))) %>% 
           filter(Pos %in% c(3,4) | as.Date(ts) == "2018-05-31") %>% 
           mutate(Pos = case_when(Sample.Name == "LiveGas" ~ 7,
                                  Sample.Name == "DeadGas" ~ 8,
                                  Sample.Name == "LiveGasHe" ~ 9,
                                  Sample.Name == "DeadGasHe" ~ 10,
                                  TRUE ~ as.numeric(Pos)),
                  dil_factor = case_when(Pos == 9 ~ (50.3 - 9.4)/9.4,
                                         Pos == 10 ~ (40 - 4.8)/4.8),
                  Num = ifelse(Pos %in% c(3, 4), "S", Num))

# Get mean ratio for solid OX-I and normalize data
stdrat <- mean(data$cor1412he[data$Num == "S" & !data$outlier])
data <- mutate(data, normFm = norm_gas(cor1412he, stdrat))

sum_hgis(data)
```

```{r}
plot_hgis(data)
```

## 2019-05-31 - startup tests

Reinstalled gas delivery arm after cleaning and polishing ball end.

Using 50um x .5m capillary source to cage and 50um x 2.5m to open split.
Open split flowing modern co2 at 10mL/min. Using same targets and wheel
as last time. Surprisingly good performance, including no-gas base
current, hgis current and solid samples.

```{r}
flowcalc(pres = 100, r = 2.5E-5, u = 1.49E-5, l = 3.0)
```

Currents up to 15uA, source vac 8.3E-7 with gas on, 5.5E-7 with gas off.

No data recorded

## 2019-06-03 - dillution tests

Spun up inlet turbopump. Looked at dilution and current.

```{r}
dil <- tibble(
  CO2 = c(10, 10, 10, 10, 10),
  Tot = c(10, 20, 30, 43.5, 10),
  vac = c(1.24E-6, 9.52E-7, 8.3E-7, 8E-7, 1.07E-6),
  cur = c(10, 9.6, 7, 5, 10)
) %>% 
  mutate(He = Tot - CO2,
         ratio = He/CO2)
dil
```

```{r}
ggplot(dil, aes(ratio, cur)) +
  geom_line() +
  ggtitle("Current vs dillution ratio, CO2 at 10ml/min")
```

# Parameterization of source and open split

## 2020-09-10 Starting up HGIS

Solid run on pos 1:4 on USAMS012220.

## 2020-09-29 Dilution vs Ratio

Testing effect of helium dilution on ratio. Used remaining targets on
USAMS012220. Reterminated capillaries at cage wall and at open split.

Used positions 7-10, results in USAMS012220. Low currents: \~6uA pure
CO2, 3uA diluted. Flows: He 10, deadCO2 11, liveCO2 10.

```{r}
# Load data
data <- get_hgis_data(here("data/USAMS012220R.txt"), 
                      as.Date(c("2020-09-10", "2020-09-29"))) %>% 
           mutate(dil_factor = ifelse(str_starts(Sample.Name, "Dil"), 1, 0), 
                  Num = ifelse(Pos == 7, "U", Num))
stdrat <- mean(data$cor1412he[data$Num == "S" & !data$outlier])
data <- mutate(data, normFm = norm_gas(cor1412he, stdrat))
sum_hgis(data)
```

```{r}
plot_hgis(data)
```

Looks like diluted and undiluted live gas have the same ratio.

## 2020-10-15 New wheel, low currents

Loaded a new wheel (USAMS101320) with 4 solid targets and 6 gas targets.
Currents were low (3uA gas, 40uA solid) and didn't increase with tune,
target position or gas arm position. Will clean source and try again.

## 2020-11-03 Continued dillution tests

Still very low currents, \~2uA for live CO2. Ran a dillution test anyway
with positions 5-8 on USAMS101320.

**Flows**

| Gas  | Flow |
|------|------|
| live | 10   |
| dead | 11.1 |
| He   | 30.2 |

```{r}
data <- get_hgis_data(here("data/USAMS101320R.txt"), as.Date("2020-11-03"))
sum_hgis(data)
```

```{r}
plot_hgis(data)
```

## 2020-11-17 Dillution testing - back to good currents

Finally back to testing after figuring out current issue. Rerunning
dilution test. Flows: live 11.6mL/min, dead 11.1, He 30.2. Pressures:
cryo base 8E-8, turbo base 3E-7, w/live CO2 1.2E-6. Source producing
\~8uA HE12C with live CO2. Data in USAMS101320.

Running positions 9 and 10 as 1:1 He dilutions of live and dead,
respectively.

```{r}
data <- get_hgis_data(here("data/USAMS101320R.txt"), as.Date("2020-11-17"))
sum_hgis(data)
```

Look at counts per second to think about needed acquisition time.

```{r}
data %>% group_by(Sample.Name) %>% 
  summarize(cps = mean(CntTotGT/(Cycles/10)))
```

```{r}
plot_hgis(data)
```

```{r}
data %>% filter(Pos %in% 5:10) %>% 
  ggplot(aes(he12C, normFm, color = Sample.Name)) +
  geom_point()
```

Doesn't seem to follow curves expected for a MBC with nearly live blank
at about 1% of gas current.

## 2020-12-04 Blank tests

Ran a series of blank tests before exposing graphite samples or
introducing live co2.

Capillary break between Target blank and Helium blank. Replaced with SGE
100um x 52cm. Initial currents of targets are high, between 10-25uA.
Drops asymptotically to base target current in 5-10 minutes.

```{r}
data <- get_hgis_data(here("data/USAMS120320R.txt"), as.Date("2020-12-04"))
sum_hgis(data)
```

```{r}
plot_hgis(data)
```

### Modeling Blank

With these data we can start to think about modeling a mass balance, or
more accurately, a current balance blank. It's not initially clear
whether the blank comes from the target/source, diluent helium, or a
combination of the two.

```{r}
data %>% filter(Pos %in% 9:12) %>% 
  ggplot(aes(he12C, normFm, color = Sample.Name)) +
  geom_point() +
  ggtitle("Blank Fm vs. current")
```

If this blank follows the constant contamination model, plotting Fm vs.
inverse current should give a linear fit. Need more data here to confirm
fit and determine current and Fm of blank.

```{r}
# subset data
gas <- data %>% filter(Pos %in% 9:12) %>% mutate(icur = 1/he12C) 
# fit a linear model
fitlm <- lm(normFm ~ icur, data = gas)

# Add predicted values to data
gas$Fm = predict(fitlm)

ggplot(gas) +
  geom_line(aes(1/he12C, y = Fm)) +
  geom_point(aes(1/he12C, normFm, color = Sample.Name)) +
  ggtitle("Blank Fm vs. inverse current")

```

TODO: Fit mass balance curve using currents TODO: Use bottle gas to
produce live/dead curves TODO: produce live/dead curves using helium
dilution

## 2020-12-11 open split flow reduction

Testing reduced open split flows. Worked with flow controller (helium)
and adjusted stops on live and dead co2 needle valves to allow lower
flows. Tested for breakthrough by flowing helium through split and into
source while blowing dead CO2 across the opening of the split at
\~10mL/min. CO2 seen in source somewhere between 100 and 300 uL/min.
Hard to measure with this setup due to slow response time and lack of
sensitivity of capillary/ion source. Need to try this with leak checker.

Reran blank tests at lower flows. Results seem as good or better than
last week's tests at high flows. This may also be due to targets being
more pumped down or burned in.

TODO: get flows from notes.

```{r}
data <- get_hgis_data(here("data/USAMS120320R.txt")) %>% 
  filter(Pos %in% 1:4 | as.Date(ts) == "2020-12-11") %>% 
  mutate(he = case_when(Pos == 9 ~ (1059 - 383),
                        Pos == 10 ~ 0,
                        Pos == 11 ~ 1370,
                        Pos == 12 ~ 0),
         co2 = case_when(Pos == 9 ~ 383,
                        Pos == 10 ~ 383,
                        Pos == 11 ~ 0,
                        Pos == 12 ~ 0),
         dil_rat = he/co2)

data %>% 
  group_by(Pos, Sample.Name, dil_rat) %>%
  summarise(Cur = mean(he12C),
            Cur.sd = sd(he12C),
            mean = mean(normFm),
            sd = sd(normFm),
            interr = 1/sqrt(sum(CntTotGT)),
            acqtime = sum(Cycles)/10,
            N_acq = n()) 
```

```{r}
plot_hgis(data)
```

```{r}
# subset data
gas <- data %>% filter(Pos %in% 9:12) %>% mutate(icur = 1/he12C) 
# fit a linear model
fitlm <- lm(normFm ~ icur, data = gas)

# Add predicted values to data
gas$Fm = predict(fitlm)

ggplot(gas) +
  geom_line(aes(1/he12C, Fm)) +
  geom_point(aes(1/he12C, normFm, color = Sample.Name)) +
  ggtitle("Blank Fm vs. inverse current")
```

## 2020-12-18 More testing of low open split flow rates

Tested open split breakthrough at low flows using leak checker. Measured
helium flow through capillary using leak checker. Ran target blanks,
dead and live samples at low flows. Started building double needle
testbed.

### Leak checker flow rate

Connected cage-split capillary to leak checker. Base leak rate with
capillary in air: 1.7E-9 mbarLs-1. Leak rate with helium in open split:
3.5E-4 mbarls-1.

### Leak checker breakthrough tests

Flowing dead CO2 through split. Flowing helium across mouth of split at
8mLmin-1. Detection of helium means atmosphere is getting into split.

```{r}
lc_breakthrough <- tribble(~co2, ~leakrate,
                          730, 6E-9,
                          735, 6.5E-9,
                          525, 7.9E-9,
                          178, 1.3E-7,
                          150, 2E-7,
                          0, 3E-6)
lc_breakthrough
ggplot(lc_breakthrough, aes(co2, leakrate)) + geom_line() + geom_point() +
  ggtitle("Helium leak rate vs. CO2 flow to split")
```

### Data acquisitions at low(er) flows

Source off base: 4E-8 Cryo base (HV + Ion): 6E-8 Turbo base: 2E-7

Set SRS to 10uA/V and 50nA/V for 12 and 13C.

Flows:

```{r}
flows <- tribble(~pos, ~Sample.Name, ~co2, ~he,
                10, "DeadGas", 270, 0,
                9, "DilDeadGas", 112, 288,
                8, "DilDeadGas6", 102, 0,
                8, "DeadGas", 102, 572-102,
                7, "HeliumBlank", 0, 572-102,
                7, "DilLiveGas6", 117, 576-117) %>% 
  mutate(dil_ratio = he/co2)
flows
```

### Data

```{r}
data <- get_hgis_data(here("data/USAMS120320R.txt"), as.Date("2020-12-18"))
sum_hgis(data)
```

```{r}
plot_hgis(data)
```

```{r}
data %>% filter(Pos %in% 7:10) %>% 
  ggplot(aes(1/he12C, normFm, color = Sample.Name)) +
  geom_point() +
  ggtitle("Fm of live and dead samples vs. inverse current")
```

## 2021-01-08 Pure gas precison testing

Plumbed long 50um capillary from live CO2 regulator to cage wall
fitting. This allows controlling flow of CO2 to source using regulator
pressure.

### Current vs flow and pressure

First test was with 748cm x 50 um capillary. Source base pressure using
the cryopump was 2E-7 Torr. Base current for Pos 7 was 300uA. Initial
max current \@ 200kPA CO2 was 9uA. Did not return to this level in the
pressure test following, not sure why, source pressure shows CO2 was
getting to source. Typically best currents with pure CO2 are at a source
pressure of around 1E-6 Torr. Reg pressure in kPa, source in Torr x 10-7
and current in uA.

```{r}
cur_vs_pres <- tribble(~regp, ~sourcep, ~c12cur,
                       0, 2.8, 0.8,
                       50, 3.3, 1,
                       100, 5.3, 1.2,
                       150, 8.2, 1.4,
                       200, 12, 1.8,
                       250, 17, 2.4,
                       300, 22, 3.1,
                       350, 28, 3.7,
                       400, 35, 4.5)
cur_vs_pres
```

Tested again with new capillaries after a clog in regulator-cage
capillary and spark damage to cage-source capillary. New capillaries:
reg-cage- 639cm x 50um, cage-source- 50cm x 50um. Base current 250nA.

```{r}
cur_vs_pres2 <- tribble(~regp, ~sourcep, ~c12cur,
                       25, 2, .25,
                       100, 4, 1.3,
                       150, 8.4, 6.3,
                       200, 12, 6.2,
                       250, 17, 6.2)
cur_vs_pres2
```

### Data vs flow

Took data on targets 6 and 7 at various CO2 flow rates.

#### Data summary

```{r}
data <- get_hgis_data(here("data/USAMS120320R.txt"), as.Date("2021-01-08"))
ds <- sum_hgis(data)
ds
```

```{r}
plot_hgis(data)
```

#### Ratio and current vs flow

Took data for 4 different flows on target 6. CO2 flow data show ratio
dependence on flow. Need to fill in data to get shape of curve. Current
vs flow

```{r}
ds %>% 
  filter(Pos == 6) %>% 
  mutate(Pressure = as.numeric(str_extract(Sample.Name, "\\d+"))) %>% 
  ggplot(aes(Pressure, Cur)) +
  geom_pointrange(aes(ymin = Cur - Cur.sd, ymax = Cur + Cur.sd)) + 
  geom_point(size = 4) +
  ggtitle("Fm of live CO2 vs. C12 current")

```

Fm vs. Flow. Note that solid standards had a lot of variability, so
treat ratio as relative, not well normalized.

```{r}
ds %>% 
  filter(Pos == 6) %>% 
  mutate(Pressure = str_extract(Sample.Name, "\\d+")) %>% 
  ggplot(aes(Cur, mean, shape = Pressure, color = Pressure)) +
  geom_pointrange(aes(ymin = mean - sd, ymax = mean + sd)) + 
  geom_point(size = 4) +
  ggtitle("Fm vs. C12 current for varying CO2 regulator pressure")
```

## 2021-01-15 Precision and vial tests

Used test vial setup to test pure CO2 precision and introduction of CO2
from vial by displacing with helium.

### Data

```{r}
df <- get_hgis_data(here("data/USAMS120320R.txt"), as.Date("2021-01-15")) 
stdrat <- mean(df$cor1412he[df$pos_name == "8 - LiveGasOS" & !df$outlier])
df <- mutate(df, normFm = norm_gas(cor1412he, stdrat))

sum_hgis(df)
```

### Ratio and current during runs

```{r}
plot_hgis_time(df) +
  facet_wrap(vars(pos_name), scales = "fixed")

ggplot(df, aes(cum_acqtime, he12C )) +
  geom_point() + 
  facet_wrap(vars(pos_name), scales = "fixed")
```

#### Ratio dependence on current

```{r}
ggplot(df, aes(he12C, normFm, color = pos_name)) +
  geom_point() +
  ggtitle("Fm of modern CO2 vs. C12 current")
```

```{r}
ggplot(df, aes(1/he12C, normFm, color = pos_name)) +
  geom_point() +
  ggtitle("Fm of modern CO2 vs. inverse C12 current")
```

Question: Is ratio current depletion change in CO2 from vial or target
degradation? Long pure CO2 run to test.

## 2021-01-22 Pure gas precison testing

Plumbed long 50um capillary from Valco valve to cage wall fitting.
Plumbed dead CO2 to Valco with 1/16" ss.

Took data with dead CO2 on target 5 at various CO2 flow rates. Same with
liveCO2 on target 6. Source not performing well, low currents and
current did not correspond well to CO2 pressure/flow. Current vs ratio
data may still be useful for blank model. Solid currents good, so this
is a gas issue. Gas blank also not great.

### Data summary

```{r}
data <- get_hgis_data(here("data/USAMS012121R.txt"), as.Date("2021-01-22"))
ds <- sum_hgis(data)
ds
```

Note that solid standards had a lot of variability, so treat ratio as
relative, not well normalized.

```{r}
plot_hgis(data)
```

#### Relationship of current and ratio to CO2 flow rate

##### Current vs flow

Supply pressure is proportional to flow rate.

```{r}
ds <- ds %>% 
  filter(Pos %in% c(5, 6)) %>% 
  mutate(Cur_inv = 1/Cur,
         Pressure = as.numeric(str_extract(Sample.Name, "\\d+")),
         Gas = ifelse(Pos == 5, "DeadCO2", "LiveCO2"))

ggplot(ds, aes(Pressure, Cur, color = Gas)) +
  geom_pointrange(aes(ymin = Cur - Cur.sd, ymax = Cur + Cur.sd)) + 
  geom_point(size = 4) +
  ggtitle("C12 current vs. CO2 regulator pressure")

```

##### Fm vs. current and flow rate

```{r}
ggplot(ds, aes(Cur, mean, color = Pos)) +
  geom_pointrange(aes(shape = factor(Pressure), ymin = mean - sd, ymax = mean + sd)) +
  ggtitle("Fm of live and dead gas vs C12 current")
```

### Modelling current dependent blank

#### Fit linear models

```{r}
fits <- ds %>% 
  ungroup() %>% 
  nest(data = -Gas) %>% 
  mutate(fit = map(data, ~lm(mean ~ Cur_inv, data = .x)),
         tidied = map(fit, tidy)) %>% 
  unnest(tidied)

fits %>% 
  select(Gas, term, estimate) %>% 
  pivot_wider(names_from = c(Gas, term), values_from = estimate) %>% 
  mutate(inv_m_blank = -(`DeadCO2_(Intercept)` - `LiveCO2_(Intercept)`)/(DeadCO2_Cur_inv - LiveCO2_Cur_inv),
         Fm_blank = `DeadCO2_(Intercept)` + DeadCO2_Cur_inv * inv_m_blank,
         m_blank = 1/inv_m_blank) %>% 
  select(Fm_blank, m_blank)
```

#### Fm vs. inverse mass with fits.

```{r}
ggplot(ds, aes(Cur_inv, mean, color = Pos)) +
  geom_pointrange(aes(shape = factor(Pressure), ymin = mean - sd, ymax = mean + sd)) + 
  geom_smooth(method = "lm", fullrange = TRUE) +
  xlim(0, 5) +
  ggtitle("Fm of live and dead CO2 vs. inverse 12C current",
          subtitle = "Fit lines should cross at Fm and current of blank")
```

For these measurements, it looks like the blank is roughly 225nA at half
modern. This is higher current and lower Fm than expected based on
previous measurements. This is in line with the high blanks seen when
measuring dead carbon.

# Gas delivery from vials

## 2021-01-29 Reproducibility between multiple vials

Measured live and dead CO2 from vials. "STD" samples were displaced to
split with flowing CO2, while the "V" samples were displaced by helium
as they would be for unknowns. All samples measured on fresh targets
after a 5 minute "burn in" to reduce blank.

### Gas flows

Replaced capillary from Valco valve to double needle with longer piece
of 50um capillary. Set helium and CO2 flows.

+----------------------+-----------------------+-----------------------+
| Gas                  | Pressure (kPa)        | Flow (uL/min)         |
+======================+=======================+=======================+
| LiveCO2              | 165                   | 136                   |
+----------------------+-----------------------+-----------------------+
| DeadCO2              | 160                   | 120                   |
+----------------------+-----------------------+-----------------------+
| Helium               | 165                   | 115                   |
+----------------------+-----------------------+-----------------------+

: Gas flow rates

Measured split to cage wall capillary flow with leak checker. Leak rate:
1.0E-5 Pa m\^3/s.

### Data

Normalizing to solid OX-I's.

```{r}
data <- get_hgis_data(here("data/USAMS012121R.txt"), as.Date("2021-01-29")) %>% 
  mutate(Num = ifelse(Pos %in% c(7, 9), "U", Num))
```

#### Currents and ratios vs time.

Sample ratio vs. time

```{r}
plot_hgis_time(data, cor1412he, ce*cor1412he)
```

Sample current vs time

```{r}
plot_hgis_time(data, he12C)
```

```{r}
# Get mean ratio for solid OX-I and normalize data
stdrat <- mean(data$cor1412he[data$Num == "S" & !data$outlier])
data <- mutate(data, normFm = norm_gas(cor1412he, stdrat))
```

#### Per-sample summary

```{r}
ds <- data %>% 
  filter(!outlier) %>% 
  sum_hgis()
ds
```

```{r}
plot_hgis(data)
```

### Agreement of replicates

Summary statistics to look at reproducibility between targets/vials.

```{r}
ds %>% 
  filter(as.numeric(Pos) > 4,
         Pos != 9) %>%  # Don't use bad target
  mutate(gas = ifelse(str_starts(Sample.Name, "Live"), "Live", "Dead"),
         helium = ifelse(str_detect(Sample.Name, "V\\d$"), TRUE, FALSE)) %>% 
  group_by(gas, helium) %>% 
  summarize(Fm.mean = mean(mean),
            Fm.sd = sd(mean),
            N = n())
```

It looks like the SD of modern samples is around 1%, and that pure CO2
performs slightly better than helium displacement.

## 2021-02-05 Carbonate samples

Measured carbonate samples prepared on gasbench. Samples included C-1,
TIRI-F, TIRI-I, C-2, and NOSAMS2. Normalized to gas standards or solid
OX-I.

First carbonate sample clogged capillary. Capillary replaced, moved
needle up to avoid sucking in acid. Was not able to get good currents
after replacing capillary, even trying several length/capillary
combinations. Source vacuum indicates normal amount of gas entering
source. May need to consider a water trap. All samples measured on fresh
targets after a 5 minute "burn in" to reduce blank.

### Samples

```{r}
carb_data <- tribble(~vial, ~name, ~rec_num, ~mass,
                     1, "C-1", 83028, 32.05,
                     2, "TIRI-F", 2138, 33.56,
                     3, "TIRI-I", 17185, 32.67,
                     4, "TIRI-I", 17185, 32.35,
                     5, "C-2", 1082, 31.35,
                     6, "C-2", 1082, 34.11,
                     7, "NOSAMS2", 38809, 32.30,
                     8, "NOSAMS2", 38809, 34.00
                     )

write_csv(carb_data, here("data/carb_data_2021-02-05.csv"))
carb_data
```

### Gas flows

Set helium and CO2 flows.

+----------------------+-----------------------+-----------------------+
| Gas                  | Pressure (kPa)        | Flow (uL/min)         |
+======================+=======================+=======================+
| LiveCO2              | 150                   | 95                    |
+----------------------+-----------------------+-----------------------+
| Helium               | 150                   | 97                    |
+----------------------+-----------------------+-----------------------+

: Gas flow rates

### Data

Normalizing to gas standards in pos 5 and 8.

```{r}
data <- get_hgis_data(here("data/USAMS020521R.txt")) %>% 
  mutate(Num = ifelse(Pos %in% c(2, 4), "U", Num))
stdrat <- mean(data$cor1412he[data$Num == "S" & !data$outlier])
data <- data %>% 
  mutate(normFm = norm_gas(cor1412he, stdrat))
```

#### Currents and ratios vs time.

Sample ratio vs. time

```{r}
plot_hgis_time(data, normFm, ce*normFm)
```

Sample current vs time

```{r}
plot_hgis_time(data, he12C)
```

#### Summary of normalized data per sample

```{r}
ds <- data %>% 
  filter(!outlier) %>% 
  sum_hgis() %>% 
  mutate(rec_num = case_when(str_starts(Sample.Name, "LiveGas") ~ 101730,
                             Sample.Name == "C-1" ~ 83028,
                             Sample.Name == "TIRI-F" ~ 2138,
                             Sample.Name == "TIRI-I" ~ 17185,
                             Sample.Name == "C-2" ~ 1082,
                             Sample.Name == "NOSAMS2" ~ 38809))

std <- getStdTable()

ds <- left_join(ds, select(std, rec_num, fm_consensus)) %>% 
  mutate(fm_consensus = ifelse(rec_num == 101730, 1.0398, fm_consensus))

ds[-(1:4),-3]
```

```{r}
data %>% 
  filter(as.numeric(Pos) > 4) %>% 
  ggplot(aes(pos_name, normFm)) +
  geom_boxplot(fill = "slategray1") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) 
```

### Agreement of replicates

Summary statistics to look at reproducibility, accuracy between
targets/vials.

```{r}
ds %>% 
  filter(as.numeric(Pos) > 4,
         Pos != 6) %>% # Don't use bad target
  mutate(Name = ifelse(Pos %in% c(5,8), "LiveGas", Sample.Name)) %>% 
  group_by(Name, fm_consensus) %>% 
  summarize(Fm.mean = mean(mean),
            Fm.sd = sd(mean),
            Cur.mean = mean(Cur),
            N = n()) 
```

It looks like the SD of replicates of modern samples with current \> 3uA
is around 0.5%, which is better than measurement statistics. Fm of
NOSAMS 2 is a bit high. Need to investigate current-dependent blank.

#### Blank correction

Testing correcting for blank using a simple, mass-independent large
blank correction.

LBC does not work well for this set of samples.

```{r}
fmblank <- ds$mean[ds$Sample.Name == "TIRI-F"]
fmblank
ds <- ds %>% 
  mutate(fm_lbc = doLBC(mean, fmblank, 1.0398))

ds %>%
  filter(as.numeric(Pos) > 4) %>% 
select(Sample.Name, Cur, fm_consensus, mean, fm_lbc)

ds %>% 
write_csv(here("data_analysed/USAMS020521.csv"))
```

## 2021-03-01 Live gas and solid samples

Ran live CO2 from vial to check performance after source cleaning and
cryopump issues. Slow to get going, so only one modern gas standard and
solid samples run. Tried second modern gas standard on pos 16, but no
current from gas that worked on pos 15. Bad target?

```{r}
data <- get_hgis_data(here("data/USAMS020521R.txt"), as.Date("2021-03-01"))
```

### Currents and ratios vs time.

Sample ratio vs. time

```{r}
plot_hgis_time(data, normFm, ce*normFm)
```

Sample current vs time

```{r}
plot_hgis_time(data, he12C)
```

### Normalization

```{r}
# Get mean ratio for solid OX-I and normalize data
stdrat <- mean(data$cor1412he[data$Num == "S" & !data$outlier])
data <- data %>% 
  mutate(normFm = norm_gas(cor1412he, stdrat))
```

### Summary of normalized data per sample

```{r}
ds <- data %>% 
  filter(!outlier) %>% 
  sum_hgis() %>% 
  mutate(rec_num = case_when(str_starts(Sample.Name, "LiveGas") ~ 101730,
                             Sample.Name == "C-1" ~ 83028,
                             Sample.Name == "TIRI-F" ~ 2138,
                             Sample.Name == "TIRI-I" ~ 17185,
                             Sample.Name == "C-2" ~ 1082,
                             Sample.Name == "NOSAMS2" ~ 38809))

std <- getStdTable()

ds <- left_join(ds, select(std, rec_num, fm_consensus)) %>% 
  mutate(fm_consensus = ifelse(rec_num == 101730, 1.0398, fm_consensus))

ds
```

```{r}
plot_hgis(data)
```

## 2021-03-05 Carbonate samples

Measured carbonate samples prepared on gasbench. Samples included C-1,
C-2, and NOSAMS2. Normalized to gas standards or solid OX-I.

All samples measured on fresh targets after a 5 minute "burn in" to
reduce blank. Gas flows same as 2021-02-05.

### Samples

```{r}
carb_data <- tribble(~vial, ~name, ~rec_num, ~mass,
                     1, "C-1", 83028, 33.69,
                     2, "C-1", 83028, 31.62,
                     3, "C-2", 1082, 32.70,
                     4, "C-2", 1082, 32.08,
                     5, "NOSAMS2", 38809, 33.23,
                     6, "NOSAMS2", 38809, 33.95
                     )

write_csv(carb_data, here("data/carb_data_2021-03-05.csv"))
carb_data
```

### Data

Normalizing to gas standards in pos 5 and 8.

```{r}
data <- get_hgis_data(here("data/USAMS030421R.txt"), as.Date("2021-03-05"))
```

#### Currents and ratios vs time.

Sample ratio vs. time

```{r}
plot_hgis_time(data, normFm, ce*normFm)
```

Sample current vs time

```{r}
plot_hgis_time(data, he12C)
```

#### Summary of normalized data per sample

```{r}
std <- getStdTable()

ds <- data %>% 
  filter(!outlier) %>% 
  sum_hgis() %>% 
  mutate(rec_num = case_when(str_starts(Sample.Name, "LiveGas") ~ 101730,
                             str_starts(Sample.Name, "C1") ~ 83028,
                             str_starts(Sample.Name, "TIRI-F") ~ 2138,
                             str_starts(Sample.Name, "TIRI-I") ~ 17185,
                             str_starts(Sample.Name, "C2") ~ 1082,
                             str_starts(Sample.Name, "NOSAMS2") ~ 38809,
                             str_starts(Sample.Name, "DeadGas") ~ 72446)) %>% 
 left_join(select(std, rec_num, fm_consensus), by = "rec_num") %>% 
 mutate(fm_consensus = case_when(rec_num == 101730 ~ 1.0398,
                                 rec_num == 72446 ~ 0.0013,
                                 TRUE ~ fm_consensus))

ds[-(1:4),-3]
```

```{r}
data %>% 
  filter(as.numeric(Pos) > 4) %>% 
  ggplot(aes(pos_name, normFm)) +
  geom_boxplot(fill = "slategray1") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) 
```

### Agreement of replicates

Summary statistics to look at reproducibility, accuracy between
targets/vials.

Mean of modern CO2 filled on HGIS setup (LiveGas) and Gilson
(LiveGasGil) were used for normalization but were fairly different on
this run, so shown separately.

```{r}
ds %>% 
  filter(as.numeric(Pos) > 4) %>% 
  mutate(Name = str_remove(Sample.Name, "-.$")) %>% 
  group_by(Name, fm_consensus) %>% 
  summarize(Fm.mean = mean(mean),
            Fm.sd = sd(mean),
            Cur.mean = mean(Cur),
            N = n()) 
```

Agreement of replicates is better than 2% for all samples. Samples have
higher Fm than consensus. I think most of this is due to blanks, but the
low gas standard also would push unknown Fm's up.

#### Blank correction

Testing correcting for blank using a simple, mass-independent large
blank correction.

Using dead gas blank, naive with no adjustment for graphite value of
blank. This pushes Fm in the right direction, but comparing samples of
the same type, there's a lot of variability that seems to be tied to
sample current.

```{r}
fmblank <- ds$mean[ds$Sample.Name == "DeadGas"]
fmblank
ds <- ds %>% 
  mutate(fm_lbc = doLBC(mean, fmblank, 1.0398))

ds %>%
  filter(as.numeric(Pos) > 4) %>% 
select(Sample.Name, Cur, fm_consensus, mean, fm_lbc)

ds %>% 
write_csv(here("data_analysed/USAMS030421.csv"))
```

## 2021-03-26 High dillution test

Determine length of 75um capillary for \~100ulm to source.

```{r}
r <- 7.5E-5 / 2
u <- 1.49E-5
x <- 3
flowcalc(100, r, u, x)
```

Startup and test after opening tank to refill terminal argon bottle.
Testing helium dilution by starting with helium in vial and displacing
with CO2. Used 2.5m of 75um capillary to provide 100ulm to source, but
initial source pressure of 3E-7 Torr indicates lower flows. Shortened
split-cage capillary to about 150cm, without a huge effect on pressure.
Flow may be defined by cage-source capillary.

LiveCO2 flow set to 244 uL/m.

Using position 14 on USAMS030421. Initial current after 10min: 150nA.

Currents rose to a stable 10uA and stayed there for more than 90 min!

### Data

```{r}
data <- get_hgis_data(here("data/USAMS030421R.txt"), as.Date("2021-03-26")) %>% 
  mutate(Num = ifelse(Pos %in% c(2,4), "S", Pos))
stdrat <- mean(data$cor1412he[data$Num == "S" & !data$outlier])
data <- data %>% 
  mutate(normFm = norm_gas(cor1412he, stdrat))
```

#### Currents and ratios vs time.

```{r}
data %>% 
  filter(Pos == 14) %>% 
  plot_hgis_time(normFm, ce*normFm) +
  ggtitle("Sample ratio vs. time")
```

```{r}
data %>% 
  filter(Pos == 14) %>% 
  plot_hgis_time(he12C) +
  ggtitle("Sample current vs. time")
```

### Predicted flow of CO2 to source for test flow conditions

```{r}
co2flow <- 244 / 60 #flow in ul/s

data.frame(x = 0:8000) %>% 
  ggplot(aes(x)) +
  stat_function(fun=function(x) concCO2(x, r = co2flow) * 7, aes(color = "He")) +
  stat_function(fun=function(x) 7 - concCO2(x, r = co2flow) * 7, aes(color = "CO2")) +
  scale_color_manual("Gas", values = c("green", "blue")) +
  labs(title = "CO2 in vial over time",
       subtitle = "7mL vial, 244uL/min CO2",
       x = "Time (s)",
       y = "Volume (mL)")
```

Stable currents reached at around 5mL CO2 in vial, but stable ratios
reached at around 2mL. Contrast with helium dilution test from 4/2
below.

### Ratio vs Current

```{r}
data %>% 
  filter(Pos == 14) %>% 
  ggplot(aes(he12C, normFm)) +
  geom_point()
```

## 2021-04-02 High dillution test

Testing performance at higher dilution ratio. Started with pure live CO2
in vial and did long acquisition with high displacement helium flow to
check current as CO2 concentration decreases.

Helium flow set to 244 uL/m.

Using position 25 on USAMS040321. Shifted targets to higher positions to
avoid wearing only low wheel positions. Initial current after 10min:
150nA.

### Data

```{r}
data <- get_hgis_data(here("data/USAMS040121R.txt"), as.Date("2021-04-02")) %>% 
  mutate(Num = ifelse(Pos == 5, "S", Pos))
stdrat <- mean(data$cor1412he[data$Num == "S" & !data$outlier])
data <- data %>% 
  mutate(normFm = norm_gas(cor1412he, stdrat))
```

#### Currents and ratios vs time.

Sample ratio vs. time

```{r}
data %>% 
  filter(Pos == 5) %>% 
  plot_hgis_time(normFm, ce*normFm)
```

Sample current vs time

```{r}
data %>% 
  filter(Pos == 5) %>% 
  plot_hgis_time(he12C)
```

### Expected CO2 ratio/flow

Plot volume of CO2 in sample vial during displacement by helium. It
looks like optimal currents are maintained until about 2000s, which
corresponds to about 2mL of CO2. Stable ratios until about 1.25 mL CO2.
Can we back calculate flow to source if current falls off below
\~2ul/min CO2?

```{r}
heflow <- 244 / 60 #flow in ul/s

data.frame(x = 0:8000) %>% 
  ggplot(aes(x)) +
  stat_function(fun=function(x) concCO2(x, r = heflow) * 7, aes(color = "CO2")) +
  scale_color_manual("Gas", values = c("green", "blue")) +
  labs(title = "CO2 in vial over time",
       subtitle = "7mL vial, 244uL/min helium",
       x = "Time (s)",
       y = "CO2 (mL)")
```

```{r}
data.frame(x = 0:8000) %>% 
  ggplot(aes(x)) +
  stat_function(fun=function(x) concCO2(x, r = heflow) * 30, aes(color = "CO2")) +
  scale_color_manual("Gas", values = c("green", "blue")) +
  labs(title = "CO2 flow over time",
       subtitle = "7mL vial, 244uL/min helium",
       x = "Time (s)",
       y = "CO2 (uL/s)")
```

Add flow at time to data and plot current vs flow. 30 ul/m is estimate
of capillary flow based on past results.

```{r}
data <- data %>% 
  mutate(co2flow = concCO2(cum_acqtime, r = heflow) * 30) #30ul/min

ggplot(data, aes(co2flow, he12C)) +
  geom_smooth(span = .3, se=FALSE) +
  geom_point() +
  labs(title = "Vial dillution current",
       subtitle = "250uL/min displacement, 30ul/min delivery",
       x = "CO2 flow (ul/m)",
       y = "12C current (uA)")
```

### Ratio vs Current

```{r}
data %>% 
  filter(Pos == 5) %>% 
  ggplot(aes(he12C, normFm)) +
  geom_point()
```

## 2021-04-09 Carbonate samples

Measured carbonate samples prepared on gasbench. Samples included
TIRI-F, TIRI-I, and NOSAMS2. Normalized to gas standards or solid OX-I.

All samples measured on fresh targets after a 10 minute "burn in" to
reduce blanks.

Using short capillaries from 2021-04-02. He flow set to 250ul/min.

### Samples

```{r}
carb_data <- tribble(~vial, ~name, ~rec_num, ~mass,
                     1, "NOSAMS2_1", 38809, 31.8,
                     2, "NOSAMS2_2", 38809, 32.3,
                     3, "NOSAMS2_3", 38809, 33.5,
                     4, "TIRI-I_1", 17185, 33.1,
                     5, "TIRI-I_2", 17185, 31.8,
                     6, "TIRI-I_3", 17185, 31.8,
                     7, "TIRI-F_1", 2138, 31.9,
                     8, "TIRI-F_2", 2138, 33.4,
                     9, "TIRI-F_3", 2138, 32.4
                     )

write_csv(carb_data, here("data/carb_data_2021-04-08.csv"))
carb_data
```

### Data

Normalizing to gas standards in pos 26 and 37.

```{r}
data <- get_hgis_data(here("data/USAMS040121R.txt"), as.Date("2021-04-09")) %>% 
  mutate(Num = ifelse(Pos %in% c(22, 24, 25, 27, 38), "U", Num))
stdrat <- mean(data$cor1412he[data$Num == "S" & !data$outlier])
data <- data %>% 
  mutate(normFm = norm_gas(cor1412he, stdrat))
```

#### Currents and ratios vs time.

```{r}
plot_hgis_time(data, normFm, ce*normFm) +
  ggtitle("Sample ratio vs. time")
```

```{r}
plot_hgis_time(data, he12C) +
  ggtitle("Sample current vs. time")
```

#### Summary of normalized data per sample

```{r}
std <- getStdTable()

ds <- data %>% 
  filter(!outlier) %>% 
  sum_hgis() %>% 
  mutate(rec_num = case_when(str_starts(Sample.Name, "LiveGas") ~ 101730,
                             str_starts(Sample.Name, "C1") ~ 83028,
                             str_starts(Sample.Name, "TIRI-F") ~ 2138,
                             str_starts(Sample.Name, "TIRI-I") ~ 17185,
                             str_starts(Sample.Name, "C2") ~ 1082,
                             str_starts(Sample.Name, "NOSAMS2") ~ 38809,
                             str_starts(Sample.Name, "DeadGas") ~ 72446)) %>% 
 left_join(select(std, rec_num, fm_consensus), by = "rec_num") %>% 
 mutate(fm_consensus = case_when(rec_num == 101730 ~ 1.0398,
                                 rec_num == 72446 ~ 0.0013,
                                 TRUE ~ fm_consensus))

ds[-(1:4),-3]
```

```{r}
data %>% 
  filter(as.numeric(Pos) > 5) %>% 
  ggplot(aes(pos_name, normFm)) +
  geom_boxplot(fill = "slategray1") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) 
```

### Agreement of replicates

Summary statistics to look at reproducibility, accuracy between
targets/vials.

```{r}
ds %>% 
  filter(as.numeric(Pos) > 5) %>% 
  mutate(Name = str_remove(Sample.Name, "_.$")) %>% 
  group_by(Name, fm_consensus) %>% 
  summarize(Fm.mean = mean(mean),
            Fm.sd = sd(mean),
            Cur.mean = mean(Cur),
            N = n()) 
```

Agreement of replicates is better than 2% for all samples. Samples have
higher Fm than consensus. I think most of this is due to blanks, but the
low gas standard also would push unknown Fm's up.

#### Blank correction

Testing correcting for blank using a mass-independent large blank
correction.

Using TIRI-F as blank, naive with no adjustment for graphite value of
blank. This pushes Fm in the right direction, but comparing samples of
the same type, there's a lot of variability that seems to be tied to
sample current.

```{r}
fmblank <- ds %>%
  filter(str_starts(Sample.Name, "TIRI-F")) %>% 
  pull(mean) %>% 
  mean()

fmblank
ds <- ds %>% 
  mutate(fm_lbc = doLBC(mean, fmblank, 1.0398),
         diff = fm_lbc - fm_consensus,
         normFm = normFm(fm_lbc, fm_consensus),
         sigma = sigma(fm_lbc, fm_consensus, merr))

ds %>%
  filter(as.numeric(Pos) > 4) %>% 
  select(Sample.Name, Cur, fm_consensus, mean, fm_lbc, diff)

ds %>% 
write_csv(here("data_analysed/USAMS040121.csv"))
```

```{r}
ggplot(ds, aes(fm_consensus, diff)) +
  geom_hline(yintercept = 0) +
  geom_errorbar(aes(ymin = diff - merr, ymax = diff + merr)) +
  geom_point() +
  ggtitle("Difference in LBC-corrected vs consensus")
```

```{r}
ggplot(ds, aes(fm_consensus, sigma)) +
  geom_hline(yintercept = 0) +
  geom_point() +
  ggtitle("Sigma of LBC-corrected vs consensus")
```

It looks like the normalization may be off a bit. Note positive linear
correlation of Fm consensus difference with Fm of sample. Gas standards
run at end of day were higher than those from the beginning. NOSAMS-2
was run early in the day, so closer to standards with lower ratios.

## 2021-04-16 Carbonate samples

Measured carbonate samples prepared on gasbench. Samples included C-1,
C-2, and NOSAMS2. Normalized to gas standards or solid OX-I.

All samples measured on fresh targets after a 10 minute "burn in" to
reduce blank.

Using short capillaries from 2021-04-02. He flow set to 250ul/min. Had
to shorten delivery capillary by \~10cm at start of first C-2 due to
acid clog. Needle blown out with compressed air.

### Samples

```{r}
carb_data <- tribble(~vial, ~name, ~rec_num, ~mass,
                     1, "C-1_1", 83028, 30.05,
                     2, "C-1_2", 83028, 36.00,
                     3, "C-1_3", 83028, 31.88,
                     4, "C-2_1", 1082, 32.57,
                     5, "C-2_2", 1082, 35.50,
                     6, "C-2_3", 1082, 34.10,
                     7, "NOSAMS-2_1", 38809, 35.20,
                     8, "NOSAMS-2_2", 38809, 32.19,
                     9, "NOSAMS-2_3", 38809, 33.20
                     )

write_csv(carb_data, here("data/carb_data_2021-04-15.csv"))
carb_data
```

### Data

Normalizing to gas standards in pos 25, 26, 36 and 37.

```{r}
data <- get_hgis_data(here("data/USAMS041521R.txt"), as.Date("2021-04-16")) 
```

#### Currents and ratios vs time.

```{r}
data %>% 
  filter(!(Pos %in% 21:24)) %>% 
  plot_hgis_time(normFm, ce*normFm) +
  ggtitle("Sample ratio vs. time")
```

```{r}
data %>% 
  filter(!(Pos %in% 21:24)) %>% 
  plot_hgis_time(he12C)
  ggtitle("Sample current vs time")
```

#### Summary of normalized data per sample

```{r}
std <- getStdTable()

ds <- data %>% 
  filter(!outlier) %>% 
  sum_hgis() %>% 
  mutate(rec_num = case_when(str_starts(Sample.Name, "LiveGas") ~ 101730,
                             str_starts(Sample.Name, "C-1") ~ 83028,
                             str_starts(Sample.Name, "TIRI-F") ~ 2138,
                             str_starts(Sample.Name, "TIRI-I") ~ 17185,
                             str_starts(Sample.Name, "C-2") ~ 1082,
                             str_starts(Sample.Name, "NOSAMS-2") ~ 38809,
                             str_starts(Sample.Name, "DeadGas") ~ 72446)) %>% 
 left_join(select(std, rec_num, fm_consensus), by = "rec_num") %>% 
 mutate(fm_consensus = case_when(rec_num == 101730 ~ 1.0398,
                                 rec_num == 72446 ~ 0.0013,
                                 TRUE ~ fm_consensus))

ds[-(1:4),-3]
```

```{r}
data %>% 
  filter(as.numeric(Pos) > 4) %>% 
  ggplot(aes(pos_name, normFm)) +
  geom_boxplot(fill = "slategray1") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) 
```

### Agreement of replicates

Summary statistics to look at reproducibility, accuracy between
targets/vials.

```{r}
ds %>% 
  filter(as.numeric(Pos) > 4) %>% 
  mutate(Name = str_remove(Sample.Name, "_.$")) %>% 
  group_by(Name, fm_consensus) %>% 
  summarize(Fm.mean = mean(mean),
            Fm.sd = sd(mean),
            Cur.mean = mean(Cur),
            N = n()) 
```

#### Blank correction

Testing correcting for blank using mass-independent large blank
correction.

Using C-1 with no adjustment for graphite value of blank.

```{r}
fmblank <- ds %>%
  filter(str_starts(Sample.Name, "C-1")) %>% 
  pull(mean) %>% 
  mean()

fmblank
ds <- ds %>% 
  mutate(fm_lbc = doLBC(mean, fmblank, 1.0398),
         diff = fm_lbc - fm_consensus,
         normFm = normFm(fm_lbc, fm_consensus),
         sigma = sigma(fm_lbc, fm_consensus, merr))

ds %>%
  filter(as.numeric(Pos) > 4) %>% 
  select(Sample.Name, Cur, fm_consensus, mean, merr, fm_lbc, diff)

ds %>% 
write_csv(here("data_analysed/USAMS041521.csv"))
```

```{r}
ggplot(ds, aes(fm_consensus, diff)) +
  geom_hline(yintercept = 0) +
  geom_errorbar(aes(ymin = diff - merr, ymax = diff + merr)) +
  geom_point() +
  ggtitle("Difference in LBC-corrected vs consensus")
```

```{r}
ggplot(ds, aes(fm_consensus, sigma)) +
  geom_hline(yintercept = 0) +
  geom_point() +
  ggtitle("Sigma of LBC-corrected vs consensus")
```

There may be some issues with normalization. The initial two standards
on both of the last carbonate runs were lower than the two at the end of
the run. On 4/9, first two standards were Fm 1.03 and 0.99, last two
were 1.05 and 1.07. On 4/16, first two were 1.02 and 1.04, last two were
both 1.05. NOSAMS2 were run first on 4/9 and were high after blank
correction. On 4/16 they were run last and were low.

# Work for SNEAP 2021 presentation

## Combining data

A weak attempt at combining data. The solution is an analysis script to
normalize, average runs, and blank correct by wheel.

ADD MOAR DATA

```{r}
data<- get_hgis_data(here("data/USAMS101320R.txt"), as.Date("2020-11-17")) %>% 
  mutate(Num = ifelse(Pos %in% c(2, 4), "S", ifelse(Num == "S", "U", Num)))
stdrat <- mean(data$cor1412he[data$Num == "S" & !data$outlier])
data1013 <- data %>% mutate(normFm = norm_gas(cor1412he, stdrat))

data <- get_hgis_data(here("data/USAMS120320R.txt"), as.Date("2020-12-04")) %>% 
  mutate(Num = ifelse(Pos %in% c(2, 4), "S", ifelse(Num == "S", "U", Num)))
stdrat <- mean(data$cor1412he[data$Num == "S" & !data$outlier])
data1204 <- data %>% mutate(normFm = norm_gas(cor1412he, stdrat))

data1211 <- get_hgis_data(here("data/USAMS120320R.txt")) %>% 
  filter(Pos %in% 1:4 | as.Date(ts) == "2020-12-11")

data <- get_hgis_data(here("data/USAMS120320R.txt"), as.Date("2020-12-18")) %>% 
  filter(Pos != 0) %>% 
  mutate(Num = ifelse(Pos %in% c(2, 4), "S", ifelse(Num == "S", "U", Num)))
stdrat <- mean(data$cor1412he[data$Num == "S" & !data$outlier])
data1218 <- data %>% mutate(normFm = norm_gas(cor1412he, stdrat))

data0108 <- get_hgis_data(here("data/USAMS120320R.txt"), as.Date("2021-01-08"))

# no solid standards
# df <- get_hgis_data(here("data/USAMS120320R.txt"), as.Date("2021-01-15")) 
# stdrat <- mean(df$cor1412he[df$pos_name == "8 - LiveGasOS" & !df$outlier])
# data0115 <- mutate(df, normFm = norm_gas(cor1412he, stdrat))

data <- get_hgis_data(here("data/USAMS020521R.txt")) %>% 
  mutate(Num = ifelse(Pos %in% c(2, 4), "S", ifelse(Num == "S", "U", Num)))
stdrat <- mean(data$cor1412he[data$Num == "S" & !data$outlier])
data205 <- data %>% mutate(normFm = norm_gas(cor1412he, stdrat))

data <- get_hgis_data(here("data/USAMS030421R.txt"), as.Date("2021-03-05")) %>% 
  mutate(Num = ifelse(Pos %in% c(2, 4), "S", ifelse(Num == "S", "U", Num)))
stdrat <- mean(data$cor1412he[data$Num == "S" & !data$outlier])
data304 <- data %>% mutate(normFm = norm_gas(cor1412he, stdrat))

data <- get_hgis_data(here("data/USAMS040121R.txt"), as.Date("2021-04-09")) %>% 
  mutate(Num = ifelse(Pos %in% c(22, 24), "S", ifelse(Num == "S", "U", Num)))
stdrat <- mean(data$cor1412he[data$Num == "S" & !data$outlier])
data409 <- data %>% mutate(normFm = norm_gas(cor1412he, stdrat))

data <- get_hgis_data(here("data/USAMS041521R.txt"), as.Date("2021-04-16"))  %>% 
  mutate(Num = ifelse(Pos %in% c(22, 24), "S", ifelse(Num == "S", "U", Num)))
stdrat <- mean(data$cor1412he[data$Num == "S" & !data$outlier])
data415 <- data %>% mutate(normFm = norm_gas(cor1412he, stdrat))

std <- getStdTable()

data <- rbind(data1013, data1204, data1211, data1218, data205, data304, data409, data415) %>% 
  mutate(rec_num = case_when(str_detect(Sample.Name, "LiveGas") ~ 101730,
                             str_starts(Sample.Name, "C-1") ~ 83028,
                             str_starts(Sample.Name, "C1") ~ 83028,
                             str_starts(Sample.Name, "TIRI-F") ~ 2138,
                             str_starts(Sample.Name, "TIRI-I") ~ 17185,
                             str_starts(Sample.Name, "C-2") ~ 1082,
                             str_starts(Sample.Name, "C2") ~ 1082,
                             str_starts(Sample.Name, "NOSAMS") ~ 38809,
                             str_detect(Sample.Name, "DeadGas") ~ 72446)) %>% 
 left_join(select(std, rec_num, fm_consensus), by = "rec_num") %>% 
 mutate(fm_consensus = case_when(rec_num == 101730 ~ 1.0398,
                                 rec_num == 72446 ~ 0.0013,
                                 TRUE ~ fm_consensus))

data %>% distinct(wheel, Pos, Sample.Name, fm_consensus, rec_num) %>% filter(fm_consensus > 0.9)
```

### Current dependent blank with combined data

```{r}
unique(data$Sample.Name)
```

Modern

```{r}
data %>% 
  filter(!(he12C < 2.5 & normFm > 1.05) & fm_consensus > 0.9) %>% 
  #filter(fm_consensus > 0.9) %>% 
  #filter(str_detect(Sample.Name, "LiveGas")) %>% 
  ggplot(aes(he12C, normFm)) + 
  #geom_smooth(method = "lm") +
  geom_point()

data %>% select(wheel, Pos, Sample.Name, he12C, fm_consensus, normFm) %>% filter(he12C < 2.5 & normFm > 1.05, fm_consensus > 0.9)
```

Dead

```{r}
data %>% 
  #filter(str_detect(Sample.Name, "DeadGas")) %>% 
  filter(fm_consensus < 0.1 & normFm < 0.3 & Sample.Name != "DilDeadGas6") %>% 
  ggplot(aes(he12C, normFm)) + 
  geom_point()

data %>% select(wheel, Pos, Sample.Name, he12C, fm_consensus, normFm) %>% filter(he12C < 1 & normFm < .1, fm_consensus < 0.3)

filter(data, Sample.Name == "DilDeadGas6")
```

Summarize by sample

```{r}
ds <- sum_hgis(data) %>% 
  filter((str_detect(Sample.Name, "LiveGas") & Cur > 1) |
         (str_detect(Sample.Name, "DeadGas") & Cur > 1) |
          str_detect(Sample.Name, "Blank")) %>% 
  mutate(Gas = case_when(str_detect(Sample.Name, "LiveGas") ~ "LiveGas",
                         str_detect(Sample.Name, "DeadGas") ~ "DeadGas",
                         TRUE ~ "Blank"),
         Cur_inv = 1/Cur)
```

Modern

```{r}
live <- ds %>% 
  filter(str_detect(Sample.Name, "LiveGas") & Cur > 1) %>% 
  ggplot(aes(Cur, mean)) + 
  # geom_smooth(method = "lm") +
  geom_pointrange(aes(ymin = mean - merr, ymax = mean + merr)) +
  geom_point()
```

Dead

```{r}
dead <- ds %>% 
  filter(str_detect(Sample.Name, "DeadGas")) %>% 
  ggplot(aes(Cur, mean)) + 
  #geom_smooth(method = "lm") +
  geom_pointrange(aes(ymin = mean - merr, ymax = mean + merr)) +
  geom_point()

filter(ds, Cur < 1)
```

```{r}
live / dead
```

### Modelling current dependent blank

#### Fit linear models

```{r}
fits <- ds %>% 
  ungroup() %>% 
  nest(data = -Gas) %>% 
  mutate(fit = map(data, ~lm(mean ~ Cur_inv, data = .x)),
         tidied = map(fit, tidy)) %>% 
  unnest(tidied)

blankfit <- fits %>% 
  select(Gas, term, estimate) %>% 
  pivot_wider(names_from = c(Gas, term), values_from = estimate) %>% 
  mutate(inv_m_blank = -(`DeadGas_(Intercept)` - `LiveGas_(Intercept)`)/(DeadGas_Cur_inv - LiveGas_Cur_inv),
         Fm_blank = `DeadGas_(Intercept)` + DeadGas_Cur_inv * inv_m_blank,
         m_blank = 1/inv_m_blank)

blankfit %>% 
  select(Fm_blank, m_blank)

blankfit
```

#### Fm vs. inverse mass with fits.

```{r}
ggplot(ds, aes(Cur_inv, mean, color = Gas)) +
  geom_abline(slope = blankfit$LiveGas_Cur_inv, 
              intercept = blankfit$`LiveGas_(Intercept)`) +
  geom_abline(slope = blankfit$DeadGas_Cur_inv, 
              intercept = blankfit$`DeadGas_(Intercept)`) +
  geom_pointrange(aes(ymin = mean - merr, ymax = mean + merr), size = .3) + 
  xlim(0, 10) +
  #ylim(0, 1.2)
  ggtitle("Fm of live and dead CO2 vs. inverse 12C current",
          subtitle = "Fit lines should cross at Fm and current of blank")
```

### Gas standards

Use above data with gas standards normalized to solid OX-I

```{r}
rownames(ds)
ds %>% 
  filter(Gas == "LiveGas") %>% 
  ggplot(aes(x = as.factor(as.numeric(rownames(.))), y = mean)) + 
  geom_pointrange(aes(ymin = mean - merr, ymax = mean + merr))
```

## 2021-05-12 Vial dillution tests

### Data

```{r}
data <- get_hgis_data(here("data/USAMS041521R.txt"), as.Date("2021-05-12"))
```

#### Currents and ratios vs time.

```{r}
data %>% 
  plot_hgis_time(cor1412he, ce*cor1412he) +
  ggtitle("Sample ratio vs. time")
```

```{r}
data %>% 
  plot_hgis_time(he12C) +
  ggtitle("Sample current vs. time")
```

## 2021-05-14 Carbonate samples

### Samples

```{r}
carb_data <- tribble(~vial, ~name, ~rec_num, ~mass,
                     1, "C-1_1", 83028, 31.61,
                     2, "C-1_2", 83028, 34.30,
                     2, "C-1_3", 83028, 33.17,
                     4, "TIRI-I_1", 17185, 32.74,
                     5, "TIRI-I_2", 17185, 31.58,
                     6, "TIRI-I_3", 17185, 33.25,
                     7, "NOSAMS-2_1", 38809, 33.31,
                     8, "NOSAMS-2_2", 38809, 31.89,
                     9, "NOSAMS-2_3", 38809, 36.47
                     )

write_csv(carb_data, here("data/carb_data_2021-05-14.csv"))
carb_data
```

### Data

Normalizing to gas standards in pos 25, 26, 36 and 37.

```{r}
data <- get_hgis_data(here("data/USAMS051421R.txt"), as.Date("2021-05-14")) %>% 
  mutate(Num = ifelse(Pos %in% c(15, 65), "U", Num))
stdrat <- mean(data$cor1412he[data$Num == "S" & !data$outlier])
data <- data %>% 
  mutate(normFm = norm_gas(cor1412he, stdrat))
```

#### Currents and ratios vs time.

```{r}
data %>% 
  filter(!(Pos %in% 1:4)) %>% 
  plot_hgis_time(normFm, ce*normFm) +
  ggtitle("Sample ratio vs. time")
```

```{r}
data %>% 
  filter(!(Pos %in% 1:4)) %>% 
  plot_hgis_time(he12C)
  ggtitle("Sample current vs time")
```

#### Summary of normalized data per sample

```{r}
std <- getStdTable()

ds <- data %>% 
  filter(!outlier) %>% 
  sum_hgis() %>% 
  mutate(rec_num = case_when(str_starts(Sample.Name, "LiveGas") ~ 101730,
                             str_starts(Sample.Name, "C1") ~ 83028,
                             str_starts(Sample.Name, "TIRI-F") ~ 2138,
                             str_starts(Sample.Name, "TIRI-I") ~ 17185,
                             str_starts(Sample.Name, "C-2") ~ 1082,
                             str_starts(Sample.Name, "NOSAMS2") ~ 38809,
                             str_starts(Sample.Name, "DeadGas") ~ 72446)) %>% 
 left_join(select(std, rec_num, fm_consensus), by = "rec_num") %>% 
 mutate(fm_consensus = case_when(rec_num == 101730 ~ 1.0398,
                                 rec_num == 72446 ~ 0.0013,
                                 TRUE ~ fm_consensus))

ds[-(1:4),]
```

```{r}
data %>% 
  filter(as.numeric(Pos) > 4) %>% 
  ggplot(aes(pos_name, normFm)) +
  geom_boxplot(fill = "slategray1") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) 
```

### Agreement of replicates

Summary statistics to look at reproducibility, accuracy between
targets/vials.

```{r}
ds %>% 
  filter(as.numeric(Pos) > 4) %>% 
  mutate(Name = str_remove(Sample.Name, "_.$")) %>% 
  group_by(Name, fm_consensus) %>% 
  summarize(Fm.mean = mean(mean),
            Fm.sd = sd(mean),
            Cur.mean = mean(Cur),
            N = n()) 
```

#### Blank correction

Testing correcting for blank using mass-independent large blank
correction.

Using C-1 with no adjustment for graphite value of blank.

```{r}
fmblank <- ds %>%
  filter(str_starts(Sample.Name, "C1")) %>% 
  pull(mean) %>% 
  mean()

fmblank
ds <- ds %>% 
  mutate(fm_lbc = doLBC(mean, fmblank, 1.0398),
         diff = fm_lbc - fm_consensus,
         normFm = normFm(fm_lbc, fm_consensus),
         sigma = sigma(fm_lbc, fm_consensus, merr))

ds %>%
  filter(as.numeric(Pos) > 4) %>% 
  select(Sample.Name, Cur, fm_consensus, mean, merr, fm_lbc, diff)

ds %>% 
write_csv(here("data_analysed/USAMS051421.csv"))
```

```{r}
ggplot(ds, aes(fm_consensus, diff)) +
  geom_hline(yintercept = 0) +
  geom_errorbar(aes(ymin = diff - merr, ymax = diff + merr)) +
  geom_point() +
  ggtitle("Difference in LBC-corrected vs consensus")
```

```{r}
ggplot(ds, aes(fm_consensus, sigma)) +
  geom_hline(yintercept = 0) +
  geom_point() +
  ggtitle("Sigma of LBC-corrected vs consensus")
```

## 2021-05-20 Dillution tests

### Data

```{r}
data <- get_hgis_data(here("data/USAMS051421R.txt"), as.Date("2021-05-20"))
```

#### Currents and ratios vs time.

```{r}
data %>% 
  plot_hgis_time(cor1412he, ce*cor1412he) +
  ggtitle("Sample ratio vs. time")
```

```{r}
data %>% 
  plot_hgis_time(he12C) +
  ggtitle("Sample current vs. time")
```

## 2021-05-28 Pendleton samples

First run of Simon Pendleton's samples, along with C-1, C-2 and modern standard. Reworked data reduction and did analysis in separate file.

## 2021-06-04 Small(er) sample tests

Tested 5, 10, and 15 mg of C-2 carbonate in vials backfilled with Helium.

### Method

1. Weigh sample into vial
2. Evacuate and acidify as normal with gilson
3. Backfill to 1 ATM with He-filled syringe 2h after acidification.

Allowed vacuum in vial to pull in He from syringe, then added 1 additional mL He. Should calculate volume CO2 in vial and measure He added to check that things make sense.

Used same capillaries as usual, but decreased He displacement flow to 140 ul/min. Source capillary flow assumed to be around 30ul/min.

### Samples

```{r}
carb_data <- tribble(~vial, ~name, ~rec_num, ~mass,
                     1, "C-2_1", 1082, 6.03,
                     2, "C-2_2", 1082, 11.10,
                     3, "C-2_3", 1082, 15.52
                     )

write_csv(carb_data, here("data/carb_data_2021-06-04.csv"))
carb_data
```

### Data

```{r}
df <- get_hgis_data(here("data/USAMS052621R.txt"), as.Date("2021-06-04")) #, standards)
```

#### Currents and ratios vs time.

```{r}
df %>% 
  filter(!(Pos %in% 1:4)) %>% 
  plot_hgis_time(normFm, ce*normFm) +
  ggtitle("Sample ratio vs. time")
```

```{r}
df %>% 
  filter(!(Pos %in% 1:4)) %>% 
  plot_hgis_time(he12C) +
  ggtitle("Sample current vs time")
```

### Results

Note: no blanks run, so need to use average blank if LBC needed.

```{r}
df_sum <- sum_hgis_targets(df) %>% 
  norm_hgis()
```

```{r}
df_sum %>% 
  filter(as.numeric(Pos) > 4) %>% 
  mutate(sig_normFm = max_err * normFm) %>% 
  select(Pos, Sample.Name, he12C, n_runs, norm_ratio, sig_norm_ratio) %>% 
  arrange(Sample.Name)
```

```{r}
df_sum %>% 
  filter(as.numeric(Pos) > 4) %>% 
  ggplot(aes(Sample.Name, norm_ratio)) +
    geom_pointrange(aes(ymin = norm_ratio - sig_norm_ratio, ymax = norm_ratio + sig_norm_ratio),
                    size = .5) + 
    theme(axis.text.x = element_text(angle = 45, hjust=1)) +
      labs(x = NULL,
           y = "Fraction Modern")
plot_hgis_summary()
```

## Agreement of replicates

Compare results for samples with more than one replicate.

```{r}
df_sum %>% 
  filter(as.numeric(Pos) > 4) %>% 
  compare_replicates()
```

## Agreement with consensus

```{r}
df_sum %>% 
  filter(as.numeric(Pos) > 4,
         !is.na(fm_consensus)) %>% 
  select(Sample.Name, fm_consensus, norm_ratio, sig_norm_ratio) %>% 
  mutate(Fm_diff = norm_ratio - fm_consensus,
         sigma = amstools::sigma(norm_ratio, fm_consensus, sig_norm_ratio)) %>% 
  arrange(Sample.Name)
```

### 2021-06-10 inst/rmd/opensplittests.Rmd
Add section for new sample run.

@@
 ```{r}
-carb_data <- tribble(~vial, ~name, ~rec_num, ~mass,
+carb_data <- tribble(~, ~name, ~rec_num, ~mass,
                      1, "C-2_1", 1082, 6.03,
@@
 ```
+
+## 2021-06-11 Small(er) sample tests
+
+Tested 8mg of C-1, C-2, and NOSAMS2 carbonate in vials backfilled with Helium.
+
+### Method
+
+1. Weigh sample into vial
+2. Evacuate and acidify as normal with gilson
+3. Backfill to 1 ATM with He-filled syringe 2h after acidification.
+
+Allowed vacuum in vial to pull in He from syringe, then added 1 additional mL He. Should calculate volume CO2 in vial and measure He added to check that things make sense.
+
+Used same capillaries as usual, but decreased He displacement flow to 140 ul/min. Source capillary flow assumed to be around 30ul/min.
+
+### Samples
+
+```{r}
+carb_data <- tribble(~name, ~rec_num, ~carbonate_mass, ~helium_added,
+                     "C-1_1", 83028, 8.29,
+                     "C-1_2", 83028, 9.01,
+                     "C-1_3", 83028, 9.00,
+                     "C-2_1", 1082, 8.23,
+                     "C-2_2", 1082, 8.77,
+                     "C-2_3", 1082, 7.34,
+                     "NOSAMS-2_1", 38809, 7.72,
+                     "NOSAMS-2_2", 38809, 7.69,
+                     "NOSAMS-2_3", 38809, 8.86,
+                     )
+
+write_csv(carb_data, here("data/carb_data_2021-06-04.csv"))
+insert_samples(carb_data)
+carb_data
+```
+
+### Data
+
+```{r}
+df <- process_hgis_results(here("data/USAMS061121R.txt"), as.Date("2021-06-11")) #, standards)
+
+insert_raw(df)
+```
+
+#### Currents and ratios vs time.
+
+```{r}
+df %>% 
+  filter(!(Pos %in% 1:4)) %>% 
+  plot_hgis_time(normFm, ce*normFm) +
+  ggtitle("Sample ratio vs. time")
+```
+
+```{r}
+df %>% 
+  filter(!(Pos %in% 1:4)) %>% 
+  plot_hgis_time(he12C) +
+  ggtitle("Sample current vs time")
+```
+
+### Results
+
+Note: no blanks run, so need to use average blank if LBC needed.
+
+```{r}
+df_sum <- sum_hgis_targets(df) %>% 
+  norm_hgis() %>% 
+  blank_cor_hgis()
+
+insert_results(df_sum)
+```
+
+```{r}
+df_sum %>% 
+  filter(as.numeric(Pos) > 4) %>% 
+  mutate(sig_normFm = max_err * normFm) %>% 
+  select(Pos, Sample.Name, he12C, n_runs, norm_ratio, sig_norm_ratio) %>% 
+  arrange(Sample.Name)
+```
+
+```{r}
+df_sum %>% 
+  filter(as.numeric(Pos) > 4) %>% 
+  ggplot(aes(Sample.Name, norm_ratio)) +
+    geom_pointrange(aes(ymin = norm_ratio - sig_norm_ratio, ymax = norm_ratio + sig_norm_ratio),
+                    size = .5) + 
+    theme(axis.text.x = element_text(angle = 45, hjust=1)) +
+      labs(x = NULL,
+           y = "Fraction Modern")
+plot_hgis_summary()
+```
+
+## Agreement of replicates
+
+Compare results for samples with more than one replicate.
+
+```{r}
+df_sum %>% 
+  filter(as.numeric(Pos) > 4) %>% 
+  compare_replicates()
+```
+
+## Agreement with consensus
+
+```{r}
+df_sum %>% 
+  filter(as.numeric(Pos) > 4,
+         !is.na(fm_consensus)) %>% 
+  select(Sample.Name, fm_consensus, norm_ratio, sig_norm_ratio) %>% 
+  mutate(Fm_diff = norm_ratio - fm_consensus,
+         sigma = amstools::sigma(norm_ratio, fm_consensus, sig_norm_ratio)) %>% 
+  arrange(Sample.Name)
+```

### 2021-06-10 inst/rmd/opensplittests.Rmd
Finish sample data.

@@
 carb_data <- tribble(~name,         ~rec_num, ~carbonate_mass, ~helium_added,
-                     "C-1_1", 83028, 8.29,
-                     "C-1_2", 83028, 9.01,
-                     "C-1_3", 83028, 9.00,
-                     "C-2_1", 1082, 8.23,
-                     "C-2_2", 1082, 8.77,
-                     "C-2_3", 1082, 7.34,
-                     "NOSAMS-2_1", 38809, 7.72,
-                     "NOSAMS-2_2", 38809, 7.69,
-                     "NOSAMS-2_3", 38809, 8.86,
+                     "C-1_1",       83028,    8.29,            5,
+                     "C-1_2",       83028,    9.01,            5,
+                     "C-1_3",       83028,    9.00,            5,
+                     "C-2_1",        1082,    8.23,            5,
+                     "C-2_2",        1082,    8.77,            5,
+                     "C-2_3",        1082,    7.34,            5,
+                     "NOSAMS-2_1",  38809,    7.72,            5,
+                     "NOSAMS-2_2",  38809,    7.69,            5,
+                     "NOSAMS-2_3",  38809,    8.86,            5
                      )
@@
 df_sum %>% 
-  filter(as.numeric(Pos) > 4) %>% 
-  ggplot(aes(Sample.Name, norm_ratio)) +
-    geom_pointrange(aes(ymin = norm_ratio - sig_norm_ratio, ymax = norm_ratio + sig_norm_ratio),
-                    size = .5) + 
-    theme(axis.text.x = element_text(angle = 45, hjust=1)) +
-      labs(x = NULL,
-           y = "Fraction Modern")
   plot_hgis_summary()
@@
 ```{r}
-df_sum %>% 
+cons <- df_sum %>% 
   filter(as.numeric(Pos) > 4,
          !is.na(fm_consensus)) %>% 
-  select(Sample.Name, fm_consensus, norm_ratio, sig_norm_ratio) %>% 
-  mutate(Fm_diff = norm_ratio - fm_consensus,
-         sigma = amstools::sigma(norm_ratio, fm_consensus, sig_norm_ratio)) %>% 
+  select(Sample.Name, fm_consensus, fm_corr, sig_fm_corr) %>% 
+  mutate(Fm_diff = fm_corr - fm_consensus,
+         sigma = amstools::sigma(fm_corr, fm_consensus, sig_fm_corr)) %>% 
   arrange(Sample.Name)
-```
+
+knitr::kable(cons)
+```

### 2021-06-13 inst/rmd/opensplittests.Rmd
Add latest run.

@@
 ```{r}
-carb_data <- tribble(~, ~name, ~rec_num, ~mass,
+carb_data <- tribble(~number, ~name, ~rec_num, ~mass,
                      1, "C-2_1", 1082, 6.03,
@@
 
-insert_raw(df)
+#insert_raw(df)
 ```
@@
 df %>% 
-  filter(!(Pos %in% 1:4)) %>% 
-  plot_hgis_time(normFm, ce*normFm) +
+  filter(pos > 2) %>% 
+  plot_hgis_time(norm_ratio, sig_norm_ratio, outlier = !.$ok_calc) +
   ggtitle("Sample ratio vs. time")
@@
 df %>% 
-  filter(!(Pos %in% 1:4)) %>% 
-  plot_hgis_time(he12C) +
+  filter(pos > 2) %>% 
+  plot_hgis_time(he12C, outlier = !.$ok_calc) +
   ggtitle("Sample current vs time")
@@
 
-insert_results(df_sum)
+#insert_results(df_sum)
 ```
@@
 df_sum %>% 
-  filter(as.numeric(Pos) > 4) %>% 
-  mutate(sig_normFm = max_err * normFm) %>% 
-  select(Pos, Sample.Name, he12C, n_runs, norm_ratio, sig_norm_ratio) %>% 
-  arrange(Sample.Name)
+  filter(pos > 2) %>% 
+  mutate(he12C = he12C*1E6) %>% 
+  select(Positon = pos, `Sample Name` = sample_name, `12C Current (uA)` = he12C, n_runs, fm_corr, sig_fm_corr) %>% 
+  arrange(`Sample Name`)
 ```
@@
 df_sum %>% 
-  filter(as.numeric(Pos) > 4) %>% 
+  filter(pos > 2) %>% 
   compare_replicates()
@@
 cons <- df_sum %>% 
-  filter(as.numeric(Pos) > 4,
+  filter(pos > 2,
          !is.na(fm_consensus)) %>% 
-  select(Sample.Name, fm_consensus, fm_corr, sig_fm_corr) %>% 
-  mutate(Fm_diff = fm_corr - fm_consensus,
+  select(sample_name, fm_consensus, fm_corr, sig_fm_corr) %>% 
+  mutate(fm_diff = fm_corr - fm_consensus,
          sigma = amstools::sigma(fm_corr, fm_consensus, sig_fm_corr)) %>% 
-  arrange(Sample.Name)
-
+  arrange(sample_name)
+cons  
 knitr::kable(cons)

### 2021-06-14 doc/hgis_data_reduction.Rmd
Update methods.

@@
 
-### Read raw file
+## Read raw file
 
@@
 
-### Apply 13C correction to each run of a target
+## Apply 13C correction to each run of a target
 
@@
 
-### Flag outliers by interquartile distance
+## Flag outliers by interquartile distance
 
-Outliers are defined as points 1.5 * interquartile range above the 3rd quartile or below the 1st quartile, and are dropped from the analysis.
+Outliers are defined as points 1.5 * interquartile range above the 3rd quartile or below the 1st quartile.
+25% of values are expected to lie below the first quartile and 25% above the 3rd quartile.
+Interquartile range (IQR) is the difference between the 3rd and 1st quartile.
+Outliers are not currently dropped from the analysis.
 
-### Determine per-target means, internal and external errors
+## Determine per-target means, internal and external errors
 
@@
 
-### Normalize targets using mean of standards and propagate errors
+## Normalize targets using mean of standards and propagate errors
 
@@
 
-$$ R_{norm} = \frac{R_{sample}}{R_{std}}*1.0398$$
+$$ R_{norm} = \frac{R_{sample}\cdot R_{exp}}{R_{std}}$$
 
-Normalization errors are propagated using this equation:
+If the expected ratio ($R_{exp}$) of the standard is taken as a measured value with an error, rather than a constant or assigned ratio, normalization errors are propagated using this equation:
 
-$$ \sigma_{norm} = \sqrt{\frac{\sigma_{sample}^2}{R_{sample}^2} + \frac{\sigma_{std}^2}{R_{std}^2}} $$
+$$ \sigma_{norm} = \sqrt{\frac{\sigma_{sample}^2}{R_{sample}^2} + \frac{\sigma_{std}^2}{R_{std}^2} +
+\frac{\sigma_{exp}^2}{R_{exp}^2}} $$
 
-### Apply a large blank correction using the mean of blanks and propagate errors.
+## Apply a large blank correction using the mean of blanks and propagate errors.
 

### 2021-06-14 inst/rmd/HGIS_pendleton_2021-05-28.Rmd
Update methods.

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(message = FALSE)
+knitr::opts_chunk$set(message = FALSE, echo = FALSE)
 ```
@@
 - Using blanks in positions `r blanks` for blank correction.
-- Outliers are defined as points 1.5 * interquartile range above the 3rd quartile or below the 1st quartile, and are dropped from the analysis.
+- Outliers are defined as points 1.5 * interquartile range above the 3rd quartile or below the 1st quartile.
+- Outliers are NOT dropped from the analysis.
 
@@
 1. Normalize targets using mean of standards and propagate errors
-1. Apply a large blank correction using the mean of blanks and propagate errors.
+1. Apply a large blank correction using the mean of blanks and propagate errors
 
@@
 ```{r}
-df <- get_hgis_data(file, date, standards)
+df <- process_hgis_results(file, date, standards)
 write_csv(df, here("data_analysed/HGIS_2021-05-28_raw.csv"))
@@
 df %>% 
-  filter(!(Pos %in% 1:4)) %>% 
-  plot_hgis_time(normFm, ce*normFm) +
-  ggtitle("Sample ratio vs. time")
+  filter(!(pos %in% 1:4)) %>% 
+  plot_hgis_time(norm_ratio, sig_norm_ratio, outlier = !.$ok_calc) +
+  labs(title = "Sample ratio vs. time",
+       y = "Fraction modern")
 ```
@@
 df %>% 
-  filter(!(Pos %in% 1:4)) %>% 
-  plot_hgis_time(he12C) +
-  ggtitle("Sample current vs time")
+  filter(!(pos %in% 1:4)) %>% 
+  plot_hgis_time(he12C * 1E6, outlier = !.$ok_calc) +
+  labs(title = "Sample current vs time",
+       x = "Time (s)",
+       y = "Current (uA)")
 ```
@@
 ```{r}
-df_sum <- sum_hgis_targets(df) %>% 
+df_sum <- sum_hgis_targets(df, remove_outliers = FALSE) %>% 
   norm_hgis() %>% 
@@
 df_sum %>% 
-  filter(as.numeric(Pos) > 4) %>% 
-  mutate(sig_normFm = max_err * normFm) %>% 
-  select(Pos, Sample.Name, he12C, n_runs, fm_corr, sig_fm_corr) %>% 
-  arrange(Sample.Name) %>% 
+  filter(as.numeric(pos) > 4) %>% 
+  mutate(he12C = he12C * 1E6) %>% 
+  select(Pos = pos, Name = sample_name, `Current (uA)` = he12C, `N runs` = n_runs, Fm = fm_corr, `Fm error` = sig_fm_corr) %>% 
+  arrange(Name) %>% 
   knitr::kable()
@@
 df_sum %>% 
-  filter(as.numeric(Pos) > 4) %>% 
+  filter(as.numeric(pos) > 4) %>% 
 plot_hgis_summary()
@@
 df_sum %>% 
-  filter(as.numeric(Pos) > 4) %>% 
+  filter(as.numeric(pos) > 4) %>% 
   compare_replicates() %>% 
+  select(Name, `Mean current (uA)` = he12C_mean, 
+         `SD of current` = he12C_sd, `Mean Fm` = fm_corr_mean, 
+         `SD of Fm` = fm_corr_sd, `Mean sample error` = sig_fm_corr_mean) %>% 
   knitr::kable()
@@
 cons <- df_sum %>% 
-  filter(as.numeric(Pos) > 4,
+  filter(as.numeric(pos) > 4,
          !is.na(fm_consensus)) %>% 
-  select(Sample.Name, fm_consensus, fm_corr, sig_fm_corr) %>% 
+  select(sample_name, fm_consensus, fm_corr, sig_fm_corr) %>% 
   mutate(Fm_diff = fm_corr - fm_consensus,
          sigma = amstools::sigma(fm_corr, fm_consensus, sig_fm_corr)) %>% 
-  arrange(Sample.Name)
+  arrange(sample_name) %>% 
+  select(Name = sample_name, 
+         #`Mean current (uA)` = he12C_mean, 
+         #`SD of current` = he12C_sd, 
+         `Consensus Fm` = fm_consensus,
+         Fm = fm_corr, 
+         `Fm error` = sig_fm_corr,
+         `Difference` = Fm_diff,
+         Sigma = sigma)
 

### 2021-06-14 inst/rmd/carbonate_reanalysis.Rmd
Add reanalysis script.

---
title: "Carbonate reanalysis"
author: "Brett Longworth"
date: "6/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Reanalysis of carbonate data with most current data reduction method.

```{r}
library(tidyverse)
library(here)
library(amstools)
library(HybridGIS)
```

```{r}
wheels <- c("020521", "030421", "040121", "041521", "051421", "052621", "061121")
files <- here("data", paste0("USAMS", wheels, "R.txt"))
dates <- as.Date(c("2021-02-05", "2021-03-05", "2021-04-09", "2021-04-16", "2021-05-14", "2021-05-28", "2021-06-11"))
stds <- list(c(5,8), NULL, c(26,37), NULL, c(25,26,36,37), NULL, NULL)
blanks <- list(7, c(7,8), c(28,29,30), c(27,28,29), c(20,35,50), NULL, NULL)

results <- pmap(list(files, dates, stds, blanks), reduce_hgis)
raw_df <- map_dfr(results, 1)
results_df <- map_dfr(results, 2)
```

```{r}
```

### 2021-06-15 inst/rmd/HGIS_small_2021-06-11.Rmd
Add analysis of small sample run.

---
title: "HGIS 2021-06-11 Carbonates"
author: "Brett Longworth"
date: "6/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, echo = FALSE)
```

```{r}
library(tidyverse)
library(here)
library(openxlsx)
library(amstools)
library(HybridGIS)

options(digits = 4)
options(scipen = 3)
```

Data reduction and analysis for carbonate samples run on June 11, 2021. Test run of smaller samples. All samples ~8mg carbonate. Includes C-1, C-2, and NOSAMS2.

Outliers are defined as points 1.5 * interquartile range above the 3rd quartile or below the 1st quartile.

Data reduction proceeds as follows:

1. Read raw file
1. Apply 13C correction to each run of a target
1. Flag outliers by interquartile distance. Outliers are flagged, but NOT removed from the analysis.
1. Determine per-target means, internal and external errors
1. Normalize targets using mean of standards and propagate errors
1. Apply a large blank correction using the mean of blanks and propagate errors.

## raw results

```{r}
df <- process_hgis_results(here("data/USAMS061121R.txt"))
write_csv(df, here("data_analysed/HGIS_2021-06-11_raw.csv"))
write.xlsx(df, here("data_analysed/HGIS_2021-06-11_raw.xlsx"))
```

#### Currents and ratios vs time.

```{r}
df %>% 
  filter(pos > 2) %>% 
  plot_hgis_time(norm_ratio, sig_norm_ratio, outlier = !.$ok_calc) +
  labs(title = "Sample ratio vs. time",
       y = "Fraction modern")
```

```{r}
df %>% 
  filter(pos > 2) %>% 
  plot_hgis_time(he12C * 1E6, outlier = !.$ok_calc) +
  labs(title = "Sample current vs time",
       x = "Time (s)",
       y = "Current (uA)")
```

```{r}
df_sum <- sum_hgis_targets(df, remove_outliers = FALSE) %>% 
  norm_hgis() %>% 
  blank_cor_hgis()

write_csv(df_sum, here("data_analysed/HGIS_2021-06-11_results.csv"))
write.xlsx(df_sum, here("data_analysed/HGIS_2021-06-11_results.xlsx"))
```

## Results

Fm and Fm error are blank corrected Fm and error, respectively.

```{r}
df_sum %>% 
  filter(pos > 2) %>% 
  mutate(he12C = he12C * 1E6) %>% 
  select(Pos = pos, Name = sample_name, `Current (uA)` = he12C, `N runs` = n_runs, Fm = fm_corr, `Fm error` = sig_fm_corr) %>% 
  arrange(Name) %>% 
  knitr::kable()
  # gt::gt()
```

```{r}
df_sum %>% 
  filter(pos > 2) %>% 
plot_hgis_summary()
```

## Agreement of replicates

Compare results for samples with more than one replicate.

```{r}
df_sum %>% 
  filter(pos > 2) %>% 
  compare_replicates() %>% 
  select(Name, `Mean current (uA)` = he12C_mean, 
         `SD of current` = he12C_sd, `Mean Fm` = fm_corr_mean, 
         `SD of Fm` = fm_corr_sd, `Mean sample error` = sig_fm_corr_mean) %>% 
  knitr::kable()
```

## Agreement with consensus

Comparison to expected Fm for sample types with a consensus value.

```{r}
cons <- df_sum %>% 
  filter(pos > 2,
         !is.na(fm_consensus)) %>% 
  select(sample_name, fm_consensus, fm_corr, sig_fm_corr) %>% 
  mutate(Fm_diff = fm_corr - fm_consensus,
         sigma = amstools::sigma(fm_corr, fm_consensus, sig_fm_corr)) %>% 
  arrange(sample_name) %>% 
  select(Name = sample_name, 
         #`Mean current (uA)` = he12C_mean, 
         #`SD of current` = he12C_sd, 
         `Consensus Fm` = fm_consensus,
         Fm = fm_corr, 
         `Fm error` = sig_fm_corr,
         `Difference` = Fm_diff,
         Sigma = sigma)

knitr::kable(cons)
```

### 2021-06-16 inst/rmd/carbonate_reanalysis.Rmd
Pass data reduction options into functions.
fix stds, don't remove outliers, better summaries

@@
 dates <- as.Date(c("2021-02-05", "2021-03-05", "2021-04-09", "2021-04-16", "2021-05-14", "2021-05-28", "2021-06-11"))
-stds <- list(c(5,8), NULL, c(26,37), NULL, c(25,26,36,37), NULL, NULL)
+stds <- list(c(5,8), NULL, c(26,37), NULL, c(10,15,65,70), NULL, NULL)
 blanks <- list(7, c(7,8), c(28,29,30), c(27,28,29), c(20,35,50), NULL, NULL)
 
-results <- pmap(list(files, dates, stds, blanks), reduce_hgis)
+results <- pmap(list(files, dates, stds, blanks), reduce_hgis, remove_outliers = FALSE)
 raw_df <- map_dfr(results, 1)
@@
 ```{r}
+results_df %>% 
+  filter(!is.na(rec_num),
+         sample_type == "B") %>% 
+  select(sample_name, rec_num, wheel, sample_type, fm_consensus, fm_corr, sig_fm_corr) %>% 
+  arrange(pos)
 ```
+
+```{r}
+results_df %>% 
+  filter(!is.na(rec_num),
+         !is.na(sig_fm_corr)) %>% 
+  compare_replicates() %>% 
+  arrange(rec_num)
+```

### 2021-06-16 inst/rmd/HGIS_pendleton_2021-05-28.Rmd
Add comparison to graphite

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(message = FALSE, echo = FALSE)
+knitr::opts_chunk$set(message = FALSE)
 ```
@@
 df <- process_hgis_results(file, date, standards)
-write_csv(df, here("data_analysed/HGIS_2021-05-28_raw.csv"))
-write.xlsx(df, here("data_analysed/HGIS_2021-05-28_raw.xlsx"))
+#write_csv(df, here("data_analysed/HGIS_2021-05-28_raw.csv"))
+#write.xlsx(df, here("data_analysed/HGIS_2021-05-28_raw.xlsx"))
 ```
@@
 
-write_csv(df_sum, here("data_analysed/HGIS_2021-05-28_results.csv"))
-write.xlsx(df_sum, here("data_analysed/HGIS_2021-05-28_results.xlsx"))
+# write_csv(df_sum, here("data_analysed/HGIS_2021-05-28_results.csv"))
+# write.xlsx(df_sum, here("data_analysed/HGIS_2021-05-28_results.xlsx"))
 ```
@@
   mutate(he12C = he12C * 1E6) %>% 
-  select(Pos = pos, Name = sample_name, `Current (uA)` = he12C, `N runs` = n_runs, Fm = fm_corr, `Fm error` = sig_fm_corr) %>% 
+  select(Pos = pos, Name = sample_name, `Current (uA)` = he12C, 
+         `N runs` = n_runs, Fm = fm_corr, `Fm error` = sig_fm_corr) %>% 
   arrange(Name) %>% 
@@
 ```
+
+## Agreement with graphite
+
+Agreement of Haiti core samples. Graphite samples were pretreated with HCl to leach 30% of the sample. HGIS samples not pretreated.
+
+Load Simon's data and compare
+
+```{r}
+haiti <- filter(df_sum, str_detect(sample_name, "HATC1")) %>% 
+  mutate(rec_num = case_when(pos == 14 ~ 171989,
+                             pos == 15 ~ 171995,
+                             pos == 16 ~ 171999))
+
+graphite_results <- getRecOS(c(171999,171995,171989)) %>% 
+  select(rec_num, fm_corr = f_modern, sig_fm_corr = f_ext_error)
+
+combined_results <- haiti %>% 
+  select(sample_name, rec_num, fm_corr, sig_fm_corr) %>% 
+  left_join(graphite_results,
+            by = "rec_num", 
+            suffix = c("_hgis", "_graphite")) %>% 
+  mutate(fm_diff = fm_corr_graphite - fm_corr_hgis,
+         fm_mean = mean(c(fm_corr_hgis, fm_corr_graphite))) %>% 
+  pivot_longer(c(fm_corr_graphite, sig_fm_corr_graphite, 
+                 fm_corr_hgis, sig_fm_corr_hgis), 
+               names_to = c(".value", "method"),
+               names_pattern = "(.*)_(.*)") %>% 
+  mutate(fm_diff = fm_corr - fm_mean)
+
+combined_results %>% 
+  select(sample_name, rec_num, method, fm_corr, sig_fm_corr) %>% 
+knitr::kable()
+```
+
+Difference plot
+
+```{r}
+combined_results %>% 
+  ggplot(aes(fm_corr, fm_diff, color = method, shape = method)) +
+  geom_hline(yintercept = 0) +
+  geom_pointrange(aes(ymin = fm_diff - sig_fm_corr, 
+                      ymax = fm_diff + sig_fm_corr)) +
+                  #position = position_dodge2(width = 0.05, padding = 0.1)) +
+  ylim(-0.02, 0.02) +
+  labs(title = "Haiti Core Graphite vs HGIS",
+       x = "Fraction modern",
+       y = "Fm - mean Fm") + theme_classic()
+```

### 2021-06-17 inst/rmd/carbonate_reanalysis.Rmd
Add manual selection of outliers, plotting of raw data, plot and table of agreement with consensus.

@@
 blanks <- list(7, c(7,8), c(28,29,30), c(27,28,29), c(20,35,50), NULL, NULL)
+outliers <- list(NULL, tibble(pos = c(6, 8, 8, 8), meas = c(18, 1, 2, 3)),
+                 tibble(pos = c(28,29), meas=c(1, 1)),
+                 NULL,
+                 tibble(pos=10, meas=1),
+                 tibble(pos=c(8,9,13), meas=c(6,1,1)),
+                 tibble(pos=c(4,5,5,6,7,8,15), meas=c(1,1,4,10,3,3,2)))
 
-results <- pmap(list(files, dates, stds, blanks), reduce_hgis, remove_outliers = FALSE)
+results <- pmap(list(files, dates, stds, blanks, outliers), reduce_hgis)
 raw_df <- map_dfr(results, 1)
 results_df <- map_dfr(results, 2)
+```
+
+Plot raw data
+
+```{r}
+map(results, 1) %>% 
+  map(plot_hgis_time, y_var = norm_ratio, outlier = ok_calc)
+```
+
+```{r}
+map(results, 1) %>% 
+  map(plot_hgis_time, y_var = he12C, outlier = ok_calc)
 ```
@@
   compare_replicates() %>% 
-  arrange(rec_num)
+  arrange(rec_num) %>% 
+  select(Name, N)
 ```
+
+```{r}
+cons_df <- compare_consensus(results_df)
+
+cons_df %>% 
+  select(wheel, pos, sample_name, fm_consensus, fm_corr, sig_fm_corr, Fm_diff, sigma) %>% 
+  arrange(desc(Fm_diff))
+```
+
+```{r}
+cons_df %>% 
+  filter(wheel != "USAMS020521") %>% 
+plot_hgis_consensus()
+```
+
+```{r}
+cons_df %>% 
+  filter(wheel != "USAMS020521") %>% 
+  group_by(Name, fm_consensus) %>% 
+  summarize(across(c(fm_corr, sig_fm_corr, Fm_diff, sigma), list(mean = mean, sd = sd)))
+```
+
+```{r}
+cons_df %>% 
+  ungroup() %>% 
+  filter(wheel != "USAMS020521") %>% 
+  summarize(across(c(sig_fm_corr, Fm_diff, sigma), list(mean = mean, sd = sd)))
+```

### 2021-06-20 inst/rmd/carbonate_reanalysis.Rmd
Add latest wheel

@@
 library(HybridGIS)
+
+theme_set(theme_bw())
+options(digits = 3)
+options(scipen = 4)
 ```
@@
 ```{r}
-wheels <- c("020521", "030421", "040121", "041521", "051421", "052621", "061121")
+wheels <- c("020521", "030421", "040121", "041521", "051421", "052621", "061121", "061821")
 files <- here("data", paste0("USAMS", wheels, "R.txt"))
-dates <- as.Date(c("2021-02-05", "2021-03-05", "2021-04-09", "2021-04-16", "2021-05-14", "2021-05-28", "2021-06-11"))
-stds <- list(c(5,8), NULL, c(26,37), NULL, c(10,15,65,70), NULL, NULL)
-blanks <- list(7, c(7,8), c(28,29,30), c(27,28,29), c(20,35,50), NULL, NULL)
+dates <- as.Date(c("2021-02-05", "2021-03-05", "2021-04-09", "2021-04-16", "2021-05-14", "2021-05-28", "2021-06-11", "2021-06-18"))
+stds <- list(c(5,8), NULL, c(26,37), NULL, c(10,15,65,70), NULL, NULL, NULL)
+blanks <- list(7, c(7,8), c(28,29,30), c(27,28,29), c(20,35,50), NULL, NULL, 23)
 outliers <- list(NULL, tibble(pos = c(6, 8, 8, 8), meas = c(18, 1, 2, 3)),
@@
                  tibble(pos=c(8,9,13), meas=c(6,1,1)),
-                 tibble(pos=c(4,5,5,6,7,8,15), meas=c(1,1,4,10,3,3,2)))
+                 tibble(pos=c(4,5,5,6,7,8,15), meas=c(1,1,4,10,3,3,2)),
+                 NULL)
 
@@
   select(wheel, pos, sample_name, fm_consensus, fm_corr, sig_fm_corr, Fm_diff, sigma) %>% 
+  filter(wheel == "USAMS061821")
   arrange(desc(Fm_diff))

### 2021-06-20 inst/rmd/carbonate_reanalysis.Rmd
Write analysed data.

@@
 ```{r}
+# TODO: Handle case of only one blank
+
 wheels <- c("020521", "030421", "040121", "041521", "051421", "052621", "061121", "061821")
@@
 results_df <- map_dfr(results, 2)
+
+write_csv(raw_df, here("data_analysed/carb_all_raw.csv"))
+write_csv(results_df, here("data_analysed/carb_all_results.csv"))
 ```

### 2021-06-20 doc/data_meeting_2021-06-22/data_meeting_2021-06-22.Rmd
Add data meeting talk

---
title: "HGIS Carbonate Results"
subtitle: "NOSAMS Data Meeting"
author: 
  - "Brett Longworth"
date: '2021-06-22'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  fig.show = TRUE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#1381B0",
  secondary_color = "#FF961C",
  inverse_header_color = "#FFFFFF"
)
```

```{r}
library(tidyverse)
library(HybridGIS)
library(here)
library(gt)

options(digits = 3)
options(scipen = 4)
options(decimals = 4)
theme_set(theme_classic())
```

## Method

.pull-left[
1. Acidify
  - ~30mg Carbonate produces 6mL CO<sub>2</sub> @ ~1ATM
1. Displace gas to open split with 100 uL/min He.
1. Introduce mixture to source @ 30 uL/min
1. Acquire data: 90s x 10 measurements.

- Standard: live gas in vials
- Blank: C-1 or TIRI-F in vials
]

.pull-right[![HGIS Figure](images/HGIS_open_split_drawing.svg)]
---

## Data reduction

1. Correct for fractionation using online <sup>13</sup>C.
1. Determine per-sample mean and error.
1. Normalize to mean of Gas standard.
1. Blank correct: LBC using mean of carbonate blanks

- Errors propagated at each step. 
- Per-sample error agrees well with variability of replicates.

Data reduction done using an R package maintained on github:

https://github.com/blongworth/hgis

---

## Results: Agreement with Consensus

```{r}
# Read data. Don't use first carbonate test or smaller samples.
results_df <- read_csv(here("data_analysed/carb_all_results.csv"))
cons_df <- results_df %>% 
  filter(!is.na(sig_fm_corr),
         wheel != "USAMS061121") %>% 
  mutate(he12C = he12C * 1E6) %>% 
  compare_consensus()
```

```{r}
plot_hgis_consensus(cons_df) +
  scale_color_continuous(name = "12C- (μA)")
```
---

.left-column[
## Results: Method precision and accuracy
]

.right-column[
```{r}
cons_df_sum <- cons_df %>% 
  group_by(Name, fm_consensus) %>% 
  summarize(across(c(fm_corr, sig_fm_corr, Fm_diff, sigma),
                   list(mean = mean, sd = sd)),
            N = n()) %>% 
  ungroup()

cons_df_sum %>% 
  arrange(fm_consensus) %>% 
  select(Name, `Consensus Fm` = fm_consensus,
         `Mean Fm` = fm_corr_mean,
         `Mean Fm error` = sig_fm_corr_mean,
         `SD of measurements` = fm_corr_sd,
         N) %>% 
gt() %>% 
  tab_header(title = "Summary statistics by sample type") %>% 
  fmt_number(3:5, decimals = 4)
```
<br>
<br>

```{r}
cons_df %>% 
  filter(Name %in% c("TIRI-I", "C-2", "NOSAMS2")) %>% 
  ungroup() %>% 
  summarize(across(c(sig_fm_corr, Fm_diff, sigma),
                   list(mean = mean, sd = sd)),
            N = n()) %>% 
  select(`Mean Fm difference` = Fm_diff_mean,
         `SD of mean difference` = Fm_diff_sd,
         `Mean Fm error` = sig_fm_corr_mean,
         `Mean Sigma` = sigma_mean,
         `SD of Sigma` = sigma_sd,
         N
         ) %>%  
  #pivot_longer(everything()) %>% 
  gt() %>% 
  tab_header(title = "Summary statisics for reference materials",
             subtitle = "TIRI-I, C-2, and NOSAMS2") %>% 
  fmt_number(3:5, decimals = 4)
```
]

---

## Results: By sample type

```{r}
ggplot(cons_df, aes(x = Fm_diff, y = fm_corr, color = he12C)) +
  geom_hline(aes(yintercept = fm_consensus), color = "lightblue") +
  geom_hline(data = cons_df_sum, aes(yintercept = fm_corr_mean), color = "darkgray") +
  geom_hline(data = cons_df_sum, aes(yintercept = fm_corr_mean + fm_corr_sd), 
             color = "grey", linetype = "dashed") +
  geom_hline(data = cons_df_sum, aes(yintercept = fm_corr_mean - fm_corr_sd),
             color = "grey", linetype = "dashed") +

  #geom_boxplot(outlier.shape = NA) +
  geom_pointrange(aes(ymin = fm_corr - sig_fm_corr, 
                      ymax = fm_corr + sig_fm_corr), 
                  size = .2) +
  labs(subtitle = "Reference and normalization standards",
       y = "Fraction modern (Fm)") +
  facet_wrap(facets = vars(Name), scales = "free") +
  theme_bw() +
  scale_x_continuous(breaks = NULL) +
  xlab(NULL) +
  scale_color_continuous(name = "12C- (μA)") +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
```

### 2021-06-21 doc/data_meeting_2021-06-22/data_meeting_2021-06-22.Rmd
Minor fix
Add Haiti carbonate comparison

@@
 library(tidyverse)
+library(amstools)
 library(HybridGIS)
@@
 1. Acidify
-  - ~30mg Carbonate produces 6mL CO<sub>2</sub> @ ~1ATM
+  - ~30mg Carbonate 
+  - produces 6mL CO<sub>2</sub> @ ~1ATM
 1. Displace gas to open split with 100 uL/min He.
@@
 ```
+
+---
+
+## Pendleton Haiti Core Carbonates
+
+```{r}
+haiti <- filter(results_df, str_detect(sample_name, "HATC1")) %>% 
+  mutate(rec_num = case_when(pos == 14 ~ 171989,
+                             pos == 15 ~ 171995,
+                             pos == 16 ~ 171999))
+
+graphite_results <- getRecOS(c(171999,171995,171989)) %>% 
+  select(rec_num, fm_corr = f_modern, sig_fm_corr = f_ext_error)
+
+combined_results <- haiti %>% 
+  select(sample_name, rec_num, fm_corr, sig_fm_corr) %>% 
+  left_join(graphite_results,
+            by = "rec_num", 
+            suffix = c("_hgis", "_graphite")) %>% 
+  mutate(fm_mean = (fm_corr_hgis + fm_corr_graphite)/2) %>% 
+  pivot_longer(c(fm_corr_graphite, sig_fm_corr_graphite, 
+                 fm_corr_hgis, sig_fm_corr_hgis), 
+               names_to = c(".value", "method"),
+               names_pattern = "(.*)_(.*)") %>% 
+  mutate(fm_diff = fm_corr - fm_mean)
+
+combined_results %>% 
+  ggplot(aes(fm_corr, fm_diff, color = method, shape = method)) +
+  geom_hline(yintercept = 0) +
+  geom_pointrange(aes(ymin = fm_diff - sig_fm_corr, 
+                      ymax = fm_diff + sig_fm_corr),
+                  size = 1,
+                  position = position_dodge2(width = 0.02)) +
+  labs(title = "Haiti Core Graphite vs HGIS",
+       x = "Fraction modern",
+       y = "Fm - mean Fm") + theme_classic()
+```

### 2021-06-21 inst/rmd/opensplittests.Rmd
Minor fixes
Fix errors

@@
 
-ds[-(1:4),-3]
+ds[-(1:4),-4]
 ```
@@
 ```{r}
-df <- get_hgis_data(here("data/USAMS052621R.txt"), as.Date("2021-06-04")) #, standards)
+df <- process_hgis_results(here("data/USAMS052621R.txt"), as.Date("2021-06-04")) #, standards)
 ```
@@
 df %>% 
-  filter(!(Pos %in% 1:4)) %>% 
-  plot_hgis_time(normFm, ce*normFm) +
+  filter(!(pos %in% 1:4)) %>% 
+  plot_hgis_time(norm_ratio, sig_norm_ratio) +
   ggtitle("Sample ratio vs. time")
@@
 df %>% 
-  filter(!(Pos %in% 1:4)) %>% 
+  filter(!(pos %in% 1:4)) %>% 
   plot_hgis_time(he12C) +
@@
 df_sum %>% 
-  filter(as.numeric(Pos) > 4) %>% 
-  mutate(sig_normFm = max_err * normFm) %>% 
-  select(Pos, Sample.Name, he12C, n_runs, norm_ratio, sig_norm_ratio) %>% 
-  arrange(Sample.Name)
+  filter(as.numeric(pos) > 4) %>% 
+  select(pos, sample_name, he12C, n_runs, norm_ratio, sig_norm_ratio) %>% 
+  arrange(sample_name)
 ```
@@
 df_sum %>% 
-  filter(as.numeric(Pos) > 4) %>% 
-  ggplot(aes(Sample.Name, norm_ratio)) +
+  filter(as.numeric(pos) > 4) %>% 
+  ggplot(aes(sample_name, norm_ratio)) +
     geom_pointrange(aes(ymin = norm_ratio - sig_norm_ratio, ymax = norm_ratio + sig_norm_ratio),
@@
            y = "Fraction Modern")
-plot_hgis_summary()
 ```
@@
 df_sum %>% 
-  filter(as.numeric(Pos) > 4) %>% 
+  filter(as.numeric(pos) > 4) %>% 
   compare_replicates()
@@
 df_sum %>% 
-  filter(as.numeric(Pos) > 4,
+  filter(as.numeric(pos) > 4,
          !is.na(fm_consensus)) %>% 
-  select(Sample.Name, fm_consensus, norm_ratio, sig_norm_ratio) %>% 
+  select(sample_name, fm_consensus, norm_ratio, sig_norm_ratio) %>% 
   mutate(Fm_diff = norm_ratio - fm_consensus,
          sigma = amstools::sigma(norm_ratio, fm_consensus, sig_norm_ratio)) %>% 
-  arrange(Sample.Name)
+  arrange(sample_name)
 ```
@@
 write_csv(carb_data, here("data/carb_data_2021-06-04.csv"))
-insert_samples(carb_data)
+#insert_samples(carb_data)
 carb_data

### 2021-06-23 inst/rmd/carbonate_samples.R
Add script to combine carbonate sample data.

@@ -1,0 +1,128 @@
+# Combine all carbonate sample data
+
+library(tidyverse)
+library(here)
+
+# 2021-02-05
+cd1 <- tribble(~pos, ~name, ~rec_num, ~mass,
+                      6, "C-1", 83028, 32.05,
+                      7, "TIRI-F", 2138, 33.56,
+                      9, "TIRI-I", 17185, 32.67,
+                     10, "TIRI-I", 17185, 32.35,
+                     11, "C-2", 1082, 31.35,
+                     12, "C-2", 1082, 34.11,
+                     13, "NOSAMS2", 38809, 32.30,
+                     14, "NOSAMS2", 38809, 34.00
+                     ) %>% 
+  mutate(wheel = "USAMS020521")
+
+## 2021-03-05 Carbonate samples
+cd2 <- tribble(~pos, ~name, ~rec_num, ~mass,
+                      7, "C1-1", 83028, 33.69,
+                      8, "C1-2", 83028, 31.62,
+                      9, "C2-1", 1082, 32.70,
+                     10, "C2-2", 1082, 32.08,
+                     11, "NOSAMS2-1", 38809, 33.23,
+                     12, "NOSAMS2-2", 38809, 33.95
+                     ) %>% 
+  mutate(wheel = "USAMS030421")
+
+## 2021-04-09 Carbonate samples
+cd3 <- tribble(~pos, ~name, ~rec_num, ~mass,
+                     31, "NOSAMS2_1", 38809, 31.8,
+                     32, "NOSAMS2_2", 38809, 32.3,
+                     33, "NOSAMS2_3", 38809, 33.5,
+                     34, "TIRI-I_1", 17185, 33.1,
+                     35, "TIRI-I_2", 17185, 31.8,
+                     36, "TIRI-I_3", 17185, 31.8,
+                     28, "TIRI-F_1", 2138, 31.9,
+                     29, "TIRI-F_2", 2138, 33.4,
+                     30, "TIRI-F_3", 2138, 32.4
+                     ) %>% 
+  mutate(wheel = "USAMS040121")
+
+## 2021-04-16 Carbonate samples
+cd4 <- tribble(~pos, ~name, ~rec_num, ~mass,
+                     27, "C-1_1", 83028, 30.05,
+                     28, "C-1_2", 83028, 36.00,
+                     29, "C-1_3", 83028, 31.88,
+                     30, "C-2_1", 1082, 32.57,
+                     31, "C-2_2", 1082, 35.50,
[…]

### 2021-06-23 doc/data_meeting_2021-06-22/data_meeting_2021-06-22.Rmd
Clean up output dataframes.

@@
   group_by(Name, fm_consensus) %>% 
-  summarize(across(c(fm_corr, sig_fm_corr, Fm_diff, sigma),
+  summarize(across(c(fm_corr, sig_fm_corr, fm_diff, sigma),
                    list(mean = mean, sd = sd)),
@@
   ungroup() %>% 
-  summarize(across(c(sig_fm_corr, Fm_diff, sigma),
+  summarize(across(c(sig_fm_corr, fm_diff, sigma),
                    list(mean = mean, sd = sd)),
             N = n()) %>% 
-  select(`Mean Fm difference` = Fm_diff_mean,
-         `SD of mean difference` = Fm_diff_sd,
+  select(`Mean Fm difference` = fm_diff_mean,
+         `SD of mean difference` = fm_diff_sd,
          `Mean Fm error` = sig_fm_corr_mean,
@@
 ```{r}
-ggplot(cons_df, aes(x = Fm_diff, y = fm_corr, color = he12C)) +
+ggplot(cons_df, aes(x = fm_diff, y = fm_corr, color = he12C)) +
   geom_hline(aes(yintercept = fm_consensus), color = "lightblue") +

### 2021-06-23 inst/rmd/analyse_carbonates.R
Add carbonate analysis script

@@ -1,0 +1,33 @@
+  # Reduce HGIS data
+# Script to reduce all available wheels of HIGS data using
+# most current data reduction method.
+
+library(tidyverse)
+library(here)
+library(amstools)
+library(HybridGIS)
+
+
+# TODO: Handle case of only one blank
+
+# Analysis metadata: wheels to analyse, outliers, positions of standards and blanks
+wheels <- c("020521", "030421", "040121", "041521", "051421", "052621", "061121", "061821")
+files <- here("data", paste0("USAMS", wheels, "R.txt"))
+dates <- as.Date(c("2021-02-05", "2021-03-05", "2021-04-09", "2021-04-16", "2021-05-14", "2021-05-28", "2021-06-11", "2021-06-18"))
+stds <- list(c(5,8), NULL, c(26,37), NULL, c(10,15,65,70), NULL, NULL, NULL)
+blanks <- list(7, c(7,8), c(28,29,30), c(27,28,29), c(20,35,50), NULL, NULL, 23)
+outliers <- list(NULL, tibble(pos = c(6, 8, 8, 8), meas = c(18, 1, 2, 3)),
+                 tibble(pos = c(28,29), meas=c(1, 1)),
+                 NULL,
+                 tibble(pos=10, meas=1),
+                 tibble(pos=c(8,9,13), meas=c(6,1,1)),
+                 tibble(pos=c(4,5,5,6,7,8,15), meas=c(1,1,4,10,3,3,2)),
+                 NULL)
+                 
+
+results <- pmap(list(files, dates, stds, blanks, outliers), reduce_hgis)
+raw_df <- map_dfr(results, 1)
+results_df <- map_dfr(results, 2)
+
+write_csv(raw_df, here("data_analysed/carb_all_raw.csv"))
+write_csv(results_df, here("data_analysed/carb_all_results.csv"))

### 2021-06-23 doc/data_meeting_2021-06-22/data_meeting_2021-06-22.Rmd
Update to allow blank correction with only one blank.

@@
 cons_df <- results_df %>% 
-  filter(!is.na(sig_fm_corr),
-         wheel != "USAMS061121") %>% 
+  filter(!(wheel %in% c("USAMS020521", "USAMS061121", "USAMS061821"))) %>% 
   mutate(he12C = he12C * 1E6) %>% 

### 2021-06-24 R/HGISFirstTest.Rmd
reorganize

---
title: "First Hybrid GIS Tests"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
date: "03-28-2017"
---

First tests of the USAMS SNICS-134 Hybrid gas ion source. Where not specified, flows are given in $\frac{\mu L}{m}$ and pressure in kPa, and ion currents are in uA of 12C.

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(amstools)
library(here)
library(HybridGIS)

theme_set(theme_bw())
options(digits = 3)
```

# Flow rates

## Calculated flow

To model laminar flow in a capillary, Poiseuille's equation is used.

$$
Q = \frac{\Delta P \pi R^4}{8 \mu \Delta x}
$$

Where $\Delta P$ is change in pressure (Pa ($\frac{kg}{m \cdot s^2}$)), $R$ is
radius of pipe (m), $\mu$ is dynamic viscosity ( $P \cdot s$ ($\frac{kg}{s
\cdot m}$)), and $\Delta x$ is the length of the pipe (m). Q is in
$\frac{m^3}{s}$

The following functions calculate flow using this equation and convert base
units to more convenient flow in $\frac{\mu L}{m}$ and pressure in kPa.

```{r}
# Poiseuille's; convert delta pressure in Pa to flow in m^3/s
prestoflow <- function(dp, r, u, l) {
  # r = radius of capillary
  # l = length of capillary
  # u = viscosity kg/m
  flow <- dp * pi * r^4 / 8 / u / l
  flow
}

# convert gauge kPa to uL/min
flowcalc <- function(pres, ...) {
  dppa <- pres * 1E3 #convert kPa to Pa
  flowcms <- prestoflow(dppa, ...)
  flowuls <- flowcms * 1E6 * 1E3
  flowulm <- flowuls * 60
  flowulm
}
```

Initial tests used a 0.050mm x 8.2m deactivated silica capillary. For this
capillary, parameters for Poiseuille's equation are as follows:

```{r}
r <- 2.5E-5
u <- 1.49E-5
x <- 8.2
factor <- flowcalc(1, r, u, x)
```

$\Delta P$ is regulator pressure

R is `r r*1E6` $\mu m$

$\mu$ is `r u*1E6` $\mu Pa\cdot s$

x is `r x`m

Given these,

Q($\frac{\mu L}{m}$) = `r factor` x P (kPa)

```{r}
flow <- function(f) {
  flowcalc(f, r, u, x)
}
ggplot(data = data.frame(pressure = 0, flow = 0),
       mapping = aes(x = pressure, y = flow)) +
  stat_function(fun = flow) +
  xlim(0,400) +
  ggtitle("Theoretical flow for given pressure",
          subtitle = "8.2m x 50um capillary")
```

## Measured flow

Kalina and I measured flow at the end of the capillary using two methods;
direct measurement with a bubble flow meter and as a helium leak rate with the
leak detector.

### Bubble meter

Kalina made measurements with both helium and CO2. The measurements below are
with helium and converted to CO2 by the ratio of their viscosities.

```{r}
bflow <- read.csv(here("data/bubbleflow.csv")) %>%
  mutate(heflow = volume/time,
         flow = heflow * 1.49E-5/1.90E-5, # co2/He viscosity
         pressure = pressure - 14.7, # subract 1 Atm for delta P
         pressure = pressure * 6.8948, # convert psi to kPa
         calcflow = flowcalc(pressure, r, u, x)) # calculate flow with Poiseuille's

ggplot(bflow, aes(pressure, flow)) + 
  stat_function(fun = flow, color = "blue") +
  geom_point() +
  geom_line() +
  xlim(0,400) + ylim(0,80) +
  ggtitle("Flow vs pressure",
          subtitle = "measured by gauge and bubble flow meter")

ggplot(bflow, aes(calcflow, flow)) + 
  geom_point() +
  geom_line() +
  ggtitle("Measured (bubble meter) vs predicted flow")
```

### Leak detector

#### 2017-10-17?

Measured leak rate using Helium leak detector. Ran capillary to KF adapter
flange. Pressures in kPa, leak rates in $\frac{Pa\cdot m^3}{s}$. Leak detector
switched to gross leak mode between 150 kPa and 200 kPa.

```{r}
ldflow <- read.csv(here("data/HeLeakCapillaryFlow.csv"), skip = 3)

ldflow <- ldflow %>% 
  mutate(sccm = Leakrate / 0.0016, # convert to sccm
         co2sccm = sccm * 1.49E-5/1.90E-5, # co2/He viscosity
         flow = co2sccm * 1E3, # convert to ul/m
         calcflow = flowcalc(Regulator, r, u, x)) # calculate flow with poiseulle's

ggplot(ldflow, aes(Regulator, flow)) + 
  stat_function(fun = flow, color = "blue") +
  geom_point() +
  geom_line() +
  ggtitle("Flow (measured by leak detector) vs pressure")

ggplot(ldflow, aes(calcflow, flow)) + 
  geom_point() +
  geom_line() +
  ggtitle("Measured (by leak detector) vs predicted flow")

```

## Efficiency functions

These functions convert volume of CO2 and coulombs of charge to atoms of C. The
ratio of these is the measurement efficiency of a gas ion source system. These are used for later calculations of efficiency given flow rate and ion current.

```{r}
# convert L CO2 to atoms C
lCO2toatC <- function (vol) {
  molC <- vol * 1/22.4 # 22.4 L of gas per mole at stp
  atC <- molC * 6.022E23
  atC
}

# convert current in A to atoms C/sec
CcurtoatC <- function(cur, cs) {
  at12C <- cur * 6.25E18 # electrons per coulomb
  at12C <- at12C / cs # divide by charge state
  at12C
}

# calculate efficiency from ml/m CO2 and uA C12
hgis_eff <- function(ulm, uA, cs = 3) {

  # if flow is zero or less, efficiency doesn't make sense
  ulm <- ifelse(ulm <= 0, NA, ulm)

  vol <- ulm * 1E-6 # conv to L
  cur <- uA * 1E-6 # conv to A
  atCin <- lCO2toatC(vol) / 60 # gives atoms/s
  atC12out <- CcurtoatC(cur, cs)
  atCout <- atC12out / .99 # 99% of C is 12C, we need total C
  atC12out / atCin
}
```

# Test Wheels

Two days of testing to date. Same wheel both days. Wheel had 5 AA, 4 OX-I, and
10 gas targets.

## USAMS032817

### Currents

For the first target, we turned up the gas supply pressure while measuring
source current at the high energy FC. These currents are roughly equivalent to
le12C (~30% transmission @ 3+ charge state).

Load and massage the data...

```{r}
cur <- read_tsv(here("data/2017-03-28Currents.tsv")) %>%
  mutate(flow = flowcalc(regp, r, u, x),
         eff = hgis_eff(flow, he12c, cs = 1)) %>% # cs 1 to estimate LE current
  filter(t > 5) 
```

```{r}
ggplot(cur, aes(regp, he12c)) +
  geom_line() +
  geom_point() +
  ggtitle("Current vs Pressure",
          subtitle = "2017-03-28")

ggplot(cur, aes(flow, he12c)) +
  geom_line() +
  geom_point() +
  ggtitle("Current vs Flow",
          subtitle = "2017-03-28")
```

```{r}
data <- get_hgis_data(here("data/USAMS032817R.txt")) %>%
           mutate(Num = ifelse(Pos == 11, "U", Num))
stdrat <- mean(data$cor1412he[data$Num == "S"])
data <- mutate(data, normFm = norm_gas(cor1412he, stdrat))
# data <- readResfile("/mnt/shared/USAMS/Results/2017 Results/USAMS032817R.txt") %>%
#   mungeResfile() 
```

Currents were very stable after an initial burn in of ~10m. Initial current
started at ~12uA without gas on target and decayed (roughly) exponentially to a
minimum around 1.2uA.

```{r}
data %>% 
  filter(Pos %in% c(11,12)) %>%
  ggplot(aes(ts, he12C, color = Pos)) +
   	geom_line() + geom_point() + ylim(c(0,8))
```

### Efficiency

Using efficiency functions defined above. Efficiency drops with increasing flow
as currents go up.

```{r, warning = FALSE}
ggplot(cur, aes(flow, eff)) + geom_line() + geom_point()
```

### Ratios

Ratios for live and dead CO2 were recorded in 180 blocks of 10s each. Solid AA
and OX-I were run with 10 blocks of 180s. These results are summarized and roughly normalized to solid OX-I below.

#### Summary table

```{r}
sum_hgis(data)
```

#### Plots

Boxplots of modern and dead samples. Ratios were essentially stable over time.
Positions < 10 are solid samples run for comparison.

Modern samples.

```{r}

data %>% 
  filter(X14.12he > .4,
         Pos %in% c(7, 11)) %>%
  mutate(type = ifelse(Pos == 7, "solid", "gas")) %>%
       	ggplot(aes(type, X14.12he)) + 
        geom_boxplot() +
        labs(x = "Sample type", y = "Fraction modern")
```

Dead gas and alpha aesar.

```{r}
data %>% filter(X14.12he < .02,
                X14.12he > 0.0005,
                Pos %in% c(3, 12)) %>%
  mutate(type = ifelse(Pos == 3, "solid", "gas")) %>%
           ggplot(aes(type, X14.12he)) + 
        geom_boxplot() +
        labs(x = "Sample type", y = "Fraction modern")

```

## USAMS040317

The same target wheel was loaded again. Karl and I experimented with the source
tune and target position to try and get more current. We were able to increase
the currents by about 1/3 by moving the target closer to the immersion lens.
Flow rates were lower for a given current. After tuning, I looked at flow vs.
ratio and current. The gas used was the modern standard, but the flask may not
have been fully flushed.

```{r, message = FALSE}
data <- get_hgis_data(here("data/USAMS040317R.txt")) %>% 
  filter(Pos == 13)

sum_hgis(data)

# data <- readResfile(here("data/USAMS040317R.txt")) %>%
#   filter(Pos == 13) %>%
#   mungeResfile()

# Assign Run numbers
counter <- 0
Run <- c(1:190)
for (i in 1:nrow(data)) {
  if (data$Meas[i] == 1) {
    counter <- counter + 1
  }
  Run[i] <- counter
} 
data$Run <- Run

# Load pressure data
pres <- read.csv(here("data/USAMS040317Pressure.csv"))
pres$Run <- as.numeric(row.names(pres))

# Join to run data
data <- inner_join(data, pres) %>%
  mutate(Source = Source * 1E-7, # I think the source pressures were to the -7, but unsure...
         flow = flowcalc(Bottle, r, u, x),
         eff = hgis_eff(flow, he12C, cs = 1)) # cs 1 to estimate LE current

# summarize
d.s <- data %>% group_by(Run, Bottle, Source, flow) %>% summarize(le12C = mean(le12C),
                                              he12C = mean(he12C),
                                              counts = sum(CntTotGT),
                                              interr = 1/sqrt(counts),
                                              C1412he = mean(X14.12he),
                                              C1412he_sd = sd(X14.12he),
                                              ferr = interr * C1412he,
                                              cor1412he = mean(cor1412he),
                                              eff = mean(eff))
d.s
```

#### Gas flow vs source pressure

```{r}
ggplot(d.s, aes(Bottle, Source)) + 
  geom_smooth() +
  geom_point() 
ggplot(d.s, aes(flow, Source)) + 
  geom_smooth() +
  geom_point() 
```

#### Currents

```{r, message = FALSE, warning = FALSE}
ggplot(d.s, aes(Bottle, he12C)) +
  geom_smooth() +
  geom_point() 
filter(d.s, flow > 0) %>%
  ggplot(aes(flow, le12C)) +
  geom_smooth() +
  geom_point() +
  labs(x = expression(CO[2]~flow~rate~~mu~L %.% min^{-1}),
       y = bquote('12C-  ' *mu~ 'A')) +
  scale_y_reverse() +
  ylim(0,-10) +
  xlim(0,15.5)
```

#### Ratio vs current

```{r, message = FALSE, warning = FALSE}
ggplot(d.s, aes(he12C, C1412he)) + 
  geom_linerange(aes(ymin = C1412he - ferr, ymax = C1412he + ferr)) +
  geom_smooth() +
  geom_point() 
```

#### Efficiency

```{r, message = FALSE, warning = FALSE}
ggplot(d.s, aes(Bottle, eff)) + 
  geom_smooth() +
  geom_point() 
ggplot(d.s, aes(Source, eff)) + 
  geom_smooth() +
  geom_point() 
ggplot(d.s, aes(flow, eff*100)) + 
  geom_smooth() +
  geom_point() +
  labs(x = expression(CO[2]~flow~rate~~mu~L %.% min^{-1}), y = "Measurement efficiency  %") +
  xlim(0,15.5)
```

### 2021-06-24 R/HGISFlowCalcs.Rmd
reorganize

---
title: "HGIS flow calcs"
output:
  html_document: default
  html_notebook: default
---

```{r}
library(here)
library(HybridGIS)
```

# Calculated Flow rates

To model laminar flow in a capillary, Poiseuille's equation is used.

$$
Q = \frac{\Delta P \pi R^4}{8 \mu \Delta x}
$$

Where $\Delta P$ is change in pressure (Pa ($\frac{kg}{m \cdot s^2}$)), $R$ is
radius of pipe (m), $\mu$ is dynamic viscosity ( $P \cdot s$ ($\frac{kg}{s
\cdot m}$)), and $\Delta x$ is the length of the pipe (m). Q is in
$\frac{m^3}{s}$

  The functions loaded above calculate flow using this equation and convert base
units to more convenient flow in $\frac{\mu L}{min}$ and pressure in kPa.

## Initial NOSAMS parameters

Initial tests used a 0.050mm x 8.2m deactivated silica caplillary. For this
capillary, parameters for Poiseuille's equation are as follows:

```{r}
r <- 2.5E-5
u <- 1.49E-5
x <- 8.2
flowcalc(100, r, u, x)
```

## Parameters for published hybrid GIS systems

Calculate flow for Xu et al 2006...

```{r}
flowcalc(1.500, 0.000125, u, 1.25)
```
Calculate flow for ETH...

```{r}
flowcalc(15.6, 3E-5, u, 2)
flowcalc(31.3, 3E-5, u, 2)

```
Calculate flow for CAMS...

```{r}
flowcalc(42, 7.5E-5, u, 6)

```
Calculate flow for Oxford...

```{r}
flowcalc(200, 3E-5, u, 1.25)

```

## Parameters for next USAMS tests

### Short capillary at 1 atm

40um capillary

```{r}
r <- 2E-5
u <- 1.49E-5
x <- 3.5
flowcalc(100, r, u, x)
```
```{r}
r <- 2E-5
u <- 1.49E-5
x <- 2
flowcalc(100, r, u, x)
```
50um capillary
```{r}
r <- 2.5E-5
u <- 1.49E-5
x <- 5
flowcalc(101, r, u, x)
```
75um capillary
```{r}
r <- 3.75E-5
u <- 1.49E-5
x <- 2.8
flowcalc(101, r, u, x)
```

```{r}
r <- 2.5E-5
u <- 1.49E-5
x <- 7
flowcalc(100, r, u, x)
```
```{r}
r <- 5E-5
u <- 1.49E-5
x <- 3
flowcalc(20, r, u, x)
```
### Low pressure 15 mbar

100um capillary

```{r}
r <- 5E-5
u <- 1.49E-5
x <- 1.58
flowcalc(1.5, r, u, x)
```

### Medium pressure 150mbar

75um capillary

```{r}
r <- 3.75E-5
u <- 1.49E-5
x <- 4
flowcalc(15, r, u, x)
```

100um capillary

```{r}
r <- 5E-5
u <- 1.49E-5
x <- 2.0
flowcalc(5, r, u, x)
```
Medium pressure 150mbar 100um x 1.58m PEEK capillary
```{r}
r <- 5E-5
u <- 1.49E-5
x <- 1.58
flowcalc(20, r, u, x)
```

10x dillution
```{r}
r <- 5E-5
u <- 1.49E-5
x <- 1.58
flowcalc(15, r, u, x)
```

Theoretical flows with pressure adjusted for best current.

```{r}
r <- 5E-5
u <- 1.49E-5
x <- 1.58
flowcalc(60, r, u, x)
```
Actual performance of "100um"" x 1.58m PEEK capillary *IF* flow is ~10uL/min. Looks like
true diameter after crimping is around 40um.

```{r}
r <- 2.0E-5
u <- 1.49E-5
x <- 1.58
flowcalc(60, r, u, x)
```

# CO2 quantities

Calculate CO2 needed for an analysis at given supply pressure.

### 2021-06-24 R/HGIS_pendleton_2021-05-28.Rmd
reorganize

--- 2021-06-16 inst/rmd/HGIS_pendleton_2021-05-28.Rmd
@@
 
+### Manual outlier flagging
+
 ```{r}
-df_sum <- sum_hgis_targets(df, remove_outliers = FALSE) %>% 
+outliers <- tribble(~pos, ~meas,
+                    8, 6,
+                    9, 1,
+                    13, 1)
+df_clean <- flag_outliers(df, outliers)
+```
+
+```{r}
+df_clean %>% 
+  filter(!(pos %in% 1:4)) %>% 
+  plot_hgis_time(norm_ratio, sig_norm_ratio, outlier = !.$ok_calc) +
+  labs(title = "Sample ratio vs. time",
+       y = "Fraction modern")
+```
+
+```{r}
+df_clean %>% 
+  filter(!(pos %in% 1:4)) %>% 
+  plot_hgis_time(he12C * 1E6, outlier = !.$ok_calc) +
+  labs(title = "Sample current vs time",
+       x = "Time (s)",
+       y = "Current (uA)")
+```
+
+## Data Reduction
+
+```{r}
+df_sum <- sum_hgis_targets(df_clean) %>% 
   norm_hgis() %>% 
@@
 
-# write_csv(df_sum, here("data_analysed/HGIS_2021-05-28_results.csv"))
+write_csv(df_sum, here("data_analysed/HGIS_2021-05-28_results.csv"))
 # write.xlsx(df_sum, here("data_analysed/HGIS_2021-05-28_results.xlsx"))
@@
   geom_pointrange(aes(ymin = fm_diff - sig_fm_corr, 
-                      ymax = fm_diff + sig_fm_corr)) +
-                  #position = position_dodge2(width = 0.05, padding = 0.1)) +
+                      ymax = fm_diff + sig_fm_corr),
+                  size = 1,
+                  position = position_dodge2(width = 0.02)) +
   ylim(-0.02, 0.02) +

### 2021-06-24 R/HGIS_pendleton_2021-05-28_outliers.Rmd
reorganize

---
title: "HGIS 2021-05-28 Carbonates"
subtitle: "Comparison of data with and without outliers"
author: "Brett Longworth"
date: "6/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

```{r}
library(tidyverse)
library(here)
library(amstools)
library(HybridGIS)

options(digits = 4)
options(scipen = 3)
```

Data reduction and analysis for carbonate samples run on May 28, 2021. Includes C-1, C-2, and test shell carbonates from Simon Pendleton. This version compares the effect of including or discarding outliers.

```{r}

#Raw AMS data results file path
file <- here("data/USAMS052621R.txt")

# Date samples run (needed to separate other experiments with this wheel/targets)
date <- as.Date("2021-05-28")

# Positions of normalizing standards
standards <- c(5, 8, 13, 17)

# Positions of blanks for blank correction
blanks <- c(6, 9, 11)
```

- Using standards in positions `r standards` for normalization.
- Using blanks in positions `r blanks` for blank correction.
- Outliers are defined as points 1.5 * interquartile range above the 3rd quartile or below the 1st quartile, and are dropped from the analysis.

Data reduction proceeds as follows:

1. Read raw file
1. Apply 13C correction to each run of a target
1. Flag outliers by interquartile distance
1. Determine per-target means, internal and external errors
1. Normalize targets using mean of standards and propagate errors
1. Apply a large blank correction using the mean of blanks and propagate errors.

## raw results

```{r}
df <- get_hgis_data(file, date, standards)
```

#### Currents and ratios vs time.

```{r}
df %>% 
  filter(!(Pos %in% 1:4)) %>% 
  plot_hgis_time(normFm, ce*normFm) +
  ggtitle("Sample ratio vs. time")
```

```{r}
df %>% 
  filter(!(Pos %in% 1:4)) %>% 
  plot_hgis_time(he12C) +
  ggtitle("Sample current vs time")
```

```{r}

# The data without outliers
df_sum <- sum_hgis_targets(df) %>% 
  norm_hgis() %>% 
  blank_cor_hgis(blanks = c(6, 9, 11)) %>% 
  mutate(outliers_dropped = TRUE)

# The data with outliers
df_sum_ol <- sum_hgis_targets(df, remove_outliers = FALSE) %>% 
  norm_hgis() %>% 
  blank_cor_hgis(blanks = c(6, 9, 11)) %>% 
  mutate(outliers_dropped = FALSE)

df_all <- rbind(df_sum, df_sum_ol)
```

## Results

```{r}
df_all %>% 
  filter(as.numeric(Pos) > 4) %>% 
  mutate(sig_normFm = max_err * normFm) %>% 
  select(Pos, Sample.Name, outliers_dropped, he12C, n_runs, fm_corr, sig_fm_corr) %>% 
  arrange(Sample.Name) %>% 
  knitr::kable()
  # gt::gt()
```

```{r}
df_all %>% 
  filter(as.numeric(Pos) > 4) %>% 
  ggplot(aes(Sample.Name, fm_corr, color = outliers_dropped)) +
    geom_pointrange(aes(ymin = fm_corr - sig_fm_corr, ymax = fm_corr + sig_fm_corr),
                    size = 0.1) + 
    theme(axis.text.x = element_text(angle = 45, hjust=1)) +
      labs(x = NULL,
           y = "Fraction Modern")
```

## Agreement of replicates

Compare results for samples with more than one replicate.

```{r}
df_all %>% 
  filter(as.numeric(Pos) > 4) %>% 
   mutate(Name = str_remove(Sample.Name, "_.$")) %>% 
    group_by(Name, outliers_dropped) %>% 
    filter(n() > 1) %>% 
    summarize(across(c(he12C, fm_corr, sig_fm_corr),
                     list(mean = mean, sd = sd)),
              N = n()) %>% 
  knitr::kable()
```

## Agreement with consensus

```{r}
cons <- df_all %>% 
  filter(as.numeric(Pos) > 4,
         !is.na(fm_consensus)) %>% 
  select(Sample.Name, outliers_dropped, fm_consensus, fm_corr, sig_fm_corr) %>% 
  mutate(Fm_diff = fm_corr - fm_consensus,
         sigma = amstools::sigma(fm_corr, fm_consensus, sig_fm_corr)) %>% 
  arrange(Sample.Name)

knitr::kable(cons)
```

Mean error, difference from consensus, and sigma with and without outliers.

```{r}
cons %>% 
  group_by(outliers_dropped) %>% 
  summarize(across(c(sig_fm_corr, Fm_diff, sigma), mean))
```

### 2021-06-24 R/carbonate_reanalysis.Rmd
reorganize

--- 2021-06-20 inst/rmd/carbonate_reanalysis.Rmd
@@
 cons_df %>% 
-  select(wheel, pos, sample_name, fm_consensus, fm_corr, sig_fm_corr, Fm_diff, sigma) %>% 
-  filter(wheel == "USAMS061821")
-  arrange(desc(Fm_diff))
+  select(wheel, pos, sample_name, fm_consensus, fm_corr, sig_fm_corr, fm_diff, sigma) %>% 
+  filter(wheel == "USAMS061821") %>% 
+  arrange(desc(fm_diff))
 ```
@@
   group_by(Name, fm_consensus) %>% 
-  summarize(across(c(fm_corr, sig_fm_corr, Fm_diff, sigma), list(mean = mean, sd = sd)))
+  summarize(across(c(fm_corr, sig_fm_corr, fm_diff, sigma), list(mean = mean, sd = sd)))
 ```
@@
   filter(wheel != "USAMS020521") %>% 
-  summarize(across(c(sig_fm_corr, Fm_diff, sigma), list(mean = mean, sd = sd)))
+  summarize(across(c(sig_fm_corr, fm_diff, sigma), list(mean = mean, sd = sd)))
 ```

### 2021-06-24 R/dualbellowstests.Rmd
reorganize

---
title: "Dual Bellows Tests"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

Load libraries and functions

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(amstools)
library(here)
library(HybridGIS)

theme_set(theme_bw())
options(digits = 3)
```

## 2017-07-31 Test flow rates using helium leak checker

Used dual bellows sample inlet block. Attached capillary to sample-mass spec outlet (valve SM). Loaded Helium at SI, measured pressure with inlet gauge. Used two PEEK capillaries- 65um and 100um ID x 158cm long.

Presures in mBar, leak rate in mbar l/s.

```{r}
r <- factor(c("5E-5", "5E-5", "5E-5", "3.3E-5", "3.3E-5", "3.3E-5"))
press <- c(18, 30, 60, 18, 60, 103)
leakrate <- c(6.3E-7, 3.8E-6, 1.8E-5, 5.7E-7, 8.8E-6, 2.5E-5)
data <- data.frame(r, press, leakrate)

#convert leak rate to Pa m3/s
data <- mutate(data, 
               leakrate = leakrate * 100 / 1000, # convert to Pa m3 s-1
                sccm = leakrate / 0.0016, # convert to sccm
                 co2sccm = sccm * 1.49E-5/1.90E-5, # co2/He viscosity
                 flow = co2sccm * 1E3, # convert to ul/m
                 press = press / 10) # convert to kPa

ggplot(data, aes(press, flow, color = r)) + 
  geom_line() + 
  geom_point() +
  scale_color_discrete(name = "Diameter (um)",
                       breaks = c("3.3E-5", "5E-5"),
                       labels = c("65", "100")) +
  ggtitle("Flow vs Pressure for 65um and 100um PEEK capillaries")
```

## 2017-10-09 Test PEEK capillaries

No sign of discharge with PEEK capillaries
Flow in 100um capillary seems low

## 2017-10-10 Source flow tests

Using 100um x 178mm glass capillary. Inlet to source, measured source pressure with source iso GV closed. No discharge observed for any capillary.

### Source pressure vs supply pressure

```{r}
r <- 5E-5
press <- c(10, 25, 55, 112, 205, 312, 408, 500, 607, 702, 802, 1000)
source <- c(7.1E-8, 7.2E-8, 7.6E-8, 8.5E-8, 1.6E-7, 6.2E-7, 1.2E-6, 1.9E-6, 2.8E-6, 3.8E-6, 4.8E-6, 7.3E-6)
data <- data.frame(r, press, source)

data <- mutate(data,
               press = press/10, # convert to kPa
               flow = flowcalc(press, 5E-5, 1.49E-5, 1.58)
)

ggplot(data, aes(press, source)) + geom_line() + geom_point() +
  ggtitle("Source pressure vs inlet pressure, 100um Glass Capillary")
```

Calculated flow vs source pressure. Something fishy here.

```{r}
ggplot(data, aes(flow, source)) + geom_line() + geom_point() +
  ggtitle("Source pressure vs CO2 flow, 100um Glass Capillary")
```

If the diameter of the capillary is adjusted to 50um, things look more realistic...

```{r}
data <- mutate(data, flow = flowcalc(press, 2.5E-5, 1.49E-5, 1.58))
ggplot(data, aes(flow, source)) + geom_line() + geom_point() +
  ggtitle("Source pressure vs CO2 flow, if capillary is 50um id")
```

## 2017-10-16 2017-10-17 - Leak rates of PEEK capillaries with leak checker

Black PEEK capillary, 100um x 158mm
Connected between helium regulator and leak detector.

```{r}
r <- 5E-5
press <- c(0, 35, 50, 75, 100, 140, 159, 190, 222, 20, 8, 0)

#press <- press/10 # convert to kPa
leakrate <- c(3.6E-11, 7.2E-7, 1.5E-6, 3.1E-6, 5.1E-6, 9.4E-6, 1.2E-5, 1.5E-5, 2.2E-5, 2.2E-7, 1.5E-8,1.4E-9)

data <- data.frame(r, press, leakrate)

data <- mutate(data,
                sccm = leakrate / 0.0016, # convert to sccm
                 co2sccm = sccm * 1.49E-5/1.90E-5, # co2/He viscosity
                 flow = co2sccm * 1E3, # convert to ul/m
                 press = press / 10,
                 calflow = flowcalc(press, 5E-5, 1.49E-5, 1.58),
                 corcalflow = flowcalc(press, 2.5E-5, 1.49E-5, 1.58) ) # convert to kPa

ggplot(data, aes(press, flow)) + 
  geom_line() + 
  geom_point() +
  ggtitle("Flow vs Pressure for 100um PEEK capillary")
```

Looks like calculated flows don't match up with measured. Is diameter of capillary wrong? Using a diameter of 50um in poiselle's gives better numbers...

```{r}
ggplot(data, aes(flow, calflow)) + geom_line() + geom_point() +
  ggtitle("Predicted vs actual flow, 100um PEEK Capillary")

ggplot(data, aes(flow, corcalflow)) + geom_line() + geom_point() +
  ggtitle("Predicted vs actual flow, if diameter = 50um")
```

### 2017-10-20 Glass capillary

Leak checker test of 100um x 178mm glass capillary.

```{r}
r <- 5E-5
press <- c(0, 32, 64, 80, 110, 121, 150, 15, 10, 4)
leakrate <- c(2E-9, 1.1E-6, 2.8E-6, 3.7E-6, 6.1E-6, 7.5E-6, 1.1E-5, 4.5E-7, 2.7E-7, 8.5E-8)

data <- data.frame(r, press, leakrate)

data <- mutate(data,
                sccm = leakrate / 0.0016, # convert to sccm
                 co2sccm = sccm * 1.49E-5/1.90E-5, # co2/He viscosity
                 flow = co2sccm * 1E3, # convert to ul/m
                 press = press / 10, # convert to kPa
                 calflow = flowcalc(press, 5E-5, 1.49E-5, 1.78))

ggplot(data, aes(press, flow)) + 
  geom_line() + 
  geom_point() + 
  ggtitle("100um Glass Capillary, flow vs pressure")
ggplot(data, aes(calflow, flow)) + 
  geom_line() + 
  geom_point() + 
  ggtitle("100um Glass Capillary, measured vs predicted flow")
```

## 2017-11-14 Beam tests

Used black PEEK capillary. 
```{r}
lpdata <- read.csv(here("data/14Nov2017LPtest.csv"), skip = 4) %>%
  mutate(supply = supply / 10,
         source = source * 1E-7,
         flow = flowcalc(supply, 2E-5, 1.49E-5, 1.58))

ggplot(lpdata, aes(supply, current)) + 
  geom_smooth() + geom_point() +
  labs(x = "Inlet pressure (kPa)", y = "C12 Current")
ggplot(lpdata, aes(supply, source)) + 
  geom_smooth() + geom_point() +
  labs(x = "Inlet pressure (kPa)", y = "Source pressure (Pa)")
ggplot(lpdata, aes(flow, source)) + 
  geom_smooth() + geom_point() +
  labs(x = "Flow (ul/min)", y = "Source pressure (Pa)")
ggplot(lpdata, aes(flow, current)) + geom_smooth() + geom_point() +
  labs(x = "Flow (ul/min)", y = "Source current (uA)")
ggplot(lpdata, aes(source, current)) + geom_smooth(span = 1) + geom_point() +
  labs(x = "Source pressure (Torr)", y = "Source current (uA)")

```

## 2018-04-25 Testing bellows compression

Wired up stepper drives. Tested currents and flows with black PEEK capillary
(nominally 100um id). Source works fine after cleaning clog from delivery tube
at last source cleaning. Ran scan of inlet pressures, saw best currents at
~600mbar. Reduced pressure to 300mbar, closed SF, used bellows to return flow
to optimum.

Using 100um x 178cm black PEEK capillary.

inletp in mbar, sourcep in Torr, current in uA.

```{r}
# load data
beldat <- read_csv(here("data/bellowsApr2518.csv"))
beldat
```

```{r}
belopen <- beldat %>% filter(bellows == 0)

ggplot(belopen, aes(inletp, sourcep)) + 
    geom_point() + 
    geom_line()
```

Inlet pressure/source pressure relationship is about the same as last time. Still way more than calculated inlet pressure for a given source pressure. Capillary pinching at the swagelok connection or bad calculations?

Based on later alignment testing, flows for good current are reduced by about 3x with better alignment of delivery arm with target.

```{r}
ggplot(belopen, aes(sourcep, c12cur)) +
  geom_line() + geom_point() +
  labs(x = "Source pressure (Torr)", y = "Source current (uA)")
```

Looks like we have about the same source pressure/current optima as before.

```{r}

bellows <- beldat %>% filter(is.na(inletp) | inletp == 300)

ggplot(bellows, aes(bellows, c12cur)) +
  geom_smooth(method = "lm") + geom_point() +
  labs(x = "Bellows position (0-120k)", y = "Source current (uA)")
```

As expected, source pressure and current follow bellows position, such that bellows position is proportional to inlet pressure.

## 2018-05-23 Bellows with test samples

Initially tried 2m x 100um glass capillary. There were two sparks with two different capillaries. Both sparks broke capillary at what is presumably spark exit point. Later open split tests with glass capillaries did not have sparking issue. Three differences with open split:

1. Pressure. Open split should be at higher pressure along cap length.
2. Open split capillary is terminated in grounded SS bulkhead connector at cage.
3. A teflon jacket was added to the capillary inside the cage.

### Load data from tests & runs

Ran OX-I's and OX-II's from bellows end archive tubes. All ~100umol.

3 gas samples:
OX-I GS-9394 Pos 15
OX-II GS-10051 Pos 17
C-1 HY-31410 Pos 18

pos 8 is ox-i graphite

```{r}
data <- get_hgis_data(here("data/USAMS052318RHIGS.txt")) %>% 
                  mutate(Num = ifelse(Pos == 15, "S", "U"))
stdrat <- mean(data$cor1412he[data$Num == "S" & data$outlier == FALSE])
data <- mutate(data, normFm = norm_gas(cor1412he, stdrat))

plot_hgis_time(data)
```

The C-1 has some burn in that should be trimmed.

```{r}
data <- filter(data, !(Pos == 18 & normFm > 0.06))
sum_hgis(data)
```

```{r}
plot_hgis(data)
```

## 2018-05-30 Helium Dilution tests

Cleaned and reloaded sample wheel:
2x ceylon, 2x ox-I, 8 gas cathodes. USAMS053018R.txt

Using Black PEEK capillary, 158mm x 100um.

Added needle valve for UHP helium to inlet. Using turbopump instead of cryo. Base pressures are about 1E-6 initially, 6E-7 after a few hours, and 4E-7 overnight. Cryo base pressures are around 1E-7.

Procedure:

1. Admit sample to inlet, measure pressure.
2. Freeze sample to CF.
3. Admit He to inlet to give desired dilution
4. Close SF
5. Warm sample
6. Adjust bellows for best current
7. Run sample

### Flow tests

Evacuate inlet. Add He and check source pressure vs. inlet pressure. Inlet pressure in mbar, source pressure in Torr.

```{r}
# data from He pressure test
he_press_test <- tribble(~inletp, ~sourcep,
                         0, 6.77,
                         52, 6.77,
                         102, 6.8,
                         202, 7,
                         303, 7.4,
                         400, 8,
                         516, 8.9,
                         603, 9.7,
                         705, 11,
                         806, 12,
                         905, 13) %>% 
  mutate(sourcep = sourcep * 1E-7)

ggplot(he_press_test, aes(inletp, sourcep)) + 
    geom_point() + 
    geom_line() +
  ggtitle("Source vs inlet pressure, Helium")
```

### Dillution test, modern tank gas and OX-II breakseal

23 mbar CO2 frozen in inlet. Added 400 mbar He. Source pressure 8E-7, C12 .4uA. Compress to give source pressure of 4.8E-6, C12 of .6uA. Currents too low.

100mbar CO2 in inlet-SI frozen, 972mbar helium added. Source pressure 9E-6, 12C current 7.5uA. Made a 30m run on position 5 with this sample.

OX-II breakseal (GS-10016) run on position 6. 83.82umol CO2, 83mbar in inlet-SI. Froze sample and added 1035mbar helium. Base source pressure 1E-6, rose to 7E-6 with sample. Initial current 7.5uA 12C.

Load data from tests & runs

```{r}
data <- get_hgis_data(here("data/USAMS053018R.txt"), as.Date("2018-05-30")) %>% 
           mutate(Pos = ifelse(Sample.Name == "OX-II", 6, Pos),
                  pos_name = paste(Pos, Sample.Name, sep = " - "),
                  dil_factor = case_when(Pos == 5 ~ 972/100,
                                         Pos == 6 ~ 1035/83),
                  Num = ifelse(Pos == 5, "S", "U"))
stdrat <- mean(data$cor1412he[data$Num == "S"])

data <- mutate(data, normFm = norm_gas(cor1412he, stdrat))
plot_hgis_time(data)
```

```{r}
data %>% 
  filter(outlier == FALSE) %>% 
sum_hgis()
```

```{r}
 data %>% 
    mutate(Fm = ifelse(normFm > .15, "modern", "dead"),
           Name = factor(pos_name, levels = unique(pos_name[order(Fm, dil_factor)])),
           Fm = ordered(Fm, levels = c("modern", "dead")),
           dil_factor = factor(dil_factor)) %>%  
    ggplot(aes(Name, normFm, fill = dil_factor)) +
      geom_boxplot() +
      geom_hline(data = data.frame(yint=1,Fm=ordered("modern", levels = c("modern", "dead"))),
                 aes(yintercept = yint)) + 
      geom_hline(data = data.frame(yint=0,Fm=ordered("dead", levels = c("modern", "dead"))), 
                 aes(yintercept = yint)) + 
      theme(axis.text.x = element_text(angle = 45, hjust=1)) +
      labs(x = NULL,
           y = "Fraction Modern") +
      facet_grid(Fm~., scales = "free_y")
```

## 2018-06-06 diagnosing flow problems

Lowered arm by a full turn, moved actuator clevis down by one turn. Some additional alignment testing: Move arm, test current, repeat. Best flow/current- 9uA at source pressure of 7E-7 Torr.

Reterminated black PEEK capillary.

Calc for best flow:

```{r}
r <- 5E-5
u <- 1.49E-5
x <- 1.58
flowcalc(20, r, u, x)
```

12C current vs source pressure before and after delivery arm realignment.

```{r}
flowdat <- read.csv(here("data/DIflowJun618.csv"))
ggplot(flowdat, aes(source, current, color = Alignment)) +
  geom_line() + geom_point() +
  labs(x = "Source pressure (Torr)", y = "Source current (uA)")
str(flowdat)
```

Data from runs is in USAMS060618R.txt. Not 100% certain which samples are in which positions. Data look like 3 modern tank gas.

```{r}
data <- get_hgis_data(here("data/USAMS060618R.txt")) %>% 
  filter(Pos != 1,                           #remove spurious pos1 runs
         !(Pos == 3 & X13.12he > 0.011)) %>%     #remove solid OX-I fliers
                  mutate(Num = ifelse(Pos == 5, "S", "U"))

# Get mean ratio for solid OX-I and normalize data
stdrat <- mean(data$cor1412he[data$Num == "S" & data$outlier == FALSE])
data <- mutate(data, normFm = norm_gas(cor1412he, stdrat))

plot_hgis_time(data)
```

```{r}
sum_hgis(data)
```

```{r}
plot_hgis(data)
```

### 2021-06-24 R/smaller_samples.Rmd
reorganize

---
title: "Small(er) Samples"
author: "Brett Longworth"
date: "6/23/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(here)
library(amstools)
library(HybridGIS)

options(digits = 3)
options(scipen = 4)

df <- read_csv(here("data_analysed/carb_all_results.csv")) %>% 
  compare_consensus() %>% 
  mutate(c12ua = he12C * 1E6)
samples <- read_csv(here("data/carb_data.csv")) %>% 
  select(wheel, pos, mass, he_added)

df <- left_join(df, samples, by = c("wheel", "pos")) %>% 
  filter(wheel != "USAMS020521")
```

```{r}
df %>% 
ggplot(aes(mass, c12ua, color = as.factor(rec_num))) +
  geom_point()
```

```{r}
df %>% 
ggplot(aes(mass, fm_diff, color = wheel)) +
  geom_point()
```

```{r}
df %>% 
ggplot(aes(mass, fm_diff, color = as.factor(rec_num))) +
  geom_point()
```

```{r}
df %>% 
ggplot(aes(c12ua, abs(fm_diff))) +
  geom_smooth() +
  geom_point(aes(color = wheel))
```

```{r}
df %>% 
  filter(sample_type == "S") %>% 
ggplot(aes(c12ua, abs(fm_diff))) +
  geom_point(aes(color = wheel))
```

### 2021-06-24 doc/Radiocarbon2018/abstract/RC-18HGISAbst.Rmd
reorganize

--- 2018-04-03 Radiocarbon2018/abstract/RC-18HGISAbst.Rmd
@@
 ---
-title: "Development of a Hybrid Gas Ion Source and Gas Inlets at NOSAMS"
+title: "Development of a Hybrid Gas Ion Source and Gas Inlet Systems at NOSAMS"
 author:
@@
     affiliation: Woods Hole Oceanographic Institution 
-  - name: Kalina D. Godispodinova
+  - name: Kalina D. Gospodinova
     affiliation: Woods Hole Oceanographic Institution 
@@
 [-@Roberts2010].
-Gas delivery to the source is via a short glass capillary which crosses the source HV potential.
+Gas delivery to the source is via a short glass capillary which crosses the source high voltage potential.
 
-  We are developing two inlet technologies, one delivering gas to the source at low pressure using modified gas handling hardware from a VG stable isotope mass spectrometer and the other sampling gas at atmospheric pressure using an open split.
+  We are developing two inlet technologies, (1) delivering gas to the source at low pressure using modified gas handling hardware from a VG stable isotope ratio mass spectrometer (SIRMS), and (2) sampling gas at atmospheric pressure using an open split.
 The gas handling system designed for use with SIRMS is nearly ideal for our purposes, providing low volume, low pressure delivery of gas, and control of mixing ratios and delivery pressure using variable volumes.
-We will use the open split with large samples, focussing first on carbonate samples, hydrolysed and delivered using the gasbench system developed for use with the CFAMS microwave plasma GIS [@Roberts2013].
+We will initially use the open split for larger carbonate samples (>2mgC), which will be hydrolysed and delivered using the gas preparation system developed as an interface for our microwave plasma gas ion source [@Roberts2013].
 

### 2021-06-24 README.Rmd
Update README

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# HybridGIS

<!-- badges: start -->
<!-- badges: end -->

The HybridGIS project contains analysis scripts and data for the NOSAMS hybrid gas ion source.

## Setting up

This project uses the amstools and hgis packages, both available from github. They may be installed as follows:

```{r, message=FALSE}
remotes::install_github(c("blongworth/amstools", "blongworth/hgis"))
```

Running `make` will download the NOSAMS results files if a connection to the NOSAMS file server is available. This currently uses an absolute path to the server of `/mnt/shared/USAMS/Results`. This should be converted to a URL, but until then, modify `RES_DIR` in `Makefile` to match it's location on your system. `make` will eventually run the analysis scripts as well.

## Functions

The `amstools` package provides general functions for ams data retrieval and reduction, and the `hgis` package has functions for data reduction, flow calculation, efficiency, data import and plotting of HGIS data.

```{r example, message=FALSE}
library(amstools)
library(hgis)
```

# Project Summary

The information below summarizes the operating parameters, development, and questions
for the hybrid gas ion source on USAMS.

```{r child='summary.Rmd'}
```

# Presentations

Slides for a presentation given at SNEAP 2021:

### [SNEAP 2021 Presentation](doc/SNEAP2021/SNEAP2021_HGIS_presentation/SNEAP2021_HGIS_presentation.html)

### 2021-06-27 R/carbonate_samples.R
Add sample data from 2021-06-25. Same wheel so not added to all data.

@@ -122,6 +122,18 @@ cd9 <- tribble(~pos, ~name, ~rec_num, ~mass, ~he_added,
                      ) %>% 
   mutate(wheel = "USAMS061821")
 
+cd10 <- tribble(~pos, ~name, ~rec_num, ~mass, ~he_added,
+                     23, "C-1_2", 83028, 8.78, 5.2,
+                     27, "C-1_3", 83028, 16.32, 3.2,
+                     24, "TIRI-I_1", 17185, 5.40, 5.8,
+                     28, "TIRI-I_2", 17185, 8.46, 5.2,
+                     31, "TIRI-I_3", 17185, 16.39, 3.2,
+                     25, "NOSAMS2_1", 38809, 5.32, 5.8,
+                     29, "NOSAMS2_2", 38809, 15.57, 3.6,
+                     32, "NOSAMS2_3", 38809, 9.32, 5.2
+                     ) %>% 
+  mutate(wheel = "USAMS061821")
+
 carb_data <- bind_rows(cd1, cd2, cd3, cd4, cd5, cd6, cd7, cd8, cd9) %>% 
   select(wheel, pos, name, rec_num, mass, he_added)
 

### 2021-06-27 R/smaller_samples.Rmd
Add data and analysis for samples run 2021-06-25.

@@
 library(amstools)
-library(HybridGIS)
+library(hgis)
 
@@
   select(wheel, pos, mass, he_added)
-
 df <- left_join(df, samples, by = c("wheel", "pos")) %>% 
@@
 df %>% 
+ggplot(aes(mass, c12ua, color = as.factor(rec_num))) +
+  geom_point()
+```
+
+```{r}
+df %>% 
+  filter(wheel == "USAMS062521") %>% 
 ggplot(aes(mass, c12ua, color = as.factor(rec_num))) +

### 2021-06-27 R/analyse_carbonates.R

@@ -5,23 +5,24 @@
 library(tidyverse)
 library(here)
 library(amstools)
-library(HybridGIS)
+library(hgis)
 
 
 # TODO: Handle case of only one blank
 
 # Analysis metadata: wheels to analyse, outliers, positions of standards and blanks
-wheels <- c("020521", "030421", "040121", "041521", "051421", "052621", "061121", "061821")
+wheels <- c("020521", "030421", "040121", "041521", "051421", "052621", "061121", "061821", "062521")
 files <- here("data", paste0("USAMS", wheels, "R.txt"))
-dates <- as.Date(c("2021-02-05", "2021-03-05", "2021-04-09", "2021-04-16", "2021-05-14", "2021-05-28", "2021-06-11", "2021-06-18"))
-stds <- list(c(5,8), NULL, c(26,37), NULL, c(10,15,65,70), NULL, NULL, NULL)
-blanks <- list(7, c(7,8), c(28,29,30), c(27,28,29), c(20,35,50), NULL, NULL, 23)
+dates <- as.Date(c("2021-02-05", "2021-03-05", "2021-04-09", "2021-04-16", "2021-05-14", "2021-05-28", "2021-06-11", "2021-06-18", "2021-06-25"))
+stds <- list(c(5,8), NULL, c(26,37), NULL, c(10,15,65,70), NULL, NULL, NULL, NULL)
+blanks <- list(7, c(7,8), c(28,29,30), c(27,28,29), c(20,35,50), NULL, NULL, 23, NULL)
 outliers <- list(NULL, tibble(pos = c(6, 8, 8, 8), meas = c(18, 1, 2, 3)),
                  tibble(pos = c(28,29), meas=c(1, 1)),
                  NULL,
                  tibble(pos=10, meas=1),
                  tibble(pos=c(8,9,13), meas=c(6,1,1)),
                  tibble(pos=c(4,5,5,6,7,8,15), meas=c(1,1,4,10,3,3,2)),
+                 NULL,
                  NULL)
                  
 

### 2021-06-27 R/carbonate_samples.R

@@ -132,9 +132,10 @@ cd10 <- tribble(~pos, ~name, ~rec_num, ~mass, ~he_added,
                      29, "NOSAMS2_2", 38809, 15.57, 3.6,
                      32, "NOSAMS2_3", 38809, 9.32, 5.2
                      ) %>% 
-  mutate(wheel = "USAMS061821")
+  mutate(wheel = "USAMS062521")
 
-carb_data <- bind_rows(cd1, cd2, cd3, cd4, cd5, cd6, cd7, cd8, cd9) %>% 
+
+carb_data <- bind_rows(cd1, cd2, cd3, cd4, cd5, cd6, cd7, cd8, cd9, cd10) %>% 
   select(wheel, pos, name, rec_num, mass, he_added)
 
 write_csv(carb_data, here("data/carb_data.csv"))

### 2021-07-03 R/analyse_carbonates.R
Add wheel

@@ -11,17 +11,18 @@ library(hgis)
 # TODO: Handle case of only one blank
 
 # Analysis metadata: wheels to analyse, outliers, positions of standards and blanks
-wheels <- c("020521", "030421", "040121", "041521", "051421", "052621", "061121", "061821", "062521")
+wheels <- c("020521", "030421", "040121", "041521", "051421", "052621", "061121", "061821", "062521", "062921")
 files <- here("data", paste0("USAMS", wheels, "R.txt"))
-dates <- as.Date(c("2021-02-05", "2021-03-05", "2021-04-09", "2021-04-16", "2021-05-14", "2021-05-28", "2021-06-11", "2021-06-18", "2021-06-25"))
-stds <- list(c(5,8), NULL, c(26,37), NULL, c(10,15,65,70), NULL, NULL, NULL, NULL)
-blanks <- list(7, c(7,8), c(28,29,30), c(27,28,29), c(20,35,50), NULL, NULL, 23, NULL)
+dates <- as.Date(c("2021-02-05", "2021-03-05", "2021-04-09", "2021-04-16", "2021-05-14", "2021-05-28", "2021-06-11", "2021-06-18", "2021-06-25", "2021-07-02"))
+stds <- list(c(5,8), NULL, c(26,37), NULL, c(10,15,65,70), NULL, NULL, NULL, NULL, NULL)
+blanks <- list(7, c(7,8), c(28,29,30), c(27,28,29), c(20,35,50), NULL, NULL, 23, NULL, NULL)
 outliers <- list(NULL, tibble(pos = c(6, 8, 8, 8), meas = c(18, 1, 2, 3)),
                  tibble(pos = c(28,29), meas=c(1, 1)),
                  NULL,
                  tibble(pos=10, meas=1),
                  tibble(pos=c(8,9,13), meas=c(6,1,1)),
                  tibble(pos=c(4,5,5,6,7,8,15), meas=c(1,1,4,10,3,3,2)),
+                 NULL,
                  NULL,
                  NULL)
                  

### 2021-07-03 R/carbonate_reanalysis.Rmd
Move analysis to analyse_carbonates.R

@@
 
+Run analysis script
+
 ```{r}
-# TODO: Handle case of only one blank
-
-wheels <- c("020521", "030421", "040121", "041521", "051421", "052621", "061121", "061821")
-files <- here("data", paste0("USAMS", wheels, "R.txt"))
-dates <- as.Date(c("2021-02-05", "2021-03-05", "2021-04-09", "2021-04-16", "2021-05-14", "2021-05-28", "2021-06-11", "2021-06-18"))
-stds <- list(c(5,8), NULL, c(26,37), NULL, c(10,15,65,70), NULL, NULL, NULL)
-blanks <- list(7, c(7,8), c(28,29,30), c(27,28,29), c(20,35,50), NULL, NULL, 23)
-outliers <- list(NULL, tibble(pos = c(6, 8, 8, 8), meas = c(18, 1, 2, 3)),
-                 tibble(pos = c(28,29), meas=c(1, 1)),
-                 NULL,
-                 tibble(pos=10, meas=1),
-                 tibble(pos=c(8,9,13), meas=c(6,1,1)),
-                 tibble(pos=c(4,5,5,6,7,8,15), meas=c(1,1,4,10,3,3,2)),
-                 NULL)
-
-results <- pmap(list(files, dates, stds, blanks, outliers), reduce_hgis)
-raw_df <- map_dfr(results, 1)
-results_df <- map_dfr(results, 2)
-
-write_csv(raw_df, here("data_analysed/carb_all_raw.csv"))
-write_csv(results_df, here("data_analysed/carb_all_results.csv"))
+source(here("R/analyse_carbonates.R"))
 ```
@@
   select(wheel, pos, sample_name, fm_consensus, fm_corr, sig_fm_corr, fm_diff, sigma) %>% 
-  filter(wheel == "USAMS061821") %>% 
   arrange(desc(fm_diff))

### 2021-07-04 R/carbonate_samples.R
Add haiti samples

@@ -134,8 +134,28 @@ cd10 <- tribble(~pos, ~name, ~rec_num, ~mass, ~he_added,
                      ) %>% 
   mutate(wheel = "USAMS062521")
 
+cd11 <- tribble(~pos, ~name, ~rec_num, ~mass, ~he_added,
+                    10, "C-1_1", 83028, 15.01, 4,      
+                    22, "C-1_2", 83028, 15.13, 4,      
+                     6, "6:7_39-40", 171996, 16.12, 1,  
+                     7, "4:7_10-11", 171992, 16.54, 3,  
+                     8, "1:1_12-13", 171987, 18.33, 2.2,  
+                     9, "1:1_81_82", 171988, 15.89, 3.2,  
+                    12, "2:7_90-91", 171989, 16.97, 3,  
+                    13, "3:7_25-26", 171990, 16.12, 3.2,  
+                    14, "3:7_73-74", 171991, 17.37, 2.8,  
+                    15, "4:7_95-96", 171993, 17.32, 3,  
+                    17, "5:7_57-58", 171994, 14.91, 4,  
+                    18, "5:7_108-109", 171995, 16.08, 3,
+                    19, "7:7_8-9", 172000, 15.95, 2,    
+                    20, "7:7_47-48", 172001, 16.09, 2,  
+                    23, "6:7_60-61", 171997, 16.09, 3.6,  
+                    24, "6:7_90-91", 171998, 18.37, 2.6,  
+                    25, "6:7_120-121", 171999, 15.36, 4) %>% 
+  mutate(wheel = "USAMS062921")
+                
 
-carb_data <- bind_rows(cd1, cd2, cd3, cd4, cd5, cd6, cd7, cd8, cd9, cd10) %>% 
+carb_data <- bind_rows(cd1, cd2, cd3, cd4, cd5, cd6, cd7, cd8, cd9, cd10, cd11) %>% 
   select(wheel, pos, name, rec_num, mass, he_added)
 
 write_csv(carb_data, here("data/carb_data.csv"))

### 2021-07-04 R/analyse_carbonates.R
Save RDS of results

@@ -26,10 +26,13 @@ outliers <- list(NULL, tibble(pos = c(6, 8, 8, 8), meas = c(18, 1, 2, 3)),
                  NULL,
                  NULL)
                  
-
+#carb_metadata <- tibble(files, dates, stds, blanks, outliers)
+#write_csv(carb_metadata, here("data/carb_metadata.csv"))
 results <- pmap(list(files, dates, stds, blanks, outliers), reduce_hgis)
 raw_df <- map_dfr(results, 1)
 results_df <- map_dfr(results, 2)
+
+saveRDS(results, here("data_analysed/carb_all.Rdata"))
 
 write_csv(raw_df, here("data_analysed/carb_all_raw.csv"))
 write_csv(results_df, here("data_analysed/carb_all_results.csv"))

### 2021-07-04 R/HGIS_pendleton_all.Rmd
Add notebook looking at all haiti carbonates.

--- 2021-06-24 R/HGIS_pendleton_2021-05-28.Rmd
@@
 ---
-title: "HGIS 2021-05-28 Carbonates"
+title: "HGIS Pendleton Carbonates"
 author: "Brett Longworth"
-date: "6/5/2021"
+date: "2021-07-05"
 output: html_document
@@
 
-Data reduction and analysis for carbonate samples run on May 28, 2021. Includes C-1, C-2, and test shell carbonates from Simon Pendleton.
-
-```{r}
-
-#Raw AMS data results file path
-file <- here("data/USAMS052621R.txt")
-
-# Date samples run (needed to separate other experiments with this wheel/targets)
-date <- as.Date("2021-05-28")
-
-# Positions of normalizing standards
-standards <- c(5, 8, 13, 17)
-
-# Positions of blanks for blank correction
-blanks <- c(6, 9, 11)
-```
+Data analysis for Haiti core carbonates from Simon Pendleton. Data reduction is handled by `carbonate_samples.R` using the method and functions from the `hgis` package.
 
@@
 
-## raw results
+## Load results
 
 ```{r}
-df <- process_hgis_results(file, date, standards)
+df <- readRDS(here("data_analysed/carb_all.Rdata"))
 #write_csv(df, here("data_analysed/HGIS_2021-05-28_raw.csv"))

### 2021-07-05 R/HGIS_pendleton_all.Rmd
Add new Haiti data. Comparison looks great!

@@
 
-- Using standards in positions `r standards` for normalization.
-- Using blanks in positions `r blanks` for blank correction.
+- Using standards in positions XXX for normalization.
+- Using blanks in positions XXX for blank correction.
 - Outliers are defined as points 1.5 * interquartile range above the 3rd quartile or below the 1st quartile.
@@
 ```{r}
-df <- readRDS(here("data_analysed/carb_all.Rdata"))
-#write_csv(df, here("data_analysed/HGIS_2021-05-28_raw.csv"))
-#write.xlsx(df, here("data_analysed/HGIS_2021-05-28_raw.xlsx"))
+data <- readRDS(here("data_analysed/carb_all.Rdata"))
+sample_data <- read_csv(here("data/carb_data.csv")) #%>% 
+  #select(wheel, pos, mass, he_added)
+
+raw <- map_dfr(data, 1) %>%  
+  filter(wheel %in% c("USAMS052621", "USAMS062921"),
+         pos > 4)
+
+results <- map_dfr(data, 2) %>% 
+  filter(wheel %in% c("USAMS052621", "USAMS062921"),
+         pos > 4) %>% 
+  left_join(sample_data, by = c("wheel", "pos")) %>% 
+  mutate(rec_num = ifelse(is.na(rec_num.y), rec_num.x, rec_num.y)) %>% 
+  select(-rec_num.x, -rec_num.y)
 ```
@@
 ```{r}
-df %>% 
-  filter(!(pos %in% 1:4)) %>% 
-  plot_hgis_time(norm_ratio, sig_norm_ratio, outlier = !.$ok_calc) +
+raw %>% 
+  plot_hgis_time(norm_ratio, outlier = ok_calc) +
   labs(title = "Sample ratio vs. time",
@@
 ```{r}
-df %>% 
-  filter(!(pos %in% 1:4)) %>% 
+raw %>% 
   plot_hgis_time(he12C * 1E6, outlier = !.$ok_calc) +
@@
 
-### Manual outlier flagging
-
-```{r}
-outliers <- tribble(~pos, ~meas,
-                    8, 6,
-                    9, 1,
-                    13, 1)
-df_clean <- flag_outliers(df, outliers)
-```
-
-```{r}
-df_clean %>% 
-  filter(!(pos %in% 1:4)) %>% 
-  plot_hgis_time(norm_ratio, sig_norm_ratio, outlier = !.$ok_calc) +
-  labs(title = "Sample ratio vs. time",
-       y = "Fraction modern")
-```
-
-```{r}
-df_clean %>% 
-  filter(!(pos %in% 1:4)) %>% 
-  plot_hgis_time(he12C * 1E6, outlier = !.$ok_calc) +
-  labs(title = "Sample current vs time",
-       x = "Time (s)",
-       y = "Current (uA)")
-```
-
-## Data Reduction
-
-```{r}
-df_sum <- sum_hgis_targets(df_clean) %>% 
-  norm_hgis() %>% 
-  blank_cor_hgis(blanks = c(6, 9, 11))
-
-write_csv(df_sum, here("data_analysed/HGIS_2021-05-28_results.csv"))
-# write.xlsx(df_sum, here("data_analysed/HGIS_2021-05-28_results.xlsx"))
-```
-
 ## Results
@@
 ```{r}
-df_sum %>% 
-  filter(as.numeric(pos) > 4) %>% 
+results %>% 
   mutate(he12C = he12C * 1E6) %>% 
@@
 ```{r}
-df_sum %>% 
-  filter(as.numeric(pos) > 4) %>% 
-plot_hgis_summary()
+plot_hgis_summary(results)
 ```
@@
 ```{r}
-df_sum %>% 
-  filter(as.numeric(pos) > 4) %>% 
+results %>% 
   compare_replicates() %>% 
@@
 ```{r}
-cons <- df_sum %>% 
-  filter(as.numeric(pos) > 4,
-         !is.na(fm_consensus)) %>% 
+cons <- results %>% 
+  filter(!is.na(fm_consensus)) %>% 
   select(sample_name, fm_consensus, fm_corr, sig_fm_corr) %>% 
@@
 ```{r}
-haiti <- filter(df_sum, str_detect(sample_name, "HATC1")) %>% 
-  mutate(rec_num = case_when(pos == 14 ~ 171989,
-                             pos == 15 ~ 171995,
-                             pos == 16 ~ 171999))
-
-graphite_results <- getRecOS(c(171999,171995,171989)) %>% 
+graphite_results <- getRecOS(results$rec_num) %>% 
   select(rec_num, fm_corr = f_modern, sig_fm_corr = f_ext_error)
 
-combined_results <- haiti %>% 
+combined_results <- results %>% 
   select(sample_name, rec_num, fm_corr, sig_fm_corr) %>% 
-  left_join(graphite_results,
+  inner_join(graphite_results,
             by = "rec_num", 
@@
   select(sample_name, rec_num, method, fm_corr, sig_fm_corr) %>% 
+  arrange(rec_num) %>% 
 knitr::kable()

### 2021-07-06 R/HGIS_pendleton_all.Rmd
Update Haiti carbonate analysis

@@
 raw %>% 
-  plot_hgis_time(norm_ratio, outlier = ok_calc) +
+  filter(str_starts(sample_name, "Live")) %>% 
+  group_by(pos) %>% 
+  mutate(outlier = ifelse(is.na(removeOutliers(norm_ratio)), TRUE, FALSE)) %>%
+  plot_hgis_time(norm_ratio, outlier = outlier) +
   labs(title = "Sample ratio vs. time",
@@
 raw %>% 
+  filter(str_starts(sample_name, "Live")) %>% 
   plot_hgis_time(he12C * 1E6, outlier = !.$ok_calc) +
@@
 combined_results <- results %>% 
-  select(sample_name, rec_num, fm_corr, sig_fm_corr) %>% 
+  select(wheel, sample_name, rec_num, fm_corr, sig_fm_corr) %>% 
   inner_join(graphite_results,
@@
 ```
+```{r}
+combined_results %>% 
+  filter(wheel == "USAMS062921", method == 'hgis') %>% 
+  select(sample_name, rec_num, fm_corr, sig_fm_corr, fm_diff) %>% 
+  pull(fm_diff) %>% 
+  mean()
+```
 
@@
                   position = position_dodge2(width = 0.02)) +
-  ylim(-0.02, 0.02) +
+  #ylim(-0.02, 0.02) +
   labs(title = "Haiti Core Graphite vs HGIS",
@@
 ```
+
+```{r}
+combined_results %>% 
+  filter(fm_diff > 0.02)
+```
+
+Sample 6:7_39-40 had an issue during prep. Vial was found to be at ~1 ATM during backfilling with helium. This means the vial had a leaky seal and would have entrained atmosphere during prep.

### 2021-07-06 R/analyse_carbonates.R

@@ -24,7 +24,7 @@ outliers <- list(NULL, tibble(pos = c(6, 8, 8, 8), meas = c(18, 1, 2, 3)),
                  tibble(pos=c(4,5,5,6,7,8,15), meas=c(1,1,4,10,3,3,2)),
                  NULL,
                  NULL,
-                 NULL)
+                 tibble(pos=c(16), meas=c(6)))
                  
 #carb_metadata <- tibble(files, dates, stds, blanks, outliers)
 #write_csv(carb_metadata, here("data/carb_metadata.csv"))

### 2021-07-09 R/HGIS_pendleton_all.Rmd
Add joins to sample data and depth,plots, etc.

--- 2021-07-05 R/HGIS_pendleton_all.Rmd
@@
 library(hgis)
+library(rbacon)
 
@@
 
-Data analysis for Haiti core carbonates from Simon Pendleton. Data reduction is handled by `carbonate_samples.R` using the method and functions from the `hgis` package.
+Data analysis for Haiti core carbonates from Simon Pendleton.
+Data reduction is handled by the `analyse_carbonates.R` script in the `HybridGIS_work` project using the method and functions from the `hgis` package.
 
-- Using standards in positions XXX for normalization.
-- Using blanks in positions XXX for blank correction.
-- Outliers are defined as points 1.5 * interquartile range above the 3rd quartile or below the 1st quartile.
-- Outliers are NOT dropped from the analysis.
+- Data are normalized using a CO2 tank standard in sample vials.
+- Method blank correction uses carbonate C-1 in vials and assumes that all samples have the same blank, including standards.
 
@@
 1. Apply 13C correction to each run of a target
-1. Flag outliers by interquartile distance
 1. Determine per-target means, internal and external errors
@@
 ```{r}
+# Read raw and reduced data as produced by analyse_carbonates.R
 data <- readRDS(here("data_analysed/carb_all.Rdata"))
-sample_data <- read_csv(here("data/carb_data.csv")) #%>% 
-  #select(wheel, pos, mass, he_added)
 
+# Read sample data
+sample_data <- read_csv(here("data/carb_data.csv"))
+
+# Select relevant raw data
 raw <- map_dfr(data, 1) %>%  
@@
 
+# Select relevant reduced data, join with sample data
 results <- map_dfr(data, 2) %>% 
@@
   select(-rec_num.x, -rec_num.y)
+
+#write results to excel
+#write.xlsx(results, here("data_analysed/haiti_carbonates.xlsx"))
 ```
 
-#### Currents and ratios vs time.
+### Currents and ratios vs time.
 
-```{r}
+```{r, fig.height=10, fig.width = 10}
 raw %>% 
-  plot_hgis_time(norm_ratio, outlier = ok_calc) +
+  plot_hgis_time(norm_ratio, outlier = !.$ok_calc) +
   labs(title = "Sample ratio vs. time",
@@
 
-```{r}
+```{r, fig.height=10, fig.width = 10}
 raw %>% 
@@
   knitr::kable()
-  # gt::gt()
-```
-
-```{r}
-plot_hgis_summary(results)
 ```
@@
 ```{r}
+names <- results %>% 
+  ungroup() %>% 
+  filter(wheel == "USAMS062921") %>% 
+  select(rec_num, sample_name)
+
 graphite_results <- getRecOS(results$rec_num) %>% 
-  select(rec_num, fm_corr = f_modern, sig_fm_corr = f_ext_error)
+  filter(!is.na(reported)) %>% 
+  select(rec_num, fm_corr = f_modern, sig_fm_corr = f_ext_error) %>% 
+  mutate(method = "graphite") %>% 
+  left_join(names, by = "rec_num")
 
 combined_results <- results %>% 
-  select(sample_name, rec_num, fm_corr, sig_fm_corr) %>% 
-  inner_join(graphite_results,
-            by = "rec_num", 
-            suffix = c("_hgis", "_graphite")) %>% 
-  mutate(fm_diff = fm_corr_graphite - fm_corr_hgis,
-         fm_mean = mean(c(fm_corr_hgis, fm_corr_graphite))) %>% 
-  pivot_longer(c(fm_corr_graphite, sig_fm_corr_graphite, 
-                 fm_corr_hgis, sig_fm_corr_hgis), 
-               names_to = c(".value", "method"),
-               names_pattern = "(.*)_(.*)") %>% 
-  mutate(fm_diff = fm_corr - fm_mean)
+  filter(str_starts(sample_name, "HATC") | str_starts(sample_name, "\\d")) %>% 
+  mutate(method = "hgis") %>% 
+  select(wheel, sample_name, method, rec_num, fm_corr, sig_fm_corr) %>% 
+  bind_rows(graphite_results) %>% 
+  group_by(rec_num) %>% 
+  mutate(fm_mean = mean(fm_corr),
+         fm_diff = fm_corr - fm_mean) %>% 
+  left_join(rename(names, Name = sample_name), by = "rec_num") %>% 
+  ungroup()
 
@@
 
-Difference plot
+### Difference plot
 
@@
                       ymax = fm_diff + sig_fm_corr),
-                  size = 1,
-                  position = position_dodge2(width = 0.02)) +
-  ylim(-0.02, 0.02) +
+                  size = 1) +#,
+                  #position = position_dodge2(width = 0.02)) +
+  #ylim(-0.02, 0.02) +
   labs(title = "Haiti Core Graphite vs HGIS",
@@
 ```
+
+```{r}
+combined_results %>% 
+  ggplot(aes(rec_num, fm_corr, color = method, shape = method)) +
+  geom_pointrange(aes(ymin = fm_corr - sig_fm_corr, 
+                      ymax = fm_corr + sig_fm_corr),
+                  size = .5,
+                  position=position_dodge(width = .5)) +#,
+                  #position = position_dodge2(width = 0.02)) +
+  #ylim(-0.02, 0.02) +
+  labs(title = "Haiti Core Graphite vs HGIS",
+       y = "Fraction modern") +
+  theme_classic() +
+  theme(axis.text.x = element_blank(),
+        axis.ticks.x = element_blank()) +
+  facet_wrap("Name", scales = "free")
+```
+
+### Outlier
+
+```{r}
+combined_results %>% 
+  filter(abs(fm_diff) > 0.02) %>% 
+  select(sample_name, method, fm_corr, sig_fm_corr)
+```
+
+Sample 6:7_39-40 had an issue during prep.
+Vial was found to be at ~1 ATM during backfilling with helium.
+This means the vial had a leaky seal and would have entrained atmosphere (Fm ~0.9) during prep.
+This is the likely cause of a HGIS Fm 5.2% higher than graphite.
+
+## Age vs depth
+
+Add depth data and radiocarbon age
+
+```{r}
+depth <- read_csv(here("data/haiti_depth.csv"))
+
+combined_results <- left_join(combined_results, depth, by = "rec_num") %>% 
+  mutate(rc_age = -8033 * log(fm_corr),
+         sig_rc_age = rc_age - -8033 * log(fm_corr + sig_fm_corr))
+```
+
+Fm depth plot
+
+```{r}
+combined_results %>% 
+  filter(sample_name != "6:7_39-40") %>% 
+ggplot(aes(depth, rc_age, color = method)) +
+  geom_pointrange(aes(ymin = rc_age - sig_rc_age, 
+                      ymax = rc_age + sig_rc_age),
+                  size = 0.5) +
+  geom_smooth() +
+  ggtitle("Age vs depth for graphite and HGIS")
+```
+
+Mmmm.... Bacon
+
+```{r}
+# Bacon data format
+# graphite
+combined_results %>% 
+  filter(method == "graphite",
+         rc_age > 0) %>% 
+  select(labID = Name, age = rc_age, error = sig_rc_age, depth) %>% 
+  write_csv(here("Bacon_runs/haiti_graphite/haiti_graphite.csv"))
+
+# hgis
+combined_results %>% 
+  filter(method == "hgis",
+         sample_name != "6:7_39-40",
+         !str_starts(sample_name, "HATC"),
+         rc_age > 0) %>% 
+  select(labID = Name, age = rc_age, error = sig_rc_age, depth) %>% 
+  write_csv(here("Bacon_runs/haiti_hgis/haiti_hgis.csv"))
+```

### 2021-07-26 doc/method_validation/method_validation.Rmd
Add work on method validation and error propagation.

---
title: "Error propagation error"
output: html_notebook
---

Trying to find the source of differences between MR's excel error prop and mine in R.

```{r}
library(tidyverse)
library(hgis)
library(readxl)
library(here)
library(errors)
```

Load results from R data reduction
```{r}
results <- readRDS(here("data_analysed/carb_all.Rdata"))
df <- map_dfr(results[10], 2)
View(df)
```

Load MR's results. I think we'd have to grab individual cells to do this.

```{r}
dfm <- read_excel(here("doc/method_validation/HGIS\ Results.xlsx"))
```

Compare raw ratios and errors

Looking at the data, the difference is in the 2 methods for calculating the error in the Fm of normalizing standards.

MR's method:

```{r}
stds <- df %>% 
  ungroup() %>% 
  filter(sample_type == "S") %>% 
  select(pos, corr_14_12, int_err, ext_err, max_err)

stds
```

```{r}
summarize(stds, ratio_std = mean(corr_14_12),
            sig_ratio_std = sqrt(sum(max_err^2))/n())
```

BL's method

```{r}
summarize(stds, ratio_std = mean(corr_14_12),
            sig_ratio_std = amstools::se(corr_14_12))
```

Propagated error of averaging using the `errors` package.

```{r}
stds <- stds %>% 
  mutate(corr_14_12_err = set_errors(corr_14_12, max_err))

ratio_std <- stds%>% 
  summarize(ratio_std = mean(corr_14_12_err)) %>% 
  pull(ratio_std)
```

Trying to get Fm using errors package propagation

```{r}
fm_std <- set_errors(1.03098, 0.0006)

df <- df %>% 
  mutate(corr_14_12_err = set_errors(corr_14_12, max_err),
         norm_ratio_err = norm_gas(corr_14_12_err, ratio_std, fm_std))

df %>% select(pos, sample_name, norm_ratio, sig_norm_ratio, norm_ratio_err)

```

### 2021-07-26 doc/method_validation/normalization error.Rmd
Add work on method validation and error propagation.

---
title: "Normalization Error"
author: "Brett Longworth"
date: "7/14/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Mark's method

Unknowns are normalized to the mean ratio of standards and then converted to Fm using the Fm of the normalizing standard:

$$ R_{norm} = \frac{R_{sample}}{R_{std}}$$

$$Fm_{sample} = R_{sample}\cdot Fm_{std}$$

With error propagation in 2 steps:

$$ \sigma_{R_{norm}} = R_{norm} \sqrt{\frac{\sigma_{sample}^2}{R_{sample}^2} + \frac{\sigma_{std}^2}{R_{std}^2}}$$

$$ \sigma_{Fm_{sample}} = Fm_{sample} \sqrt{\frac{\sigma_{R_{norm}}^2}{R_{norm}^2} + \frac{\sigma_{Fm_{std}}^2}{Fm_{std}^2}} $$

## Normalize targets using mean of standards and propagate errors

Unknowns are normalized using the mean measured value of gas standards using this equation:

$$ Fm_{sample} = \frac{R_{sample}\cdot Fm_{std}}{R_{std}}$$

If the expected fraction modern ($Fm_{std}$) of the standard is taken as a measured value with an error, rather than a constant or assigned ratio, normalization errors are propagated using this equation:

$$ \sigma_{Fm_{sample}} = Fm_{sample} \sqrt{\frac{\sigma_{sample}^2}{R_{sample}^2} + \frac{\sigma_{std}^2}{R_{std}^2} +
\frac{\sigma_{Fm_{std}}^2}{Fm_{std}^2}} $$

### 2021-08-04 R/haiti_core_depths.R
Add Haiti core depths

@@ -1,0 +1,19 @@
+
+cd <- tribble(~depth, ~rec_num,
+               764.5,	171997,
+               824.5,	171999,
+                12.5,	171987,
+                81.5,	171988,
+               214.5,	171989,
+               294.5,	171990,
+               343.5,	171991,
+               424.5,	171992,
+               504.5,	171993,
+               616.5,	171994,
+               667.5,	171995,
+               743.5,	171996,
+               794.5,	171998,
+               855.5,	172000,
+               894.5,	172001)
+
+write_csv(cd, here("data/haiti_depth.csv"))

### 2021-08-04 doc/method_validation/method_validation.Rmd
Add results comparison

@@
 title: "Error propagation error"
-output: html_notebook
+output:
+  html_document:
+    df_print: paged
 ---
@@
 library(errors)
+library(janitor)
 ```
@@
 df <- map_dfr(results[10], 2)
-View(df)
 ```
 
-Load MR's results. I think we'd have to grab individual cells to do this.
+Load MR's results.
 
 ```{r}
-dfm <- read_excel(here("doc/method_validation/HGIS\ Results.xlsx"))
+dfm <- read_excel(here("doc/method_validation/HGIS\ Results.xlsx"), skip = 2) %>% 
+  select(pos = Pos, sample_name = `Sample Name`, sample_type = Type, 
+         norm_ratio = Fm...5, sig_norm_ratio = dFm...6,
+         fm_corr = Fm...11, sig_fm_corr = dFm...12) %>% 
+  remove_empty("rows")
 ```
@@
 summarize(stds, ratio_std = mean(corr_14_12),
-            sig_ratio_std = sqrt(sum(max_err^2))/n())
+            sig_ratio_std = sqrt(sum(max_err^2))/n(),
+          se_ratio_std = amstools::se(corr_14_12))
 ```
@@
 
-df <- df %>% 
+dfe <- df %>% 
   mutate(corr_14_12_err = set_errors(corr_14_12, max_err),
@@
 
-df %>% select(pos, sample_name, norm_ratio, sig_norm_ratio, norm_ratio_err)
+dfe %>% select(pos, sample_name, norm_ratio, sig_norm_ratio, norm_ratio_err)
 
 ```
+
+# Compare data
+
+Compare data after applying new normalizing standard error propagation...
+
+### Norm Ratio
+
+```{r}
+df %>% 
+  select(pos, sample_name, norm_ratio, sig_norm_ratio) %>% 
+  inner_join(select(dfm, pos, norm_ratio, sig_norm_ratio), by = "pos", suffix = c(".bl", ".mr")) %>% 
+  mutate(norm_ratio_diff = norm_ratio.bl - norm_ratio.mr,
+         sig_norm_ratio_diff = sig_norm_ratio.bl - sig_norm_ratio.mr)
+```
+
+### Blank corrected Fm
+
+```{r}
+df %>% 
+  select(pos, sample_name, fm_corr, sig_fm_corr) %>% 
+  inner_join(select(dfm, pos, fm_corr, sig_fm_corr), by = "pos", suffix = c(".bl", ".mr")) %>% 
+  mutate(fm_diff = fm_corr.bl - fm_corr.mr,
+         sig_fm_diff = sig_fm_corr.bl - sig_fm_corr.mr)
+```

### 2021-08-09 doc/AMS-15/AMS15_abstract.Rmd
First commit.

---
title: "Measurement of CO2 from carbonate minerals using a hybrid sputter gas ion source with an open split interface"
author:
- Brett Longworth
- Josh Burton
- Simon Pendleton
- Josh Hlavenka
- Mark Roberts
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

### National Ocean Sciences AMS Facility, Woods Hole Oceanographic Institution, Woods Hole, MA, USA

We have developed a method for radiocarbon measurement of carbonate minerals as CO~2~ gas via a hybrid sputter gas ion source. Our method uses a carrier gas to displace CO~2~ from sample vials to an open split, where a capillary samples the mixture for delivery directly to a hybrid sputter gas ion source (HGIS). This method skips the gas transfer and quantification steps used in a closed inlet HGIS system, simplifying sample measurement. Results from measurements of consensus standards and samples from a marine core show that the method agrees well with traditional AMS measurement of the same samples, and that the method error is less than 1.5%. We discuss advantanges and disadvantanges of continuous flow sample introduction.

### 2021-08-17 doc/AMS-15/AMS15_abstract.Rmd
update abstract

@@
 - Josh Hlavenka
+- Sydney Moser
 - Mark Roberts
 output:
+  word_document: default
+  html_document: default
   pdf_document: default
-  html_document: default
-  word_document: default
 ---
@@
 
-We have developed a method for radiocarbon measurement of carbonate minerals as CO~2~ gas via a hybrid sputter gas ion source.
-Our method uses a carrier gas to displace CO~2~ from sample vials to an open split, where a capillary samples the mixture for delivery directly to a hybrid sputter gas ion source (HGIS).
-This method skips the gas transfer and quantification steps used in a closed inlet HGIS system, simplifying sample measurement.
-Results from measurements of consensus standards and samples from a marine core show that the method agrees well with traditional AMS measurement of the same samples, and that the method error is less than 1.5%.
-We discuss advantanges and disadvantanges of continuous flow sample introduction.
+We have developed a method for measurement of radiocarbon in carbonate minerals
+as CO~2~ gas via a NEC MCSNICS hybrid sputter gas ion source. Our method uses a
+carrier gas to displace CO~2~ from sample vials to an open split, where a
+capillary samples the mixture for delivery directly to a hybrid sputter gas ion
+source (HGIS).  This method skips the gas transfer and quantification steps
+used in a closed inlet HGIS system, simplifying sample measurement. Results
+from measurements of consensus standards (TIRI-I, IAEA C2, and an internal
+modern shell standard), and samples from a marine core show that the method
+agrees well with traditional AMS measurement of the same samples, and that the
+method error is less than 1.5%. We discuss advantanges and disadvantanges of
+continuous flow sample introduction.

### 2021-08-18 doc/AMS-15/AMS15_abstract.Rmd
Add author

@@
 - Mark Roberts
+- Mark Kurz
 output:

### 2021-10-08 R/HGIS_pendleton_all.Rmd
Playing with Bchron for calibration
reformat age-depth plot

@@
 options(scipen = 3)
+theme_set(theme_classic())
 ```
@@
   geom_smooth() +
+  coord_flip() +
+  scale_y_reverse() +
+  scale_x_reverse() +
   ggtitle("Age vs depth for graphite and HGIS")
@@
 ```
+
+Trying Bchron here instead. Use is a bit friendlier in Rmarkdown.
+
+```{r}
+library(Bchron)
+
+cal_sub <- combined_results %>% 
+  filter(rc_age > 0) %>% 
+  mutate(id = paste(sample_name, "-", method)) %>% 
+  select(id, depth, rc_age, sig_rc_age)
+
+cal_dates <- with(cal_sub,
+  BchronCalibrate(rc_age, sig_rc_age, positions = depth, ids = id))
+plot(cal_dates)
+```

### 2021-10-08 doc/AMS-15/AMS15_abstract.Rmd
Update abstract

--- 2021-08-09 doc/AMS-15/AMS15_abstract.Rmd
@@
 ---
-title: "Measurement of CO2 from carbonate minerals using a hybrid sputter gas ion source with an open split interface"
+title: "Radiocarbon measurement of CO~2~ from carbonate minerals using a gas ion source with an open split interface"
 author:
@@
 - Josh Hlavenka
+- Sydney Moser
 - Mark Roberts
+- Mark Kurz
 output:
+  word_document: default
+  html_document: default
   pdf_document: default
-  html_document: default
-  word_document: default
 ---
@@
 
-We have developed a method for radiocarbon measurement of carbonate minerals as CO~2~ gas via a hybrid sputter gas ion source.
-Our method uses a carrier gas to displace CO~2~ from sample vials to an open split, where a capillary samples the mixture for delivery directly to a hybrid sputter gas ion source (HGIS).
-This method skips the gas transfer and quantification steps used in a closed inlet HGIS system, simplifying sample measurement.
-Results from measurements of consensus standards and samples from a marine core show that the method agrees well with traditional AMS measurement of the same samples, and that the method error is less than 1.5%.
-We discuss advantanges and disadvantanges of continuous flow sample introduction.
+We have developed a method for measurement of radiocarbon in carbonate minerals
+as CO~2~ gas via a NEC MCSNICS hybrid sputter gas ion source (HGIS). The method
+uses helium as a carrier gas to displace CO~2~ from sample vials to an open
+split, where a glass capillary samples the mixture for delivery directly to the
+HGIS. This method skips the gas transfer and quantification steps used in a
+closed inlet HGIS system, simplifying sample measurement. Samples larger than 10
+mg carbonate can currently be measured. Results from measurements of consensus
+standards (TIRI-I, IAEA C2, and an internal modern shell standard), and samples
+from a marine core (40-115 pMC) show that the method agrees well with
+traditional AMS measurement of the same samples as graphite, and that the method
+error is less than 1.5 pMC. We discuss advantages and disadvantages of
+continuous flow sample introduction, and the effect of reduced precision on
+calibrated age-depth models produced using gas-source data.

### 2021-10-08 doc/NAPB2021/NAPB2021.Rmd
Figures for NAPB2021

---
title: "HGIS Carbonate Results"
subtitle: "NAPB 2021"
author: 
  - "Brett Longworth"
date: '2021-10-08'
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=6, fig.height=3, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  fig.show = TRUE,
  hiline = TRUE
)
```

```{r}
library(tidyverse)
library(amstools)
library(hgis)
library(here)
library(gt)
library(Bchron)

options(digits = 3)
options(scipen = 4)
options(decimals = 4)
theme_set(theme_classic())
```

## Method

1. Acidify
  - ~30mg Carbonate 
  - produces 6mL CO<sub>2</sub> @ ~1ATM
1. Displace gas to open split with 100 uL/min He.
1. Introduce mixture to source @ 30 uL/min
1. Acquire data: 90s x 10 measurements.

- Standard: live gas in vials
- Blank: C-1 or TIRI-F in vials

![HGIS Figure](images/HGIS_open_split_drawing.svg)

## Data reduction

1. Correct for fractionation using online <sup>13</sup>C.
1. Determine per-sample mean and error.
1. Normalize to mean of Gas standard.
1. Blank correct: LBC using mean of carbonate blanks

- Errors propagated at each step. 
- Per-sample error agrees well with variability of replicates.

Data reduction done using an R package maintained on github:

https://github.com/blongworth/hgis

## Results: Agreement with Consensus

```{r}
# Read data. Don't use first carbonate test, smaller samples, or gas standards.
results_df <- read_csv(here("data_analysed/carb_all_results.csv"))

cons_df <- results_df %>% 
  filter(!(wheel %in% c("USAMS020521", "USAMS061121", "USAMS061821"))) %>% 
  mutate(he12C = he12C * 1E6) %>% 
  compare_consensus() %>% 
  filter(!str_detect(Name, "Gas"))
```

## Results: Method precision and accuracy

```{r}
cons_df_sum <- cons_df %>% 
  group_by(Name, fm_consensus) %>% 
  summarize(across(c(fm_corr, sig_fm_corr, fm_diff, sigma),
                   list(mean = mean, sd = sd)),
            N = n()) %>% 
  ungroup()

cons_df_sum %>% 
  arrange(fm_consensus) %>% 
  select(Name, `Consensus Fm` = fm_consensus,
         `Mean Fm` = fm_corr_mean,
         `Mean Fm error` = sig_fm_corr_mean,
         `SD of measurements` = fm_corr_sd,
         N) %>% 
gt() %>% 
  tab_header(title = "Summary statistics by sample type") %>% 
  fmt_number(3:5, decimals = 4)
```

```{r}
cons_df %>% 
  filter(Name %in% c("TIRI-I", "C-2", "NOSAMS2")) %>% 
  ungroup() %>% 
  summarize(across(c(sig_fm_corr, fm_diff, sigma),
                   list(mean = mean, sd = sd)),
            N = n()) %>% 
  select(`Mean Fm difference` = fm_diff_mean,
         `SD of mean difference` = fm_diff_sd,
         `Mean Fm error` = sig_fm_corr_mean,
         `Mean Sigma` = sigma_mean,
         `SD of Sigma` = sigma_sd,
         N
         ) %>%  
  gt() %>% 
  tab_header(title = "Summary statisics for reference materials",
             subtitle = "TIRI-I, C-2, and NOSAMS2") %>% 
  fmt_number(3:5, decimals = 4)
```

```{r}
mean_diff <- mean(cons_df$fm_diff)
sd_diff <- sd(cons_df$fm_diff)

cons_df %>% 
ggplot(aes(fm_consensus, fm_diff, color = Name)) + 
  geom_hline(yintercept = 0) +
  geom_hline(yintercept = mean_diff, color = "blue") +
  geom_hline(yintercept = mean_diff + sd_diff, color = "lightblue") +
  geom_hline(yintercept = mean_diff - sd_diff, color = "lightblue") +
  geom_smooth(method = "lm",  se = FALSE) + 
  geom_pointrange(aes(ymin = fm_diff - sig_fm_corr, 
                      ymax = fm_diff + sig_fm_corr), 
                  position = position_dodge2(width = 0.1), 
                  size = 0.2) + 
  labs(subtitle = "Blank corrected", x = "Fm expected (points jittered for clarity)", 
       y = "Fm difference (cons - measured)")
```

## Results: Method precision and accuracy

```{r}
cons_df_sum <- cons_df %>% 
  group_by(Name, fm_consensus) %>% 
  summarize(across(c(fm_corr, sig_fm_corr, fm_diff, sigma),
                   list(mean = mean, sd = sd)),
            N = n()) %>% 
  ungroup()

cons_df_sum %>% 
  arrange(fm_consensus) %>% 
  select(Name, `Consensus Fm` = fm_consensus,
         `Mean Fm` = fm_corr_mean,
         `Mean Fm error` = sig_fm_corr_mean,
         `SD of measurements` = fm_corr_sd,
         N) %>% 
gt() %>% 
  tab_header(title = "Summary statistics by sample type") %>% 
  fmt_number(3:5, decimals = 4)
```

```{r}
cons_df %>% 
  filter(Name %in% c("TIRI-I", "C-2", "NOSAMS2")) %>% 
  ungroup() %>% 
  summarize(across(c(sig_fm_corr, fm_diff, sigma),
                   list(mean = mean, sd = sd)),
            N = n()) %>% 
  select(`Mean Fm difference` = fm_diff_mean,
         `SD of mean difference` = fm_diff_sd,
         `Mean Fm error` = sig_fm_corr_mean,
         `Mean Sigma` = sigma_mean,
         `SD of Sigma` = sigma_sd,
         N
         ) %>%  
  gt() %>% 
  tab_header(title = "Summary statisics for reference materials",
             subtitle = "TIRI-I, C-2, and NOSAMS2") %>% 
  fmt_number(3:5, decimals = 4)
```

## Results: By sample type

```{r}
ggplot(cons_df, aes(x = fm_diff, y = fm_corr, color = he12C)) +
  geom_hline(aes(yintercept = fm_consensus), color = "lightblue") +
  geom_hline(data = cons_df_sum, aes(yintercept = fm_corr_mean), color = "darkgray") +
  geom_hline(data = cons_df_sum, aes(yintercept = fm_corr_mean + fm_corr_sd), 
             color = "grey", linetype = "dashed") +
  geom_hline(data = cons_df_sum, aes(yintercept = fm_corr_mean - fm_corr_sd),
             color = "grey", linetype = "dashed") +

  geom_pointrange(aes(ymin = fm_corr - sig_fm_corr, 
                      ymax = fm_corr + sig_fm_corr), 
                  size = .2) +
  labs(subtitle = "Reference and normalization standards",
       y = "Fraction modern (Fm)") +
  facet_wrap(facets = vars(Name), scales = "free") +
  theme_bw() +
  scale_x_continuous(breaks = NULL) +
  xlab(NULL) +
  scale_color_continuous(name = "12C- (μA)") +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
```

## Pendleton Haiti Core Carbonates

### Read data, pull graphite results and sample depth, combine

```{r}
# Read raw and reduced data as produced by analyse_carbonates.R
data <- readRDS(here("data_analysed/carb_all.Rdata"))

# Read sample data
sample_data <- read_csv(here("data/carb_data.csv"))

# Select relevant reduced data, join with sample data
results <- map_dfr(data, 2) %>% 
  filter(wheel %in% c("USAMS052621", "USAMS062921"),
         pos > 4) %>% 
  left_join(sample_data, by = c("wheel", "pos")) %>% 
  mutate(rec_num = ifelse(is.na(rec_num.y), rec_num.x, rec_num.y)) %>% 
  select(-rec_num.x, -rec_num.y)

names <- results %>% 
  ungroup() %>% 
  filter(wheel == "USAMS062921") %>% 
  select(rec_num, sample_name)

graphite_results <- getRecOS(results$rec_num) %>% 
  filter(!is.na(reported)) %>% 
  select(rec_num, fm_corr = f_modern, sig_fm_corr = f_ext_error) %>% 
  mutate(method = "graphite") %>% 
  left_join(names, by = "rec_num")

combined_results <- results %>% 
  filter(str_starts(sample_name, "HATC") | str_starts(sample_name, "\\d")) %>% 
  mutate(method = "hgis") %>% 
  select(wheel, sample_name, method, rec_num, fm_corr, sig_fm_corr) %>% 
  bind_rows(graphite_results) %>% 
  group_by(rec_num) %>% 
  mutate(fm_mean = mean(fm_corr),
         fm_diff = fm_corr - fm_mean) %>% 
  left_join(rename(names, Name = sample_name), by = "rec_num") %>% 
  ungroup()

depth <- read_csv(here("data/haiti_depth.csv"))

combined_results <- left_join(combined_results, depth, by = "rec_num") %>% 
  mutate(rc_age = -8033 * log(fm_corr),
         sig_rc_age = rc_age - -8033 * log(fm_corr + sig_fm_corr))
```

### Outlier

```{r}
combined_results %>% 
  filter(abs(fm_diff) > 0.02) %>% 
  select(sample_name, method, fm_corr, sig_fm_corr)
```

Sample 6:7_39-40 had an issue during prep. Vial was found to be at ~1 ATM during backfilling with helium. This means the vial had a leaky seal and would have entrained atmosphere (Fm ~0.9) during prep. This is the likely cause of a HGIS Fm 5.2% higher than graphite.

## Difference plot

```{r}
outlier_df <- combined_results %>% 
  filter(abs(fm_diff) > 0.02)

combined_results %>% 
  ggplot() +
  geom_hline(yintercept = 0) +
  geom_pointrange(aes(fm_corr, fm_diff, color = method, shape = method,
                      ymin = fm_diff - sig_fm_corr, 
                      ymax = fm_diff + sig_fm_corr),
                  size = 1) +#,
  #gghighlight::gghighlight(abs(fm_diff > 0.02)) +
  geom_pointrange(data = outlier_df,
                  aes(fm_corr, fm_diff, shape = method, color = method,
                      ymin = fm_diff - sig_fm_corr,
                      ymax = fm_diff + sig_fm_corr),
                  color = "grey",
                  size = 1) +#,
                  #position = position_dodge2(width = 0.02)) +
  #ylim(-0.02, 0.02) +
  labs(title = "Haiti Core Graphite vs HGIS",
       x = "Fraction modern",
       y = "Fm - mean Fm") + 
  theme(legend.position = "none")
```

## Fm depth plot

```{r}
combined_results %>% 
  filter(sample_name != "6:7_39-40") %>% 
ggplot(aes(depth, rc_age, color = method)) +
  geom_pointrange(aes(ymin = rc_age - sig_rc_age, 
                      ymax = rc_age + sig_rc_age),
                  size = 0.5) +
  geom_smooth(alpha = .2) +
  coord_flip() +
  scale_y_reverse() +
  scale_x_reverse() +
  labs(title = "Age vs depth for graphite and HGIS",
       y = "Radiocarbon years (BP)",
       x = "Core depth (cm)")
```

## BChron model

```{r}

cal_sub <- combined_results %>% 
  filter(sample_name != "6:7_39-40") %>% 
  filter(!str_starts(sample_name, "HATC")) %>% 
  filter(rc_age > 0) %>% 
  mutate(id = paste(sample_name, "-", method)) %>% 
  select(id, depth, rc_age, sig_rc_age, method)

cal_dates <- with(cal_sub,
  BchronCalibrate(rc_age, sig_rc_age, positions = depth, ids = id))

cal_dates_gr <- cal_sub %>% 
  filter(method == "graphite") %>% 
  with(
  BchronCalibrate(rc_age, sig_rc_age, positions = depth, ids = id))

cal_dates_gas <- cal_sub %>% 
  filter(method == "hgis") %>%
  with(
  BchronCalibrate(rc_age, sig_rc_age, positions = depth, ids = id))

```

```{r}
plot(cal_dates)
```

```{r}
plot(cal_dates_gr)
```

```{r}
plot(cal_dates_gas)
```

### 2021-10-11 doc/NAPB2021/NAPB2021.Rmd
Use gghighlight for showing outlier

@@
 library(Bchron)
+library(gghighlight)
 
@@
 ```{r}
-outlier_df <- combined_results %>% 
-  filter(abs(fm_diff) > 0.02)
+ggplot(combined_results) +
+  geom_hline(yintercept = 0) +
+  geom_pointrange(aes(fm_corr, fm_diff, color = method, shape = method,
+                      ymin = fm_diff - sig_fm_corr, 
+                      ymax = fm_diff + sig_fm_corr),
+                  size = 1) +
+  labs(title = "Haiti Core Graphite vs HGIS",
+       x = "Fraction modern",
+       y = "Fm - mean Fm")
+```
 
+No outliers
+
+```{r}
 combined_results %>% 
+  filter(abs(fm_diff) < 0.02) %>% 
 ggplot() +
@@
                       ymax = fm_diff + sig_fm_corr),
-                  size = 1) +#,
-  #gghighlight::gghighlight(abs(fm_diff > 0.02)) +
-  geom_pointrange(data = outlier_df,
-                  aes(fm_corr, fm_diff, shape = method, color = method,
+                  size = 1) +
+  labs(title = "Haiti Core Graphite vs HGIS",
+       x = "Fraction modern",
+       y = "Fm - mean Fm")
+```
+
+Highlight outliers
+
+```{r}
+ggplot(combined_results) +
+  geom_hline(yintercept = 0) +
+  geom_pointrange(aes(fm_corr, fm_diff, color = method, shape = method,
                       ymin = fm_diff - sig_fm_corr, 
                       ymax = fm_diff + sig_fm_corr),
-                  color = "grey",
-                  size = 1) +#,
-                  #position = position_dodge2(width = 0.02)) +
-  #ylim(-0.02, 0.02) +
+                  size = 1) +
+  gghighlight(abs(fm_diff) > 0.02,
+              #unhighlighted_params = list(colour = NULL),
+              keep_scales = TRUE) +
   labs(title = "Haiti Core Graphite vs HGIS",
        x = "Fraction modern",
-       y = "Fm - mean Fm") + 
-  theme(legend.position = "none")
+       y = "Fm - mean Fm")
+```
+
+All in one.
+
+```{r}
+outlier_df <- combined_results %>% 
+  filter(abs(fm_diff) > 0.02)
+
+ggplot(combined_results) +
+  geom_hline(yintercept = 0) +
+  geom_pointrange(aes(fm_corr, fm_diff, color = method, shape = method,
+                      ymin = fm_diff - sig_fm_corr, 
+                      ymax = fm_diff + sig_fm_corr),
+                  size = 1,
+                  color = "grey") +#,
+  gghighlight(abs(fm_diff) > 0.02,
+              unhighlighted_params = list(colour = NULL),
+              keep_scales = TRUE) +
+  labs(title = "Haiti Core Graphite vs HGIS",
+       x = "Fraction modern",
+       y = "Fm - mean Fm")
 ```

### 2021-10-12 R/smaller_samples.Rmd
remove samples with no mass.

@@
 df <- left_join(df, samples, by = c("wheel", "pos")) %>% 
-  filter(wheel != "USAMS020521")
+  filter(wheel != "USAMS020521",
+         !is.na(mass))
 ```
+
+### Size and Current
+
+Until we get below 10mg of carbonate, any size-current relationship is mostly obscured by other factors.
 
@@
 df %>% 
-ggplot(aes(mass, c12ua, color = as.factor(rec_num))) +
-  geom_point()
+ggplot(aes(mass, c12ua, color = Name)) +
+  geom_point() +
+  ggtitle()
 ```

### 2021-10-12 doc/AMS-15/AMS-15_poster.Rmd
Put together figures for AMS-15 poster

--- 2021-10-11 doc/NAPB2021/NAPB2021.Rmd
@@
 ---
-title: "HGIS Carbonate Results"
-subtitle: "NAPB 2021"
+title: "Radiocarbon measurement of CO~2~ from carbonate minerals using a gas ion source with an open split interface"
 author:
-  - "Brett Longworth"
-date: '2021-10-08'
+- Brett Longworth
+- Josh Burton
+- Simon Pendleton
+- Josh Hlavenka
+- Sydney Moser
+- Mark Roberts
+- Mark Kurz
+output:
+  html_document: default
+  word_document: default
+  pdf_document: default
 ---
@@
 library(gghighlight)
+library(patchwork)
 
@@
 
-## Method
+Text and figures for AMS-15 poster.
+
+# Abstract
+
+We have developed a method for measurement of radiocarbon in carbonate minerals
+as CO~2~ gas via a NEC MCSNICS hybrid sputter gas ion source (HGIS). The method
+uses helium as a carrier gas to displace CO~2~ from sample vials to an open
+split, where a glass capillary samples the mixture for delivery directly to the
+HGIS. This method skips the gas transfer and quantification steps used in a
+closed inlet HGIS system, simplifying sample measurement. Samples larger than 10
+mg carbonate can currently be measured. Results from measurements of consensus
+standards (TIRI-I, IAEA C2, and an internal modern shell standard), and samples
+from a marine core (40-115 pMC) show that the method agrees well with
+traditional AMS measurement of the same samples as graphite, and that the method
+error is less than 1.5 pMC. We discuss advantages and disadvantages of
+continuous flow sample introduction, and the effect of reduced precision on
+calibrated age-depth models produced using gas-source data.
+
+# Introduction
+
+# Method
 
@@
 
-## Results: Agreement with Consensus
+# Results
+
+Data used for current, blanks, consensus
 
@@
 ```
+## Initial measurements
+
+### Flow vs current
+
+```{r cur-time-os, fig.height=7, dev='svg'}
+data <- get_hgis_data(here("data/USAMS040121R.txt"), as.Date("2021-04-02")) %>% 
+  filter(Pos == 5) %>% 
+  mutate(time = cum_acqtime/60,
+         co2flow = concCO2(time, r = 244) * 30) #30ul/min
+
+# flow vs time subplot
+flow_time <- data.frame(x = 0:125) %>% 
+  ggplot(aes(x)) +
+  stat_function(fun=function(x) concCO2(x, r = 244) * 30, aes(color = "CO2"), size = 1.5) +
+  stat_function(fun=function(x) 30 - (concCO2(x, r = 244) * 30), aes(color = "Helium"), size = 1.5) +
+  scale_color_manual("Gas", values = c("#00b7bd", "#b7bf10")) +
+  labs(title = "Gas flows to source",
+       subtitle = "7mL vial, 244μL/min helium, 30μL/min to source",
+       y = "Gas flow (μL/min)") +
+  theme_classic() +
+  theme(axis.title.x = element_blank(),
+        axis.text.x = element_blank(),
+        legend.position = c(0.87, 0.65),
+        legend.background = element_rect(fill = "white", color = "black"))
+
+# current vs time subplot
+cur_time <- ggplot(data, aes(time, he12C)) +
+  geom_smooth(span = .4, se=FALSE, color = "#00b7bd") +
+  geom_point(size = 3, color = "#0069b1") +
+  xlim(0, 125) +
+  labs(title = "Ion current",
+       x = "Time (min)",
+       y = "12C current (μA)") +
+  theme_classic()
+
+flow_time / cur_time
+```
+
+```{r cur-flow-os, fig.height=6, dev='svg'}
+ggplot(data, aes(co2flow, he12C)) +
+  geom_smooth(span = .3, se=FALSE, color = "#00b7bd") +
+  geom_point(size = 3, color = "#0069b1") +
+  labs(title = "Vial dilution current",
+       subtitle = "250μL/min displacement, 30μl/min delivery",
+       x = bquote('CO'[2]~'Flow (μl/min)'), 
+       y = "12C current (μA)") +
+  theme_classic()
+```
+
+compare to first tests
+
+Note that our flat-topped current curve allows variable flow of the vial system. MICADAS has a more pronounced peak so this may not work as well. (Check papers for peak shape)
+
+Also note current dependent ratio (blank) as a complicating factor
+
+Lead into blank on next slide
+
+### Blanks and blank correction
+
+# Current independent blank
+
+If **standards**, **blanks**, and **unknowns** have the **same** blank and current, current drops out.
+
+#### Linear fit used to correct:
+
+$$Fm_s = Fm_m - Fm_{blk}\frac{Fm_{std} - Fm_m}{Fm_{std}}$$
+
+#### Mean Fm of blanks: `r cons_df %>% filter(fm_consensus < 0.1 & he12C > 2) %>% pull(norm_ratio) %>% mean() %>% format(digits = 3)`
+
+```{r blanks, fig.height=5, dev='svg'}
+cons_df %>% 
+  filter(fm_consensus < 0.1) %>% 
+  ggplot(aes(he12C, norm_ratio)) +
+  geom_errorbar(aes(ymin = norm_ratio - sig_norm_ratio, ymax = norm_ratio  + sig_norm_ratio)) +
+  geom_errorbarh(aes(xmin = he12C - sig_he12C, xmax = he12C  + sig_he12C)) +
+  geom_point(aes(color = wheel), size = 3) +
+  #scale_color_manual("wheel", values = c("#b7bf10", "#00a9e0", "#0069b1")) +
+  labs(title = "Blanks",
+       x = "12C Current (μA)",
+       y = "Fraction modern") +
+  theme_classic() +
+  theme(legend.position = c(0.82, 0.76),
+        legend.background = element_rect(fill = "white", color = "black")) 
+```
+
+## Results: Consensus standards
 

### 2021-10-14 doc/AMS-15/AMS-15_poster.Rmd
Hide some plots

@@
 
-```{r}
+```{r, eval = FALSE}
 ggplot(combined_results) +
@@
 
-No outliers
-
 ```{r}
+# No outliers
 combined_results %>% 
@@
   geom_hline(yintercept = 0) +
-  geom_pointrange(aes(fm_corr, fm_diff, color = method, shape = method,
+  geom_pointrange(aes(depth, fm_diff, color = method, shape = method,
                       ymin = fm_diff - sig_fm_corr, 
                       ymax = fm_diff + sig_fm_corr),
-                  size = 1) +
+                  position = "jitter") +
   labs(title = "Haiti Core Graphite vs HGIS",
-       x = "Fraction modern",
+       x = "Core depth (cm)",
        y = "Fm - mean Fm")
@@
 
-Highlight outliers
-
-```{r}
+```{r, eval=FALSE}
+# Highlight outliers
 ggplot(combined_results) +
@@
 
-All in one.
+```{r eval=FALSE}
+#All in one.
 
-```{r}
 outlier_df <- combined_results %>% 

### 2021-10-15 doc/AMS-15/AMS-15_poster.Rmd
Add text. Revert to hgis - graphite for Haiti comparison plot. Fix colors. Add core top to BChron.

@@
 
-1. Acidify
-  - ~30mg Carbonate 
-  - produces 6mL CO<sub>2</sub> @ ~1ATM
+### Sample Preparation
+
+1. 10-30mg Carbonate in septa-sealed 7mL vials
+  - produces 2-6mL CO<sub>2</sub> @ ~1ATM
+1. Evacuate and Acidify with 1 mL H3PO4 using automated Gilson prep bench
+1. Add helium to 1 ATM
+
+### Measurement
+
 1. Displace gas to open split with 100 uL/min He.
@@
 
-- Standard: live gas in vials
-- Blank: C-1 or TIRI-F in vials
+- Normalizing standards: live tank gas in vials
+- Blanks: C-1 or TIRI-F in vials
 
@@
 
+![HGIS Photo](images/gas_control.jpg)
+![HGIS Photo](images/gilson.png)
+
 ## Data reduction
 
-1. Correct for fractionation using online <sup>13</sup>C.
-1. Determine per-sample mean and error.
+1. Correct for fractionation using online AMS <sup>13</sup>C measurement.
+1. Determine per-sample mean and error (repeatabilty and counting error).
 1. Normalize to mean of Gas standard.
-1. Blank correct: LBC using mean of carbonate blanks
+1. Blank correct: Blank subtraction using mean of carbonate blanks
 
@@
 ### Flow vs current
+
+Introducing sample through an open split using helium displacement results in a decreasing flow of CO2 to the source over time.
+Measured ratio can depend on sample current, so CO2 flow rate must be kept in a range where ion currents are relatively constant with changing flow.
+We evaluated C- current over a range of CO2 flows and found that a range of flows produce a "flat top" in current.
+Less CO2 will starve the source of available material to dissociate and ionize, and too much CO2 also suppresses ionization.
 
@@
 
-# Current independent blank
+#### Current dependent blank
+
+Measurements of live and dead gas over a range of C- currents show that there is a current-dependent component to the blank.
+We believe this is due to a combination of a constant amount of contaminant CO2 and a "target blank" current.
+This current dependent blank follows the constant contamination model and can be represented with a current balance between the sample and the constant blank.
+
+#### Current independent blank
 
@@
 
-#### Linear fit used to correct:
+While this current-dependent blank is important to consider for samples producing varying ion current, we simplified our assessment of blank by keeping sample currents as constant as possible for blanks, normalizing standards, and unknowns, which allows a simpler correction depending only on the Fm of the blank, standard, and unknown.
+
+This leads to the following model for blank correction:
 
 $$Fm_s = Fm_m - Fm_{blk}\frac{Fm_{std} - Fm_m}{Fm_{std}}$$
+
+Data from the measurement of consensus standards over a range of fraction modern shows that this simple correction is effective for removing blank contribution without bias.
 
@@
   mutate(fm_mean = mean(fm_corr),
-         fm_diff = fm_corr - fm_mean) %>% 
+         fm_diff_mean = fm_corr - fm_mean) %>% 
   left_join(rename(names, Name = sample_name), by = "rec_num") %>% 
+  left_join(select(graphite_results, rec_num, fm_corr_gr = fm_corr)) %>% 
+  mutate(fm_diff = fm_corr - fm_corr_gr) %>% 
   ungroup()
@@
   filter(abs(fm_diff) > 0.02) %>% 
-  select(sample_name, method, fm_corr, sig_fm_corr)
+  select(sample_name, method, fm_corr, sig_fm_corr, fm_diff)
 ```
@@
                   size = 1) +
+  scale_color_manual(values = c("#00b7bd", "#b7bf10")) +
   labs(title = "Haiti Core Graphite vs HGIS",
@@
 ```{r}
-# No outliers
+# No outliers, vs depth
 combined_results %>% 
@@
                   position = "jitter") +
+  scale_color_manual(values = c("#00b7bd", "#b7bf10")) +
   labs(title = "Haiti Core Graphite vs HGIS",
        x = "Core depth (cm)",
+       y = "Fm - mean Fm")
+```
+
+```{r}
+# No outliers, vs mean_fm
+combined_results %>% 
+  filter(abs(fm_diff) < 0.02) %>% 
+ggplot() +
+  geom_hline(yintercept = 0) +
+  geom_pointrange(aes(fm_mean, fm_diff, color = method, shape = method,
+                      ymin = fm_diff - sig_fm_corr, 
+                      ymax = fm_diff + sig_fm_corr),
+                  position = "jitter") +
+  geom_smooth(method = "lm", aes(fm_mean, fm_diff, color = method)) +
+  scale_color_manual(values = c("#00b7bd", "#b7bf10")) +
+  labs(title = "Haiti Core Graphite vs HGIS",
+       x = "Mean Fm",
        y = "Fm - mean Fm")
@@
   filter(rc_age > 0) %>% 
-  mutate(id = paste(sample_name, "-", method)) %>% 
-  select(id, depth, rc_age, sig_rc_age, method)
+  mutate(id = paste(sample_name, "-", method),
+         cal_curve = "intcal20") %>% 
+  select(id, depth, rc_age, sig_rc_age, cal_curve, method) %>%
+  add_row(id = "top-1", depth = 0, rc_age = 0, sig_rc_age = 1, cal_curve = "normal", method = "top")
 
 cal_dates <- with(cal_sub,
-  BchronCalibrate(rc_age, sig_rc_age, positions = depth, ids = id))
+  BchronCalibrate(rc_age, sig_rc_age, positions = depth, ids = id, calCurves = cal_curve))
 
 cal_dates_gr <- cal_sub %>% 
-  filter(method == "graphite") %>% 
+  filter(method != "hgis") %>% 
   with(
-  BchronCalibrate(rc_age, sig_rc_age, positions = depth, ids = id))
+  BchronCalibrate(rc_age, sig_rc_age, positions = depth, ids = id, calCurves = cal_curve))
 
 cal_dates_gas <- cal_sub %>% 
-  filter(method == "hgis") %>%
+  filter(method != "graphite") %>%
   with(
-  BchronCalibrate(rc_age, sig_rc_age, positions = depth, ids = id))
+  BchronCalibrate(rc_age, sig_rc_age, positions = depth, ids = id, calCurves = cal_curve))
 
@@
 
-```{r}
+```{r, eval=FALSE}
 plot(cal_dates)

### 2021-10-15 doc/AMS-15/AMS-15_poster.Rmd
Add/edit text. Remove graphite value for hgis outlier.

@@
 
+![HGIS Photo](images/gas_control.jpg)
+
+Our method uses an open split and glass capillary to control gas flow to the ion source.
+Compared to closed inlet systems which regulate introduction of a CO2/Helium mixture to the ion source by regulating sample pressure, our method is simpler and faster, but does not maintain a constant flow of CO2 to the source.
+We compensate for this by limiting CO2 flow to a range that produces relatively constant C- ion current and stable ratios.
+
+![HGIS Figure](images/HGIS_open_split_drawing.svg)
+
 ### Sample Preparation
@@
 1. Add helium to 1 ATM
+
+![HGIS Photo](images/gilson.png)
 
@@
 
-![HGIS Figure](images/HGIS_open_split_drawing.svg)
+## Data reduction
 
-![HGIS Photo](images/gas_control.jpg)
-![HGIS Photo](images/gilson.png)
-
-## Data reduction
+Normalization to standard, blank correction, and estimation of measurement error are done with the following steps:
 
@@
 
-- Errors propagated at each step. 
-- Per-sample error agrees well with variability of replicates.
+Errors from normalization and blank correction are propagated at each step.
+The per-sample error agrees well with variability of replicates, and measurement of known-value standards shows that the method has no significant bias.
 
-Data reduction done using an R package maintained on github:
-
-https://github.com/blongworth/hgis
+Data reduction is done using an R package maintained on github (https://github.com/blongworth/hgis).
 
 # Results
-
-Data used for current, blanks, consensus
 
@@
 ```
-## Initial measurements
 
@@
 
-compare to first tests
-
-Note that our flat-topped current curve allows variable flow of the vial system. MICADAS has a more pronounced peak so this may not work as well. (Check papers for peak shape)
-
-Also note current dependent ratio (blank) as a complicating factor
-
-Lead into blank on next slide
-
 ### Blanks and blank correction
-
-#### Current dependent blank
 
@@
 This current dependent blank follows the constant contamination model and can be represented with a current balance between the sample and the constant blank.
-
-#### Current independent blank
-
-If **standards**, **blanks**, and **unknowns** have the **same** blank and current, current drops out.
 
@@
 Data from the measurement of consensus standards over a range of fraction modern shows that this simple correction is effective for removing blank contribution without bias.
-
-#### Mean Fm of blanks: `r cons_df %>% filter(fm_consensus < 0.1 & he12C > 2) %>% pull(norm_ratio) %>% mean() %>% format(digits = 3)`
+Our mean measured fraction modern of C-1 and TIRI-F carbonate blanks is `r cons_df %>% filter(fm_consensus < 0.1 & he12C > 2) %>% pull(norm_ratio) %>% mean() %>% format(digits = 3)`.
 
@@
 ```
-
-## Results: Consensus standards
 
@@
 
-## Results: Method precision and accuracy
-
 ```{r}
@@
 
-## Results: By sample type
+## Marine Sediment Core Carbonates
 
-```{r}
-ggplot(cons_df, aes(x = fm_diff, y = fm_corr, color = he12C)) +
-  geom_hline(aes(yintercept = fm_consensus), color = "lightblue") +
-  geom_hline(data = cons_df_sum, aes(yintercept = fm_corr_mean), color = "darkgray") +
-  geom_hline(data = cons_df_sum, aes(yintercept = fm_corr_mean + fm_corr_sd), 
-             color = "grey", linetype = "dashed") +
-  geom_hline(data = cons_df_sum, aes(yintercept = fm_corr_mean - fm_corr_sd),
-             color = "grey", linetype = "dashed") +
-
-  geom_pointrange(aes(ymin = fm_corr - sig_fm_corr, 
-                      ymax = fm_corr + sig_fm_corr), 
-                  size = .2) +
-  labs(subtitle = "Reference and normalization standards",
-       y = "Fraction modern (Fm)") +
-  facet_wrap(facets = vars(Name), scales = "free") +
-  theme_bw() +
-  scale_x_continuous(breaks = NULL) +
-  xlab(NULL) +
-  scale_color_continuous(name = "12C- (μA)") +
-  theme(
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank()
-  )
-```
-
-## Pendleton Haiti Core Carbonates
-
-### Read data, pull graphite results and sample depth, combine
+To further test the method and a potential application, we measured 15 shell and coral macrofossils from a subtropical coastal marine sediment core using our gas method compared to the same samples measured using the traditional graphite method.
+We found the measurements to be in good agreement, with no significant bias and errors averaging XXX compared with XXX for graphite measurements.
 
@@
 
-### Outlier
-
-```{r}
-combined_results %>% 
-  filter(abs(fm_diff) > 0.02) %>% 
-  select(sample_name, method, fm_corr, sig_fm_corr, fm_diff)
-```
-
-Sample 6:7_39-40 had an issue during prep.
-Vial was found to be at ~1 ATM during backfilling with helium.
-This means the vial had a leaky seal and would have entrained atmosphere (Fm ~0.9) during prep.
-This is the likely cause of a HGIS Fm 5.2% higher than graphite.
-
-## Difference plot
-
 ```{r, eval = FALSE}
@@
 combined_results %>% 
-  filter(abs(fm_diff) < 0.02) %>% 
+  filter(abs(fm_diff_mean) < 0.02) %>% 
 ggplot() +
@@
 
-```{r}
+```{r, eval = FALSE}
 # No outliers, vs mean_fm
 combined_results %>% 
-  filter(abs(fm_diff) < 0.02) %>% 
+  filter(abs(fm_diff_mean) < 0.02) %>% 
 ggplot() +
@@
 ```
+
+### Outlier
+
+Sample 6:7_39-40 had an issue during prep.
+Vial was found to be at ~1 ATM during backfilling with helium.
+This means the vial had a leaky seal and would have entrained atmosphere (Fm ~0.9) during prep.
+This is the likely cause of a HGIS Fm 5.2% higher than graphite.
+This sample was dropped from the comparison.
+Not sure whether to mention this on poster.
+
+```{r}
+combined_results %>% 
+  filter(abs(fm_diff) > 0.02) %>% 
+  select(sample_name, method, fm_corr, sig_fm_corr, fm_diff)
+```

### 2021-10-15 doc/AMS-15/AMS-15_poster.Rmd
Improve fig captions. Text edits. Remove extra plots.

@@
 
-![HGIS Photo](images/gas_control.jpg)
+![Experimental setup. Helium and standard gas regulators at left, sample vial and open split at top right.](images/gas_control.jpg)
 
@@
 
-![HGIS Figure](images/HGIS_open_split_drawing.svg)
+![Inlet layout.
+Helium displacement gas is delivered to vial through a double needle, mixed sample and helium flows to an open split, and a glass capillary delivers gas to the ion source.](images/HGIS_open_split_drawing.svg)
 
@@
 
-1. 10-30mg Carbonate in septa-sealed 7mL vials
-  - produces 2-6mL CO<sub>2</sub> @ ~1ATM
+1. Weigh 10-30mg Carbonate in septa-sealed 7mL vials produces 2-6mL CO<sub>2</sub> @ ~1ATM
 1. Evacuate and Acidify with 1 mL H3PO4 using automated Gilson prep bench
-1. Add helium to 1 ATM
+1. After reaction is complete, add helium to bring vials to ambient pressure.
 
-![HGIS Photo](images/gilson.png)
+![A modified liquid handling system is used to evacuate and add acid to sample vials.](images/gilson.png)
 
@@
 
-```{r cur-flow-os, fig.height=6, dev='svg'}
+```{r cur-flow-os, fig.height=4, dev='svg'}
 ggplot(data, aes(co2flow, he12C)) +
@@
 ## Results: Method precision and accuracy
+
+Replicate measurements of known-value carbonate reference materials were used to establish the precision and accuracy of the method.
 
@@
        y = "Fm difference (cons - measured)")
-```
-
-```{r}
-cons_df_sum <- cons_df %>% 
-  group_by(Name, fm_consensus) %>% 
-  summarize(across(c(fm_corr, sig_fm_corr, fm_diff, sigma),
-                   list(mean = mean, sd = sd)),
-            N = n()) %>% 
-  ungroup()
-
-cons_df_sum %>% 
-  arrange(fm_consensus) %>% 
-  select(Name, `Consensus Fm` = fm_consensus,
-         `Mean Fm` = fm_corr_mean,
-         `Mean Fm error` = sig_fm_corr_mean,
-         `SD of measurements` = fm_corr_sd,
-         N) %>% 
-gt() %>% 
-  tab_header(title = "Summary statistics by sample type") %>% 
-  fmt_number(3:5, decimals = 4)
-```
-
-```{r}
-cons_df %>% 
-  filter(Name %in% c("TIRI-I", "C-2", "NOSAMS2")) %>% 
-  ungroup() %>% 
-  summarize(across(c(sig_fm_corr, fm_diff, sigma),
-                   list(mean = mean, sd = sd)),
-            N = n()) %>% 
-  select(`Mean Fm difference` = fm_diff_mean,
-         `SD of mean difference` = fm_diff_sd,
-         `Mean Fm error` = sig_fm_corr_mean,
-         `Mean Sigma` = sigma_mean,
-         `SD of Sigma` = sigma_sd,
-         N
-         ) %>%  
-  gt() %>% 
-  tab_header(title = "Summary statisics for reference materials",
-             subtitle = "TIRI-I, C-2, and NOSAMS2") %>% 
-  fmt_number(3:5, decimals = 4)
 ```
@@
 
-```{r, eval=FALSE}
-# Highlight outliers
-ggplot(combined_results) +
-  geom_hline(yintercept = 0) +
-  geom_pointrange(aes(fm_corr, fm_diff, color = method, shape = method,
-                      ymin = fm_diff - sig_fm_corr, 
-                      ymax = fm_diff + sig_fm_corr),
-                  size = 1) +
-  gghighlight(abs(fm_diff) > 0.02,
-              #unhighlighted_params = list(colour = NULL),
-              keep_scales = TRUE) +
-  labs(title = "Haiti Core Graphite vs HGIS",
-       x = "Fraction modern",
-       y = "Fm - mean Fm")
-```
-
-```{r eval=FALSE}
-#All in one.
-
-outlier_df <- combined_results %>% 
-  filter(abs(fm_diff) > 0.02)
-
-ggplot(combined_results) +
-  geom_hline(yintercept = 0) +
-  geom_pointrange(aes(fm_corr, fm_diff, color = method, shape = method,
-                      ymin = fm_diff - sig_fm_corr, 
-                      ymax = fm_diff + sig_fm_corr),
-                  size = 1,
-                  color = "grey") +#,
-  gghighlight(abs(fm_diff) > 0.02,
-              unhighlighted_params = list(colour = NULL),
-              keep_scales = TRUE) +
-  labs(title = "Haiti Core Graphite vs HGIS",
-       x = "Fraction modern",
-       y = "Fm - mean Fm")
-```
-
 ## Fm depth plot
@@
 ```{r}
-
 cal_sub <- combined_results %>% 

### 2021-10-18 doc/AMS-15/AMS-15_poster.Rmd
Improve figure captions and formatting. Add and rearrange text. Remove extra figures.

@@
 
-# Results
+## Ion currents and method blank
 
@@
 
-### Flow vs current
+### Ion current vs CO2 flow
 
@@
   geom_point(size = 3, color = "#0069b1") +
-  labs(title = "Vial dilution current",
-       subtitle = "250μL/min displacement, 30μl/min delivery",
+  labs(title = "Ion current is stable for a range of CO2 flow",
+       subtitle = "Current vs. CO2 flow during vial dillution",
        x = bquote('CO'[2]~'Flow (μl/min)'), 
@@
 
-```{r blanks, fig.height=5, dev='svg'}
+```{r blanks, fig.height=4, dev='svg'}
 cons_df %>% 
@@
   #scale_color_manual("wheel", values = c("#b7bf10", "#00a9e0", "#0069b1")) +
-  labs(title = "Blanks",
+  labs(title = "Contaminant carbon is related to current",
+       subtitle = "Fm and current of radiocarbon-free carbonates",
        x = "12C Current (μA)",
@@
 
-## Results: Method precision and accuracy
+## Consensus standards
 
@@
                   size = 0.2) + 
-  labs(subtitle = "Blank corrected", x = "Fm expected (points jittered for clarity)", 
-       y = "Fm difference (cons - measured)")
+  labs(title = "Agreement of HGIS with graphite measurements",
+       subtitle = "Blank corrected HGIS measurements of consensus carbonate standards",
+       x = "Fm expected (points jittered for clarity)", 
+       y = "Fm difference (HGIS - expected)")
 ```
@@
 ## Marine Sediment Core Carbonates
-
-To further test the method and a potential application, we measured 15 shell and coral macrofossils from a subtropical coastal marine sediment core using our gas method compared to the same samples measured using the traditional graphite method.
-We found the measurements to be in good agreement, with no significant bias and errors averaging XXX compared with XXX for graphite measurements.
 
@@
          sig_rc_age = rc_age - -8033 * log(fm_corr + sig_fm_corr))
+
+# No outliers
+cr_no <- combined_results %>% 
+  filter(abs(fm_diff_mean) < 0.02)
+
+mean_diff <- cr_no %>% 
+  filter(method == "hgis") %>% 
+  summarise(across(fm_diff, list(mean = mean, sd = sd)))
+
+mean_errs <- cr_no %>% 
+  group_by(method) %>% 
+  summarise(mean_sig = mean(sig_fm_corr)) %>% 
+  pivot_wider(names_from = method,
+              values_from = mean_sig)
 ```
 
-```{r, eval = FALSE}
-ggplot(combined_results) +
-  geom_hline(yintercept = 0) +
-  geom_pointrange(aes(fm_corr, fm_diff, color = method, shape = method,
-                      ymin = fm_diff - sig_fm_corr, 
-                      ymax = fm_diff + sig_fm_corr),
-                  size = 1) +
-  scale_color_manual(values = c("#00b7bd", "#b7bf10")) +
-  labs(title = "Haiti Core Graphite vs HGIS",
-       x = "Fraction modern",
-       y = "Fm - mean Fm")
-```
+To further test the method and a potential application, we measured 15 shell and coral macrofossils from a subtropical coastal marine sediment core using our gas method compared to the same samples measured using the traditional graphite method.
+We found the measurements to be in good agreement, with no significant bias and errors averaging Fm `r mean_errs$hgis` compared with Fm `r mean_errs$graphite` for graphite measurements.
 
 ```{r}
-# No outliers, vs depth
-combined_results %>% 
-  filter(abs(fm_diff_mean) < 0.02) %>% 
-ggplot() +
+ggplot(cr_no) +
   geom_hline(yintercept = 0) +
+  #geom_hline(yintercept = mean_diff$fm_diff_mean, color = "blue") +
+  #geom_hline(yintercept = mean_diff$fm_diff_mean + mean_diff$fm_diff_sd, color = "lightblue") +
+  #geom_hline(yintercept = mean_diff$fm_diff_mean - mean_diff$fm_diff_sd, color = "lightblue") +
   geom_pointrange(aes(depth, fm_diff, color = method, shape = method,
@@
   scale_color_manual(values = c("#00b7bd", "#b7bf10")) +
-  labs(title = "Haiti Core Graphite vs HGIS",
+  labs(title = "HGIS measurements agree with graphite",
+       subtitle = "Sediment core macrofossils measured via HGIS and graphite",
        x = "Core depth (cm)",
-       y = "Fm - mean Fm")
-```
-
-```{r, eval = FALSE}
-# No outliers, vs mean_fm
-combined_results %>% 
-  filter(abs(fm_diff_mean) < 0.02) %>% 
-ggplot() +
-  geom_hline(yintercept = 0) +
-  geom_pointrange(aes(fm_mean, fm_diff, color = method, shape = method,
-                      ymin = fm_diff - sig_fm_corr, 
-                      ymax = fm_diff + sig_fm_corr),
-                  position = "jitter") +
-  geom_smooth(method = "lm", aes(fm_mean, fm_diff, color = method)) +
-  scale_color_manual(values = c("#00b7bd", "#b7bf10")) +
-  labs(title = "Haiti Core Graphite vs HGIS",
-       x = "Mean Fm",
-       y = "Fm - mean Fm")
+       y = "HGIS Fm - graphite Fm") +
+  theme(legend.position = c(0.80, 0.85),
+        legend.direction = "horizontal",
+        legend.background = element_rect(fill = "white", color = "black")) 
 ```
@@
   filter(sample_name != "6:7_39-40") %>% 
-ggplot(aes(depth, rc_age, color = method)) +
+ggplot(aes(depth, rc_age, color = method, shape = method)) +
   geom_pointrange(aes(ymin = rc_age - sig_rc_age, 
@@
   scale_x_reverse() +
-  labs(title = "Age vs depth for graphite and HGIS",
+  scale_color_manual(values = c("#00b7bd", "#b7bf10")) +
+  labs(title = "HGIS age-depth is similar to graphite",
+       subtitle = "Age vs. depth for graphite and HGIS",
        y = "Radiocarbon years (BP)",
-       x = "Core depth (cm)")
+       x = "Core depth (cm)") +
+  theme(legend.position = c(0.85, 0.30),
+        legend.background = element_rect(fill = "white", color = "black")) 
 ```
@@
 
-```{r}
-plot(cal_dates_gr)
-```
+```{r, fig.height=5}
+bc_gr <- plot(cal_dates_gr) +
+  xlim(8000, 0) +
+  ylim(1000, 0) +
+  labs(title = "Graphite measurements",
+       xlab = "Age (yBP)",
+       ylab = "Core depth (cm)")
 
-```{r}
-plot(cal_dates_gas)
+bc_gas <- plot(cal_dates_gas) +
+  xlim(8000, 0) +
+  ylim(1000, 0) +
+  labs(title = "HGIS measurements",
+       xlab = "Age (yBP)",
+       ylab = "Core depth (cm)")
+
+bchron <- bc_gr / bc_gas
+
+bchron +
+  plot_annotation(title = "HGIS and graphite age models agree",
+                  subtitle = "Calibrated ages vs core depth")
 ```

### 2021-10-20 doc/AMS-15/AMS-15_poster.Rmd
scale to pMC, use dark color palette.

@@
 library(patchwork)
+library(scales)
 
@@
                   position = position_dodge2(width = 0.1), 
-                  size = 0.2) + 
+                  size = 0.5) + 
+  scale_color_brewer(palette = "Dark2") +
+  scale_x_continuous(labels = label_percent(suffix = "")) +
+  scale_y_continuous(breaks = breaks_extended(7),
+                     labels = label_percent(suffix = "",
+                                            accuracy = 1)) +
   labs(title = "Agreement of HGIS with graphite measurements",
        subtitle = "Blank corrected HGIS measurements of consensus carbonate standards",
-       x = "Fm expected (points jittered for clarity)", 
-       y = "Fm difference (HGIS - expected)")
+       x = "pMC expected (points jittered for clarity)", 
+       y = "pMC difference (HGIS - expected)")
 ```

### 2021-10-21 doc/AMS-15/AMS-15_poster.Rmd
improve figure sizes, colors.

--- 2021-10-18 doc/AMS-15/AMS-15_poster.Rmd
@@
 
-We have developed a method for measurement of radiocarbon in carbonate minerals
-as CO~2~ gas via a NEC MCSNICS hybrid sputter gas ion source (HGIS). The method
-uses helium as a carrier gas to displace CO~2~ from sample vials to an open
-split, where a glass capillary samples the mixture for delivery directly to the
-HGIS. This method skips the gas transfer and quantification steps used in a
-closed inlet HGIS system, simplifying sample measurement. Samples larger than 10
-mg carbonate can currently be measured. Results from measurements of consensus
-standards (TIRI-I, IAEA C2, and an internal modern shell standard), and samples
-from a marine core (40-115 pMC) show that the method agrees well with
-traditional AMS measurement of the same samples as graphite, and that the method
-error is less than 1.5 pMC. We discuss advantages and disadvantages of
-continuous flow sample introduction, and the effect of reduced precision on
-calibrated age-depth models produced using gas-source data.
+We have developed a method for measurement of radiocarbon in carbonate minerals as CO~2~ gas via a NEC MCSNICS hybrid sputter gas ion source (HGIS).
+The method uses helium as a carrier gas to displace CO~2~ from sample vials to an open split, where a glass capillary samples the mixture for delivery directly to the HGIS.
+This method skips the gas transfer and quantification steps used in a closed inlet HGIS system, simplifying sample measurement.
+Samples larger than 10 mg carbonate can currently be measured.
+Results from measurements of consensus standards (TIRI-I, IAEA C2, and an internal modern shell standard), and samples from a marine core (40-115 pMC) show that the method agrees well with traditional AMS measurement of the same samples as graphite, and that the method error is less than 1.5 pMC.
+We discuss advantages and disadvantages of continuous flow sample introduction, and the effect of reduced precision on calibrated age-depth models produced using gas-source data.
 
@@
 
-```{r cur-flow-os, fig.height=4, dev='svg'}
+```{r cur-flow-os}
 ggplot(data, aes(co2flow, he12C)) +
   geom_smooth(span = .3, se=FALSE, color = "#00b7bd") +
-  geom_point(size = 3, color = "#0069b1") +
+  geom_point(size = 2, color = "#0069b1") +
   labs(title = "Ion current is stable for a range of CO2 flow",
@@
 
-```{r blanks, fig.height=4, dev='svg'}
+```{r blanks}
 cons_df %>% 
@@
   geom_errorbarh(aes(xmin = he12C - sig_he12C, xmax = he12C  + sig_he12C)) +
-  geom_point(aes(color = wheel), size = 3) +
+  geom_point(size = 2, color = "#0069b1") + #"#b7bf10"
   #scale_color_manual("wheel", values = c("#b7bf10", "#00a9e0", "#0069b1")) +
@@
 ```
+
+<br><br>
 
@@
   labs(title = "Graphite measurements",
-       xlab = "Age (yBP)",
-       ylab = "Core depth (cm)")
+       x = "Age (yBP)",
+       y = "Core depth (cm)")
 
@@
   labs(title = "HGIS measurements",
-       xlab = "Age (yBP)",
-       ylab = "Core depth (cm)")
+       x = "Age (yBP)",
+       y = "Core depth (cm)")
 

### 2021-10-21 doc/AMS-15/AMS-15_poster.Rmd
Finish pMC conversion. Clean up tables and figures.

@@
   geom_point(size = 2, color = "#0069b1") +
-  labs(title = "Ion current is stable for a range of CO2 flow",
-       subtitle = "Current vs. CO2 flow during vial dillution",
+  labs(title = bquote('Ion current is stable for a range of CO'[2]~'flow'),
+       subtitle = bquote('Current vs. CO'[2]~'flow during vial dillution'),
        x = bquote('CO'[2]~'Flow (μl/min)'), 
@@
   #scale_color_manual("wheel", values = c("#b7bf10", "#00a9e0", "#0069b1")) +
+  scale_y_continuous(breaks = breaks_extended(7),
+                     labels = label_percent(suffix = "",
+                                            accuracy = 1)) +
   labs(title = "Contaminant carbon is related to current",
@@
        x = "12C Current (μA)",
-       y = "Fraction modern") +
+       y = "pMC") +
   theme_classic() +
@@
   arrange(fm_consensus) %>% 
-  select(Name, `Consensus Fm` = fm_consensus,
-         `Mean Fm` = fm_corr_mean,
-         `Mean Fm error` = sig_fm_corr_mean,
-         `SD of measurements` = fm_corr_sd,
+  select(Name, `Consensus pMC` = fm_consensus,
+         `Mean pMC` = fm_corr_mean,
+         `Mean pMC error` = sig_fm_corr_mean,
+         `Std. dev. of measurements` = fm_corr_sd,
          N) %>% 
@@
   tab_header(title = "Summary statistics by sample type") %>% 
-  fmt_number(3:5, decimals = 4)
+  fmt_number(2:5, 
+             scale_by = 100,
+             decimals = 2)
 ```
@@
   ungroup() %>% 
-  summarize(across(c(sig_fm_corr, fm_diff, sigma),
+  summarize(across(c(sig_fm_corr, fm_diff),
                    list(mean = mean, sd = sd)),
             N = n()) %>% 
-  select(`Mean Fm difference` = fm_diff_mean,
-         `SD of mean difference` = fm_diff_sd,
-         `Mean Fm error` = sig_fm_corr_mean,
-         `Mean Sigma` = sigma_mean,
-         `SD of Sigma` = sigma_sd,
+  select(`Consensus difference` = fm_diff_mean,
+         `Std. Dev. of consensus difference` = fm_diff_sd,
+         `per-sample error` = sig_fm_corr_mean,
          N
@@
   tab_header(title = "Summary statisics for reference materials",
-             subtitle = "TIRI-I, C-2, and NOSAMS2") %>% 
-  fmt_number(3:5, decimals = 4)
+             subtitle = "TIRI-I, C-2, and NOSAMS2. Values in pMC") %>% 
+  fmt_number(1:3, 
+             scale_by = 100,
+             decimals = 2)
 ```
@@
   scale_color_manual(values = c("#00b7bd", "#b7bf10")) +
+  scale_y_continuous(breaks = breaks_extended(7),
+                     labels = label_percent(suffix = "",
+                                            accuracy = 1)) +
   labs(title = "HGIS measurements agree with graphite",
@@
        x = "Core depth (cm)",
-       y = "HGIS Fm - graphite Fm") +
+       y = "HGIS - graphite (pMC)") +
   theme(legend.position = c(0.80, 0.85),

### 2021-10-21 doc/AMS-15/AMS-15_poster.Rmd
Fix superscripts in plot axis labels

@@
 library(scales)
+library(latex2exp)
 
@@
        x = bquote('CO'[2]~'Flow (μl/min)'), 
-       y = "12C current (μA)") +
+       y = TeX(r'(^{12}C^- current (μA))')) +
   theme_classic()
@@
        subtitle = "Fm and current of radiocarbon-free carbonates",
-       x = "12C Current (μA)",
+       x = TeX(r'(^{12}C^- current (μA))'),
        y = "pMC") +

### 2021-10-22 doc/AMS-15/AMS-15_poster.Rmd
Improve table titles

@@
 gt() %>% 
-  tab_header(title = "Summary statistics by sample type") %>% 
+  tab_header(title = "Reference material precision and accuracy") %>% 
   fmt_number(2:5, 
@@
   gt() %>% 
-  tab_header(title = "Summary statisics for reference materials",
+  tab_header(title = "Reference material performance summary",
              subtitle = "TIRI-I, C-2, and NOSAMS2. Values in pMC") %>% 

### 2021-10-27 doc/AMS-15/AMS-15_poster.Rmd
Add errors to tables

@@
 library(latex2exp)
+library(errors)
 
@@
 options(decimals = 4)
+options(errors.notation = "plus-minus")
 theme_set(theme_classic())
@@
   arrange(fm_consensus) %>% 
-  select(Name, `Consensus pMC` = fm_consensus,
-         `Mean pMC` = fm_corr_mean,
-         `Mean pMC error` = sig_fm_corr_mean,
-         `Std. dev. of measurements` = fm_corr_sd,
+  mutate(across(c(fm_consensus, fm_corr_mean, fm_corr_sd, sig_fm_corr_mean), 
+         function(x) x * 100),
+         fm_corr_err = set_errors(fm_corr_mean, fm_corr_sd)) %>% 
+  select(Name, `Consensus` = fm_consensus,
+         `HGIS ± SD` = fm_corr_err,
+         #`HGIS` = fm_corr_mean,
+         #`Std. dev. of measurements` = fm_corr_sd,
+         `Mean per-sample error` = sig_fm_corr_mean,
          N) %>% 
 gt() %>% 
-  tab_header(title = "Reference material precision and accuracy") %>% 
-  fmt_number(2:5, 
-             scale_by = 100,
-             decimals = 2)
+  tab_header(title = "Reference material precision and accuracy",
+             subtitle = "Values in pMC, Errors are SD of measurements") %>% 
+  fmt_number(4, 
+             decimals = 1) %>% 
+  cols_align("right", 3)
 ```
@@
                   position = position_dodge2(width = 0.1), 
-                  size = 0.5) + 
+                  ) + #size = 0.5) + 
    scale_color_brewer(palette = "Dark2") +
+  # scale_color_manual(values = c("#00a9e0", "#0069b1", "#00b7bd", "#b7bf10", "#00b7bd")) +
   scale_x_continuous(labels = label_percent(suffix = "")) +

### 2021-12-28 doc/AMS-15/AMS-15_paper.Rmd
add method paper outline

---
title: Analysis of Carbonates as CO2 using an open split interface to a hybrid sputter
  gas ion source
author: "Brett Longworth"
date: "7/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

NEC hybrid gas ion source

NEC inlets

MICADAS source and inlets

Comparison of open split to closed method

Pros/cons

# Methods

## Inlet description

## Sample preparation

## Data acquisition

## Data reduction

# Results

## System performance

### Currents

### Blanks

### Sample size

## Reference materials

## Haiti Core

# Discussion

Quantification of method

# Conclusion

We developed a method for the analysis of carbonates using an open-split interface to a hybrid sputter gas ion source. Sample requirements, analysis time, precision, accuracy.

### 2021-12-29 doc/AMS-15/AMS-15_paper.Rmd
working on paper

@@
 ---
-title: Analysis of Carbonates as CO2 using an open split interface to a hybrid sputter
-  gas ion source
-author: "Brett Longworth"
-date: "7/12/2021"
-output: html_document
+title: "Radiocarbon measurement of CO~2~ from carbonate minerals using a gas ion source with an open split interface"
+author:
+- Brett Longworth
+- Josh Burton
+- Simon Pendleton
+- Josh Hlavenka
+- Sydney Moser
+- Mark Roberts
+- Mark Kurz
+output:
+  html_document: default
+  word_document: default
+  pdf_document: default
 ---
@@
 
+We have developed a method for measurement of radiocarbon in carbonate minerals
+as CO~2~ gas via a NEC MCSNICS hybrid sputter gas ion source (HGIS). The method
+uses helium as a carrier gas to displace CO~2~ from sample vials to an open
+split, where a glass capillary samples the mixture for delivery directly to the
+HGIS. This method skips the gas transfer and quantification steps used in a
+closed inlet HGIS system, simplifying sample measurement. Samples larger than 10
+mg carbonate can currently be measured. Results from measurements of consensus
+standards (TIRI-I, IAEA C2, and an internal modern shell standard), and samples
+from a marine core (40-115 pMC) show that the method agrees well with
+traditional AMS measurement of the same samples as graphite, and that the method
+error is less than 1.5 pMC. We discuss advantages and disadvantages of
+continuous flow sample introduction, and the effect of reduced precision on
+calibrated age-depth models produced using gas-source data.
+
 # Introduction
+
+The NOSAMS USAMS is an AMS system based on a 3MV Tandetron accelerator.
+A NEC 134-sample gas-capable MCSNICS source.
+Analysis of samples as CO2 simplifies sample preparation, making analysis less expensive and faster.
+
+The gas-capable 134-sample MCSNICS is produced by NEC. Sample CO2 is introduced onto a titanium frit in an aluminum target using an arm with a probe that mates with the back of the target.
 
@@
 
+Most interfaces to gas-accepting sputter ion sources are a closed inlet design.
+A mixture of sample and helium is introduced to the source through a capillary, and flow is controlled by the pressure in the sample volume and the length and diameter of the capillary.
+Flow may be adjusted by changing the pressure by varying the sample volume.
+We designed a simpler open inlet where a capillary samples gas in a flowing stream that exhausts to the atmosphere.
+This defines the delivery pressure as 1 ATM and therefore flow is controlled by the capillary and is not adjustable.
+The flow of CO2 to the source may be varied by controlling the ratio of CO2 to helium in the sample stream.
+An open inlet has the advantage of simplicity; flow of CO2 to the source is preset and there are no separate sample transfer and dilution steps.
+A disadvantage of the open split is that the flow of CO2 to the source varies with sample size (CO2:He) and changes over the course of a measurement due to dillution of the sample by displacement helium.
+Salazar and Szidat, Roberts gasbench, other open split methods.
+
 Pros/cons
 
-# Methods
+# Method
 
@@
 
+We developed a simple open-split interface that delivers a mixture of helium and CO2 to the source at atmospheric pressure.
+
+Figure 1. photo and drawing of inlet.
+
+The open split interface supplies a helium-CO2 mixture to a split where a
+capillary delivers gas to the source and the rest is vented.
+The current split is a VICI Luer adapter on the outflow side of the dual needle (Figure 1).
+The split has fairly low volume.
+Testing is needed to determine minimum sample flow to avoid atmosphere entering the split.
+
+The test setup for extracting CO2 from septa-sealed vials is as follows: a double needle is inserted into the vial through the septa.
+One side of the needle is connected to helium which is used to force CO2 from the vial through the other side of the needle.
+The outflow from the vial enters an open split where a capillary is used to admit sample to the ion source.
+The system is currently set up to run with samples starting at atmospheric pressure.
+
+Currently, 250cm of 50um capillary defines flow rate from the open split into
+the source, with a bulkhead connector at the source cage.
+
 ## Sample preparation
+
+Reference: gilson
+
+Carbonate samples are weighed into 7 ml septa-sealed Vacutainer vials (Labco # 768W).
+Sealed vials are evacuated, then 1 ml 85% H3PO4 is added.
+The samples are allowed to react at 60C for 4-6 h.
+Sample pressure was adjusted to 1 ATM by adding helium or removing sample using a syringe.
+
+Process blank samples were prepared from powdered C-1 or TIRI-F. Standards were prepared on the Gilson prep system by evacuating vials and back filling with 1 ATM of a modern standard gas.
 
@@
 
+A new target was used for each sample.
+Targets were allowed to "burn in" for 5-10 minutes with pure helium delivered to the source.
+Initial currents of > 10 uA dropped rapidly, eventually settling to 100-400 nA.
+After a stable base current was reached, the sample vial was installed on the double needle, starting sample flow to the source.
+Sample current rises to a stable level within 30 s, at which time, data acquisition is started.
+10 blocks of 90 s for a total of 15 m.
+
 ## Data reduction
+
+Normalization and blank correction of measurements
+
+A linear correction using measured values of standard gas and blank samples was used for correction of contaminant carbon and machine blank. We use equation 1 for correction
+
+LBC equation
+
+where x = y.
+
+Constant contaminant discussed below.
+
+Data reduction via R package supplementary materials
 
@@
 
+All results in supplementary materials
+
 ## System performance
+
+Best currents for the source are around 10uA 12C- or roughly 1/10 of graphite
+currents. This is in line with what other hybrid sources produce. Typical "good" currents are between 5-10uA. The source
+typically produces best current at around 5uLs-1 CO2. Literature values give
+1.5uLs-1 as optimal, but we've never seen that, and pressures, capillaries and
+flows in at least one of these papers doesn't make sense.
+
+Source currents are variable by day and by target.
+"Good" currents for a given day vary between 5-10uA, and it's not yet clear what causes this variability.
+Alignment, source condition, and settings of the ionizer and cesium are all factors.
+Current also varies by target for a given set of gas and source conditions.
+Some targets are clearly "bad", producing currents much lower than other targets for a given set of conditions.
+Capillary clogging can be an issue, especially when sampling from vials with acid.
+
+Precision taken as the standard error of the mean of repeated runs of a sample is around 1%.
+Precision is typically a little worse than predicted by counting statistics.
+More work is needed to look at whether this reflects variability between samples, but initial test of paired replicates show similar standard deviation for samples that show good stable currents.
+
+Measured value of tank modern gas normalized to solid graphite standards ranges from agreeing with solid OX-I to 7% lower than expected for low current samples.
+Normalizing to tank CO2 is probably the best method to account for variability in hybrid source performance.
+
+Precision of
+measurements of modern CO2 has been better than 1% (std error of mean of
+measurements of a target) for high current samples, with precision dropping off
+as current decreases below 6uA or so. Accuracy is reasonable within the large
+errors on gas measurements. Uncorrected measurements of dead CO2 range from 1%
+to 5% of modern depending on sample current.
+
+Measurement of a small number of replicates shows that between sample reproducibility is around 1%.
+Samples displaced by helium have slightly higher variability, but more tests are needed to better assess reproducibility.
 
@@
 
+Current vs flow modelling
+
+Figure: Current vs flow
+
 ### Blanks
+
+LBC vs constant contaminant/MBC.
+
+Figure: current dependent blank
+
+Figure: justification for large blank
+
+Data for blanks
+
+Constant current blank is likely the best model, but we simplify to a large blank correction since blanks vary by run, currents are relatively similar between blanks, unknowns and standards, and the simpler model works well within the error of the method.
 
@@
 
+Sample size tests.
+
+Figure: sample size vs current vs precision and accuracy
+
+Aimed to fill vial initially
+
+No current degradation down to X. Used Y as optimal mass and backfilled with helium.
+
+First tests of displacing pure CO2 from a vial into the open split using helium
+are promising. Helium flowing at 100uLm-1 produces current from 7mL of CO2 for
+over 2 hours. Currents do decrease with time due to dilution, but by supplying
+higher than optimal CO2 flow to the source initially, stable currents can be
+maintained for some time. Fraction of CO2 in sample gas should follow this relationship:
+
+$$C(t) = e^{-\frac{r}{V}t}$$
+
+Where $C(t)$ is the fraction CO2, $r$ is the rate of helium displacement flow, and $V$ is the initial volume of the CO2-filled vial.
+
+Setting flow to the source to 10uL/min provides excess CO2 to source until the helium dilution ratio increases to > 1:1.
+Another approach may be to backfill vials with helium and run at a higher dilution ratio and flow to the source.
+This should produce a smaller change in He:CO2 over the course of a run.
+MICADAS runs at 95% He:CO2, which would mean producing < 0.5mL CO2 in a 7mL vial and backfilling with helium before displacing to the source.
+Some intermediate dilution may be appropriate.
+Alternatively, vials could be flushed with helium prior to acidification, rather than evacuating, leading to some concentration of CO2 in He at slightly > 1 ATM.
+
+Testing long runs of dilution of CO2 with displacing He and diluting He with CO2 show that the setup is still optimized for low flow to the source, even with larger diameter delivery capillaries.
+There appears to be a constriction in the split-source tubing that is keeping flows low.
+Current focus is on the large-sample 7mL method, so this will be diagnosed as time allows.
+
+Targets have shown "pulses" of higher current.
+Not sure why this is, but I suspect it's either buildup of graphite or other available carbon on the target, or target switching to the "high state" as sometimes happens with solid targets.
+
+First tests with carbonate in vials were hit-or-miss.
+Problems with capillary clogging and low currents led to poor results for several samples.
+A pair of NOSAMS2 samples produced good currents, results agreed well, but were about 3% higher than expected Fm.
+A second set of carbonates produced better currents (~5uA) and better results.
+Agreement of replicates varied from 3 to 16 permil.
+Near-modern NOSAMS2 was very close to the consensus value, while C-2 and C-1 were elevated by 2 and 3.6% respectively, indicating that a blank correction is needed.
+
+Two more recent carbonate tests produced stable currents and backgrounds.
+Since sample, target blank, and sample blank currents were relatively constant across samples, a large blank correction can be used to correct results.
+Samples from both runs showed a linear relationship between difference from expected Fm and Fm of the sample, but in one sample set the relationship was positive, and in the other, negative.
+Since the intercept with 0-offset was close to 0 Fm, I think this is a problem with normalization, rather than blank correction.
+Standards run at the end of the day are higher than those run at the beginning, which may give clues as to what's going on.
+
 ## Reference materials
 
+Measured
+
+figure: difference plot
+
+Assessment of data quality.
+
 ## Haiti Core
+
+### Measurements
+
+Outlier
+
+Figure: bland-altman of HGIS vs graphite.
+
+Summary of agreement
+
+### Age model
+
+Figure: age model comparison graphite HGIS
+
+Age model quality changes with lower precision.
 
@@
 
+Cost vs quality discussion
+
+ideas for further development of method
+
 # Conclusion

### 2021-12-30 doc/AMS-15/AMS-15_paper.Rmd
Add more text

@@
 Reference: gilson
+\@ref(Roberts 2013)
 
@@
 
-## Haiti Core
+## Carbonate macrofossils for age-depth model
+
+A suite of carbonate macrofossils consisting of shell and coral fragements from a series of coastal marine cores REF were measured as graphite and as CO2 using the HGIS method.
 
@@
 
-Outlier
+Table of measurements (supplementary?)
+
+One HGIS sample vial was noted to be at atmospheric pressure when only XX mL CO2 should have been produced.
+The HGIS measurement for this sample was XX% higher than the measurement as graphite.
+We suspect that this sample was improperly sealed and eliminated it from the analysis as an outlier.
+SO2 story?
 
@@
 
-We developed a method for the analysis of carbonates using an open-split interface to a hybrid sputter gas ion source. Sample requirements, analysis time, precision, accuracy.
+We developed a method for the analysis of carbonates using an open-split interface to a hybrid sputter gas ion source.
+This system produces results that agree with graphite measurements to better than 1% from XXX mg of carbonate.
+Sample preparation takes about 10 min.
+of person time and 6h of automated processing and reaction time, and 20 min.
+of measurement time per sample.
+The simplicity of this approach...
+With lower cost and time per sample, the lower precision of the method may be balanced by the ability to measure more samples from a population, ie.
+triplicate measurement or more core depths.
+This may lead to more accurate results compared to a more precise graphite measurement where reproducibility is not as well constrained.
+
+Sample requirements, analysis time, precision, accuracy.
+
+# Acknowledgements
+
+This work is supported by NSF cooperative agreement. We are grateful to C. McIntyre, R. Hansman, G. Salazar, T. Ognibene, T. Broeck.

### 2022-01-10 doc/method_validation/method_validation.Rmd
Update for amsdata

@@
 library(janitor)
+library(amsdata)
 ```
@@
   mutate(corr_14_12_err = set_errors(corr_14_12, max_err),
-         norm_ratio_err = norm_gas(corr_14_12_err, ratio_std, fm_std))
+         norm_ratio_err = norm_run(corr_14_12_err, ratio_std, fm_std))
 

### 2022-01-11 doc/AMS-15/AMS-15_paper.Rmd
Add figures, more text, some things from poster.

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = FALSE)
 ```
 
-We have developed a method for measurement of radiocarbon in carbonate minerals
-as CO~2~ gas via a NEC MCSNICS hybrid sputter gas ion source (HGIS). The method
-uses helium as a carrier gas to displace CO~2~ from sample vials to an open
-split, where a glass capillary samples the mixture for delivery directly to the
-HGIS. This method skips the gas transfer and quantification steps used in a
-closed inlet HGIS system, simplifying sample measurement. Samples larger than 10
-mg carbonate can currently be measured. Results from measurements of consensus
-standards (TIRI-I, IAEA C2, and an internal modern shell standard), and samples
-from a marine core (40-115 pMC) show that the method agrees well with
-traditional AMS measurement of the same samples as graphite, and that the method
-error is less than 1.5 pMC. We discuss advantages and disadvantages of
-continuous flow sample introduction, and the effect of reduced precision on
-calibrated age-depth models produced using gas-source data.
+# Abstract
+
+We have developed a method for measurement of radiocarbon in carbonate minerals as CO~2~ gas via a NEC MCSNICS hybrid sputter gas ion source (HGIS).
+The method uses helium as a carrier gas to displace CO~2~ from sample vials to an open split, where a glass capillary samples the mixture for delivery directly to the HGIS.
+This method skips the gas transfer and quantification steps used in a closed inlet HGIS system, simplifying sample measurement.
+Samples larger than 10 mg carbonate can currently be measured.
+Results from measurements of consensus standards (TIRI-I, IAEA C2, and an internal modern shell standard), and samples from a marine core (40-115 pMC) show that the method agrees well with traditional AMS measurement of the same samples as graphite, and that the method error is less than 1.5 pMC.
+We discuss advantages and disadvantages of continuous flow sample introduction, and the effect of reduced precision on calibrated age-depth models produced using gas-source data.
 
@@
 
+Radiocarbon Accelerator Mass Spectrometry (AMS) is a sensitive method for determining the age, provenance, and composition of carbon-containing materials.
+A primary disadvantage of the method is its cost, which often limits the number of samples analyzed.
+This forces researchers to carefully choose samples to be run and limits repeat measurements for error determination.
+One driver of the cost of AMS analysis is the time and effort required for sample preparation, including reduction of the sample to graphite.
+Analysis of analyte CO2 directly via a gas-accepting AMS ion source eliminates the cost and effort of graphitization and returns results more rapidly.
 The NOSAMS USAMS is an AMS system based on a 3MV Tandetron accelerator.
@@
 An open inlet has the advantage of simplicity; flow of CO2 to the source is preset and there are no separate sample transfer and dilution steps.
-A disadvantage of the open split is that the flow of CO2 to the source varies with sample size (CO2:He) and changes over the course of a measurement due to dillution of the sample by displacement helium.
+A disadvantage of the open split is that the flow of CO2 to the source varies with sample size (CO2:He) and changes over the course of a measurement due to dilution of the sample by displacement helium.
 Salazar and Szidat, Roberts gasbench, other open split methods.
@@
 
-Figure 1. photo and drawing of inlet.
-
-The open split interface supplies a helium-CO2 mixture to a split where a
-capillary delivers gas to the source and the rest is vented.
+The open split interface supplies a helium-CO2 mixture to a split where a capillary delivers gas to the source and the rest is vented.
 The current split is a VICI Luer adapter on the outflow side of the dual needle (Figure 1).
@@
 
-Currently, 250cm of 50um capillary defines flow rate from the open split into
-the source, with a bulkhead connector at the source cage.
+Currently, 250cm of 50um capillary defines flow rate from the open split into the source, with a bulkhead connector at the source cage.
+
+![Experimental setup. Helium and standard gas regulators at left, sample vial and open split at top right.](images/gas_control.jpg)
+
+Our method uses an open split and glass capillary to control gas flow to the ion source.
+Compared to closed inlet systems which regulate introduction of a CO2/Helium mixture to the ion source by regulating sample pressure, our method is simpler and faster, but does not maintain a constant flow of CO2 to the source.
+We compensate for this by limiting CO2 flow to a range that produces relatively constant C- ion current and stable ratios.
+
+![Inlet layout.
+Helium displacement gas is delivered to vial through a double needle, mixed sample and helium flows to an open split, and a glass capillary delivers gas to the ion source.](images/HGIS_open_split_drawing.svg)
 
@@
 
-Reference: gilson
-\@ref(Roberts 2013)
+Reference: gilson \@ref(Roberts 2013)
 
-Carbonate samples are weighed into 7 ml septa-sealed Vacutainer vials (Labco # 768W).
+Carbonate samples are weighed into 7 ml septa-sealed Vacutainer vials (Labco \# 768W).
 Sealed vials are evacuated, then 1 ml 85% H3PO4 is added.
@@
 
+1.  Weigh 10-30mg Carbonate in septa-sealed 7mL vials produces 2-6mL CO<sub>2</sub> \@ \~1ATM
+2.  Evacuate and Acidify with 1 mL H3PO4 using automated Gilson prep bench
+3.  After reaction is complete, add helium to bring vials to ambient pressure.
+
+![A modified liquid handling system is used to evacuate and add acid to sample vials.](images/gilson.png)
+
+### Measurement
+
+1.  Displace gas to open split with 100 uL/min He.
+2.  Introduce mixture to source \@ 30 uL/min
+3.  Acquire data: 90s x 10 measurements.
+
+-   Normalizing standards: live tank gas in vials
+-   Blanks: C-1 or TIRI-F in vials
+
+### Data reduction
+
+Normalization to standard, blank correction, and estimation of measurement error are done with the following steps:
+
+1.  Correct for fractionation using online AMS <sup>13</sup>C measurement.
+2.  Determine per-sample mean and error (repeatabilty and counting error).
+3.  Normalize to mean of Gas standard.
+4.  Blank correct: Blank subtraction using mean of carbonate blanks
+
+Errors from normalization and blank correction are propagated at each step.
+The per-sample error agrees well with variability of replicates, and measurement of known-value standards shows that the method has no significant bias.
+
+Data reduction is done using an R package maintained on github (<https://github.com/blongworth/hgis>).
+
 ## Data acquisition
@@
 Targets were allowed to "burn in" for 5-10 minutes with pure helium delivered to the source.
-Initial currents of > 10 uA dropped rapidly, eventually settling to 100-400 nA.
+Initial currents of \> 10 uA dropped rapidly, eventually settling to 100-400 nA.
 After a stable base current was reached, the sample vial was installed on the double needle, starting sample flow to the source.
@@
 
-Best currents for the source are around 10uA 12C- or roughly 1/10 of graphite
-currents. This is in line with what other hybrid sources produce. Typical "good" currents are between 5-10uA. The source
-typically produces best current at around 5uLs-1 CO2. Literature values give
-1.5uLs-1 as optimal, but we've never seen that, and pressures, capillaries and
-flows in at least one of these papers doesn't make sense.
+Best currents for the source are around 10uA 12C- or roughly 1/10 of graphite currents.
+This is in line with what other hybrid sources produce.
+Typical "good" currents are between 5-10uA.
+The source typically produces best current at around 5uLs-1 CO2.
+Literature values give 1.5uLs-1 as optimal, but we've never seen that, and pressures, capillaries and flows in at least one of these papers doesn't make sense.
 
@@
 
-Precision of
-measurements of modern CO2 has been better than 1% (std error of mean of
-measurements of a target) for high current samples, with precision dropping off
-as current decreases below 6uA or so. Accuracy is reasonable within the large
-errors on gas measurements. Uncorrected measurements of dead CO2 range from 1%
-to 5% of modern depending on sample current.
+Precision of measurements of modern CO2 has been better than 1% (std error of mean of measurements of a target) for high current samples, with precision dropping off as current decreases below 6uA or so.
+Accuracy is reasonable within the large errors on gas measurements.
+Uncorrected measurements of dead CO2 range from 1% to 5% of modern depending on sample current.
 
@@
 
-First tests of displacing pure CO2 from a vial into the open split using helium
-are promising. Helium flowing at 100uLm-1 produces current from 7mL of CO2 for
-over 2 hours. Currents do decrease with time due to dilution, but by supplying
-higher than optimal CO2 flow to the source initially, stable currents can be
-maintained for some time. Fraction of CO2 in sample gas should follow this relationship:
+First tests of displacing pure CO2 from a vial into the open split using helium are promising.
+Helium flowing at 100uLm-1 produces current from 7mL of CO2 for over 2 hours.
+Currents do decrease with time due to dilution, but by supplying higher than optimal CO2 flow to the source initially, stable currents can be maintained for some time.
+Fraction of CO2 in sample gas should follow this relationship:
 
@@
 
-Setting flow to the source to 10uL/min provides excess CO2 to source until the helium dilution ratio increases to > 1:1.
+Setting flow to the source to 10uL/min provides excess CO2 to source until the helium dilution ratio increases to \> 1:1.
 Another approach may be to backfill vials with helium and run at a higher dilution ratio and flow to the source.
 This should produce a smaller change in He:CO2 over the course of a run.
-MICADAS runs at 95% He:CO2, which would mean producing < 0.5mL CO2 in a 7mL vial and backfilling with helium before displacing to the source.
+MICADAS runs at 95% He:CO2, which would mean producing \< 0.5mL CO2 in a 7mL vial and backfilling with helium before displacing to the source.
 Some intermediate dilution may be appropriate.
-Alternatively, vials could be flushed with helium prior to acidification, rather than evacuating, leading to some concentration of CO2 in He at slightly > 1 ATM.
+Alternatively, vials could be flushed with helium prior to acidification, rather than evacuating, leading to some concentration of CO2 in He at slightly \> 1 ATM.
 
@@
 A pair of NOSAMS2 samples produced good currents, results agreed well, but were about 3% higher than expected Fm.
-A second set of carbonates produced better currents (~5uA) and better results.
+A second set of carbonates produced better currents (\~5uA) and better results.
 Agreement of replicates varied from 3 to 16 permil.

### 2022-01-11 R/analyse_carbonates.R
change file suffix

@@ -32,7 +32,7 @@ results <- pmap(list(files, dates, stds, blanks, outliers), reduce_hgis)
 raw_df <- map_dfr(results, 1)
 results_df <- map_dfr(results, 2)
 
-saveRDS(results, here("data_analysed/carb_all.Rdata"))
+saveRDS(results, here("data_analysed/carb_all.rds"))
 
 write_csv(raw_df, here("data_analysed/carb_all_raw.csv"))
 write_csv(results_df, here("data_analysed/carb_all_results.csv"))

### 2022-01-11 doc/AMS-15/AMS-15_poster.Rmd
Change to newer data munger and change rds file suffix

@@
 ```{r cur-time-os, fig.height=7, dev='svg'}
-data <- get_hgis_data(here("data/USAMS040121R.txt"), as.Date("2021-04-02")) %>% 
-  filter(Pos == 5) %>% 
-  mutate(time = cum_acqtime/60,
+data <- process_hgis_results(here("data/USAMS040121R.txt"), as.Date("2021-04-02")) %>% 
+  filter(pos == 5) %>% 
+  mutate(time = cumsum(corr_lt) / 60,
          co2flow = concCO2(time, r = 244) * 30) #30ul/min
@@
 # Read raw and reduced data as produced by analyse_carbonates.R
-data <- readRDS(here("data_analysed/carb_all.Rdata"))
+data <- readRDS(here("data_analysed/carb_all.rds"))
 

### 2022-01-12 R/smaller_samples.Rmd
Change data source, fix plots

@@
 options(scipen = 4)
+```
 
-df <- read_csv(here("data_analysed/carb_all_results.csv")) %>% 
+```{r}
+# Read reduced data as produced by analyse_carbonates.R
+df <- readRDS(here("data_analysed/carb_all.rds")) %>% 
+  map_dfr(2) %>% 
+#df <- read_csv(here("data_analysed/carb_all_results.csv")) %>% 
   compare_consensus() %>% 
@@
 df <- left_join(df, samples, by = c("wheel", "pos")) %>% 
-  filter(wheel != "USAMS020521",
-         !is.na(mass))
+  filter(wheel != "USAMS020521")
 ```
@@
 ggplot(aes(mass, c12ua, color = Name)) +
-  geom_point() +
-  ggtitle()
+  geom_point()
 ```
@@
 ggplot(aes(c12ua, abs(fm_diff))) +
-  geom_smooth() +
+  geom_smooth(method = "lm") +
   geom_point(aes(color = wheel))
@@
 ggplot(aes(c12ua, abs(fm_diff))) +
+  geom_smooth(method = "lm") +
   geom_point(aes(color = wheel))

### 2022-02-04 doc/AMS-15/AMS-15_paper.Rmd
consolidate introduction

@@
 
-The gas-capable 134-sample MCSNICS is produced by NEC. Sample CO2 is introduced onto a titanium frit in an aluminum target using an arm with a probe that mates with the back of the target.
-
-NEC hybrid gas ion source
-
-NEC inlets
-
-MICADAS source and inlets
-
-Comparison of open split to closed method
+The gas-capable 134-sample MCSNICS from National Electrostatics Corporation (NEC) works on the same principal as hybrid sources developed by ETH (REF) and HVEE (REF).
+Sample CO2 is introduced onto a titanium frit in an aluminum target using stainless tubing mounted to an arm with a probe that mates with the back of the target.
+A CO~2~ - helium mixture is exposed to a beam of charged cesium atoms, which causes dissociation of CO~2~ and ionization of molecular and atomic species.
+Negative ions are extracted from the target and a magnetic filter picks species with masses corresponding to carbon.
 
@@
 Flow may be adjusted by changing the pressure by varying the sample volume.
-We designed a simpler open inlet where a capillary samples gas in a flowing stream that exhausts to the atmosphere.
+
+We designed a open inlet where a capillary samples gas for delivery to the ion source in a flowing stream that exhausts to the atmosphere.
 This defines the delivery pressure as 1 ATM and therefore flow is controlled by the capillary and is not adjustable.
@@
 An open inlet has the advantage of simplicity; flow of CO2 to the source is preset and there are no separate sample transfer and dilution steps.
+This may translate to greater throughput by eliminating the time required for sample transfer, dilution, and setting of delivery pressure.
 A disadvantage of the open split is that the flow of CO2 to the source varies with sample size (CO2:He) and changes over the course of a measurement due to dilution of the sample by displacement helium.
+Our method compensates for this by limiting CO2 flow to a range that produces relatively constant C- ion current and stable ratios.
 Salazar and Szidat, Roberts gasbench, other open split methods.
-
-Pros/cons
 
@@
 The open split interface supplies a helium-CO2 mixture to a split where a capillary delivers gas to the source and the rest is vented.
-The current split is a VICI Luer adapter on the outflow side of the dual needle (Figure 1).
-The split has fairly low volume.
+A flow of helium is used to displace sample CO2 to the delivery capillary in the split.
+The core of the design is a dual needle and split housed in a VICI cross.
+Displacement helium flows in through one side of the cross, into the sample vial through 1/16" stainless steel tubing, which forms the outer flow path of the concentric dual needle.
+The mixture of sample CO2 and displacement helium flows up a needle inserted through the cross and stainless tubing which forms the inner flow path of the dual needle.
+The split is made from a VICI Luer adapter fitted to the outflow of the dual needle with the source delivery capillary inserted(Figure 1).
+The split has fairly low volume (< 1 mL).
 Testing is needed to determine minimum sample flow to avoid atmosphere entering the split.
+TODO: move flow ratio test to results.
 
@@
 
-Currently, 250cm of 50um capillary defines flow rate from the open split into the source, with a bulkhead connector at the source cage.
+250cm of 50um capillary defines flow rate from the open split into the source, with a bulkhead connector at the source cage.
 
 ![Experimental setup. Helium and standard gas regulators at left, sample vial and open split at top right.](images/gas_control.jpg)
-
-Our method uses an open split and glass capillary to control gas flow to the ion source.
-Compared to closed inlet systems which regulate introduction of a CO2/Helium mixture to the ion source by regulating sample pressure, our method is simpler and faster, but does not maintain a constant flow of CO2 to the source.
-We compensate for this by limiting CO2 flow to a range that produces relatively constant C- ion current and stable ratios.
 
@@
 
-Reference: gilson \@ref(Roberts 2013)
+Reference: gilson
 
@@
 The samples are allowed to react at 60C for 4-6 h.
-Sample pressure was adjusted to 1 ATM by adding helium or removing sample using a syringe.
+Sample pressure is adjusted to 1 ATM by adding helium or removing sample using a syringe.
+Evacuation, acid addition, and heated reaction are performed by the automated gas handling system developed for the CFAMS GIS carbonate system \@ref(Roberts 2013).
 
-Process blank samples were prepared from powdered C-1 or TIRI-F. Standards were prepared on the Gilson prep system by evacuating vials and back filling with 1 ATM of a modern standard gas.
+Process blank samples were prepared from powdered C-1 or TIRI-F.
+Standards were prepared using the automated gas handling system by evacuating vials prior to filling with 1 ATM of a modern standard gas.
 

### 2022-02-04 doc/AMS-15/AMS-15_paper/AMS-15_paper.Rmd
Move paper to it's own directory Add scripts for figures

--- 2022-02-04 doc/AMS-15/AMS-15_paper.Rmd
@@
 
-We developed a simple open-split interface that delivers a mixture of helium and CO2 to the source at atmospheric pressure.
-
 The open split interface supplies a helium-CO2 mixture to a split where a capillary delivers gas to the source and the rest is vented.
 A flow of helium is used to displace sample CO2 to the delivery capillary in the split.
+Our open-split interface is constructed of simple, off-the-shelf parts.
 The core of the design is a dual needle and split housed in a VICI cross.
@@
 The mixture of sample CO2 and displacement helium flows up a needle inserted through the cross and stainless tubing which forms the inner flow path of the dual needle.
-The split is made from a VICI Luer adapter fitted to the outflow of the dual needle with the source delivery capillary inserted(Figure 1).
+The split is made from a VICI Luer adapter fitted to the outflow of the dual needle with the source delivery capillary inserted (Figure 1).
 The split has fairly low volume (< 1 mL).
@@
 TODO: move flow ratio test to results.
+The system is currently not automated, but may be easily mounted on the arm of a gasbench-style autosampler.
 
@@
 
-250cm of 50um capillary defines flow rate from the open split into the source, with a bulkhead connector at the source cage.
+flow rate from the open split into the source is defined by 250cm of 50um glass capillary (confirm this, SGE, Trajan, Agilent?), split into two sections by a zero-volume bulkhead connector at the source cage.
+The section inside the cage is jacketed in teflon tubing to help protect against spark damage.
 
@@
 
-Process blank samples were prepared from powdered C-1 or TIRI-F.
-Standards were prepared using the automated gas handling system by evacuating vials prior to filling with 1 ATM of a modern standard gas.
+Process blank samples are prepared from powdered C-1 or TIRI-F. Standards are prepared using the automated gas handling system by evacuating vials prior to filling with 1 ATM of a modern standard gas.
 
@@
 
+Samples are measured on a modified 3MV Tandetron AMS system \@ref(Longworth 2015).
+The displacement helium is allowed to flow constantly during analysis, and the sample needle is inserted into a helium-filled vial between samples.
+Measurement of a sample proceeds as follows: a clean sputter target is inserted into the source and the target is sputtered until ten minutes (check this) have passed or blank target current falls below 100nA.
+The delivery needle is then inserted into the sample vial, which starts sample delivery to the split and source.
+Currents take ~30s to stabilize, at which point the data acquisition is started.
+Data is acquired in 10 blocks of 90 s each, for 15 minutes total acquisition time.
+Normalizing standards, blanks, and secondary standards are measured in the same way.
+Total time for a measurement, including target pre-cleaning is 20-25 minutes.
+
+Discuss delivery rates and sample sizes here or in results?
+
+Initial currents of \> 10 uA dropped rapidly, eventually settling to 100-400 nA.
 1.  Displace gas to open split with 100 uL/min He.
 2.  Introduce mixture to source \@ 30 uL/min
-3.  Acquire data: 90s x 10 measurements.
-
--   Normalizing standards: live tank gas in vials
--   Blanks: C-1 or TIRI-F in vials
 
@@
 Errors from normalization and blank correction are propagated at each step.
-The per-sample error agrees well with variability of replicates, and measurement of known-value standards shows that the method has no significant bias.
 
 Data reduction is done using an R package maintained on github (<https://github.com/blongworth/hgis>).
-
-## Data acquisition
-
-A new target was used for each sample.
-Targets were allowed to "burn in" for 5-10 minutes with pure helium delivered to the source.
-Initial currents of \> 10 uA dropped rapidly, eventually settling to 100-400 nA.
-After a stable base current was reached, the sample vial was installed on the double needle, starting sample flow to the source.
-Sample current rises to a stable level within 30 s, at which time, data acquisition is started.
-10 blocks of 90 s for a total of 15 m.
-
-## Data reduction
 
@@
 
+#### Current dependent blank
+
+$$I_{sample} + I_{blank} = I_{meas}$$
+$$Fm_mI_m=Fm_bI_b+Fm_sI_s$$
+
+$$Fm_s = Fm_m + \frac{I_b(Fm_m-Fm_b)}{I_m-I_b}$$
+
+0.1 μA blank current with 10 μA total = 1% blank
 LBC vs constant contaminant/MBC.
@@
 Figure: current dependent blank
+
+#### Current independent blank
+
+We decided to use a simple linear correction. Samples kept within a narrow range of current and to a constant size.
+
+If **standards**, **blanks**, and **unknowns** have the **same** blank and current, current drops out.
+
+LBC eqn
+
+$$Fm_s = Fm_m - Fm_{blk}\frac{Fm_{std} - Fm_m}{Fm_{std}}$$
+
+```{r}
+# Fix to use current form of processed data
+carb_data <- map_dfr(list.files(here("data_analysed"), full.names = TRUE), read_csv) %>% 
+  filter(!is.na(rec_num),
+         Cur > 2) %>% 
+  mutate(fm_diff = mean - fm_consensus,
+         fm_sigma = sigma(mean, fm_consensus, merr),
+         fmbc_diff = fm_lbc - fm_consensus,
+         fmbc_sigma = sigma(fm_lbc, fm_consensus, merr))
+```
+
+Mean Fm of blanks: `r carb_data %>% filter(fm_consensus < 0.1 & Cur > 2) %>% pull(mean) %>% mean() %>% format(digits = 3)`
 
@@
 ## Reference materials
+The per-sample error agrees well with variability of replicates, and measurement of known-value standards shows that the method has no significant bias.
 

### 2022-02-04 doc/AMS-15/AMS-15_paper/figures/fig1_current_vs_flow.R

@@ -1,0 +1,57 @@
+# Figure showing relationship of current, flow, and evolution over time
+# TODO: combine into a single figure and polish
+
+library(tidyverse)
+library(here)
+library(hgis)
+library(patchwork)
+library(latex2exp)
+
+# process data from flow test samples
+data <- process_hgis_results(here("data/USAMS040121R.txt"),
+                             as.Date("2021-04-02")) %>%
+  filter(pos == 5) %>% 
+  mutate(time = cumsum(corr_lt) / 60,
+         co2flow = concCO2(time, r = 244) * 30) #30ul/min
+
+# flow vs time subplot
+flow_time <- data.frame(x = 0:125) %>% 
+  ggplot(aes(x)) +
+  stat_function(fun = function(x) concCO2(x, r = 244) * 30,
+                aes(color = "CO2"), size = 1.5) +
+  stat_function(fun = function(x) 30 - (concCO2(x, r = 244) * 30),
+                aes(color = "Helium"), size = 1.5) +
+  scale_color_manual("Gas", values = c("#00b7bd", "#b7bf10")) +
+  labs(title = "Gas flows to source",
+       subtitle = "7mL vial, 244μL/min helium, 30μL/min to source",
+       y = "Gas flow (μL/min)") +
+  theme_classic() +
+  theme(axis.title.x = element_blank(),
+        axis.text.x = element_blank(),
+        legend.position = c(0.87, 0.65),
+        legend.background = element_rect(fill = "white", color = "black"))
+
+# current vs time subplot
+cur_time <- ggplot(data, aes(time, he12C)) +
+  geom_smooth(span = .4, se = FALSE, color = "#00b7bd") +
+  geom_point(size = 3, color = "#0069b1") +
+  xlim(0, 125) +
+  labs(title = "Ion current",
+       x = "Time (min)",
+       y = "12C current (μA)") +
+  theme_classic()
+
+flow_time / cur_time
+
+ggsave("fig1a_flow_current_time.svg")
+
+ggplot(data, aes(co2flow, he12C)) +
+  geom_smooth(span = .3, se = FALSE, color = "#00b7bd") +
+  geom_point(size = 2, color = "#0069b1") +
+  labs(title = bquote('Ion current is stable for a range of CO'[2]~'flow'),
+       subtitle = bquote('Current vs. CO'[2]~'flow during vial dillution'),
[…]

### 2022-02-04 doc/AMS-15/AMS-15_paper/figures/fig2_current_dependent_blank.R

@@ -1,0 +1,110 @@
+# Current dependent blank model and data
+# Plot of data and model with measured blanks
+# This should be remade with more data and functions
+# Use sample depletion data?
+# Or skip altogether!
+
+data<- get_hgis_data(here("data/USAMS101320R.txt"), as.Date("2020-11-17")) %>% 
+  mutate(Num = ifelse(Pos %in% c(2, 4), "S", ifelse(Num == "S", "U", Num)))
+stdrat <- mean(data$cor1412he[data$Num == "S" & !data$outlier])
+data1013 <- data %>% mutate(normFm = norm_gas(cor1412he, stdrat))
+
+data <- get_hgis_data(here("data/USAMS120320R.txt"), as.Date("2020-12-04")) %>% 
+  mutate(Num = ifelse(Pos %in% c(2, 4), "S", ifelse(Num == "S", "U", Num)))
+stdrat <- mean(data$cor1412he[data$Num == "S" & !data$outlier])
+data1204 <- data %>% mutate(normFm = norm_gas(cor1412he, stdrat))
+
+data1211 <- get_hgis_data(here("data/USAMS120320R.txt")) %>% 
+  filter(Pos %in% 1:4 | as.Date(ts) == "2020-12-11")
+
+data <- get_hgis_data(here("data/USAMS120320R.txt"), as.Date("2020-12-18")) %>% 
+  filter(Pos != 0) %>% 
+  mutate(Num = ifelse(Pos %in% c(2, 4), "S", ifelse(Num == "S", "U", Num)))
+stdrat <- mean(data$cor1412he[data$Num == "S" & !data$outlier])
+data1218 <- data %>% mutate(normFm = norm_gas(cor1412he, stdrat))
+
+data0108 <- get_hgis_data(here("data/USAMS120320R.txt"), as.Date("2021-01-08"))
+
+data <- get_hgis_data(here("data/USAMS020521R.txt")) %>% 
+  mutate(Num = ifelse(Pos %in% c(2, 4), "S", ifelse(Num == "S", "U", Num)))
+stdrat <- mean(data$cor1412he[data$Num == "S" & !data$outlier])
+data205 <- data %>% mutate(normFm = norm_gas(cor1412he, stdrat))
+
+data <- get_hgis_data(here("data/USAMS030421R.txt"), as.Date("2021-03-05")) %>% 
[…]

### 2022-02-04 doc/AMS-15/AMS-15_paper/figures/fig3_curr_independent_blank.R

@@ -1,0 +1,22 @@
+# Shows uncorrected blank Fm's vs current
+# Need to investigate high outliers- bad wheel?
+
+# Get data
+carb_data <- read_csv(here("data_analysed/carb_all_results.csv"))
+
+# Plot
+carb_data %>% 
+    filter(rec_num %in% c(2138, 83028),
+                     he12C > 2E-6) %>%
+    mutate(Type = ifelse(rec_num == 2138, "TIRI-F", "IAEA C-1")) %>%
+  ggplot(aes(he12C, norm_ratio)) +
+  geom_errorbar(aes(ymin = norm_ratio - sig_norm_ratio,
+                    ymax = norm_ratio + sig_norm_ratio)) +
+  geom_errorbarh(aes(xmin = he12C - sig_he12C, xmax = he12C + sig_he12C)) +
+  geom_point(aes(color = Type), size = 3) +
+  labs(title = "Blanks",
+       x = "12C Current (μA)",
+       y = "Fraction modern") +
+  theme_classic() +
+  theme(legend.position = c(0.82, 0.76),
+        legend.background = element_rect(fill = "white", color = "black"))

### 2022-03-09 doc/AMS-15/AMS-15_paper/AMS-15_paper.Rmd
Clean up text. Add figure captions

@@
 
-The gas-capable 134-sample MCSNICS from National Electrostatics Corporation (NEC) works on the same principal as hybrid sources developed by ETH (REF) and HVEE (REF).
-Sample CO2 is introduced onto a titanium frit in an aluminum target using stainless tubing mounted to an arm with a probe that mates with the back of the target.
-A CO~2~ - helium mixture is exposed to a beam of charged cesium atoms, which causes dissociation of CO~2~ and ionization of molecular and atomic species.
-Negative ions are extracted from the target and a magnetic filter picks species with masses corresponding to carbon.
-
 Most interfaces to gas-accepting sputter ion sources are a closed inlet design.
@@
 Flow may be adjusted by changing the pressure by varying the sample volume.
-
 We designed a open inlet where a capillary samples gas for delivery to the ion source in a flowing stream that exhausts to the atmosphere.
@@
 Salazar and Szidat, Roberts gasbench, other open split methods.
+
+The gas-capable 134-sample MCSNICS from National Electrostatics Corporation (NEC) works on the same principal as hybrid sources developed by ETH (REF) and HVEE (REF).
+Sample CO2 is introduced onto a titanium frit in an aluminum target using stainless tubing mounted to an arm with a probe that mates with the back of the target.
+A CO~2~ - helium mixture is exposed to a beam of charged cesium atoms, which causes dissociation of CO~2~ and ionization of molecular and atomic species.
+Negative ions are extracted from the target and a magnetic filter picks species with masses corresponding to carbon.
 
@@
 TODO: move flow ratio test to results.
-The system is currently not automated, but may be easily mounted on the arm of a gasbench-style autosampler.
+The system is currently operated manually, but may be easily mounted on the arm of a gasbench-style autosampler to enable automated operation.
 
@@
 
-Figure: Current vs flow
+Figure1: Current vs flow
 
@@
 
-Figure: current dependent blank
+Figure2: current dependent blank
 
@@
 
-Figure: justification for large blank
+Figure3: justification for large blank
 
@@
 
-Figure: sample size vs current vs precision and accuracy
+Figure4: sample size vs current vs precision and accuracy
 
@@
 Two more recent carbonate tests produced stable currents and backgrounds.
-Since sample, target blank, and sample blank currents were relatively constant across samples, a large blank correction can be used to correct results.
+Sample, target blank, and sample blank currents were relatively constant across samples, so a large blank correction can be used to correct results.
 Samples from both runs showed a linear relationship between difference from expected Fm and Fm of the sample, but in one sample set the relationship was positive, and in the other, negative.
@@
 
-figure: difference plot
+Table: summary of reference materials
+
+figure5: difference plot
 
@@
 
-Figure: bland-altman of HGIS vs graphite.
+Figure6: bland-altman of HGIS vs graphite.
 
@@
 
-Figure: age model comparison graphite HGIS
+Figure7: age model comparison graphite HGIS
 
@@
 This work is supported by NSF cooperative agreement. We are grateful to C. McIntyre, R. Hansman, G. Salazar, T. Ognibene, T. Broeck.
+
+# Figure captions
+
+Fig. 1. Performance of the NOSAMS MCSNICS hybrid gas ion source.
+
+Fig.
+2.
+Assessment of blanks.
+(a) Current dependent blank.
+(b) Fraction modern of large, radiocarbon dead samples compared to sample current.
+Blank rises with decreasing current, as expected, but may be treated as a single blank with a simple correction for full-size samples producing greater than XX uA C- current.
+
+Fig. 3. Relationship of sample size to C- ion current. There is no strong relationship between size and current for samples larger than XX ug C.
+
+Fig 4. Agreement of standards with graphite consensus values.
+
+Fig 5. Comparison of Haiti core samples measured as graphite.

### 2022-03-09 doc/AMS-15/AMS-15_paper/tables/table_1_reference_materials.Rmd
Add summary tables. Add current vs size figure

---
title: "Table_1_ref_standards"
author: "Brett Longworth"
date: "3/9/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

```{r}
library(tidyverse)
library(hgis)
library(here)
library(gt)
library(errors)

options(digits = 3)
options(scipen = 4)
options(decimals = 4)
options(errors.notation = "plus-minus")

# Load data
results_df <- read_csv(here("data_analysed/carb_all_results.csv"))

# Format data
cons_df <- results_df %>% 
  filter(!(wheel %in% c("USAMS020521", "USAMS061121", "USAMS061821"))) %>% 
  mutate(he12C = he12C * 1E6) %>% 
  compare_consensus() %>% 
  filter(!str_detect(Name, "Gas"))

cons_df_sum <- cons_df %>% 
  group_by(Name, fm_consensus) %>% 
  summarize(across(c(fm_corr, sig_fm_corr, fm_diff, sigma),
                   list(mean = mean, sd = sd)),
            N = n()) %>% 
  ungroup()
```

```{r}

# Summary table
cons_df_sum %>% 
  arrange(fm_consensus) %>% 
  mutate(across(c(fm_consensus, fm_corr_mean, fm_corr_sd, sig_fm_corr_mean), 
         function(x) x * 100),
         fm_corr_err = set_errors(fm_corr_mean, fm_corr_sd)) %>% 
  select(Name, `Consensus` = fm_consensus,
         `HGIS ± SD` = fm_corr_err,
         #`HGIS` = fm_corr_mean,
         #`Std. dev. of measurements` = fm_corr_sd,
         `Mean per-sample error` = sig_fm_corr_mean,
         N) %>% 
gt() %>% 
  tab_header(title = "Reference material precision and accuracy",
             subtitle = "Values in pMC, Errors are SD of measurements") %>% 
  fmt_number(4, 
             decimals = 1) %>% 
  cols_align("right", 3)
```

# Consensus summary
```{r}
cons_df %>% 
  filter(Name %in% c("TIRI-I", "C-2", "NOSAMS2")) %>% 
  ungroup() %>% 
  summarize(across(c(sig_fm_corr, fm_diff),
                   list(mean = mean, sd = sd)),
            N = n()) %>% 
  select(`Consensus difference` = fm_diff_mean,
         `Std. Dev. of consensus difference` = fm_diff_sd,
         `per-sample error` = sig_fm_corr_mean,
         N
         ) %>%  
  gt() %>% 
  tab_header(title = "Reference material performance summary",
             subtitle = "TIRI-I, C-2, and NOSAMS2. Values in pMC") %>% 
  fmt_number(1:3, 
             scale_by = 100,
             decimals = 2)

```

### 2022-03-09 doc/AMS-15/AMS-15_paper/figures/fig4_current_vs_size.R

@@ -1,0 +1,18 @@
+# Relationship of sample size to ion current
+
+
+# Read reduced data as produced by analyse_carbonates.R
+df <- readRDS(here("data_analysed/carb_all.rds")) %>% 
+  map_dfr(2) %>% 
+  compare_consensus() %>% 
+  mutate(c12ua = he12C * 1E6)
+samples <- read_csv(here("data/carb_data.csv")) %>% 
+  select(wheel, pos, mass, he_added)
+df <- left_join(df, samples, by = c("wheel", "pos")) %>% 
+  filter(wheel != "USAMS020521")
+
+
+# Basic plot of current vs mass
+df %>% 
+ggplot(aes(mass, c12ua)) +
+  geom_point()

### 2022-03-09 doc/AMS-15/AMS-15_paper/tables/table_1_reference_materials.R

@@ -1,0 +1,68 @@
+# Table 1 showing performance of reference standards and performance summary
+
+
+library(tidyverse)
+library(hgis)
+library(here)
+library(gt)
+library(errors)
+
+options(digits = 3)
+options(scipen = 4)
+options(decimals = 4)
+options(errors.notation = "plus-minus")
+
+# Load data
+results_df <- read_csv(here("data_analysed/carb_all_results.csv"))
+
+# Format data
+cons_df <- results_df %>% 
+  filter(!(wheel %in% c("USAMS020521", "USAMS061121", "USAMS061821"))) %>% 
+  mutate(he12C = he12C * 1E6) %>% 
+  compare_consensus() %>% 
+  filter(!str_detect(Name, "Gas"))
+
+cons_df_sum <- cons_df %>% 
+  group_by(Name, fm_consensus) %>% 
+  summarize(across(c(fm_corr, sig_fm_corr, fm_diff, sigma),
+                   list(mean = mean, sd = sd)),
+            N = n()) %>% 
+  ungroup()
+
+# Summary table
+cons_df_sum %>% 
+  arrange(fm_consensus) %>% 
+  mutate(across(c(fm_consensus, fm_corr_mean, fm_corr_sd, sig_fm_corr_mean), 
+         function(x) x * 100),
+         fm_corr_err = set_errors(fm_corr_mean, fm_corr_sd)) %>% 
+  select(Name, `Consensus` = fm_consensus,
+         `HGIS ± SD` = fm_corr_err,
+         #`HGIS` = fm_corr_mean,
+         #`Std. dev. of measurements` = fm_corr_sd,
+         `Mean per-sample error` = sig_fm_corr_mean,
+         N) %>% 
+gt() %>% 
+  tab_header(title = "Reference material precision and accuracy",
+             subtitle = "Values in pMC, Errors are SD of measurements") %>% 
+  fmt_number(4, 
+             decimals = 1) %>% 
+  cols_align("right", 3)
+
+# Consensus summary
+cons_df %>% 
+  filter(Name %in% c("TIRI-I", "C-2", "NOSAMS2")) %>% 
+  ungroup() %>% 
+  summarize(across(c(sig_fm_corr, fm_diff),
+                   list(mean = mean, sd = sd)),
+            N = n()) %>% 
+  select(`Consensus difference` = fm_diff_mean,
+         `Std. Dev. of consensus difference` = fm_diff_sd,
[…]

### 2022-03-15 doc/AMS-15/AMS-15_paper/AMS-15_paper.Rmd
Data reduction section

@@
 The split has fairly low volume (< 1 mL).
-Testing is needed to determine minimum sample flow to avoid atmosphere entering the split.
-TODO: move flow ratio test to results.
 The system is currently operated manually, but may be easily mounted on the arm of a gasbench-style autosampler to enable automated operation.
@@
 
-flow rate from the open split into the source is defined by 250cm of 50um glass capillary (confirm this, SGE, Trajan, Agilent?), split into two sections by a zero-volume bulkhead connector at the source cage.
+The flow rate from the open split into the source is defined by 250cm of 50um glass capillary (confirm this, SGE, Trajan, Agilent?), split into two sections by a zero-volume bulkhead connector at the source cage.
 The section inside the cage is jacketed in teflon tubing to help protect against spark damage.
-
-![Experimental setup. Helium and standard gas regulators at left, sample vial and open split at top right.](images/gas_control.jpg)
 
@@
 Carbonate samples are weighed into 7 ml septa-sealed Vacutainer vials (Labco \# 768W).
-Sealed vials are evacuated, then 1 ml 85% H3PO4 is added.
+Sealed vials are evacuated, then 1 ml 85% Phosphoric acid is added.
 The samples are allowed to react at 60C for 4-6 h.
-Sample pressure is adjusted to 1 ATM by adding helium or removing sample using a syringe.
+Sample pressure is adjusted to 1 ATM by allowing the sample to equilibrate with a helium filled syringe connected through the vial septum.
 Evacuation, acid addition, and heated reaction are performed by the automated gas handling system developed for the CFAMS GIS carbonate system \@ref(Roberts 2013).
 
-Process blank samples are prepared from powdered C-1 or TIRI-F. Standards are prepared using the automated gas handling system by evacuating vials prior to filling with 1 ATM of a modern standard gas.
-
-1.  Weigh 10-30mg Carbonate in septa-sealed 7mL vials produces 2-6mL CO<sub>2</sub> \@ \~1ATM
-2.  Evacuate and Acidify with 1 mL H3PO4 using automated Gilson prep bench
-3.  After reaction is complete, add helium to bring vials to ambient pressure.
-
-![A modified liquid handling system is used to evacuate and add acid to sample vials.](images/gilson.png)
+Standards are prepared by evacuating vials and then filling with 1 ATM of a modern standard gas using the automated gas handling system.
+Process blank samples are prepared from powdered C-1 or TIRI-F, acidified in vials using the method above.
 
@@
 
-Discuss delivery rates and sample sizes here or in results?
-
-Initial currents of \> 10 uA dropped rapidly, eventually settling to 100-400 nA.
-1.  Displace gas to open split with 100 uL/min He.
-2.  Introduce mixture to source \@ 30 uL/min
-
 ### Data reduction
 
-Normalization to standard, blank correction, and estimation of measurement error are done with the following steps:
+After aquisition, data reduction, normalization and blank correction result in a final fraction modern with error.
+First, each 90s block is corrected for fractionation using the AMS <sup>13</sup>C measurement.
+The raw 14/12C ratio divided by the 13/12C ratio squared gives the corrected 14/12C ratio.
+Outlier runs, defined as those outside of 1.5 times the interquartile range, are dropped at this point.
+Remaining runs of a sample are averaged and per-sample error is taken as the greater of counting error or the standard error of the measurements.
+Ratios are then normalized using the mean of gas standards measured during the run.
+The measurement error of the standards is taken as the greater of the standard error of all standards or the errors of the standards added in quadrature.
 
-1.  Correct for fractionation using online AMS <sup>13</sup>C measurement.
-2.  Determine per-sample mean and error (repeatabilty and counting error).
-3.  Normalize to mean of Gas standard.
-4.  Blank correct: Blank subtraction using mean of carbonate blanks
+Measurements are corrected for process and measurement blank using a linear correction that assumes that the corrected fraction modern of blank samples should be zero, samples with the fraction modern of the standard need no correction, and applies a proportional correction to samples of intermediate fraction modern.
+(REF) The correction uses measured values of standard gas and blank samples for correction of contaminant carbon and machine blank.
+We use equation 1 for this correction.
 
-Errors from normalization and blank correction are propagated at each step.
+$$ R_s = R_m - \frac{R_b(R_{ox}-R_m)}{R_{ox}} $$
 
-Data reduction is done using an R package maintained on github (<https://github.com/blongworth/hgis>).
-
-Normalization and blank correction of measurements
-
-A linear correction using measured values of standard gas and blank samples was used for correction of contaminant carbon and machine blank. We use equation 1 for correction
-
-LBC equation
-
-where x = y.
-
-Constant contaminant discussed below.
-
-Data reduction via R package supplementary materials
+Where $R_s$ is the ratio of the unknown sample, $R_m$ is the measured ratio, $R_b$ is the ratio of the blank, and $R_{ox}$ is the ratio of the normalizing standard.
+The errors in the blanks and standards are propagated with errors of the samples to produce the final error estimate for the sample.
+This correction does not account for the effect of a blank with constant mass or current on samples of varying mass or current.
+For all measurements used to assess the performance of the method and for the Haiti core comparison, sample, standard, and blank masses were kept constant.
+Sample current did vary, but we believe the linear correction described above produces the most accurate results within this range.
+We discuss this constant contaminant (REF) blank in the results below.
+Data reduction is done using a package written in R (REF) which is published on github (<https://github.com/blongworth/hgis>).
 
@@
 All results in supplementary materials
+
+Determination of ideal flows.
+Testing is needed to determine minimum sample flow to avoid atmosphere entering the split. TODO: move flow ratio test to results.
 
@@
 
+Discuss delivery rates and sample sizes here or in results?
+
+Initial currents of \> 10 uA dropped rapidly, eventually settling to 100-400 nA.
+
 ### Currents
@@
 
-### Blanks
+### Contaminant carbon
 

### 2022-03-16 R/analyse_carbonates.R
Drop high blank from early wheel. Notes confirm that this sample was suspect.

@@ -15,7 +15,7 @@ wheels <- c("020521", "030421", "040121", "041521", "051421", "052621", "061121"
 files <- here("data", paste0("USAMS", wheels, "R.txt"))
 dates <- as.Date(c("2021-02-05", "2021-03-05", "2021-04-09", "2021-04-16", "2021-05-14", "2021-05-28", "2021-06-11", "2021-06-18", "2021-06-25", "2021-07-02"))
 stds <- list(c(5,8), NULL, c(26,37), NULL, c(10,15,65,70), NULL, NULL, NULL, NULL, NULL)
-blanks <- list(7, c(7,8), c(28,29,30), c(27,28,29), c(20,35,50), NULL, NULL, 23, NULL, NULL)
+blanks <- list(7, c(8), c(28,29,30), c(27,28,29), c(20,35,50), NULL, NULL, 23, NULL, NULL)
 outliers <- list(NULL, tibble(pos = c(6, 8, 8, 8), meas = c(18, 1, 2, 3)),
                  tibble(pos = c(28,29), meas=c(1, 1)),
                  NULL,

### 2022-03-16 doc/AMS-15/AMS-15_paper/figures/fig3_curr_independent_blank.R
Add libraries. Use blanks used in analysis. Save result to file.

@@ -1,22 +1,29 @@
 # Shows uncorrected blank Fm's vs current
 # Need to investigate high outliers- bad wheel?
 
+# Libraries
+library(tidyverse)
+library(here)
+
 # Get data
 carb_data <- read_csv(here("data_analysed/carb_all_results.csv"))
 
 # Plot
 carb_data %>% 
-    filter(rec_num %in% c(2138, 83028),
-                     he12C > 2E-6) %>%
-    mutate(Type = ifelse(rec_num == 2138, "TIRI-F", "IAEA C-1")) %>%
+  filter(rec_num %in% c(2138, 83028),
+         he12C > 2E-6,
+         sample_type == "B") %>%
+  mutate(Type = ifelse(rec_num == 2138, "TIRI-F", "IAEA C-1")) %>%
   ggplot(aes(he12C, norm_ratio)) +
   geom_errorbar(aes(ymin = norm_ratio - sig_norm_ratio,
                     ymax = norm_ratio + sig_norm_ratio)) +
   geom_errorbarh(aes(xmin = he12C - sig_he12C, xmax = he12C + sig_he12C)) +
-  geom_point(aes(color = Type), size = 3) +
+  geom_point(aes(color = wheel), size = 3) +
   labs(title = "Blanks",
        x = "12C Current (μA)",
        y = "Fraction modern") +
   theme_classic() +
   theme(legend.position = c(0.82, 0.76),
         legend.background = element_rect(fill = "white", color = "black"))
+
+ggsave("fig3_cur_indep_blank.svg")

### 2022-03-16 doc/AMS-15/AMS-15_paper/AMS-15_paper.Rmd
Add results section on blank correction

@@
 Testing is needed to determine minimum sample flow to avoid atmosphere entering the split. TODO: move flow ratio test to results.
+## Method development
 
@@
 
+To assess blank contribution and ratio dependence on sample current, vials of live and dead standard gas were measured as the displacement helium flow dilluted the samples, producing diminishing current over time.
+We used these data to develop a current-balance model following methods used to assess mass-dependent or constant contamination blanks in graphite targets (REFS).
+The constant contamination blank was also assessed directly by measuring pure helium in vials using the same method.
+Results from both tests are shown in figure 2.
+The inverse of the sample current is used to linearize the data, allowing assessment with a simple linear model.
+Fits from modern and dead dilution tests should trend towards the fraction modern of the blank, crossing at the inverse of blank current.
+Contaminant current measured directly agrees well with the model assessment, but has a higher fraction modern.
+This is likely due to a combination of low measurement precision for the direct measurement and high sensitivity of the model to variability in the measured fraction modern and current in the depletion samples (REF).
+This constant contamination blank is consistent with the fact that hybrid cathodes produce a minimum current with or without gas flow to the source.
+Given that this minimum current ranges from 100 to 400, this current likely makes up a large fraction of the method blank.
+
 $$I_{sample} + I_{blank} = I_{meas}$$
@@
 
+While dillution and direct measurement data show evidence
 We decided to use a simple linear correction. Samples kept within a narrow range of current and to a constant size.

### 2022-03-16 doc/AMS-15/AMS-15_paper/AMS-15_paper.Rmd
Add current independent blank justification

@@
 
-While dillution and direct measurement data show evidence
-We decided to use a simple linear correction. Samples kept within a narrow range of current and to a constant size.
-
-If **standards**, **blanks**, and **unknowns** have the **same** blank and current, current drops out.
+While dillution and direct measurement data show evidence for a constant blank with a variable effect depending on sample size, the blank correction may be simplified if blanks, standards, and unknown samples are the same size.
+If the fraction modern of the blank is 0.8 modern as estimated above, blank samples will show the most effect of a constant contamination blank with current.
+Figure 3 shows the fraction modern of measured blanks compared to their mean 12C- ion current.
+Fraction modern of measured blanks does decrease with increasing current, but the effect is small, less than Fm 0.01 over the range of currents measured.
+This change with current is roughly equal to the variablility between blanks, so the error due to the use of a current-dependent blank based on samples measured over multiple wheels will likely be larger than a current independent blank correction based on blanks measured on each wheel.
 

### 2022-03-17 doc/AMS-15/AMS-15_paper/figures/fig6_core_carbonates.R
write combined core data to file.

@@ -52,6 +52,7 @@ depth <- read_csv(here("data/haiti_depth.csv"))
 combined_results <- left_join(combined_results, depth, by = "rec_num") %>% 
   mutate(rc_age = -8033 * log(fm_corr),
          sig_rc_age = rc_age - -8033 * log(fm_corr + sig_fm_corr))
+write_csv(combined_results, here("data_analysed/haiti_combined.csv"))
 
 # No outliers
 cr_no <- combined_results %>% 

### 2022-03-18 doc/AMS-15/AMS-15_paper/AMS-15_paper.Rmd
Expand figure captions, reorganize results.

@@
 Testing is needed to determine minimum sample flow to avoid atmosphere entering the split. TODO: move flow ratio test to results.
+
 ## Method development
@@
 Figure1: Current vs flow
+
+### Sample size
+
+Sample size tests.
+
+Figure4: sample size vs current vs precision and accuracy
+
+Aimed to fill vial initially
+
+No current degradation down to X. Used Y as optimal mass and backfilled with helium.
+
+First tests of displacing pure CO2 from a vial into the open split using helium are promising.
+Helium flowing at 100uLm-1 produces current from 7mL of CO2 for over 2 hours.
+Currents do decrease with time due to dilution, but by supplying higher than optimal CO2 flow to the source initially, stable currents can be maintained for some time.
+Fraction of CO2 in sample gas should follow this relationship:
+
+$$C(t) = e^{-\frac{r}{V}t}$$
+
+Where $C(t)$ is the fraction CO2, $r$ is the rate of helium displacement flow, and $V$ is the initial volume of the CO2-filled vial.
+
+Setting flow to the source to 10uL/min provides excess CO2 to source until the helium dilution ratio increases to \> 1:1.
+Another approach may be to backfill vials with helium and run at a higher dilution ratio and flow to the source.
+This should produce a smaller change in He:CO2 over the course of a run.
+MICADAS runs at 95% He:CO2, which would mean producing \< 0.5mL CO2 in a 7mL vial and backfilling with helium before displacing to the source.
+Some intermediate dilution may be appropriate.
+Alternatively, vials could be flushed with helium prior to acidification, rather than evacuating, leading to some concentration of CO2 in He at slightly \> 1 ATM.
+
+Testing long runs of dilution of CO2 with displacing He and diluting He with CO2 show that the setup is still optimized for low flow to the source, even with larger diameter delivery capillaries.
+There appears to be a constriction in the split-source tubing that is keeping flows low.
+Current focus is on the large-sample 7mL method, so this will be diagnosed as time allows.
+
+Targets have shown "pulses" of higher current.
+Not sure why this is, but I suspect it's either buildup of graphite or other available carbon on the target, or target switching to the "high state" as sometimes happens with solid targets.
+
+First tests with carbonate in vials were hit-or-miss.
+Problems with capillary clogging and low currents led to poor results for several samples.
+A pair of NOSAMS2 samples produced good currents, results agreed well, but were about 3% higher than expected Fm.
+A second set of carbonates produced better currents (\~5uA) and better results.
+Agreement of replicates varied from 3 to 16 permil.
+Near-modern NOSAMS2 was very close to the consensus value, while C-2 and C-1 were elevated by 2 and 3.6% respectively, indicating that a blank correction is needed.
+
+Two more recent carbonate tests produced stable currents and backgrounds.
+Sample, target blank, and sample blank currents were relatively constant across samples, so a large blank correction can be used to correct results.
+Samples from both runs showed a linear relationship between difference from expected Fm and Fm of the sample, but in one sample set the relationship was positive, and in the other, negative.
+Since the intercept with 0-offset was close to 0 Fm, I think this is a problem with normalization, rather than blank correction.
+Standards run at the end of the day are higher than those run at the beginning, which may give clues as to what's going on.
 
@@
 
-### Sample size
-
-Sample size tests.
-
-Figure4: sample size vs current vs precision and accuracy
-
-Aimed to fill vial initially
-
-No current degradation down to X. Used Y as optimal mass and backfilled with helium.
-
-First tests of displacing pure CO2 from a vial into the open split using helium are promising.
-Helium flowing at 100uLm-1 produces current from 7mL of CO2 for over 2 hours.
-Currents do decrease with time due to dilution, but by supplying higher than optimal CO2 flow to the source initially, stable currents can be maintained for some time.
-Fraction of CO2 in sample gas should follow this relationship:
-
-$$C(t) = e^{-\frac{r}{V}t}$$
-
-Where $C(t)$ is the fraction CO2, $r$ is the rate of helium displacement flow, and $V$ is the initial volume of the CO2-filled vial.
-
-Setting flow to the source to 10uL/min provides excess CO2 to source until the helium dilution ratio increases to \> 1:1.
-Another approach may be to backfill vials with helium and run at a higher dilution ratio and flow to the source.
-This should produce a smaller change in He:CO2 over the course of a run.
-MICADAS runs at 95% He:CO2, which would mean producing \< 0.5mL CO2 in a 7mL vial and backfilling with helium before displacing to the source.
-Some intermediate dilution may be appropriate.
-Alternatively, vials could be flushed with helium prior to acidification, rather than evacuating, leading to some concentration of CO2 in He at slightly \> 1 ATM.
-
-Testing long runs of dilution of CO2 with displacing He and diluting He with CO2 show that the setup is still optimized for low flow to the source, even with larger diameter delivery capillaries.
-There appears to be a constriction in the split-source tubing that is keeping flows low.
-Current focus is on the large-sample 7mL method, so this will be diagnosed as time allows.
-
-Targets have shown "pulses" of higher current.
-Not sure why this is, but I suspect it's either buildup of graphite or other available carbon on the target, or target switching to the "high state" as sometimes happens with solid targets.
-
-First tests with carbonate in vials were hit-or-miss.
-Problems with capillary clogging and low currents led to poor results for several samples.
-A pair of NOSAMS2 samples produced good currents, results agreed well, but were about 3% higher than expected Fm.
-A second set of carbonates produced better currents (\~5uA) and better results.
-Agreement of replicates varied from 3 to 16 permil.
-Near-modern NOSAMS2 was very close to the consensus value, while C-2 and C-1 were elevated by 2 and 3.6% respectively, indicating that a blank correction is needed.
-
-Two more recent carbonate tests produced stable currents and backgrounds.
-Sample, target blank, and sample blank currents were relatively constant across samples, so a large blank correction can be used to correct results.
-Samples from both runs showed a linear relationship between difference from expected Fm and Fm of the sample, but in one sample set the relationship was positive, and in the other, negative.
-Since the intercept with 0-offset was close to 0 Fm, I think this is a problem with normalization, rather than blank correction.
-Standards run at the end of the day are higher than those run at the beginning, which may give clues as to what's going on.
-
 ## Reference materials
@@
 
-A suite of carbonate macrofossils consisting of shell and coral fragements from a series of coastal marine cores REF were measured as graphite and as CO2 using the HGIS method.
+A suite of 15 carbonate macrofossils consisting of shell and coral fragements from a series of coastal marine cores REF were measured as graphite and as CO2 using the HGIS method to further test the method and serve to demonstrate a potential application.
+coral macrofossils from a subtropical coastal marine sediment core using our gas 
+method and compared to the same samples measured using the traditional 
+graphite method.
+
+To further test the method and a potential application, we measured 15 shell and 
+coral macrofossils from a subtropical coastal marine sediment core using our gas 
+method and compared to the same samples measured using the traditional 
+graphite method. 
+We found the HGIS and graphite measurements to be in good agreement, with no 
+significant bias (figure 6). Mean per-sample error was 1 pMC for HGIS measurements, 
+compared with 0.3 pMC for graphite.
+The accuracy of age-depth models is often driven by environmental factors and 
+sparse data, rather than the precision of individual dates [3]. The lower time and 
+cost of the hybrid method allows analysis of more samples. This may produce more 
+accurate age models, despite reduced precision of individual measurements.
 
@@
 
-Quantification of method
+## Quantification of method
 
-Cost vs quality discussion
+## Cost vs quality discussion
 
-ideas for further development of method
+The accuracy of age-depth models is often driven by environmental factors and 
+sparse data, rather than the precision of individual dates [3]. The lower time and 
+cost of the hybrid method allows analysis of more samples. This may produce more 
+accurate age models, despite reduced precision of individual measurements.
+
+## ideas for further development of method
 
@@
 
-Fig. 1. Performance of the NOSAMS MCSNICS hybrid gas ion source.
+Fig.
+1.
+Schematic of the open split inlet.
+Two valco valves provide the ability to admit helium and live or dead tank standards.
+The double needle  is housed in a VICI cross and allows displacement helium to flow into the sample vial and the sample-helium mixture to flow to the open split, where a capillary delivers the mixture to the ion source.
 
@@
 2.
+Performance of the NOSAMS MCSNICS hybrid gas ion source.
+a.
+Modelled flows of helium and CO2 to the ion source during displacement of a pure CO2 vial sample with helium.
+b.
+Measured 12C- ion current during the same dillution experiment.
+c.
+Measured ion current for modelled CO2 flow.
+The stability of the source over a range of flow produces stable current for up to 40 m of measurement time.
+
+Fig. 3. Relationship of sample size to C- ion current. There is no strong relationship between size and current for samples larger than 800 ug C.
+
+Fig.
+4.
 Assessment of blanks.
-(a) Current dependent blank.
+(a) Current dependent blank model and data.
+Fm and current of measurements from dillution tests of live and dead gas are used to generate linear fits which should cross at the Fm and current of the constant current blank.
+Direct determination of Fm and current of the blank are also shown.
 (b) Fraction modern of large, radiocarbon dead samples compared to sample current.
-Blank rises with decreasing current, as expected, but may be treated as a single blank with a simple correction for full-size samples producing greater than XX uA C- current.
+Blank rises with decreasing current, as expected, but may be treated as a single blank with a simple correction for full-size samples producing greater than 3 uA C- current.
 
-Fig. 3. Relationship of sample size to C- ion current. There is no strong relationship between size and current for samples larger than XX ug C.
+Fig 5.
+Agreement of standards with graphite consensus values.
+Difference of HGIS measurements from consensus for individual samples are plotted against their expected value.
+Expected values are spread to avoid overlapping points.
+Error bars represent one sigma sample errors.
+The black line is the mean difference from expected value for all measurements, and the blue lines are one standard deviation of all measurements.
 
-Fig 4. Agreement of standards with graphite consensus values.
-
-Fig 5. Comparison of Haiti core samples measured as graphite.
+Fig 6.
+Comparison of Haiti core samples measured as graphite.
+a.
+Difference between HGIS and graphite measurements of the same materials vs.
+core depth (should this be mean of values?).
+Samples are offset on the horizontal axis for clarity.
+Error bars are measurement error for each sample.
+b) comparison of age-depth models produced using HGIS and graphite measurements.

### 2022-03-22 doc/AMS-15/AMS-15_paper/figures/fig5_consensus_samples.R
Add aggregate data comparison for core data

@@ -42,3 +42,4 @@ ggplot(aes(fm_consensus, fm_diff, color = Name)) +
        y = "pMC difference (HGIS - expected)")
 
 ggsave("fig5_consensus_stds.svg")
+

### 2022-03-22 doc/AMS-15/AMS-15_paper/figures/fig6_core_carbonates.R

@@ -6,8 +6,7 @@ library(here)
 library(hgis)
 library(amstools)
 library(scales)
-
-
+library(gt)
 
 # Read raw and reduced data as produced by analyse_carbonates.R
 data <- readRDS(here("data_analysed/carb_all.rds"))
@@ -68,6 +67,7 @@ mean_errs <- cr_no %>%
   pivot_wider(names_from = method,
               values_from = mean_sig)
 
+# Make difference plot of core carbonates
 ggplot(cr_no) +
   geom_hline(yintercept = 0) +
   #geom_hline(yintercept = mean_diff$fm_diff_mean, color = "blue") +
@@ -91,3 +91,24 @@ ggplot(cr_no) +
 
 ggsave(here("doc/AMS-15/AMS-15_paper/figures/fig6_core_carbonates.svg"))
 
+# data for core carbonates section
+
+# mean 
+
+cr_no %>% 
+  group_by(method) %>% 
+  #filter(method == "hgis") %>% 
+  summarize(across(c(sig_fm_corr, fm_diff),
+                   list(mean = mean, sd = sd)),
+            N = n()) %>% 
+  select(`Consensus difference` = fm_diff_mean,
+         `Std. Dev. of consensus difference` = fm_diff_sd,
+         `per-sample error` = sig_fm_corr_mean,
+         N
+         ) %>%  
+  gt() %>% 
+  tab_header(title = "Haiti carbonates performance summary",
+             subtitle = "TIRI-I, C-2, and NOSAMS2. Values in pMC") %>% 
+  fmt_number(1:3, 
+             scale_by = 100,
+             decimals = 2)

### 2022-03-22 doc/AMS-15/AMS-15_paper/AMS-15_paper.Rmd
Add consensus standards and core samples sections. Reorganize.

@@
 Best currents for the source are around 10uA 12C- or roughly 1/10 of graphite currents.
-This is in line with what other hybrid sources produce.
-Typical "good" currents are between 5-10uA.
+This is generally comparable with other hybrid sources REFS.
+Typical currents for this study were between 4-10uA???.
 The source typically produces best current at around 5uLs-1 CO2.
@@
 ## Reference materials
+
+Precision and accuracy of the method were assessed by measuring carbonate standards with consensus values from graphite measurements.
+Summary data for these samples are listed in table 1, and measurement data for all samples is in supplementary information.
+Measurement, data reduction, and blank correction was performed as described above.
+
+Overall, reference materials run with the hybrid gas method were in good agreement with consensus values.
+Mean difference from consenus for all measurements was -0.04 pMC, and the standard deviation of the difference was 1.10 pMC.
+Mean propagated measurement error was 1.13 pMC for these samples.
+The hybrid method has no bias in agregate.
+Method error is larger than graphite, but calculated per-sample error is similar to standard deviation of replicate measurements.
+A summary of accuracy and precision of these measurements is in Table 1.
+
+The difference beween reference samples and graphite consensus does not vary systematically by fraction modern, indicating that the method has no significant bias by fraction modern (Figure 6).
+
 The per-sample error agrees well with variability of replicates, and measurement of known-value standards shows that the method has no significant bias.
-
-Measured
-
-Table: summary of reference materials
-
-figure5: difference plot
-
-Assessment of data quality.
 
@@
 graphite method.
+
 We found the HGIS and graphite measurements to be in good agreement, with no 
-significant bias (figure 6). Mean per-sample error was 1 pMC for HGIS measurements, 
-compared with 0.3 pMC for graphite.
+significant bias (figure 6).
+Mean difference between hybrid and graphite measurements was -0.35 pMC and the standard deviation of this difference was 0.85 pMC.
+Mean per-sample error was 1.02 pMC for HGIS measurements, compared with 0.18 pMC for graphite.
+
+One HGIS sample vial was noted to be at atmospheric pressure when the 16mg sample should have produced only one half atmosphere of CO2.
+The HGIS measurement for this sample was 5.3 pMC higher than the measurement as graphite.
+We suspect that this sample was improperly sealed and eliminated it from the analysis as an outlier.
+
+### Age model
+
+Age models were developed for both the traditional graphite measurements of the core carbonates and the HGIS measurements (Figure 7).
+We used the Bchron Baesian radiochronology package to develop these models.
+The age-depth curves produced by the two datasets are very similar, with the largest differences being in the uncertainty of the individual dates.
+
+Mean and sd of difference of calibrated ages?
+
 The accuracy of age-depth models is often driven by environmental factors and 
@@
 accurate age models, despite reduced precision of individual measurements.
-
-### Measurements
-
-Table of measurements (supplementary?)
-
-One HGIS sample vial was noted to be at atmospheric pressure when only XX mL CO2 should have been produced.
-The HGIS measurement for this sample was XX% higher than the measurement as graphite.
-We suspect that this sample was improperly sealed and eliminated it from the analysis as an outlier.
-SO2 story?
-
-Figure6: bland-altman of HGIS vs graphite.
-
-Summary of agreement
-
-### Age model
 
@@
 
-## Quantification of method
-
-## Cost vs quality discussion
-
-The accuracy of age-depth models is often driven by environmental factors and 
-sparse data, rather than the precision of individual dates [3]. The lower time and 
-cost of the hybrid method allows analysis of more samples. This may produce more 
-accurate age models, despite reduced precision of individual measurements.
-
-## ideas for further development of method
-
-# Conclusion
-
 We developed a method for the analysis of carbonates using an open-split interface to a hybrid sputter gas ion source.
-This system produces results that agree with graphite measurements to better than 1% from XXX mg of carbonate.
+This system produces results that agree with graphite measurements to better than 1% from 8-20 mg of carbonate.
 Sample preparation takes about 10 min.
@@
 
+The accuracy of age-depth models is often driven by environmental factors and 
+sparse data, rather than the precision of individual dates [3]. The lower time and 
+cost of the hybrid method allows analysis of more samples. This may produce more 
+accurate age models, despite reduced precision of individual measurements.
+
 Sample requirements, analysis time, precision, accuracy.
+
+## ideas for further development of method
 

### 2022-03-23 doc/AMS-15/AMS-15_paper/figures/fig1_current_vs_flow.R
Add figure paths. Update for new functions.

@@ -43,7 +43,7 @@ cur_time <- ggplot(data, aes(time, he12C)) +
 
 flow_time / cur_time
 
-ggsave("fig1a_flow_current_time.svg")
+ggsave(here("doc/AMS-15/AMS-15_paper/figures/fig1a_flow_current_time.svg"))
 
 ggplot(data, aes(co2flow, he12C)) +
   geom_smooth(span = .3, se = FALSE, color = "#00b7bd") +
@@ -54,4 +54,4 @@ ggplot(data, aes(co2flow, he12C)) +
        y = TeX(r'(^{12}C^- current (μA))')) +
   theme_classic()
 
-ggsave("fig1b_current_flow.svg")
+ggsave(here("doc/AMS-15/AMS-15_paper/figures/fig1b_current_flow.svg"))

### 2022-03-23 doc/AMS-15/AMS-15_paper/figures/fig2_current_dependent_blank.R

@@ -1,18 +1,25 @@
 # Current dependent blank model and data
 # Plot of data and model with measured blanks
-# This should be remade with more data and functions
+# This should be remade with new data reduction
 # Use sample depletion data?
 # Or skip altogether!
+
+library(tidyverse)
+library(amstools)
+library(amsdata)
+library(hgis)
+library(here)
+library(broom)
 
 data<- get_hgis_data(here("data/USAMS101320R.txt"), as.Date("2020-11-17")) %>% 
   mutate(Num = ifelse(Pos %in% c(2, 4), "S", ifelse(Num == "S", "U", Num)))
 stdrat <- mean(data$cor1412he[data$Num == "S" & !data$outlier])
-data1013 <- data %>% mutate(normFm = norm_gas(cor1412he, stdrat))
+data1013 <- data %>% mutate(normFm = norm_run(cor1412he, stdrat))
 
 data <- get_hgis_data(here("data/USAMS120320R.txt"), as.Date("2020-12-04")) %>% 
   mutate(Num = ifelse(Pos %in% c(2, 4), "S", ifelse(Num == "S", "U", Num)))
 stdrat <- mean(data$cor1412he[data$Num == "S" & !data$outlier])
-data1204 <- data %>% mutate(normFm = norm_gas(cor1412he, stdrat))
+data1204 <- data %>% mutate(normFm = norm_run(cor1412he, stdrat))
 
 data1211 <- get_hgis_data(here("data/USAMS120320R.txt")) %>% 
   filter(Pos %in% 1:4 | as.Date(ts) == "2020-12-11")
@@ -21,29 +28,29 @@ data <- get_hgis_data(here("data/USAMS120320R.txt"), as.Date("2020-12-18")) %>%
   filter(Pos != 0) %>% 
   mutate(Num = ifelse(Pos %in% c(2, 4), "S", ifelse(Num == "S", "U", Num)))
 stdrat <- mean(data$cor1412he[data$Num == "S" & !data$outlier])
-data1218 <- data %>% mutate(normFm = norm_gas(cor1412he, stdrat))
+data1218 <- data %>% mutate(normFm = norm_run(cor1412he, stdrat))
 
 data0108 <- get_hgis_data(here("data/USAMS120320R.txt"), as.Date("2021-01-08"))
 
 data <- get_hgis_data(here("data/USAMS020521R.txt")) %>% 
[…]

### 2022-03-23 doc/AMS-15/AMS-15_paper/figures/fig3_curr_independent_blank.R

@@ -6,13 +6,17 @@ library(tidyverse)
 library(here)
 
 # Get data
-carb_data <- read_csv(here("data_analysed/carb_all_results.csv"))
+carb <- readRDS(here("data_analysed/carb_all.rds")) %>% 
+  map_dfr(2) 
 
-# Plot
-carb_data %>% 
+# Filter blanks used for correction
+
+blanks <- carb %>% 
   filter(rec_num %in% c(2138, 83028),
          he12C > 2E-6,
-         sample_type == "B") %>%
+         sample_type == "B")
+# Plot
+blanks %>% 
   mutate(Type = ifelse(rec_num == 2138, "TIRI-F", "IAEA C-1")) %>%
   ggplot(aes(he12C, norm_ratio)) +
   geom_errorbar(aes(ymin = norm_ratio - sig_norm_ratio,
@@ -27,3 +31,5 @@ carb_data %>%
         legend.background = element_rect(fill = "white", color = "black"))
 
 ggsave("fig3_cur_indep_blank.svg")
+
+ggsave(here("doc/AMS-15/AMS-15_paper/figures/fig3_cur_indep_blank.svg"))

### 2022-03-23 doc/AMS-15/AMS-15_paper/figures/fig5_consensus_samples.R

@@ -41,5 +41,5 @@ ggplot(aes(fm_consensus, fm_diff, color = Name)) +
        x = "pMC expected (points jittered for clarity)", 
        y = "pMC difference (HGIS - expected)")
 
-ggsave("fig5_consensus_stds.svg")
+ggsave(here("doc/AMS-15/AMS-15_paper/figures/fig5_consensus_stds.svg"))
 

### 2022-03-23 doc/AMS-15/AMS-15_paper/figures/fig5_consensus_samples.R
Change output name.

@@ -41,5 +41,5 @@ ggplot(aes(fm_consensus, fm_diff, color = Name)) +
        x = "pMC expected (points jittered for clarity)", 
        y = "pMC difference (HGIS - expected)")
 
-ggsave(here("doc/AMS-15/AMS-15_paper/figures/fig5_consensus_stds.svg"))
+ggsave(here("doc/AMS-15/AMS-15_paper/figures/fig5_consensus_samples.svg"))
 

### 2022-03-23 doc/AMS-15/AMS-15_paper/AMS-15_paper.Rmd
Add figures. Update blanks section. Other edits.

@@
 knitr::opts_chunk$set(echo = FALSE)
+```
+
+```{r, message = FALSE}
+library(tidyverse)
+library(hgis)
+library(here)
+
+carb <- readRDS(here("data_analysed/carb_all.rds")) %>% 
+  map_dfr(2)
 ```
@@
 
-![Inlet layout.
-Helium displacement gas is delivered to vial through a double needle, mixed sample and helium flows to an open split, and a glass capillary delivers gas to the ion source.](images/HGIS_open_split_drawing.svg)
-
 ## Sample preparation
-
-Reference: gilson
 
@@
 
-### Measurement
+## Measurement
 
@@
 
-### Data reduction
+## Data reduction
 
@@
 
-## System performance
+## Source performance
+
+```{r}
+carbcur <- carb %>%  
+  map_dfr(2) %>% 
+  compare_consensus() %>% 
+  mutate(c12ua = he12C * 1E6)
+samples <- read_csv(here("data/carb_data.csv")) %>% 
+  select(wheel, pos, mass, he_added)
+currents <- left_join(carbcur, samples, by = c("wheel", "pos")) %>% 
+  filter(wheel != "USAMS020521")
+
+curmean <- mean(currents$c12ua) %>% format(digits = 2) %>% paste("uA")
+cursd <- sd(currents$c12ua)
+```
 
@@
 This is generally comparable with other hybrid sources REFS.
-Typical currents for this study were between 4-10uA???.
+Typical currents for this study were between 4-9uA and average current for all samples was `r curmean`.
 The source typically produces best current at around 5uLs-1 CO2.
@@
 More work is needed to look at whether this reflects variability between samples, but initial test of paired replicates show similar standard deviation for samples that show good stable currents.
-
-Measured value of tank modern gas normalized to solid graphite standards ranges from agreeing with solid OX-I to 7% lower than expected for low current samples.
-Normalizing to tank CO2 is probably the best method to account for variability in hybrid source performance.
 
@@
 Sample size tests.
-
-Figure4: sample size vs current vs precision and accuracy
 
@@
 
-To assess blank contribution and ratio dependence on sample current, vials of live and dead standard gas were measured as the displacement helium flow dilluted the samples, producing diminishing current over time.
+Measurement of radiocarbon-free reference materials shows that the average normalized ratio for these samples is XXX +- YYY (N=XXX).
+We investigated the sources of this method blank as a current-dependent blank and as a current-independent blank.
+
+To assess blank contribution and ratio dependence on sample current, vials of live and dead standard gas were measured as the displacement helium flow diluted the samples, producing diminishing current over time.
 We used these data to develop a current-balance model following methods used to assess mass-dependent or constant contamination blanks in graphite targets (REFS).
@@
 Fits from modern and dead dilution tests should trend towards the fraction modern of the blank, crossing at the inverse of blank current.
-Contaminant current measured directly agrees well with the model assessment, but has a higher fraction modern.
+Based on this assessment, we estimate the ratio of the blank as XXX pMC with a current of XXX nA.
+
+The directly blank current (XXX nA) agrees well with the model assessment, but has a higher fraction modern (XXX pMC).
 This is likely due to a combination of low measurement precision for the direct measurement and high sensitivity of the model to variability in the measured fraction modern and current in the depletion samples (REF).
 This constant contamination blank is consistent with the fact that hybrid cathodes produce a minimum current with or without gas flow to the source.
-Given that this minimum current ranges from 100 to 400, this current likely makes up a large fraction of the method blank.
+Given that this minimum current ranges from 100 nA to 400 nA, this current likely makes up a large fraction of the method blank.
 
@@
 
-Figure2: current dependent blank
-
 #### Current independent blank
 
-While dillution and direct measurement data show evidence for a constant blank with a variable effect depending on sample size, the blank correction may be simplified if blanks, standards, and unknown samples are the same size.
+While dilution and direct measurement data show evidence for a constant blank with a variable effect depending on sample size, the blank correction may be simplified if blanks, standards, and unknown samples are the same size.
 If the fraction modern of the blank is 0.8 modern as estimated above, blank samples will show the most effect of a constant contamination blank with current.
@@
 Fraction modern of measured blanks does decrease with increasing current, but the effect is small, less than Fm 0.01 over the range of currents measured.
-This change with current is roughly equal to the variablility between blanks, so the error due to the use of a current-dependent blank based on samples measured over multiple wheels will likely be larger than a current independent blank correction based on blanks measured on each wheel.
+This change with current is roughly equal to the variability between blanks, so the error due to the use of a current-dependent blank based on samples measured over multiple wheels will likely be larger than a current independent blank correction based on blanks measured on each wheel.
 
@@
 # Fix to use current form of processed data
-carb_data <- map_dfr(list.files(here("data_analysed"), full.names = TRUE), read_csv) %>% 
-  filter(!is.na(rec_num),
-         Cur > 2) %>% 
-  mutate(fm_diff = mean - fm_consensus,
-         fm_sigma = sigma(mean, fm_consensus, merr),
-         fmbc_diff = fm_lbc - fm_consensus,
-         fmbc_sigma = sigma(fm_lbc, fm_consensus, merr))
+blanks <- carb %>% 
+  filter(rec_num %in% c(2138, 83028),
+         he12C > 2E-6,
+         sample_type == "B")
 ```
 
-Mean Fm of blanks: `r carb_data %>% filter(fm_consensus < 0.1 & Cur > 2) %>% pull(mean) %>% mean() %>% format(digits = 3)`
+Mean Fm of blanks: `r mean(blanks$norm_ratio) %>% format(digits = 3)`
+SD of Fm of blanks: `r sd(blanks$norm_ratio) %>% format(digits = 3)`
 
@@
 
+# Figures
+
+![Figure 1.
+Inlet layout.
+Helium displacement gas is delivered to vial through a double needle, mixed sample and helium flows to an open split, and a glass capillary delivers gas to the ion source.](images/HGIS_open_split_drawing.svg)
+
+![Figure 2](figures/fig1a_flow_current_time.svg)
+
+![Figure 2b](figures/fig1b_current_flow.svg)
+
+![Figure 4](figures/fig2_current_dependent_blank.svg)
+
+![Figure 4b](figures/fig3_cur_indep_blank.svg)
+
+![Figure 5](figures/fig5_consensus_samples.svg)
+
+![Figure 6](figures/fig6_core_carbonates.svg)
+
 # Figure captions
 
-Fig.
-1.
+Figure 1.
 Schematic of the open split inlet.
@@
 
-Fig.
-2.
+Figure 2.
 Performance of the NOSAMS MCSNICS hybrid gas ion source.
@@
 
-Fig. 3. Relationship of sample size to C- ion current. There is no strong relationship between size and current for samples larger than 800 ug C.
+Will probably skip this figure:
+Figure 3. Relationship of sample size to C- ion current. There is no strong relationship between size and current for samples larger than 800 ug C.
 
-Fig.
-4.
+Figure 4.
 Assessment of blanks.
@@
 
-Fig 5.
+Figure 5.
 Agreement of standards with graphite consensus values.
@@
 
-Fig 6.
+Figure 6.
 Comparison of Haiti core samples measured as graphite.

### 2022-03-24 doc/AMS-15/AMS-15_paper/AMS-15_paper.Rmd
Edits to introduction and beginning of results.

@@
 Most interfaces to gas-accepting sputter ion sources are a closed inlet design.
-A mixture of sample and helium is introduced to the source through a capillary, and flow is controlled by the pressure in the sample volume and the length and diameter of the capillary.
-Flow may be adjusted by changing the pressure by varying the sample volume.
-We designed a open inlet where a capillary samples gas for delivery to the ion source in a flowing stream that exhausts to the atmosphere.
-This defines the delivery pressure as 1 ATM and therefore flow is controlled by the capillary and is not adjustable.
+A mixture of sample and helium is introduced to the source through a capillary, and flow is controlled by the pressure in the inlet volume and the length and diameter of the capillary.
+Pressure is adjusted by changing either the inlet volume or the amount of gas in the inlet.
+We designed a open inlet where a capillary samples gas for delivery to the ion source in a flowing stream that exhausts to the atmosphere, often called an open split.
+This defines the delivery pressure as 1 ATM and therefore flow is controlled by the capillary parameters and is not easily adjustable during measurement.
 The flow of CO2 to the source may be varied by controlling the ratio of CO2 to helium in the sample stream.
+Changing the ratio of CO2 to helium changes the partial pressure of CO2 and therefore the carbon flow rate to the source.
+
 An open inlet has the advantage of simplicity; flow of CO2 to the source is preset and there are no separate sample transfer and dilution steps.
@@
 
+Our initial work with the NEC source established optimal source parameters for consistent source performance.
+We used tank CO2 connected through a regulator to the ion source to measure source performance over a range of CO2 flows.
+Optimal flow for this source is 7.5uLs-1 CO2, but the source produces near optimal currents over a wide range of flows.
+Adding helium to the gas mixture produced similar currents to pure CO2 at the same CO2 flow rates.
+
 Best currents for the source are around 10uA 12C- or roughly 1/10 of graphite currents.
@@
 
-Source currents are variable by day and by target.
-"Good" currents for a given day vary between 5-10uA, and it's not yet clear what causes this variability.
-Alignment, source condition, and settings of the ionizer and cesium are all factors.
+Source currents are variable by day and by target under similar source, inlet, and sample conditions.
+Capillary condition, source and gas delivery arm alignment, slight variability in target geometry, source condition, and performance of the ionizer and cesium may all be factors.
 Current also varies by target for a given set of gas and source conditions.
@@
 Uncorrected measurements of dead CO2 range from 1% to 5% of modern depending on sample current.
-
-Measurement of a small number of replicates shows that between sample reproducibility is around 1%.
-Samples displaced by helium have slightly higher variability, but more tests are needed to better assess reproducibility.
 

### 2022-03-24 doc/AMS-15/AMS-15_paper/figures/fig6_core_carbonates.R
Working on Bchron chronologies

@@ -7,6 +7,7 @@ library(hgis)
 library(amstools)
 library(scales)
 library(gt)
+library(Bchron)
 
 # Read raw and reduced data as produced by analyse_carbonates.R
 data <- readRDS(here("data_analysed/carb_all.rds"))
@@ -70,22 +71,25 @@ mean_errs <- cr_no %>%
 # Make difference plot of core carbonates
 ggplot(cr_no) +
   geom_hline(yintercept = 0) +
-  #geom_hline(yintercept = mean_diff$fm_diff_mean, color = "blue") +
-  #geom_hline(yintercept = mean_diff$fm_diff_mean + mean_diff$fm_diff_sd, color = "lightblue") +
-  #geom_hline(yintercept = mean_diff$fm_diff_mean - mean_diff$fm_diff_sd, color = "lightblue") +
-  geom_pointrange(aes(depth, fm_diff, color = method, shape = method,
+  # geom_hline(yintercept = mean_diff$fm_diff_mean, color = "blue") +
+  # geom_hline(yintercept = mean_diff$fm_diff_mean + mean_diff$fm_diff_sd, color = "lightblue") +
+  # geom_hline(yintercept = mean_diff$fm_diff_mean - mean_diff$fm_diff_sd, color = "lightblue") +
+  geom_pointrange(aes(fm_mean, fm_diff, color = method, shape = method,
                       ymin = fm_diff - sig_fm_corr, 
                       ymax = fm_diff + sig_fm_corr),
                   position = "jitter") +
   scale_color_manual(values = c("#00b7bd", "#b7bf10")) +
+  scale_x_continuous(breaks = breaks_extended(7),
+                     labels = label_percent(suffix = "",
+                                            accuracy = 1)) +
   scale_y_continuous(breaks = breaks_extended(7),
                      labels = label_percent(suffix = "",
                                             accuracy = 1)) +
   labs(title = "HGIS measurements agree with graphite",
        subtitle = "Sediment core macrofossils measured via HGIS and graphite",
-       x = "Core depth (cm)",
+       x = "Mean of measurements (pMC)",
        y = "HGIS - graphite (pMC)") +
-  theme(legend.position = c(0.80, 0.85),
+  theme(legend.position = c(0.85, 0.15),
         legend.direction = "horizontal",
[…]

### 2022-03-25 doc/AMS-15/AMS-15_paper/figures/fig3_curr_independent_blank.R
rename and combine figures

@@ -1,35 +1,0 @@
-# Shows uncorrected blank Fm's vs current
-# Need to investigate high outliers- bad wheel?
-
-# Libraries
-library(tidyverse)
-library(here)
-
-# Get data
-carb <- readRDS(here("data_analysed/carb_all.rds")) %>% 
-  map_dfr(2) 
-
-# Filter blanks used for correction
-
-blanks <- carb %>% 
-  filter(rec_num %in% c(2138, 83028),
-         he12C > 2E-6,
-         sample_type == "B")
-# Plot
-blanks %>% 
-  mutate(Type = ifelse(rec_num == 2138, "TIRI-F", "IAEA C-1")) %>%
-  ggplot(aes(he12C, norm_ratio)) +
-  geom_errorbar(aes(ymin = norm_ratio - sig_norm_ratio,
-                    ymax = norm_ratio + sig_norm_ratio)) +
-  geom_errorbarh(aes(xmin = he12C - sig_he12C, xmax = he12C + sig_he12C)) +
-  geom_point(aes(color = wheel), size = 3) +
-  labs(title = "Blanks",
-       x = "12C Current (μA)",
-       y = "Fraction modern") +
-  theme_classic() +
-  theme(legend.position = c(0.82, 0.76),
-        legend.background = element_rect(fill = "white", color = "black"))
-
-ggsave("fig3_cur_indep_blank.svg")
-
-ggsave(here("doc/AMS-15/AMS-15_paper/figures/fig3_cur_indep_blank.svg"))

### 2022-03-25 doc/AMS-15/AMS-15_paper/figures/fig2_current_flow.R
Update and combine figures

@@ -7,12 +7,17 @@ library(hgis)
 library(patchwork)
 library(latex2exp)
 
+theme_set(theme_classic())
+
+
 # process data from flow test samples
+
 data <- process_hgis_results(here("data/USAMS040121R.txt"),
                              as.Date("2021-04-02")) %>%
   filter(pos == 5) %>% 
   mutate(time = cumsum(corr_lt) / 60,
          co2flow = concCO2(time, r = 244) * 30) #30ul/min
+
 
 # flow vs time subplot
 flow_time <- data.frame(x = 0:125) %>% 
@@ -22,10 +27,10 @@ flow_time <- data.frame(x = 0:125) %>%
   stat_function(fun = function(x) 30 - (concCO2(x, r = 244) * 30),
                 aes(color = "Helium"), size = 1.5) +
   scale_color_manual("Gas", values = c("#00b7bd", "#b7bf10")) +
-  labs(title = "Gas flows to source",
-       subtitle = "7mL vial, 244μL/min helium, 30μL/min to source",
+  labs(title = "A",
+       #title = "Gas flows to source",
+       #subtitle = "7mL vial, 244μL/min helium, 30μL/min to source",
        y = "Gas flow (μL/min)") +
-  theme_classic() +
   theme(axis.title.x = element_blank(),
         axis.text.x = element_blank(),
         legend.position = c(0.87, 0.65),
@@ -36,22 +41,24 @@ cur_time <- ggplot(data, aes(time, he12C)) +
   geom_smooth(span = .4, se = FALSE, color = "#00b7bd") +
   geom_point(size = 3, color = "#0069b1") +
   xlim(0, 125) +
-  labs(title = "Ion current",
+  labs(title = "B",
+       #title = "Ion current",
        x = "Time (min)",
-       y = "12C current (μA)") +
-  theme_classic()
+       y = "12C current (μA)")
 
-flow_time / cur_time
 
-ggsave(here("doc/AMS-15/AMS-15_paper/figures/fig1a_flow_current_time.svg"))
+# Current vs flow subplot
 
-ggplot(data, aes(co2flow, he12C)) +
-  geom_smooth(span = .3, se = FALSE, color = "#00b7bd") +
-  geom_point(size = 2, color = "#0069b1") +
-  labs(title = bquote('Ion current is stable for a range of CO'[2]~'flow'),
[…]

### 2022-03-25 doc/AMS-15/AMS-15_paper/figures/fig3_blanks.R

@@ -3,6 +3,7 @@
 # This should be remade with new data reduction
 # Use sample depletion data?
 # Or skip altogether!
+# Add current error for fig A
 
 library(tidyverse)
 library(amstools)
@@ -10,6 +11,9 @@ library(amsdata)
 library(hgis)
 library(here)
 library(broom)
+library(patchwork)
+
+theme_set(theme_classic())
 
 data<- get_hgis_data(here("data/USAMS101320R.txt"), as.Date("2020-11-17")) %>% 
   mutate(Num = ifelse(Pos %in% c(2, 4), "S", ifelse(Num == "S", "U", Num)))
@@ -98,22 +102,58 @@ blankfit <- fits %>%
 blank_val <- blankfit %>% 
   select(Fm_blank, m_blank)
 
-ggplot(ds, aes(Cur_inv, mean, color = Gas)) +
+
+# Make plot of fm vs current with model fits
+
+blank_model <- ggplot(ds, aes(Cur_inv, mean, color = Gas)) +
   geom_abline(slope = blankfit$LiveGas_Cur_inv, 
               intercept = blankfit$`LiveGas_(Intercept)`,
               color = "#0069b1") +
   geom_abline(slope = blankfit$DeadGas_Cur_inv, 
               intercept = blankfit$`DeadGas_(Intercept)`,
               color = "#00a9e0") +
-  geom_pointrange(aes(ymin = mean - merr, ymax = mean + merr), size = .3) + 
+  geom_pointrange(aes(ymin = mean - merr, ymax = mean + merr)) + 
   scale_color_manual("Gas", values = c("#b7bf10", "#00a9e0", "#0069b1")) +
   xlim(0, 10) +
-  labs(title = "Modeled blank",
-       subtitle = "Fits cross at Fm and current of blank",
+  labs(title = "A",
+       #subtitle = "Fits cross at Fm and current of blank",
        x = "Inverse 12C Current (μA-1)",
-       y = "Fraction modern") +
-  theme_classic() +
+       y = "F14C") +
   theme(legend.position = "none",
         legend.background = element_rect(fill = "white", color = "black")) 
 
-ggsave(here("doc/AMS-15/AMS-15_paper/figures/fig2_current_dependent_blank.svg"))
+###
+# Direct blank measurements
+###
+
+# Shows uncorrected blank Fm's vs current
[…]

### 2022-03-25 doc/AMS-15/AMS-15_paper/figures/fig4_consensus_samples.R

@@ -6,6 +6,8 @@ library(here)
 library(hgis)
 library(scales)
 
+theme_set(theme_classic())
+
 # Read data. Don't use first carbonate test, smaller samples, or gas standards.
 results_df <- read_csv(here("data_analysed/carb_all_results.csv"))
 
@@ -15,12 +17,12 @@ cons_df <- results_df %>%
   compare_consensus() %>% 
   filter(!str_detect(Name, "Gas"))
 
-
 mean_diff <- mean(cons_df$fm_diff)
 sd_diff <- sd(cons_df$fm_diff)
 
+# Generate plot
 cons_df %>% 
-ggplot(aes(fm_consensus, fm_diff, color = Name)) + 
+ggplot(aes(fm_consensus, fm_diff, color = Name, shape = Name)) + 
   geom_hline(yintercept = 0) +
   geom_hline(yintercept = mean_diff, color = "blue") +
   geom_hline(yintercept = mean_diff + sd_diff, color = "lightblue") +
@@ -30,16 +32,14 @@ ggplot(aes(fm_consensus, fm_diff, color = Name)) +
                       ymax = fm_diff + sig_fm_corr), 
                   position = position_dodge2(width = 0.1), 
                   ) + #size = 0.5) + 
-   scale_color_brewer(palette = "Dark2") +
+  scale_color_brewer(palette = "Dark2") +
   # scale_color_manual(values = c("#00a9e0", "#0069b1", "#00b7bd", "#b7bf10", "#00b7bd")) +
-  scale_x_continuous(labels = label_percent(suffix = "")) +
-  scale_y_continuous(breaks = breaks_extended(7),
-                     labels = label_percent(suffix = "",
-                                            accuracy = 1)) +
-  labs(title = "Agreement of HGIS with graphite measurements",
-       subtitle = "Blank corrected HGIS measurements of consensus carbonate standards",
-       x = "pMC expected (points jittered for clarity)", 
-       y = "pMC difference (HGIS - expected)")
+  scale_y_continuous(breaks = seq(-0.03,0.03,0.01)) +
+  labs(#title = "Agreement of HGIS with graphite measurements",
+       #subtitle = "Blank corrected HGIS measurements of consensus carbonate standards",
+       x = "F14C expected (points jittered for clarity)", 
+       y = "F14C difference (HGIS - expected)")
 
[…]

### 2022-03-25 doc/AMS-15/AMS-15_paper/figures/fig5_core_carbonates.R

@@ -8,6 +8,9 @@ library(amstools)
 library(scales)
 library(gt)
 library(Bchron)
+library(patchwork)
+
+theme_set(theme_classic())
 
 # Read raw and reduced data as produced by analyse_carbonates.R
 data <- readRDS(here("data_analysed/carb_all.rds"))
@@ -68,8 +71,10 @@ mean_errs <- cr_no %>%
   pivot_wider(names_from = method,
               values_from = mean_sig)
 
+
 # Make difference plot of core carbonates
-ggplot(cr_no) +
+
+core_compare <- ggplot(cr_no) +
   geom_hline(yintercept = 0) +
   # geom_hline(yintercept = mean_diff$fm_diff_mean, color = "blue") +
   # geom_hline(yintercept = mean_diff$fm_diff_mean + mean_diff$fm_diff_sd, color = "lightblue") +
@@ -79,21 +84,42 @@ ggplot(cr_no) +
                       ymax = fm_diff + sig_fm_corr),
                   position = "jitter") +
   scale_color_manual(values = c("#00b7bd", "#b7bf10")) +
-  scale_x_continuous(breaks = breaks_extended(7),
-                     labels = label_percent(suffix = "",
-                                            accuracy = 1)) +
-  scale_y_continuous(breaks = breaks_extended(7),
-                     labels = label_percent(suffix = "",
-                                            accuracy = 1)) +
-  labs(title = "HGIS measurements agree with graphite",
-       subtitle = "Sediment core macrofossils measured via HGIS and graphite",
-       x = "Mean of measurements (pMC)",
-       y = "HGIS - graphite (pMC)") +
-  theme(legend.position = c(0.85, 0.15),
-        legend.direction = "horizontal",
+  labs(title = "A",
+       #subtitle = "Sediment core macrofossils measured via HGIS and graphite",
+       x = "Mean of measurements (F14C)",
+       y = "HGIS - graphite (F14C)") +
+  theme(legend.position = "none")
+  # theme(legend.position = c(0.85, 0.15),
+  #       legend.direction = "horizontal",
+  #       legend.background = element_rect(fill = "white", color = "black")) 
+
+
+# Simple age depth plot
+# If this is used, should use calibrated ages
+# or at least bomb correction
+
+age_depth <- cr_no %>% 
+  ggplot(aes(depth, rc_age, color = method, shape = method)) +
[…]

### 2022-03-25 R/HGIS_pendleton_all.Rmd
Change data source, fix plots
update library

@@
   mutate(Fm_diff = fm_corr - fm_consensus,
-         sigma = amstools::sigma(fm_corr, fm_consensus, sig_fm_corr)) %>% 
+         sigma = amsdata::sigma(fm_corr, fm_consensus, sig_fm_corr)) %>% 
   arrange(sample_name) %>% 

### 2022-03-25 doc/AMS-15/AMS-15_paper/AMS-15_paper.Rmd
Update figures

@@
 
-![Figure 2](figures/fig1a_flow_current_time.svg)
+![Figure 2](figures/fig2_current_flow.svg)
 
-![Figure 2b](figures/fig1b_current_flow.svg)
+![Figure 3](figures/fig3_blanks.svg)
 
-![Figure 4](figures/fig2_current_dependent_blank.svg)
+![Figure 4](figures/fig4_consensus_samples.svg)
 
-![Figure 4b](figures/fig3_cur_indep_blank.svg)
-
-![Figure 5](figures/fig5_consensus_samples.svg)
-
-![Figure 6](figures/fig6_core_carbonates.svg)
+![Figure 5](figures/fig5_core_carbonates.svg)
 
@@
 a.
-Modelled flows of helium and CO2 to the ion source during displacement of a pure CO2 vial sample with helium.
+Modeled flows of helium and CO2 to the ion source during displacement of a pure CO2 vial sample with helium.
 b.
-Measured 12C- ion current during the same dillution experiment.
+Measured 12C- ion current during the same dilution experiment.
 c.
-Measured ion current for modelled CO2 flow.
+Measured ion current for modeled CO2 flow.
 The stability of the source over a range of flow produces stable current for up to 40 m of measurement time.
@@
 
-Figure 4.
+Figure 3.
 Assessment of blanks.
 (a) Current dependent blank model and data.
-Fm and current of measurements from dillution tests of live and dead gas are used to generate linear fits which should cross at the Fm and current of the constant current blank.
+Fm and current of measurements from dilution tests of live and dead gas are used to generate linear fits which should cross at the Fm and current of the constant current blank.
 Direct determination of Fm and current of the blank are also shown.
@@
 
-Figure 5.
+Figure 4.
 Agreement of standards with graphite consensus values.
@@
 
-Figure 6.
-Comparison of Haiti core samples measured as graphite.
+Figure 5.
+Comparison of Haiti core samples measured as graphite and via HGIS.
 a.
-Difference between HGIS and graphite measurements of the same materials vs.
-core depth (should this be mean of values?).
+Difference between HGIS and graphite measurements of the same materials.
+Difference between HGIS and graphite measurements is shown on the vertical axis and plotted against the mean of the two measurements.
 Samples are offset on the horizontal axis for clarity.
 Error bars are measurement error for each sample.
+Graphite measurements are included to allow comparison of error estimates.
 b) comparison of age-depth models produced using HGIS and graphite measurements.

### 2022-03-25 doc/AMS-15/AMS-15_paper/AMS-15_paper.Rmd
Add values for blanks.

@@
 - Simon Pendleton
-- Josh Hlavenka
 - Sydney Moser
@@
   html_document: default
+  pdf_document: default
   word_document: default
-  pdf_document: default
 ---
@@
 library(here)
+#library(errors)
+
+#options(errors.notation = "plus-minus")
 
@@
 
+```{r}
+# Fix to use current form of processed data
+blanks <- carb %>% 
+  filter(rec_num %in% c(2138, 83028),
+         he12C > 2E-6,
+         sample_type == "B")
+blankmean <- mean(blanks$norm_ratio) %>% format(digits = 3)
+blanksd <-  sd(blanks$norm_ratio) %>% format(digits = 3)
+
+# errors should produce an "all in one" for this, but it doesn't work with inline code
+# meanblank <- as.errors(mean(blanks$norm_ratio), sd(blanks$norm_ratio))
+nblank <- length(blanks$norm_ratio)
+```
+
 #### Current dependent blank
 
-Measurement of radiocarbon-free reference materials shows that the average normalized ratio for these samples is XXX +- YYY (N=XXX).
+Measurement of radiocarbon-free reference materials shows that the average normalized ratio for these samples is `r blankmean` ± `r blanksd` (N=`r nblank`).
 We investigated the sources of this method blank as a current-dependent blank and as a current-independent blank.
@@
 Fits from modern and dead dilution tests should trend towards the fraction modern of the blank, crossing at the inverse of blank current.
-Based on this assessment, we estimate the ratio of the blank as XXX pMC with a current of XXX nA.
+Based on this assessment, we estimate the ratio of the blank as having F14C of 0.43 with a current of 110 nA.
 
-The directly blank current (XXX nA) agrees well with the model assessment, but has a higher fraction modern (XXX pMC).
+The directly measured blank current (120 nA) agrees well with the model assessment, but has a higher fraction modern (F14C = 0.82).
 This is likely due to a combination of low measurement precision for the direct measurement and high sensitivity of the model to variability in the measured fraction modern and current in the depletion samples (REF).
@@
 $$Fm_s = Fm_m - Fm_{blk}\frac{Fm_{std} - Fm_m}{Fm_{std}}$$
-
-```{r}
-# Fix to use current form of processed data
-blanks <- carb %>% 
-  filter(rec_num %in% c(2138, 83028),
-         he12C > 2E-6,
-         sample_type == "B")
-```
-
-Mean Fm of blanks: `r mean(blanks$norm_ratio) %>% format(digits = 3)`
-SD of Fm of blanks: `r sd(blanks$norm_ratio) %>% format(digits = 3)`
 

### 2022-03-25 doc/AMS-15/AMS-15_paper/figures/fig3_blanks.R

@@ -102,6 +102,12 @@ blankfit <- fits %>%
 blank_val <- blankfit %>% 
   select(Fm_blank, m_blank)
 
+# Directly measured Fm and mass of blank
+blank_meas <- ds %>% 
+  ungroup() %>% 
+  filter(Gas == "Blank") %>% 
+  summarize(mean_cur = mean(Cur),
+            mean_rat = mean(mean))
 
 # Make plot of fm vs current with model fits
 

### 2022-03-25 doc/AMS-15/AMS-15_paper/AMS-15_paper.Rmd
Many edits to text.

@@
 One driver of the cost of AMS analysis is the time and effort required for sample preparation, including reduction of the sample to graphite.
-Analysis of analyte CO2 directly via a gas-accepting AMS ion source eliminates the cost and effort of graphitization and returns results more rapidly.
-The NOSAMS USAMS is an AMS system based on a 3MV Tandetron accelerator.
-A NEC 134-sample gas-capable MCSNICS source.
-Analysis of samples as CO2 simplifies sample preparation, making analysis less expensive and faster.
+Analysis of CO2 directly via a gas-accepting AMS ion source eliminates the cost and effort of graphitization and returns results more rapidly.
+The NOSAMS USAMS is an AMS system based on a 3MV Tandetron accelerator (REF) combined with A NEC 134-sample gas-capable MCSNICS source \@ref(Longworth 2015).
 
@@
 Our method compensates for this by limiting CO2 flow to a range that produces relatively constant C- ion current and stable ratios.
-Salazar and Szidat, Roberts gasbench, other open split methods.
+In gas-source AMS, Roberts et.
+al (2013?) use an open split to introduce sample gas to a ECR ion source and Salazar and Szidat (2021?) designed a continuous flow system that maintains relatively stable pCO2 during acquisition by controlling pressure and venting excess gas to atmosphere.
 
-The gas-capable 134-sample MCSNICS from National Electrostatics Corporation (NEC) works on the same principal as hybrid sources developed by ETH (REF) and HVEE (REF).
-Sample CO2 is introduced onto a titanium frit in an aluminum target using stainless tubing mounted to an arm with a probe that mates with the back of the target.
-A CO~2~ - helium mixture is exposed to a beam of charged cesium atoms, which causes dissociation of CO~2~ and ionization of molecular and atomic species.
-Negative ions are extracted from the target and a magnetic filter picks species with masses corresponding to carbon.
+The gas-capable 134-sample MCSNICS from National Electrostatics Corporation (NEC) works on the same principle as hybrid sources developed by ETH (REF) and HVEE (REF).
+The sample CO~2~ - helium mixture is introduced onto a titanium frit in an aluminum target cathode using stainless tubing mounted to an arm with a probe that mates with the back of the cathode.
+Sample gas is exposed to a beam of charged cesium atoms, which causes dissociation of CO~2~ and ionization of molecular and atomic species.
+Negative ions are extracted from the target and carbon species are selected and interferences filtered by the USAMS Tandetron beamline.
 
@@
 
+The helium displacement flow rate was set to 250ul/min for initial tests and was reduced to 100 ul/min for later work including most reference standard comparisons and the core comparison.
 The flow rate from the open split into the source is defined by 250cm of 50um glass capillary (confirm this, SGE, Trajan, Agilent?), split into two sections by a zero-volume bulkhead connector at the source cage.
 The section inside the cage is jacketed in teflon tubing to help protect against spark damage.
+The total flow rate of gas to the source varies due to the viscosity difference between CO2 and helium and is estimated to be 30 ul/min.
+???
 
@@
 
-```{r}
+```{r, message = FALSE}
 carbcur <- carb %>%  
@@
 
-Our initial work with the NEC source established optimal source parameters for consistent source performance.
+Initial work with the NEC source established optimal source parameters for consistent source performance.
 We used tank CO2 connected through a regulator to the ion source to measure source performance over a range of CO2 flows.
@@
 
-Precision taken as the standard error of the mean of repeated runs of a sample is around 1%.
-Precision is typically a little worse than predicted by counting statistics.
-More work is needed to look at whether this reflects variability between samples, but initial test of paired replicates show similar standard deviation for samples that show good stable currents.
-
 Precision of measurements of modern CO2 has been better than 1% (std error of mean of measurements of a target) for high current samples, with precision dropping off as current decreases below 6uA or so.
@@
 Uncorrected measurements of dead CO2 range from 1% to 5% of modern depending on sample current.
-
-Discuss delivery rates and sample sizes here or in results?
-
-Initial currents of \> 10 uA dropped rapidly, eventually settling to 100-400 nA.
 
@@
 Currents do decrease with time due to dilution, but by supplying higher than optimal CO2 flow to the source initially, stable currents can be maintained for some time.
+
 Fraction of CO2 in sample gas should follow this relationship:
@@
 
+The HGIS method produces higher ratios for radiocarbon dead samples than the graphite method.
 Measurement of radiocarbon-free reference materials shows that the average normalized ratio for these samples is `r blankmean` ± `r blanksd` (N=`r nblank`).
@@
 
+HGIS cathodes produce measurable ion current, even with no gas flow.
+This current is initially > 10 uA and drops exponentially, reaching a baseline of 100-200 nA after 5-10 minutes.
+For the range of currents in this study (3-9 uA), this baseline current represents a 1-6% contribution to the total current at the F14C of the target baseline.
+This can be interpreted as a constant current dependent blank, analogous to a mass dependent "constant contaminant" blank and may be modelled with the same tools (REF).
 To assess blank contribution and ratio dependence on sample current, vials of live and dead standard gas were measured as the displacement helium flow diluted the samples, producing diminishing current over time.

### 2022-03-27 doc/AMS-15/AMS-15_paper/AMS-15_paper.Rmd
Add references

@@
   word_document: default
+bibliography: HGIS_paper.bib
 ---
@@
 Radiocarbon Accelerator Mass Spectrometry (AMS) is a sensitive method for determining the age, provenance, and composition of carbon-containing materials.
-A primary disadvantage of the method is its cost, which often limits the number of samples analyzed.
+A primary disadvantage of the method is its cost, which often limits the number of samples analyzed (REFS).
 This forces researchers to carefully choose samples to be run and limits repeat measurements for error determination.
@@
 Analysis of CO2 directly via a gas-accepting AMS ion source eliminates the cost and effort of graphitization and returns results more rapidly.
-The NOSAMS USAMS is an AMS system based on a 3MV Tandetron accelerator (REF) combined with A NEC 134-sample gas-capable MCSNICS source \@ref(Longworth 2015).
+The NOSAMS USAMS is an AMS system based on a 3MV Tandetron accelerator combined with A NEC 134-sample gas-capable MCSNICS source (@Longworth2015).
 
-Most interfaces to gas-accepting sputter ion sources are a closed inlet design.
+Most interfaces to gas-accepting sputter ion sources are a closed inlet design (@Xu2007 @Uhl2007 @ruff_-line_2010).
 A mixture of sample and helium is introduced to the source through a capillary, and flow is controlled by the pressure in the inlet volume and the length and diameter of the capillary.
@@
 Our method compensates for this by limiting CO2 flow to a range that produces relatively constant C- ion current and stable ratios.
-In gas-source AMS, Roberts et.
-al (2013?) use an open split to introduce sample gas to a ECR ion source and Salazar and Szidat (2021?) designed a continuous flow system that maintains relatively stable pCO2 during acquisition by controlling pressure and venting excess gas to atmosphere.
+In gas-source AMS, @Roberts2013 use an open split to introduce sample gas to a ECR ion source and @salazar_low-cost_2019 designed a continuous flow system that maintains relatively stable pCO2 during acquisition by controlling pressure and venting excess gas to atmosphere.
 
-The gas-capable 134-sample MCSNICS from National Electrostatics Corporation (NEC) works on the same principle as hybrid sources developed by ETH (REF) and HVEE (REF).
+The gas-capable 134-sample MCSNICS from National Electrostatics Corporation (NEC) (@Xu2007) works on the same principle as hybrid sources developed by ETH (@ruff_gas_2007, @ruff2010) and HVEE (@BronkRamsey2004).
 The sample CO~2~ - helium mixture is introduced onto a titanium frit in an aluminum target cathode using stainless tubing mounted to an arm with a probe that mates with the back of the cathode.
@@
 
-Samples are measured on a modified 3MV Tandetron AMS system \@ref(Longworth 2015).
+Samples are measured on a modified 3MV Tandetron AMS system (@Longworth2015).
 The displacement helium is allowed to flow constantly during analysis, and the sample needle is inserted into a helium-filled vial between samples.
@@
 Measurements are corrected for process and measurement blank using a linear correction that assumes that the corrected fraction modern of blank samples should be zero, samples with the fraction modern of the standard need no correction, and applies a proportional correction to samples of intermediate fraction modern.
-(REF) The correction uses measured values of standard gas and blank samples for correction of contaminant carbon and machine blank.
+(@roberts_14c_2019) The correction uses measured values of standard gas and blank samples for correction of contaminant carbon and machine blank.
 We use equation 1 for this correction.
@@
 Sample current did vary, but we believe the linear correction described above produces the most accurate results within this range.
-We discuss this constant contaminant (REF) blank in the results below.
-Data reduction is done using a package written in R (REF) which is published on github (<https://github.com/blongworth/hgis>).
+We discuss this constant contaminant blank in the results below.
+Data reduction is done using a package written in R (@r_core_team_r_2022) which is published on github (<https://github.com/blongworth/hgis>).
 
@@
 We used tank CO2 connected through a regulator to the ion source to measure source performance over a range of CO2 flows.
-Optimal flow for this source is 7.5uLs-1 CO2, but the source produces near optimal currents over a wide range of flows.
+Best currents for the source are around 10uA 12C- or roughly 1/10 of graphite currents.
+Best currents were produced at a CO2 flow rate of 7.5uLs-1 CO2, but the source produces near optimal currents with flows from 7.5- XXX ul/min CO2.
 Adding helium to the gas mixture produced similar currents to pure CO2 at the same CO2 flow rates.
-
-Best currents for the source are around 10uA 12C- or roughly 1/10 of graphite currents.
-This is generally comparable with other hybrid sources REFS.
+This is generally comparable with other hybrid sources @Xu2007, although improved target design and continued optimization of source parameters may improve performance @Salazar2013.
 Typical currents for this study were between 4-9uA and average current for all samples was `r curmean`.
-The source typically produces best current at around 5uLs-1 CO2.
-Literature values give 1.5uLs-1 as optimal, but we've never seen that, and pressures, capillaries and flows in at least one of these papers doesn't make sense.
+The source typically produces best current at around 7.5uLs-1 CO2.
+Several studies have shown best currents at lower CO2 flow rates (@Xu2007, @Ruff2010).
+While efficiency of the source is best at very low flow rates (<1.5ul/min), measuring samples with flow set for optimal current produces more precise results due to higher count rates.
 
@@
 HGIS cathodes produce measurable ion current, even with no gas flow.
-This current is initially > 10 uA and drops exponentially, reaching a baseline of 100-200 nA after 5-10 minutes.
+This current is initially > 10 uA and drops exponentially, reaching a baseline of ~100 nA after 5-10 minutes.
 For the range of currents in this study (3-9 uA), this baseline current represents a 1-6% contribution to the total current at the F14C of the target baseline.
-This can be interpreted as a constant current dependent blank, analogous to a mass dependent "constant contaminant" blank and may be modelled with the same tools (REF).
+This can be interpreted as a constant current dependent blank, analogous to a mass dependent "constant contaminant" blank and may be modeled with the same tools (@haghipour_compound-specific_2019 @hanke_comprehensive_2017 @roberts_14c_2019 @sun_14_2020).
 To assess blank contribution and ratio dependence on sample current, vials of live and dead standard gas were measured as the displacement helium flow diluted the samples, producing diminishing current over time.
@@
 Results from both tests are shown in figure 2.
-The inverse of the sample current is used to linearize the data, allowing assessment with a simple linear model.
+The inverse of the sample current is used to linearize the data, allowing assessment with a simple linear model (@shah_walter_ultra-small_2015).
 Fits from modern and dead dilution tests should trend towards the fraction modern of the blank, crossing at the inverse of blank current.
@@
 The directly measured blank current (120 nA) agrees well with the model assessment, but has a higher fraction modern (F14C = 0.82).
-This is likely due to a combination of low measurement precision for the direct measurement and high sensitivity of the model to variability in the measured fraction modern and current in the depletion samples (REF).
+This is likely due to a combination of low measurement precision for the direct measurement and high sensitivity of the model to variability in the measured fraction modern and current in the depletion samples (@sun_14_2020).
 This constant contamination blank is consistent with the fact that hybrid cathodes produce a minimum current with or without gas flow to the source.
@@
 
-While dilution and direct measurement data show evidence for a constant blank with a variable effect depending on sample size, the blank correction may be simplified if blanks, standards, and unknown samples are the same size.
+While dilution and direct measurement data show evidence for a constant blank with a variable effect depending on sample size, the blank correction may be simplified if blanks, standards, and unknown samples are the same size (@donahue_isotope-ratio_1990 @roberts_14c_2019).
 If the fraction modern of the blank is 0.8 modern as estimated above, blank samples will show the most effect of a constant contamination blank with current.
@@
 $$Fm_s = Fm_m - Fm_{blk}\frac{Fm_{std} - Fm_m}{Fm_{std}}$$
-
-Figure3: justification for large blank
 
@@
 Overall, reference materials run with the hybrid gas method were in good agreement with consensus values.
-Mean difference from consenus for all measurements was -0.04 pMC, and the standard deviation of the difference was 1.10 pMC.
+Mean difference from consensus for all measurements was -0.04 pMC, and the standard deviation of the difference was 1.10 pMC.
 Mean propagated measurement error was 1.13 pMC for these samples.
-The hybrid method has no bias in agregate.
+The hybrid method has no bias in aggregate.
 Method error is larger than graphite, but calculated per-sample error is similar to standard deviation of replicate measurements.
@@
 
-The difference beween reference samples and graphite consensus does not vary systematically by fraction modern, indicating that the method has no significant bias by fraction modern (Figure 6).
+The difference between reference samples and graphite consensus does not vary systematically by fraction modern, indicating that the method has no significant bias by fraction modern (Figure 6).
 
@@
 
-A suite of 15 carbonate macrofossils consisting of shell and coral fragements from a series of coastal marine cores REF were measured as graphite and as CO2 using the HGIS method to further test the method and serve to demonstrate a potential application.
+A suite of 15 carbonate macrofossils consisting of shell and coral fragments from a series of coastal marine cores REF were measured as graphite and as CO2 using the HGIS method to further test the method and serve to demonstrate a potential application.
 coral macrofossils from a subtropical coastal marine sediment core using our gas 
@@
 Age models were developed for both the traditional graphite measurements of the core carbonates and the HGIS measurements (Figure 7).
-We used the Bchron Baesian radiochronology package to develop these models.
+We used the Bchron Baesian radiochronology package (@haslett_simple_2008) to develop these models.
 The age-depth curves produced by the two datasets are very similar, with the largest differences being in the uncertainty of the individual dates.
@@
 The accuracy of age-depth models is often driven by environmental factors and 
-sparse data, rather than the precision of individual dates [3]. The lower time and 
+sparse data, rather than the precision of individual dates (REFS from poster). The lower time and 
 cost of the hybrid method allows analysis of more samples. This may produce more 

### 2022-03-28 doc/AMS-15/AMS-15_paper/AMS-15_paper.Rmd
add reference

@@
 The gas-capable 134-sample MCSNICS from National Electrostatics Corporation (NEC) (@Xu2007) works on the same principle as hybrid sources developed by ETH (@ruff_gas_2007, @ruff2010) and HVEE (@BronkRamsey2004).
-The sample CO~2~ - helium mixture is introduced onto a titanium frit in an aluminum target cathode using stainless tubing mounted to an arm with a probe that mates with the back of the cathode.
+The sample CO~2~ - helium mixture is introduced onto a titanium insert or frit in an aluminum target cathode using stainless tubing mounted to an arm with a probe that mates with the back of the cathode.
 Sample gas is exposed to a beam of charged cesium atoms, which causes dissociation of CO~2~ and ionization of molecular and atomic species.
@@
 The accuracy of age-depth models is often driven by environmental factors and 
-sparse data, rather than the precision of individual dates (REFS from poster). The lower time and 
+sparse data, rather than the precision of individual dates (@blaauw_double_2018 ONE MORE?). The lower time and 
 cost of the hybrid method allows analysis of more samples. This may produce more 
 accurate age models, despite reduced precision of individual measurements.
-
-Figure7: age model comparison graphite HGIS
 

### 2022-03-28 doc/AMS-15/AMS-15_paper/figures/fig5_BChron_model.R
Move Bchron to new file. Add bomb curve to bchron. Use marine curve.

@@ -1,0 +1,132 @@
+
+# Bchron plots
+# separate data and make cal curve for graphite and hgis
+# How to handle bomb dates
+# use IntCal package and assign bomb curve
+
+library(tidyverse)
+library(here)
+library(hgis)
+library(amstools)
+library(scales)
+library(gt)
+library(Bchron)
+library(IntCal)
+library(patchwork)
+
+theme_set(theme_classic())
+
+# Read raw and reduced data as produced by analyse_carbonates.R
+data <- readRDS(here("data_analysed/carb_all.rds"))
+
+# Read sample data
+sample_data <- read_csv(here("data/carb_data.csv"))
+
+# Select relevant reduced data, join with sample data
+results <- map_dfr(data, 2) %>% 
+  filter(wheel %in% c("USAMS052621", "USAMS062921"),
+         pos > 4) %>% 
+  left_join(sample_data, by = c("wheel", "pos")) %>% 
+  mutate(rec_num = ifelse(is.na(rec_num.y), rec_num.x, rec_num.y)) %>% 
+  select(-rec_num.x, -rec_num.y)
+
+names <- results %>% 
+  ungroup() %>% 
+  filter(wheel == "USAMS062921") %>% 
+  select(rec_num, sample_name)
+
+graphite_results <- getRecOS(results$rec_num) %>% 
+  filter(!is.na(reported)) %>% 
+  select(rec_num, fm_corr = f_modern, sig_fm_corr = f_ext_error) %>% 
+  mutate(method = "graphite") %>% 
+  left_join(names, by = "rec_num")
+
+combined_results <- results %>% 
+  filter(str_starts(sample_name, "HATC") | str_starts(sample_name, "\\d")) %>% 
+  mutate(method = "hgis") %>% 
+  select(wheel, sample_name, method, rec_num, fm_corr, sig_fm_corr) %>% 
+  bind_rows(graphite_results) %>% 
+  group_by(rec_num) %>% 
+  mutate(fm_mean = mean(fm_corr),
+         fm_diff_mean = fm_corr - fm_mean) %>% 
+  left_join(rename(names, Name = sample_name), by = "rec_num") %>% 
+  left_join(select(graphite_results, rec_num, fm_corr_gr = fm_corr)) %>% 
+  mutate(fm_diff = fm_corr - fm_corr_gr) %>% 
+  ungroup()
+
+depth <- read_csv(here("data/haiti_depth.csv"))
+
[…]

### 2022-03-28 doc/AMS-15/AMS-15_paper/figures/fig5_core_carbonates.R

@@ -143,57 +143,3 @@ cr_no %>%
              scale_by = 100,
              decimals = 2)
 
-# Bchron plots
-# separate data and make cal curve for graphite and hgis
-# How to handle bomb dates
-# use IntCal package and assign bomb curve
-
-cal_sub <- cr_no %>% 
-  filter(rc_age > 0) %>% 
-  mutate(id = paste(sample_name, "-", method)) %>% 
-  select(id, depth, rc_age, sig_rc_age)
-
-cal_dates <- with(cal_sub,
-  BchronCalibrate(rc_age, sig_rc_age, positions = depth, ids = id))
-plot(cal_dates)
-
-cal_sub_hgis <- cr_no %>% 
-  filter(rc_age > 0,
-         method == "hgis") %>% 
-  select(id = sample_name, depth, rc_age, sig_rc_age) %>% 
-  mutate(curve = "intcal20") %>% 
-  add_row(id = "Top", depth = 0, rc_age = 0, sig_rc_age = 1, curve = "normal")
-
-cal_dates_hgis  <- with(cal_sub_hgis,
-  BchronCalibrate(rc_age, sig_rc_age, positions = depth, ids = id, calCurves = curve))
-plot(cal_dates_hgis)
-
-chron_hgis  <- with(cal_sub_hgis,
-  Bchronology(rc_age, sig_rc_age, positions = depth, ids = id, calCurves = curve))
-plot(chron_hgis) +
-  ggtitle("HGIS chronology")
-summary(chron_hgis, "convergence")
-
-cal_sub_gr <- cr_no %>% 
-  filter(rc_age > 0,
-         method == "graphite") %>% 
-  select(id = sample_name, depth, rc_age, sig_rc_age) %>% 
-  mutate(curve = "intcal20",
-         thickness = 10) %>% 
-  add_row(id = "Top", depth = 0, rc_age = 0, sig_rc_age = 1, curve = "normal", thickness = 10)
-
-cal_dates_gr  <- with(cal_sub_gr,
-  BchronCalibrate(rc_age, sig_rc_age, 
-                  positions = depth, ids = id, 
-                  calCurves = curve))
-plot(cal_dates_gr)
-
-chron_gr  <- with(cal_sub_gr,
-  Bchronology(rc_age, sig_rc_age, 
-              positions = depth, ids = id, 
-              calCurves = curve, positionThicknesses = thickness))
-plot(chron_gr) +
[…]

### 2022-03-29 doc/AMS-15/AMS-15_paper/AMS-15_paper.Rmd
General reorg and edits.

@@
 (@roberts_14c_2019) The correction uses measured values of standard gas and blank samples for correction of contaminant carbon and machine blank.
-We use equation 1 for this correction.
-
-$$ R_s = R_m - \frac{R_b(R_{ox}-R_m)}{R_{ox}} $$
-
-Where $R_s$ is the ratio of the unknown sample, $R_m$ is the measured ratio, $R_b$ is the ratio of the blank, and $R_{ox}$ is the ratio of the normalizing standard.
 The errors in the blanks and standards are propagated with errors of the samples to produce the final error estimate for the sample.
@@
 # Results
-
-All results in supplementary materials
-
-Determination of ideal flows.
-Testing is needed to determine minimum sample flow to avoid atmosphere entering the split. TODO: move flow ratio test to results.
-
-## Method development
 
@@
 We used tank CO2 connected through a regulator to the ion source to measure source performance over a range of CO2 flows.
-Best currents for the source are around 10uA 12C- or roughly 1/10 of graphite currents.
-Best currents were produced at a CO2 flow rate of 7.5uLs-1 CO2, but the source produces near optimal currents with flows from 7.5- XXX ul/min CO2.
+Best currents for the source are around 10uA 12C- or roughly 1/10 of graphite target current.
+Best currents were produced at a CO2 flow rate of 7.5uLs-1 CO2, but the source produces near optimal currents with flows from 7.5-25 ul/min CO2.
 Adding helium to the gas mixture produced similar currents to pure CO2 at the same CO2 flow rates.
@@
 Typical currents for this study were between 4-9uA and average current for all samples was `r curmean`.
-The source typically produces best current at around 7.5uLs-1 CO2.
 Several studies have shown best currents at lower CO2 flow rates (@Xu2007, @Ruff2010).
@@
 Current also varies by target for a given set of gas and source conditions.
-Some targets are clearly "bad", producing currents much lower than other targets for a given set of conditions.
 Capillary clogging can be an issue, especially when sampling from vials with acid.
@@
 
-### Currents
+### Currents and flows
 
 Current vs flow modelling
-
-Figure1: Current vs flow
-
-### Sample size
-
-Sample size tests.
-
-Aimed to fill vial initially
-
-No current degradation down to X. Used Y as optimal mass and backfilled with helium.
-
-First tests of displacing pure CO2 from a vial into the open split using helium are promising.
-Helium flowing at 100uLm-1 produces current from 7mL of CO2 for over 2 hours.
-Currents do decrease with time due to dilution, but by supplying higher than optimal CO2 flow to the source initially, stable currents can be maintained for some time.
 
@@
 
-Setting flow to the source to 10uL/min provides excess CO2 to source until the helium dilution ratio increases to \> 1:1.
-Another approach may be to backfill vials with helium and run at a higher dilution ratio and flow to the source.
-This should produce a smaller change in He:CO2 over the course of a run.
-MICADAS runs at 95% He:CO2, which would mean producing \< 0.5mL CO2 in a 7mL vial and backfilling with helium before displacing to the source.
-Some intermediate dilution may be appropriate.
-Alternatively, vials could be flushed with helium prior to acidification, rather than evacuating, leading to some concentration of CO2 in He at slightly \> 1 ATM.
+Dilution ratio - 1:1 to 1:3 during analysis
 
-Testing long runs of dilution of CO2 with displacing He and diluting He with CO2 show that the setup is still optimized for low flow to the source, even with larger diameter delivery capillaries.
-There appears to be a constriction in the split-source tubing that is keeping flows low.
-Current focus is on the large-sample 7mL method, so this will be diagnosed as time allows.
+First tests of displacing pure CO2 from a vial into the open split using helium are promising.
+Helium flowing at 100uLm-1 produces current from 7mL of CO2 for over 2 hours.
+Currents do decrease with time due to dilution, but by supplying higher than optimal CO2 flow to the source initially, stable currents can be maintained for some time.
 
-Targets have shown "pulses" of higher current.
-Not sure why this is, but I suspect it's either buildup of graphite or other available carbon on the target, or target switching to the "high state" as sometimes happens with solid targets.
+### Sample size
 
-First tests with carbonate in vials were hit-or-miss.
-Problems with capillary clogging and low currents led to poor results for several samples.
-A pair of NOSAMS2 samples produced good currents, results agreed well, but were about 3% higher than expected Fm.
-A second set of carbonates produced better currents (\~5uA) and better results.
-Agreement of replicates varied from 3 to 16 permil.
-Near-modern NOSAMS2 was very close to the consensus value, while C-2 and C-1 were elevated by 2 and 3.6% respectively, indicating that a blank correction is needed.
+Work with currents and flows (above and figure 2) predicts that optimal currents should be produced until less than 2mL CO2 (in ug?) remains in the sample vial.
+Tests of a range of sample sizes support this estimate and show that no reduction in measured current is seen in samples larger than XXX mg C which corresponds to XXX mg CaCO3 sample material.
+To ensure ample sample for optimal measurements, we chose 15 mg as a target sample mass.
+To ensure that samples started at or slightly above atmospheric pressure, we injected clean helium with a syringe to equilibrate pressure in the vials when the reaction with acid was complete.
 
-Two more recent carbonate tests produced stable currents and backgrounds.
-Sample, target blank, and sample blank currents were relatively constant across samples, so a large blank correction can be used to correct results.
-Samples from both runs showed a linear relationship between difference from expected Fm and Fm of the sample, but in one sample set the relationship was positive, and in the other, negative.
-Since the intercept with 0-offset was close to 0 Fm, I think this is a problem with normalization, rather than blank correction.
-Standards run at the end of the day are higher than those run at the beginning, which may give clues as to what's going on.
-
-### Contaminant carbon
+## Contaminant carbon
 
@@
 
-#### Current dependent blank
+### Current dependent blank
 
@@
 
-$$I_{sample} + I_{blank} = I_{meas}$$
-$$Fm_mI_m=Fm_bI_b+Fm_sI_s$$
-
 $$Fm_s = Fm_m + \frac{I_b(Fm_m-Fm_b)}{I_m-I_b}$$
 
-0.1 μA blank current with 10 μA total = 1% blank
-LBC vs constant contaminant/MBC.
-
-#### Current independent blank
+### Current independent blank
 
 While dilution and direct measurement data show evidence for a constant blank with a variable effect depending on sample size, the blank correction may be simplified if blanks, standards, and unknown samples are the same size (@donahue_isotope-ratio_1990 @roberts_14c_2019).
-If the fraction modern of the blank is 0.8 modern as estimated above, blank samples will show the most effect of a constant contamination blank with current.
-Figure 3 shows the fraction modern of measured blanks compared to their mean 12C- ion current.
+If the fraction modern of the blank is 0.8 modern as estimated above, radiocarbon-free samples will show the most effect of a constant contamination blank with current.
+Figure 3B shows the fraction modern of measured blanks compared to their mean 12C- ion current.
 Fraction modern of measured blanks does decrease with increasing current, but the effect is small, less than Fm 0.01 over the range of currents measured.
 This change with current is roughly equal to the variability between blanks, so the error due to the use of a current-dependent blank based on samples measured over multiple wheels will likely be larger than a current independent blank correction based on blanks measured on each wheel.
+Constant current blank is likely the best model, but we simplify to a large blank correction since blanks vary by run, currents are relatively similar between blanks, unknowns and standards, and the simpler model works well within the error of the method.
 
-LBC eqn
+We use equation 1 for this correction.
 
-$$Fm_s = Fm_m - Fm_{blk}\frac{Fm_{std} - Fm_m}{Fm_{std}}$$
+$$ R_{samp} = R_{meas} - \frac{R_{blank}(R_{std}-R_{meas})}{R_{std}} $$
 
-Data for blanks
-
-Constant current blank is likely the best model, but we simplify to a large blank correction since blanks vary by run, currents are relatively similar between blanks, unknowns and standards, and the simpler model works well within the error of the method.
+Where $R_{samp}$ is the ratio of the unknown sample, $R_{meas}$ is the measured ratio, $R_{blank}$ is the ratio of the blank, and $R_{std}$ is the ratio of the normalizing standard.
 
@@
 
-The difference between reference samples and graphite consensus does not vary systematically by fraction modern, indicating that the method has no significant bias by fraction modern (Figure 6).
+The difference between reference samples and graphite consensus does not vary systematically by fraction modern, indicating that the method has no significant bias by fraction modern (Figure 4).
 
@@
 
-## ideas for further development of method
+ideas for further development of method
 
@@
 
-This work is supported by NSF cooperative agreement. We are grateful to C. McIntyre, R. Hansman, G. Salazar, T. Ognibene, T. Broeck.
+This work is supported by NSF cooperative agreement. We are grateful to C. McIntyre, R. Hansman, G. Salazar, T. Ognibene, T. Broek.
 
@@
 b) comparison of age-depth models produced using HGIS and graphite measurements.
+
+# References

### 2022-03-29 doc/AMS-15/AMS-15_paper/figures/fig5_BChron_model.R
Add surface tie point

@@ -86,45 +86,54 @@ cal_sub <- cr_no %>%
   mutate(id = paste(sample_name, "-", method)) %>% 
   select(sample_name, id, depth, rc_age, sig_rc_age, method) %>% 
   mutate(curve = if_else(rc_age > 0, "marine20", "bomb_NH1")) %>% 
-  add_row(id = "Top", depth = 0, rc_age = 0, sig_rc_age = 1, curve = "normal")
+  add_row(id = "Top", sample_name = "Top", depth = 0, rc_age = 0, sig_rc_age = 1, curve = "normal")
 
 cal_dates <- with(cal_sub,
   BchronCalibrate(rc_age, sig_rc_age, positions = depth, ids = id, calCurves = curve))
 plot(cal_dates)
 
 cal_sub_hgis <- cal_sub %>% 
-  filter(method == "hgis")
+  filter(method == "hgis" | is.na(method))
 
 cal_sub_gr <- cal_sub %>% 
-  filter(method == "graphite") 
+  filter(method == "graphite" | is.na(method)) 
 
 # Calibrate hgis ages
-cal_dates_hgis  <- with(cal_sub_hgis,
+cal_dates_hgis <- with(cal_sub_hgis,
   BchronCalibrate(rc_age, sig_rc_age, 
                   positions = depth, ids = sample_name, calCurves = curve))
-plot(cal_dates_hgis)
-
-# Make hgis chronology
-chron_hgis  <- with(cal_sub_hgis,
-                    Bchronology(rc_age, sig_rc_age, positions = depth, 
-                                ids = sample_name, calCurves = curve,
-                                artificialThickness = 0.01))
-plot(chron_hgis) +
-  ggtitle("HGIS chronology")
-summary(chron_hgis, "convergence")
 
 # calibrate graphite dates
 cal_dates_gr  <- with(cal_sub_gr,
   BchronCalibrate(rc_age, sig_rc_age, 
                   positions = depth, ids = sample_name, 
                   calCurves = curve))
-plot(cal_dates_gr)
+
+# make dataframe with cal ages
+cal_bp_gr <- map(cal_dates_gr, 1)
+sig_cal_bp_gr <- map(cal_dates_gr, 2)
+
+
+plot(cal_dates_hgis) +
+  ggtitle("HGIS dates")
+
+# Make hgis chronology
+chron_hgis  <- with(cal_sub_hgis,
+                    Bchronology(rc_age, sig_rc_age, positions = depth, 
+                                ids = sample_name, calCurves = curve,
[…]

### 2022-03-30 doc/AMS-15/AMS-15_paper/AMS-15_paper.Rmd
Editing pass, add references, clean up age depth model.

@@
 
-We have developed a method for measurement of radiocarbon in carbonate minerals as CO~2~ gas via a NEC MCSNICS hybrid sputter gas ion source (HGIS).
+We have developed a method for measuring radiocarbon in carbonate minerals as CO~2~ gas via a NEC MCSNICS hybrid sputter gas ion source (HGIS).
 The method uses helium as a carrier gas to displace CO~2~ from sample vials to an open split, where a glass capillary samples the mixture for delivery directly to the HGIS.
 This method skips the gas transfer and quantification steps used in a closed inlet HGIS system, simplifying sample measurement.
-Samples larger than 10 mg carbonate can currently be measured.
-Results from measurements of consensus standards (TIRI-I, IAEA C2, and an internal modern shell standard), and samples from a marine core (40-115 pMC) show that the method agrees well with traditional AMS measurement of the same samples as graphite, and that the method error is less than 1.5 pMC.
+Samples larger than XXX mg carbonate can be measured.
+Results from measurements of consensus standards (TIRI-I, IAEA C2, and an internal modern shell standard), and samples from a marine core (F^14^C = 0.4 - 1.15) show that the method agrees well with traditional AMS measurement of the same samples as graphite, and that the method error is less than F^14^C 0.015.
 We discuss advantages and disadvantages of continuous flow sample introduction, and the effect of reduced precision on calibrated age-depth models produced using gas-source data.
@@
 Radiocarbon Accelerator Mass Spectrometry (AMS) is a sensitive method for determining the age, provenance, and composition of carbon-containing materials.
-A primary disadvantage of the method is its cost, which often limits the number of samples analyzed (REFS).
+A primary disadvantage of the method is its cost, which often limits the number of samples analyzed (REFS?).
 This forces researchers to carefully choose samples to be run and limits repeat measurements for error determination.
 One driver of the cost of AMS analysis is the time and effort required for sample preparation, including reduction of the sample to graphite.
-Analysis of CO2 directly via a gas-accepting AMS ion source eliminates the cost and effort of graphitization and returns results more rapidly.
-The NOSAMS USAMS is an AMS system based on a 3MV Tandetron accelerator combined with A NEC 134-sample gas-capable MCSNICS source (@Longworth2015).
+Analysis of CO~2~ directly via a gas-accepting AMS ion source eliminates the cost and effort of graphitization and returns results more rapidly.
+The NOSAMS USAMS is an AMS system based on a 3MV Tandetron accelerator combined with a NEC 134-sample gas-capable MCSNICS source (@Longworth2015).
 
-Most interfaces to gas-accepting sputter ion sources are a closed inlet design (@Xu2007 @Uhl2007 @ruff_-line_2010).
+Most interfaces to gas-accepting sputter ion sources are a closed inlet design (@Xu2007 @Uhl2007 @ruff_-line_2010, @Fahrni2013, @Wacker2013, @gottschalk_radiocarbon_2018).
 A mixture of sample and helium is introduced to the source through a capillary, and flow is controlled by the pressure in the inlet volume and the length and diameter of the capillary.
@@
 We designed a open inlet where a capillary samples gas for delivery to the ion source in a flowing stream that exhausts to the atmosphere, often called an open split.
-This defines the delivery pressure as 1 ATM and therefore flow is controlled by the capillary parameters and is not easily adjustable during measurement.
-The flow of CO2 to the source may be varied by controlling the ratio of CO2 to helium in the sample stream.
-Changing the ratio of CO2 to helium changes the partial pressure of CO2 and therefore the carbon flow rate to the source.
+This defines the delivery pressure as 1 ATM, and therefore flow is controlled by the capillary parameters and is not easily adjustable during measurement.
+The flow of CO~2~ to the source may be varied by controlling the ratio of CO~2~ to helium in the sample stream.
+Changing the ratio of CO~2~ to helium changes the partial pressure of CO~2~ and therefore the carbon flow rate to the source.
 
-An open inlet has the advantage of simplicity; flow of CO2 to the source is preset and there are no separate sample transfer and dilution steps.
+An open inlet has the advantage of simplicity; flow of CO~2~ to the source is preset and there are no separate sample transfer and dilution steps.
 This may translate to greater throughput by eliminating the time required for sample transfer, dilution, and setting of delivery pressure.
-A disadvantage of the open split is that the flow of CO2 to the source varies with sample size (CO2:He) and changes over the course of a measurement due to dilution of the sample by displacement helium.
-Our method compensates for this by limiting CO2 flow to a range that produces relatively constant C- ion current and stable ratios.
-In gas-source AMS, @Roberts2013 use an open split to introduce sample gas to a ECR ion source and @salazar_low-cost_2019 designed a continuous flow system that maintains relatively stable pCO2 during acquisition by controlling pressure and venting excess gas to atmosphere.
+Reduced system complexity also minimizes points of failure and eases automation and maintenance.
 
-The gas-capable 134-sample MCSNICS from National Electrostatics Corporation (NEC) (@Xu2007) works on the same principle as hybrid sources developed by ETH (@ruff_gas_2007, @ruff2010) and HVEE (@BronkRamsey2004).
+A disadvantage of the open split is that the flow of CO~2~ to the source varies with sample size (CO~2~:He) and changes over the course of a measurement due to dilution of the sample by displacement helium.
+Our method compensates for this by limiting CO~2~ flow to a range that produces relatively constant C- ion current and stable ratios.
+In gas-source AMS, @Roberts2013 use an open split to introduce sample gas to a ECR ion source and @salazar_low-cost_2019 designed a continuous flow system that maintains relatively stable pCO~2~ during acquisition by controlling pressure and venting excess gas to atmosphere.
+
+The gas-capable 134-sample MCSNICS from National Electrostatics Corporation (NEC) (@Xu2007) works on the same principle as hybrid sources developed by ETH Zürich (@ruff_gas_2007, @ruff2010) and High Voltage Engineering Europa (HVEE) (@BronkRamsey2004).
 The sample CO~2~ - helium mixture is introduced onto a titanium insert or frit in an aluminum target cathode using stainless tubing mounted to an arm with a probe that mates with the back of the cathode.
@@
 
-The open split interface supplies a helium-CO2 mixture to a split where a capillary delivers gas to the source and the rest is vented.
-A flow of helium is used to displace sample CO2 to the delivery capillary in the split.
+The open split interface supplies a helium - CO~2~ mixture to a split where a capillary delivers gas to the source and the rest is vented.
+A flow of helium is used to displace sample CO~2~ to the delivery capillary in the split.
 Our open-split interface is constructed of simple, off-the-shelf parts.
 The core of the design is a dual needle and split housed in a VICI cross.
-Displacement helium flows in through one side of the cross, into the sample vial through 1/16" stainless steel tubing, which forms the outer flow path of the concentric dual needle.
-The mixture of sample CO2 and displacement helium flows up a needle inserted through the cross and stainless tubing which forms the inner flow path of the dual needle.
+Displacement helium flows in through one side of the cross, into the sample vial through 1/16 in.
+stainless steel tubing, which forms the outer flow path of the concentric dual needle.
+The mixture of sample CO~2~ and displacement helium flows up a needle inserted through the cross and stainless tubing which forms the inner flow path of the dual needle.
 The split is made from a VICI Luer adapter fitted to the outflow of the dual needle with the source delivery capillary inserted (Figure 1).
 The split has fairly low volume (< 1 mL).
-The system is currently operated manually, but may be easily mounted on the arm of a gasbench-style autosampler to enable automated operation.
+The system was operated manually during development, but may be easily mounted on the arm of a gasbench-style autosampler to enable automated operation.
 
-The test setup for extracting CO2 from septa-sealed vials is as follows: a double needle is inserted into the vial through the septa.
-One side of the needle is connected to helium which is used to force CO2 from the vial through the other side of the needle.
+The test setup for extracting CO~2~ from septa-sealed vials is as follows: a double needle is inserted into the vial through the septa.
+One side of the needle is connected to helium which is used to force CO~2~ from the vial through the other side of the needle.
 The outflow from the vial enters an open split where a capillary is used to admit sample to the ion source.
-The system is currently set up to run with samples starting at atmospheric pressure.
+The system is set up to run with samples starting at atmospheric pressure.
 
-The helium displacement flow rate was set to 250ul/min for initial tests and was reduced to 100 ul/min for later work including most reference standard comparisons and the core comparison.
-The flow rate from the open split into the source is defined by 250cm of 50um glass capillary (confirm this, SGE, Trajan, Agilent?), split into two sections by a zero-volume bulkhead connector at the source cage.
-The section inside the cage is jacketed in teflon tubing to help protect against spark damage.
-The total flow rate of gas to the source varies due to the viscosity difference between CO2 and helium and is estimated to be 30 ul/min.
+The helium displacement flow rate was set to 250 μl min^-1^ for initial tests, and this was reduced to 100 μl min^-1^ for later work including most reference standard comparisons and the core comparison.
+The flow rate from the open split into the source is defined by 250 cm of 50 μm glass capillary (confirm this, SGE, Trajan, Agilent?), split into two sections by a zero-volume bulkhead connector at the source cage.
+The section inside the cage is jacketed in Teflon tubing to help protect against spark damage.
+The total flow rate of gas to the source varies due to the viscosity difference between CO~2~ and helium, and is estimated to be 30 μl min^-1^.
 ???
@@
 Carbonate samples are weighed into 7 ml septa-sealed Vacutainer vials (Labco \# 768W).
-Sealed vials are evacuated, then 1 ml 85% Phosphoric acid is added.
-The samples are allowed to react at 60C for 4-6 h.
-Sample pressure is adjusted to 1 ATM by allowing the sample to equilibrate with a helium filled syringe connected through the vial septum.
+Sealed vials are evacuated, then 1 ml 85% phosphoric acid is added.
+The samples are allowed to react at 60C for 4-6 hours.
+Sample pressure is adjusted to 1 ATM by allowing the sample to equilibrate with a helium-filled syringe connected through the vial septum.
 Evacuation, acid addition, and heated reaction are performed by the automated gas handling system developed for the CFAMS GIS carbonate system \@ref(Roberts 2013).
@@
 The displacement helium is allowed to flow constantly during analysis, and the sample needle is inserted into a helium-filled vial between samples.
-Measurement of a sample proceeds as follows: a clean sputter target is inserted into the source and the target is sputtered until ten minutes (check this) have passed or blank target current falls below 100nA.
+Measurement of a sample proceeds as follows: a clean sputter target is inserted into the source and the target is sputtered until ten minutes (check this) have passed or blank target current falls below 100 nA.
 The delivery needle is then inserted into the sample vial, which starts sample delivery to the split and source.
-Currents take ~30s to stabilize, at which point the data acquisition is started.
+Currents take ~30 s to stabilize, at which point the data acquisition is started.
 Data is acquired in 10 blocks of 90 s each, for 15 minutes total acquisition time.
 Normalizing standards, blanks, and secondary standards are measured in the same way.
-Total time for a measurement, including target pre-cleaning is 20-25 minutes.
+Total time for a measurement is 20-25 minutes including target pre-cleaning.
 
@@
 After aquisition, data reduction, normalization and blank correction result in a final fraction modern with error.
-First, each 90s block is corrected for fractionation using the AMS <sup>13</sup>C measurement.
-The raw 14/12C ratio divided by the 13/12C ratio squared gives the corrected 14/12C ratio.
+First, each 90 s block is corrected for fractionation using the AMS <sup>13</sup>C measurement.
+The raw ^14/12^C ratio divided by the ^13/12^C ratio squared gives the corrected ^14/12^C ratio.
 Outlier runs, defined as those outside of 1.5 times the interquartile range, are dropped at this point.
@@
 We discuss this constant contaminant blank in the results below.
-Data reduction is done using a package written in R (@r_core_team_r_2022) which is published on github (<https://github.com/blongworth/hgis>).
+Data were reduced using a package written in R (@r_core_team_r_2022) which is published on github (<https://github.com/blongworth/hgis>).
 
@@
 
-curmean <- mean(currents$c12ua) %>% format(digits = 2) %>% paste("uA")
+curmean <- mean(currents$c12ua) %>% format(digits = 2) %>% paste("μA")
 cursd <- sd(currents$c12ua)
@@
 Initial work with the NEC source established optimal source parameters for consistent source performance.
-We used tank CO2 connected through a regulator to the ion source to measure source performance over a range of CO2 flows.
-Best currents for the source are around 10uA 12C- or roughly 1/10 of graphite target current.
-Best currents were produced at a CO2 flow rate of 7.5uLs-1 CO2, but the source produces near optimal currents with flows from 7.5-25 ul/min CO2.
-Adding helium to the gas mixture produced similar currents to pure CO2 at the same CO2 flow rates.
+We used tank CO~2~ connected through a regulator to the ion source to measure source performance over a range of CO~2~ flows.
+Best currents for the source are around 10 μA ^12^C^-^, or roughly 1/10 of graphite target current.
+Best currents were produced at a CO~2~ flow rate of 7.5 μl min^-1^ CO~2~, but the source produces near optimal currents with flows from 7.5-25 μl min^-1^ CO~2~.
+Adding helium to the gas mixture produced similar currents to pure CO~2~ at the same CO~2~ flow rates.
 This is generally comparable with other hybrid sources @Xu2007, although improved target design and continued optimization of source parameters may improve performance @Salazar2013.
-Typical currents for this study were between 4-9uA and average current for all samples was `r curmean`.
-Several studies have shown best currents at lower CO2 flow rates (@Xu2007, @Ruff2010).
-While efficiency of the source is best at very low flow rates (<1.5ul/min), measuring samples with flow set for optimal current produces more precise results due to higher count rates.
+Typical currents for this study were between 4-9 μA and average current for all samples was `r curmean`.
+Several studies have shown best currents at lower CO~2~ flow rates (@Xu2007, @Ruff2010).
+While efficiency of the source is best at very low flow rates (<1.5 μl min^-1^), measuring samples with flow set for optimal current produces more precise results due to higher count rates.
 
 Source currents are variable by day and by target under similar source, inlet, and sample conditions.
-Capillary condition, source and gas delivery arm alignment, slight variability in target geometry, source condition, and performance of the ionizer and cesium may all be factors.
+Factors affecting source current may include capillary condition, source and gas delivery arm alignment, slight variability in target geometry, source condition, and performance of the ionizer and cesium.
 Current also varies by target for a given set of gas and source conditions.
-Capillary clogging can be an issue, especially when sampling from vials with acid.
+Capillary clogging can be an issue, especially when sampling from vials containing acid.
 
-Precision of measurements of modern CO2 has been better than 1% (std error of mean of measurements of a target) for high current samples, with precision dropping off as current decreases below 6uA or so.
+Precision of measurements of modern CO~2~ has been better than 1% (std error of mean of measurements of a target) for high current samples, with precision dropping off as current decreases below 6 μA or so.
 Accuracy is reasonable within the large errors on gas measurements.
-Uncorrected measurements of dead CO2 range from 1% to 5% of modern depending on sample current.
+Uncorrected measurements of dead CO~2~ range from 1% to 5% of modern depending on sample current.
 
@@
 
-Fraction of CO2 in sample gas should follow this relationship:
+Fraction of CO~2~ in sample gas should follow this relationship:
 
@@
 
-Where $C(t)$ is the fraction CO2, $r$ is the rate of helium displacement flow, and $V$ is the initial volume of the CO2-filled vial.
+Where $C(t)$ is the fraction CO~2~, $r$ is the rate of helium displacement flow, and $V$ is the initial volume of the CO~2~-filled vial.
 
@@
 
-First tests of displacing pure CO2 from a vial into the open split using helium are promising.
-Helium flowing at 100uLm-1 produces current from 7mL of CO2 for over 2 hours.
-Currents do decrease with time due to dilution, but by supplying higher than optimal CO2 flow to the source initially, stable currents can be maintained for some time.
+First tests of displacing pure CO~2~ from a vial into the open split using helium are promising.
+Helium flowing at 100 μl min^-1^ produces current from 7 mL of CO~2~ for over 2 hours.
+Currents do decrease with time due to dilution, but by supplying higher than optimal CO~2~ flow to the source initially, stable currents can be maintained for some time.
 
@@
 
-Work with currents and flows (above and figure 2) predicts that optimal currents should be produced until less than 2mL CO2 (in ug?) remains in the sample vial.
-Tests of a range of sample sizes support this estimate and show that no reduction in measured current is seen in samples larger than XXX mg C which corresponds to XXX mg CaCO3 sample material.
+Work with currents and flows (above and Figure 2) predicts that optimal currents should be produced until less than 2 mL CO~2~ (in μg?) remains in the sample vial.
+Tests of a range of sample sizes support this estimate and show that no reduction in measured current is seen in samples larger than XXX mg C which corresponds to XXX mg CaCO~3~ sample material.
 To ensure ample sample for optimal measurements, we chose 15 mg as a target sample mass.
@@
 HGIS cathodes produce measurable ion current, even with no gas flow.
-This current is initially > 10 uA and drops exponentially, reaching a baseline of ~100 nA after 5-10 minutes.
-For the range of currents in this study (3-9 uA), this baseline current represents a 1-6% contribution to the total current at the F14C of the target baseline.
-This can be interpreted as a constant current dependent blank, analogous to a mass dependent "constant contaminant" blank and may be modeled with the same tools (@haghipour_compound-specific_2019 @hanke_comprehensive_2017 @roberts_14c_2019 @sun_14_2020).
+This current is initially > 10 μA and drops exponentially, reaching a baseline of ~100 nA after 5-10 minutes.
+For the range of currents in this study (3-9 μA), this baseline current represents a 1-6% contribution to the total current at the F^14^C of the target baseline.
+This can be interpreted as a constant current dependent blank, analogous to a mass-dependent "constant contaminant" blank and may be modeled with the same tools (@haghipour_compound-specific_2019 @hanke_comprehensive_2017 @roberts_14c_2019 @sun_14_2020).
 To assess blank contribution and ratio dependence on sample current, vials of live and dead standard gas were measured as the displacement helium flow diluted the samples, producing diminishing current over time.
@@
 The constant contamination blank was also assessed directly by measuring pure helium in vials using the same method.
-Results from both tests are shown in figure 2.
+Results from both tests are shown in Figure 3a.
 The inverse of the sample current is used to linearize the data, allowing assessment with a simple linear model (@shah_walter_ultra-small_2015).
 Fits from modern and dead dilution tests should trend towards the fraction modern of the blank, crossing at the inverse of blank current.
-Based on this assessment, we estimate the ratio of the blank as having F14C of 0.43 with a current of 110 nA.
+Based on this assessment, we estimate the ratio of the blank as having F^14^C of 0.43 with a current of 110 nA.
 
-The directly measured blank current (120 nA) agrees well with the model assessment, but has a higher fraction modern (F14C = 0.82).
+The directly measured blank current (120 nA) agrees well with the model assessment, but has a higher fraction modern (F^14^C = 0.82).
 This is likely due to a combination of low measurement precision for the direct measurement and high sensitivity of the model to variability in the measured fraction modern and current in the depletion samples (@sun_14_2020).
@@
 If the fraction modern of the blank is 0.8 modern as estimated above, radiocarbon-free samples will show the most effect of a constant contamination blank with current.
-Figure 3B shows the fraction modern of measured blanks compared to their mean 12C- ion current.
-Fraction modern of measured blanks does decrease with increasing current, but the effect is small, less than Fm 0.01 over the range of currents measured.
+Figure 3B shows the fraction modern of measured blanks compared to their mean ^12^C^-^ ion current.
+Fraction modern of measured blanks does decrease with increasing current, but the effect is small, less than F^14^C 0.01 over the range of currents measured.
 This change with current is roughly equal to the variability between blanks, so the error due to the use of a current-dependent blank based on samples measured over multiple wheels will likely be larger than a current independent blank correction based on blanks measured on each wheel.
@@
 Precision and accuracy of the method were assessed by measuring carbonate standards with consensus values from graphite measurements.
-Summary data for these samples are listed in table 1, and measurement data for all samples is in supplementary information.
+Summary data for these samples are listed in Table 1, and measurement data for all samples is in supplementary information.
 Measurement, data reduction, and blank correction was performed as described above.
@@
 Overall, reference materials run with the hybrid gas method were in good agreement with consensus values.
-Mean difference from consensus for all measurements was -0.04 pMC, and the standard deviation of the difference was 1.10 pMC.
-Mean propagated measurement error was 1.13 pMC for these samples.
+Mean difference from consensus for all measurements was -0.04 pMC, and the standard deviation of the difference was 0.011 F^14^C.
+Mean propagated measurement error was 0.011 F^14^C for these samples.
 The hybrid method has no bias in aggregate.
@@
 
-A suite of 15 carbonate macrofossils consisting of shell and coral fragments from a series of coastal marine cores REF were measured as graphite and as CO2 using the HGIS method to further test the method and serve to demonstrate a potential application.
+A suite of 15 carbonate macrofossils consisting of shell and coral fragments from a series of coastal marine cores REF were measured as graphite and as CO~2~ using the HGIS method to further test the method and serve to demonstrate a potential application.
 coral macrofossils from a subtropical coastal marine sediment core using our gas 
@@
 We found the HGIS and graphite measurements to be in good agreement, with no 
-significant bias (figure 6).
-Mean difference between hybrid and graphite measurements was -0.35 pMC and the standard deviation of this difference was 0.85 pMC.
-Mean per-sample error was 1.02 pMC for HGIS measurements, compared with 0.18 pMC for graphite.
+significant bias (Figure 5a).
+Mean difference between hybrid and graphite measurements was -0.003 F^14^C and the standard deviation of this difference was 0.009 F^14^C.
+Mean per-sample error was 0.010 F^14^C for HGIS measurements, compared with 0.002 F^14^C for graphite.
 
-One HGIS sample vial was noted to be at atmospheric pressure when the 16mg sample should have produced only one half atmosphere of CO2.
-The HGIS measurement for this sample was 5.3 pMC higher than the measurement as graphite.
+One HGIS sample vial was noted to be at atmospheric pressure when the 16 mg sample should have produced only 0.5 atmosphere of CO~2~.
+The HGIS measurement for this sample was 0.053 F^14^C higher than the measurement as graphite.
 We suspect that this sample was improperly sealed and eliminated it from the analysis as an outlier.
@@
 
-Age models were developed for both the traditional graphite measurements of the core carbonates and the HGIS measurements (Figure 7).
+Age models were developed for both the traditional graphite measurements of the core carbonates and the HGIS measurements (Figure 5b).
 We used the Bchron Baesian radiochronology package (@haslett_simple_2008) to develop these models.
@@
 Sample preparation takes about 10 min.
-of person time and 6h of automated processing and reaction time, and 20 min.
+of person time and 6 h of automated processing and reaction time, and 20 min.
 of measurement time per sample.
@@
 a.
-Modeled flows of helium and CO2 to the ion source during displacement of a pure CO2 vial sample with helium.
+Modeled flows of helium and CO~2~ to the ion source during displacement of a pure CO~2~ vial sample with helium.
 b.
@@
 c.
-Measured ion current for modeled CO2 flow.
+Measured ion current for modeled CO~2~ flow.
 The stability of the source over a range of flow produces stable current for up to 40 m of measurement time.
@@
 Will probably skip this figure:
-Figure 3. Relationship of sample size to C- ion current. There is no strong relationship between size and current for samples larger than 800 ug C.
+Figure 3. Relationship of sample size to ^12^C^-^ ion current. There is no strong relationship between size and current for samples larger than 800 μg C.
 
@@
 (a) Current dependent blank model and data.
-Fm and current of measurements from dilution tests of live and dead gas are used to generate linear fits which should cross at the Fm and current of the constant current blank.
-Direct determination of Fm and current of the blank are also shown.
+F^14^C and current of measurements from dilution tests of live and dead gas are used to generate linear fits which should cross at the F^14^C and current of the constant current blank.
+Direct determination of F^14^C and current of the blank are also shown.
 (b) Fraction modern of large, radiocarbon dead samples compared to sample current.
-Blank rises with decreasing current, as expected, but may be treated as a single blank with a simple correction for full-size samples producing greater than 3 uA C- current.
+Blank rises with decreasing current, as expected, but may be treated as a single blank with a simple correction for full-size samples producing greater than 3 μA ^12^C^-^ current.
 

### 2022-03-30 doc/AMS-15/AMS-15_paper/figures/fig5_BChron_model.R

@@ -1,20 +1,21 @@
 
 # Bchron plots
 # separate data and make cal curve for graphite and hgis
-# How to handle bomb dates
-# use IntCal package and assign bomb curve
 
 library(tidyverse)
 library(here)
-library(hgis)
+# library(hgis)
 library(amstools)
-library(scales)
-library(gt)
 library(Bchron)
 library(IntCal)
-library(patchwork)
+# library(patchwork)
 
 theme_set(theme_classic())
+
+
+#####
+# Read and combine data
+#####
 
 # Read raw and reduced data as produced by analyse_carbonates.R
 data <- readRDS(here("data_analysed/carb_all.rds"))
@@ -30,17 +31,20 @@ results <- map_dfr(data, 2) %>%
   mutate(rec_num = ifelse(is.na(rec_num.y), rec_num.x, rec_num.y)) %>% 
   select(-rec_num.x, -rec_num.y)
 
+# Make cross ref table to allow joining graphite and hgis data
 names <- results %>% 
   ungroup() %>% 
   filter(wheel == "USAMS062921") %>% 
   select(rec_num, sample_name)
 
+# Get graphite results
 graphite_results <- getRecOS(results$rec_num) %>% 
   filter(!is.na(reported)) %>% 
   select(rec_num, fm_corr = f_modern, sig_fm_corr = f_ext_error) %>% 
   mutate(method = "graphite") %>% 
   left_join(names, by = "rec_num")
 
+# Join graphite and hgis results
 combined_results <- results %>% 
   filter(str_starts(sample_name, "HATC") | str_starts(sample_name, "\\d")) %>% 
   mutate(method = "hgis") %>% 
@@ -54,19 +58,23 @@ combined_results <- results %>%
   mutate(fm_diff = fm_corr - fm_corr_gr) %>% 
   ungroup()
 
+# read core depth data
 depth <- read_csv(here("data/haiti_depth.csv"))
 
+# Join core depth data to combined results
 combined_results <- left_join(combined_results, depth, by = "rec_num") %>% 
   mutate(rc_age = -8033 * log(fm_corr),
          sig_rc_age = rc_age - -8033 * log(fm_corr + sig_fm_corr))
 write_csv(combined_results, here("data_analysed/haiti_combined.csv"))
 
-# No outliers
+# Remove outlier: HGIS sample leaked
 cr_no <- combined_results %>% 
-  filter(abs(fm_diff_mean) < 0.02)
[…]

### 2022-03-31 doc/AMS-15/AMS-15_paper/AMS-15_paper.Rmd
Fix references and add NIMB csl

@@
 bibliography: HGIS_paper.bib
+csl: elsevier-without-titles.csl
 ---
@@
 Analysis of CO~2~ directly via a gas-accepting AMS ion source eliminates the cost and effort of graphitization and returns results more rapidly.
-The NOSAMS USAMS is an AMS system based on a 3MV Tandetron accelerator combined with a NEC 134-sample gas-capable MCSNICS source (@Longworth2015).
+The NOSAMS USAMS is an AMS system based on a 3MV Tandetron accelerator combined with a NEC 134-sample gas-capable MCSNICS source [@Longworth2015].
 
-Most interfaces to gas-accepting sputter ion sources are a closed inlet design (@Xu2007 @Uhl2007 @ruff_-line_2010, @Fahrni2013, @Wacker2013, @gottschalk_radiocarbon_2018).
+Most interfaces to gas-accepting sputter ion sources are a closed inlet design [@Xu2007; @Uhl2007; @ruff_-line_2010; @Fahrni2013; @wacker_versatile_2013; @gottschalk_radiocarbon_2018].
 A mixture of sample and helium is introduced to the source through a capillary, and flow is controlled by the pressure in the inlet volume and the length and diameter of the capillary.
@@
 Our method compensates for this by limiting CO~2~ flow to a range that produces relatively constant C- ion current and stable ratios.
-In gas-source AMS, @Roberts2013 use an open split to introduce sample gas to a ECR ion source and @salazar_low-cost_2019 designed a continuous flow system that maintains relatively stable pCO~2~ during acquisition by controlling pressure and venting excess gas to atmosphere.
+In gas-source AMS, Roberts et al.
+@Roberts2013 use an open split to introduce sample gas to a ECR ion source and Salazar and Szidat @salazar_low-cost_2019 designed a continuous flow system that maintains relatively stable pCO~2~ during acquisition by controlling pressure and venting excess gas to atmosphere.
 
-The gas-capable 134-sample MCSNICS from National Electrostatics Corporation (NEC) (@Xu2007) works on the same principle as hybrid sources developed by ETH Zürich (@ruff_gas_2007, @ruff2010) and High Voltage Engineering Europa (HVEE) (@BronkRamsey2004).
+The gas-capable 134-sample MCSNICS from National Electrostatics Corporation (NEC) [@Xu2007] works on the same principle as hybrid sources developed by ETH Zürich [@ruff_gas_2007; @Ruff2010] and High Voltage Engineering Europa (HVEE) [@BronkRamsey2004].
 The sample CO~2~ - helium mixture is introduced onto a titanium insert or frit in an aluminum target cathode using stainless tubing mounted to an arm with a probe that mates with the back of the cathode.
@@
 The split is made from a VICI Luer adapter fitted to the outflow of the dual needle with the source delivery capillary inserted (Figure 1).
-The split has fairly low volume (< 1 mL).
+The split has fairly low volume (\< 1 mL).
 The system was operated manually during development, but may be easily mounted on the arm of a gasbench-style autosampler to enable automated operation.
@@
 Sample pressure is adjusted to 1 ATM by allowing the sample to equilibrate with a helium-filled syringe connected through the vial septum.
-Evacuation, acid addition, and heated reaction are performed by the automated gas handling system developed for the CFAMS GIS carbonate system \@ref(Roberts 2013).
+Evacuation, acid addition, and heated reaction are performed by the automated gas handling system developed for the CFAMS GIS carbonate system [@Roberts2013].
 
@@
 
-Samples are measured on a modified 3MV Tandetron AMS system (@Longworth2015).
+Samples are measured on a modified 3MV Tandetron AMS system [@Longworth2015].
 The displacement helium is allowed to flow constantly during analysis, and the sample needle is inserted into a helium-filled vial between samples.
@@
 The delivery needle is then inserted into the sample vial, which starts sample delivery to the split and source.
-Currents take ~30 s to stabilize, at which point the data acquisition is started.
+Currents take \~30 s to stabilize, at which point the data acquisition is started.
 Data is acquired in 10 blocks of 90 s each, for 15 minutes total acquisition time.
@@
 Measurements are corrected for process and measurement blank using a linear correction that assumes that the corrected fraction modern of blank samples should be zero, samples with the fraction modern of the standard need no correction, and applies a proportional correction to samples of intermediate fraction modern.
-(@roberts_14c_2019) The correction uses measured values of standard gas and blank samples for correction of contaminant carbon and machine blank.
+[@roberts_14c_2019] The correction uses measured values of standard gas and blank samples for correction of contaminant carbon and machine blank.
 The errors in the blanks and standards are propagated with errors of the samples to produce the final error estimate for the sample.
@@
 We discuss this constant contaminant blank in the results below.
-Data were reduced using a package written in R (@r_core_team_r_2022) which is published on github (<https://github.com/blongworth/hgis>).
+Data were reduced using a package written in R [@r_core_team_r_2022] which is published on github (<https://github.com/blongworth/hgis>).
 
@@
 Adding helium to the gas mixture produced similar currents to pure CO~2~ at the same CO~2~ flow rates.
-This is generally comparable with other hybrid sources @Xu2007, although improved target design and continued optimization of source parameters may improve performance @Salazar2013.
+This is generally comparable with other hybrid sources [@Xu2007], although improved target design and continued optimization of source parameters may improve performance [@Salazar2013; @Wacker2013].
 Typical currents for this study were between 4-9 μA and average current for all samples was `r curmean`.
-Several studies have shown best currents at lower CO~2~ flow rates (@Xu2007, @Ruff2010).
-While efficiency of the source is best at very low flow rates (<1.5 μl min^-1^), measuring samples with flow set for optimal current produces more precise results due to higher count rates.
+Several studies have shown best currents at lower CO~2~ flow rates [@Xu2007; @Ruff2010].
+While efficiency of the source is best at very low flow rates (\<1.5 μl min^-1^), measuring samples with flow set for optimal current produces more precise results due to higher count rates.
 
@@
 HGIS cathodes produce measurable ion current, even with no gas flow.
-This current is initially > 10 μA and drops exponentially, reaching a baseline of ~100 nA after 5-10 minutes.
+This current is initially \> 10 μA and drops exponentially, reaching a baseline of \~100 nA after 5-10 minutes.
 For the range of currents in this study (3-9 μA), this baseline current represents a 1-6% contribution to the total current at the F^14^C of the target baseline.
-This can be interpreted as a constant current dependent blank, analogous to a mass-dependent "constant contaminant" blank and may be modeled with the same tools (@haghipour_compound-specific_2019 @hanke_comprehensive_2017 @roberts_14c_2019 @sun_14_2020).
+This can be interpreted as a constant current dependent blank, analogous to a mass-dependent "constant contaminant" blank and may be modeled with the same tools [@haghipour_compound-specific_2019; @hanke_comprehensive_2017; @roberts_14c_2019; @sun_14_2020].
 To assess blank contribution and ratio dependence on sample current, vials of live and dead standard gas were measured as the displacement helium flow diluted the samples, producing diminishing current over time.
@@
 Results from both tests are shown in Figure 3a.
-The inverse of the sample current is used to linearize the data, allowing assessment with a simple linear model (@shah_walter_ultra-small_2015).
+The inverse of the sample current is used to linearize the data, allowing assessment with a simple linear model [@shah_walter_ultra-small_2015].
 Fits from modern and dead dilution tests should trend towards the fraction modern of the blank, crossing at the inverse of blank current.
@@
 The directly measured blank current (120 nA) agrees well with the model assessment, but has a higher fraction modern (F^14^C = 0.82).
-This is likely due to a combination of low measurement precision for the direct measurement and high sensitivity of the model to variability in the measured fraction modern and current in the depletion samples (@sun_14_2020).
+This is likely due to a combination of low measurement precision for the direct measurement and high sensitivity of the model to variability in the measured fraction modern and current in the depletion samples [@sun_14_2020].
 This constant contamination blank is consistent with the fact that hybrid cathodes produce a minimum current with or without gas flow to the source.
@@
 
-While dilution and direct measurement data show evidence for a constant blank with a variable effect depending on sample size, the blank correction may be simplified if blanks, standards, and unknown samples are the same size (@donahue_isotope-ratio_1990 @roberts_14c_2019).
+While dilution and direct measurement data show evidence for a constant blank with a variable effect depending on sample size, the blank correction may be simplified if blanks, standards, and unknown samples are the same size [@donahue_isotope-ratio_1990; @roberts_14c_2019].
 If the fraction modern of the blank is 0.8 modern as estimated above, radiocarbon-free samples will show the most effect of a constant contamination blank with current.
@@
 A suite of 15 carbonate macrofossils consisting of shell and coral fragments from a series of coastal marine cores REF were measured as graphite and as CO~2~ using the HGIS method to further test the method and serve to demonstrate a potential application.
-coral macrofossils from a subtropical coastal marine sediment core using our gas 
-method and compared to the same samples measured using the traditional 
-graphite method.
+coral macrofossils from a subtropical coastal marine sediment core using our gas method and compared to the same samples measured using the traditional graphite method.
 
-To further test the method and a potential application, we measured 15 shell and 
-coral macrofossils from a subtropical coastal marine sediment core using our gas 
-method and compared to the same samples measured using the traditional 
-graphite method.
+To further test the method and a potential application, we measured 15 shell and coral macrofossils from a subtropical coastal marine sediment core using our gas method and compared to the same samples measured using the traditional graphite method.
 
-We found the HGIS and graphite measurements to be in good agreement, with no 
-significant bias (Figure 5a).
+We found the HGIS and graphite measurements to be in good agreement, with no significant bias (Figure 5a).
 Mean difference between hybrid and graphite measurements was -0.003 F^14^C and the standard deviation of this difference was 0.009 F^14^C.
@@
 Age models were developed for both the traditional graphite measurements of the core carbonates and the HGIS measurements (Figure 5b).
-We used the Bchron Baesian radiochronology package (@haslett_simple_2008) to develop these models.
+We used the Bchron Baesian radiochronology package [@haslett_simple_2008] to develop these models.
 The age-depth curves produced by the two datasets are very similar, with the largest differences being in the uncertainty of the individual dates.
@@
 
-The accuracy of age-depth models is often driven by environmental factors and 
-sparse data, rather than the precision of individual dates (@blaauw_double_2018 ONE MORE?). The lower time and 
-cost of the hybrid method allows analysis of more samples. This may produce more 
-accurate age models, despite reduced precision of individual measurements.
+The accuracy of age-depth models is often driven by environmental factors and sparse data, rather than the precision of individual dates [@blaauw_double_2018] ONE MORE?.
+The lower time and cost of the hybrid method allows analysis of more samples.
+This may produce more accurate age models, despite reduced precision of individual measurements.
 
@@
 
-The accuracy of age-depth models is often driven by environmental factors and 
-sparse data, rather than the precision of individual dates [3]. The lower time and 
-cost of the hybrid method allows analysis of more samples. This may produce more 
-accurate age models, despite reduced precision of individual measurements.
+The accuracy of age-depth models is often driven by environmental factors and sparse data, rather than the precision of individual dates \[3\].
+The lower time and cost of the hybrid method allows analysis of more samples.
+This may produce more accurate age models, despite reduced precision of individual measurements.
 
@@
 
-Will probably skip this figure:
-Figure 3. Relationship of sample size to ^12^C^-^ ion current. There is no strong relationship between size and current for samples larger than 800 μg C.
+Will probably skip this figure: Figure 3. Relationship of sample size to ^12^C^-^ ion current. There is no strong relationship between size and current for samples larger than 800 μg C.
 

### 2022-03-31 doc/AMS-15/AMS-15_paper/AMS-15_paper.Rmd
Fix flow calcs

@@
 The split is made from a VICI Luer adapter fitted to the outflow of the dual needle with the source delivery capillary inserted (Figure 1).
-The split has fairly low volume (\< 1 mL).
+The split has fairly low volume (< 1 mL).
 The system was operated manually during development, but may be easily mounted on the arm of a gasbench-style autosampler to enable automated operation.
@@
 
-The helium displacement flow rate was set to 250 μl min^-1^ for initial tests, and this was reduced to 100 μl min^-1^ for later work including most reference standard comparisons and the core comparison.
-The flow rate from the open split into the source is defined by 250 cm of 50 μm glass capillary (confirm this, SGE, Trajan, Agilent?), split into two sections by a zero-volume bulkhead connector at the source cage.
+```{r}
+# Flow calc for CO2
+co2flow <- flowcalc(100, 2.5E-5, 1.47E-5, 1.50)
+
+# Flow calc for helium
+heflow <- flowcalc(101, 2.5E-5, 1.96E-5, 1.50)
+```
+
+The helium displacement flow rate was set to 250 μl min^-1^ for initial tests, and this was reduced to 100 μl min^-1^ for later work, including most reference standard comparisons and the core comparison.
+The flow rate from the open split into the source is defined by 1.5 m of 50 μm glass capillary (confirm this, SGE, Trajan, Agilent?), split into two sections by a zero-volume bulkhead connector at the source cage.
 The section inside the cage is jacketed in Teflon tubing to help protect against spark damage.
-The total flow rate of gas to the source varies due to the viscosity difference between CO~2~ and helium, and is estimated to be 30 μl min^-1^.
-???
+The total flow rate of gas to the source varies due to the viscosity difference between helium and CO~2~ and helium, and is estimated to be between 30 μl min^-1^ for pure helium and 40 μl min^-1^ for pure CO~2~ using the Hagen-Poiseuille equation.
 
@@
 
-Carbonate samples are weighed into 7 ml septa-sealed Vacutainer vials (Labco \# 768W).
+Carbonate samples are weighed into 7 ml septa-sealed Vacutainer vials (Labco 768W).
 Sealed vials are evacuated, then 1 ml 85% phosphoric acid is added.
@@
 Adding helium to the gas mixture produced similar currents to pure CO~2~ at the same CO~2~ flow rates.
-This is generally comparable with other hybrid sources [@Xu2007], although improved target design and continued optimization of source parameters may improve performance [@Salazar2013; @Wacker2013].
+This is generally comparable with other hybrid sources [@Xu2007], although improved target design and continued optimization of source parameters may improve performance [@Salazar2013; @wacker_versatile_2013].
 Typical currents for this study were between 4-9 μA and average current for all samples was `r curmean`.

### 2022-04-01 doc/AMS-15/AMS-15_paper/AMS-15_paper.Rmd
Finish current and flow section. Sundry edits. Convert table to F14C.

@@
 output:
+  word_document: default
+  pdf_document: default
   html_document: default
-  pdf_document: default
-  word_document: default
 bibliography: HGIS_paper.bib
@@
 This method skips the gas transfer and quantification steps used in a closed inlet HGIS system, simplifying sample measurement.
-Samples larger than XXX mg carbonate can be measured.
+Samples larger than 8 mg carbonate can be measured.
 Results from measurements of consensus standards (TIRI-I, IAEA C2, and an internal modern shell standard), and samples from a marine core (F^14^C = 0.4 - 1.15) show that the method agrees well with traditional AMS measurement of the same samples as graphite, and that the method error is less than F^14^C 0.015.
@@
 Radiocarbon Accelerator Mass Spectrometry (AMS) is a sensitive method for determining the age, provenance, and composition of carbon-containing materials.
-A primary disadvantage of the method is its cost, which often limits the number of samples analyzed (REFS?).
+A primary disadvantage of the method is its cost, which often limits the number of samples analyzed.
 This forces researchers to carefully choose samples to be run and limits repeat measurements for error determination.
@@
 Analysis of CO~2~ directly via a gas-accepting AMS ion source eliminates the cost and effort of graphitization and returns results more rapidly.
-The NOSAMS USAMS is an AMS system based on a 3MV Tandetron accelerator combined with a NEC 134-sample gas-capable MCSNICS source [@Longworth2015].
+The NOSAMS USAMS system at the National Oceanographic AMS lab (NOSAMS) is based on a 3MV Tandetron accelerator [@Longworth2015] combined with a NEC 134-sample gas-capable MCSNICS source [@Xu2007] .
+We present here a system for semi-automated analysis of carbonate minerals by acid hydrolysis in vials and introduction to the ion source in a flow of helium via an open split.
 
-Most interfaces to gas-accepting sputter ion sources are a closed inlet design [@Xu2007; @Uhl2007; @ruff_-line_2010; @Fahrni2013; @wacker_versatile_2013; @gottschalk_radiocarbon_2018].
+Most interfaces to gas-accepting sputter ion sources are a closed inlet design [@Xu2007; @Uhl2007; @ruff_gas_2007; \@ruff\_-line_2010; @Fahrni2013; @wacker_versatile_2013; @gottschalk_radiocarbon_2018].
 A mixture of sample and helium is introduced to the source through a capillary, and flow is controlled by the pressure in the inlet volume and the length and diameter of the capillary.
@@
 
-The gas-capable 134-sample MCSNICS from National Electrostatics Corporation (NEC) [@Xu2007] works on the same principle as hybrid sources developed by ETH Zürich [@ruff_gas_2007; @Ruff2010] and High Voltage Engineering Europa (HVEE) [@BronkRamsey2004].
+The gas-capable 134-sample MCSNICS from National Electrostatics Corporation (NEC) [@Xu2007\] works on the same principle as hybrid sources developed by ETH Zürich \[@ruff_gas_2007; @Ruff2010] and High Voltage Engineering Europa (HVEE) [@BronkRamsey2004].
 The sample CO~2~ - helium mixture is introduced onto a titanium insert or frit in an aluminum target cathode using stainless tubing mounted to an arm with a probe that mates with the back of the cathode.
 Sample gas is exposed to a beam of charged cesium atoms, which causes dissociation of CO~2~ and ionization of molecular and atomic species.
-Negative ions are extracted from the target and carbon species are selected and interferences filtered by the USAMS Tandetron beamline.
+Negative ions are extracted from the target, carbon species are selected, and interfering masses filtered out by the USAMS Tandetron beamline.
 
@@
 
-The open split interface supplies a helium - CO~2~ mixture to a split where a capillary delivers gas to the source and the rest is vented.
-A flow of helium is used to displace sample CO~2~ to the delivery capillary in the split.
+The open split interface (Figure 1) supplies a helium - CO~2~ mixture to a split where a capillary delivers gas to the source and the rest is vented.
+Flowing helium is used to displace sample CO~2~ from the sample vial to the delivery capillary in the split.
 Our open-split interface is constructed of simple, off-the-shelf parts.
-The core of the design is a dual needle and split housed in a VICI cross.
-Displacement helium flows in through one side of the cross, into the sample vial through 1/16 in.
+The core of the design is a dual needle and split housed in a VICI cross fitting.
+Displacement helium flows in through one side of the cross and into the sample vial through 1/16 in.
 stainless steel tubing, which forms the outer flow path of the concentric dual needle.
-The mixture of sample CO~2~ and displacement helium flows up a needle inserted through the cross and stainless tubing which forms the inner flow path of the dual needle.
-The split is made from a VICI Luer adapter fitted to the outflow of the dual needle with the source delivery capillary inserted (Figure 1).
-The split has fairly low volume (< 1 mL).
+
+The mixture of sample CO~2~ and displacement helium from the vial flows up a 21 gauge (0.819 mm) needle inserted through the cross and stainless tubing which forms the inner flow path of the dual needle.
+The open split is a VICI Luer adapter (part ZBUMLPK) with 5cm of 1/16 in.
+PFTE tubing attached.
+The split is connected to the top of the inner needle, and the source delivery capillary is inserted into the split.
+The split has fairly low volume (\< 1 mL).
 The system was operated manually during development, but may be easily mounted on the arm of a gasbench-style autosampler to enable automated operation.
-
-The test setup for extracting CO~2~ from septa-sealed vials is as follows: a double needle is inserted into the vial through the septa.
-One side of the needle is connected to helium which is used to force CO~2~ from the vial through the other side of the needle.
-The outflow from the vial enters an open split where a capillary is used to admit sample to the ion source.
-The system is set up to run with samples starting at atmospheric pressure.
 
@@
 The helium displacement flow rate was set to 250 μl min^-1^ for initial tests, and this was reduced to 100 μl min^-1^ for later work, including most reference standard comparisons and the core comparison.
-The flow rate from the open split into the source is defined by 1.5 m of 50 μm glass capillary (confirm this, SGE, Trajan, Agilent?), split into two sections by a zero-volume bulkhead connector at the source cage.
-The section inside the cage is jacketed in Teflon tubing to help protect against spark damage.
-The total flow rate of gas to the source varies due to the viscosity difference between helium and CO~2~ and helium, and is estimated to be between 30 μl min^-1^ for pure helium and 40 μl min^-1^ for pure CO~2~ using the Hagen-Poiseuille equation.
+The flow rate from the open split into the source was defined by 150 cm of 50 μm fused silica capillary, split into two sections by a zero-volume bulkhead connector at the source high-voltage cage.
+The section inside the cage was jacketed in 1/16 in.
+PFTE tubing to help protect against spark damage.
+The total flow rate of gas to the source varied due to the viscosity difference between helium and CO~2~ and helium, and was estimated to be between 30 μl min^-1^ for pure helium and 40 μl min^-1^ for pure CO~2~ using the Hagen-Poiseuille equation.
 
@@
 
-Carbonate samples are weighed into 7 ml septa-sealed Vacutainer vials (Labco 768W).
-Sealed vials are evacuated, then 1 ml 85% phosphoric acid is added.
-The samples are allowed to react at 60C for 4-6 hours.
-Sample pressure is adjusted to 1 ATM by allowing the sample to equilibrate with a helium-filled syringe connected through the vial septum.
-Evacuation, acid addition, and heated reaction are performed by the automated gas handling system developed for the CFAMS GIS carbonate system [@Roberts2013].
+Carbonate samples were weighed into 7 ml septa-sealed Vacutainer vials (Labco 768W).
+Sealed vials were evacuated, then 1 ml 85% phosphoric acid was added.
+The samples were allowed to react at 60 C for 4-6 h.
+Sample pressure was adjusted to 1 atm by allowing the sample to equilibrate with a helium-filled syringe connected through the vial septum.
+Evacuation, acid addition, and heated reaction were performed by the automated gas handling system developed for the CFAMS GIS carbonate system [@Roberts2013].
 
-Standards are prepared by evacuating vials and then filling with 1 ATM of a modern standard gas using the automated gas handling system.
-Process blank samples are prepared from powdered C-1 or TIRI-F, acidified in vials using the method above.
+Standards were prepared by evacuating vials and then filling with a modern standard CO~2~ using the automated gas handling system.
+Standards were adjusted to the same size as other samples by removing CO~2~ and replacing with helium using a syringe.
+Process blank samples were prepared from powdered C-1 or TIRI-F, acidified in vials using the method above.
 
@@
 
-Samples are measured on a modified 3MV Tandetron AMS system [@Longworth2015].
-The displacement helium is allowed to flow constantly during analysis, and the sample needle is inserted into a helium-filled vial between samples.
-Measurement of a sample proceeds as follows: a clean sputter target is inserted into the source and the target is sputtered until ten minutes (check this) have passed or blank target current falls below 100 nA.
-The delivery needle is then inserted into the sample vial, which starts sample delivery to the split and source.
-Currents take \~30 s to stabilize, at which point the data acquisition is started.
-Data is acquired in 10 blocks of 90 s each, for 15 minutes total acquisition time.
-Normalizing standards, blanks, and secondary standards are measured in the same way.
-Total time for a measurement is 20-25 minutes including target pre-cleaning.
+Samples were measured on a modified 3MV Tandetron AMS system [@Longworth2015].
+The displacement helium was allowed to flow constantly during analysis, and the sample needle was inserted into a helium-filled vial between samples.
+Measurement of a sample proceeded as follows: a clean sputter target cathode was rotated into position and the target was pre-cleaned by sputtering for 5 minutes or until cathode ^12^C^-^ current fell below 100 nA.
+The delivery needle was then transferred to the sample vial, which started sample delivery to the split and source.
+Currents were allowed to stabilize for 30 s , at which point the data acquisition was started.
+Data was acquired in 10 blocks of 90 s each, for 15 min total acquisition time.
+Normalizing standards, blanks, and secondary standards were measured identically to unknowns.
+Total time for a measurement was 20-25 min including target pre-cleaning and sample changes.
 
@@
 
-After aquisition, data reduction, normalization and blank correction result in a final fraction modern with error.
-First, each 90 s block is corrected for fractionation using the AMS <sup>13</sup>C measurement.
+After acquisition, data reduction, normalization and blank correction result in a final fraction modern with error.
+First, each 90 s block was corrected for fractionation using the AMS ^13/12^C measurement.
 The raw ^14/12^C ratio divided by the ^13/12^C ratio squared gives the corrected ^14/12^C ratio.
-Outlier runs, defined as those outside of 1.5 times the interquartile range, are dropped at this point.
-Remaining runs of a sample are averaged and per-sample error is taken as the greater of counting error or the standard error of the measurements.
-Ratios are then normalized using the mean of gas standards measured during the run.
-The measurement error of the standards is taken as the greater of the standard error of all standards or the errors of the standards added in quadrature.
+Blocks that are more than 1.5 times the interquartile range (IQR) below the first quartile of all measurements of a sample or more than 1.5 IQR above the third quartile are considered outliers and were dropped at this point.
+Remaining runs of a sample were averaged and per-sample error was taken as the greater of counting error or the standard error of the measurements.
+The sample ratio was then normalized using the mean of gas standards measured during the run.
+The measurement error of the standards was taken as the greater of the standard error of all standards or the errors of the standards added in quadrature.
 
-Measurements are corrected for process and measurement blank using a linear correction that assumes that the corrected fraction modern of blank samples should be zero, samples with the fraction modern of the standard need no correction, and applies a proportional correction to samples of intermediate fraction modern.
-[@roberts_14c_2019] The correction uses measured values of standard gas and blank samples for correction of contaminant carbon and machine blank.
-The errors in the blanks and standards are propagated with errors of the samples to produce the final error estimate for the sample.
+Measurements were corrected for process and measurement blank using a linear correction that assumes that the corrected fraction modern of blank samples should be zero, samples with the fraction modern of the standard need no correction, and applies a proportional correction to samples of intermediate fraction modern \[@roberts_14c_2019\].
+The correction uses measured values of standard gas and blank samples for correction of contaminant carbon and machine blank.
+The errors in the blanks and standards were propagated with errors of the samples to produce the final error estimate for the sample.
 This correction does not account for the effect of a blank with constant mass or current on samples of varying mass or current.
@@
 Initial work with the NEC source established optimal source parameters for consistent source performance.
-We used tank CO~2~ connected through a regulator to the ion source to measure source performance over a range of CO~2~ flows.
-Best currents for the source are around 10 μA ^12^C^-^, or roughly 1/10 of graphite target current.
-Best currents were produced at a CO~2~ flow rate of 7.5 μl min^-1^ CO~2~, but the source produces near optimal currents with flows from 7.5-25 μl min^-1^ CO~2~.
-Adding helium to the gas mixture produced similar currents to pure CO~2~ at the same CO~2~ flow rates.
-This is generally comparable with other hybrid sources [@Xu2007], although improved target design and continued optimization of source parameters may improve performance [@Salazar2013; @wacker_versatile_2013].
+We used tank CO~2~ connected through a regulator and glass capillary to the ion source to measure source performance over a range of CO~2~ flows.
+Best currents for the source was 10 μA ^12^C^-^, or roughly 1/10 of graphite target current.
+Peak current was produced at a CO~2~ flow rate of 7.5 μl min^-1^ CO~2~, but the source produced near optimal currents with flows from 7.5-25 μl min^-1^ CO~2~.
+This is comparable with other hybrid sources [@Xu2007], although improved target design and continued optimization of source parameters may improve performance \[@Salazar2013; \@Fahrni2013\].
 Typical currents for this study were between 4-9 μA and average current for all samples was `r curmean`.
@@
 
-Current vs flow modelling
-
-Fraction of CO~2~ in sample gas should follow this relationship:
+One of the greatest contrasts of the open split approach with closed-source HGIS methods is changing CO~2~ flow and helium dilution ratio over the period of measurement due to the dilution of the sample by the displacement helium.
+We examined the effect of this variable carbon delivery and dilution by measuring ^12^C^-^ current evolution during helium displacement/dilution of vial CO~2~ (Figure 2b) and comparing this to the composition of the gas delivered to the source.
+We estimated gas composition (Figure 2a) using a simple mixing model:
 
@@
 Where $C(t)$ is the fraction CO~2~, $r$ is the rate of helium displacement flow, and $V$ is the initial volume of the CO~2~-filled vial.
-
-Dilution ratio - 1:1 to 1:3 during analysis
-
-First tests of displacing pure CO~2~ from a vial into the open split using helium are promising.
-Helium flowing at 100 μl min^-1^ produces current from 7 mL of CO~2~ for over 2 hours.
-Currents do decrease with time due to dilution, but by supplying higher than optimal CO~2~ flow to the source initially, stable currents can be maintained for some time.
+Figure 2c shows the resulting relationship of sample ion current to CO~2~ flow rate.
+The most important feature of the current-flow relationship is the wide range of flows under which stable ion current is measured, which in turn produces relatively stable currents and ratios despite a changing carbon flow rate during measurement.
 
@@
 
-Work with currents and flows (above and Figure 2) predicts that optimal currents should be produced until less than 2 mL CO~2~ (in μg?) remains in the sample vial.
-Tests of a range of sample sizes support this estimate and show that no reduction in measured current is seen in samples larger than XXX mg C which corresponds to XXX mg CaCO~3~ sample material.
+Work with currents and flows (above and Figure 2) predicts that optimal currents should be produced until less than 2 ml CO~2~ (equivalent to 0.8 mg C) remains in the sample vial.
+Tests of a range of sample sizes support this estimate and show that no reduction in measured current is seen in samples larger than 1 mg C which corresponds to 8 mg CaCO~3~ sample material.
 To ensure ample sample for optimal measurements, we chose 15 mg as a target sample mass.
@@
 
-### Current dependent blank
-
 The HGIS method produces higher ratios for radiocarbon dead samples than the graphite method.
@@
 This current is initially \> 10 μA and drops exponentially, reaching a baseline of \~100 nA after 5-10 minutes.
-For the range of currents in this study (3-9 μA), this baseline current represents a 1-6% contribution to the total current at the F^14^C of the target baseline.
-This can be interpreted as a constant current dependent blank, analogous to a mass-dependent "constant contaminant" blank and may be modeled with the same tools [@haghipour_compound-specific_2019; @hanke_comprehensive_2017; @roberts_14c_2019; @sun_14_2020].
+For the range of currents in this study (3-9 μA), this baseline current represents a 1-6% contribution to the total measured sample current at the F^14^C of the target current baseline.
+This can be interpreted as a constant current dependent blank, analogous to a mass-dependent "constant contaminant" blank and may be modeled with the same tools [@haghipour_compound-specific_2019; @hanke_comprehensive_2017; @roberts_14c_2019; @sun_14_2020\].
 To assess blank contribution and ratio dependence on sample current, vials of live and dead standard gas were measured as the displacement helium flow diluted the samples, producing diminishing current over time.
-We used these data to develop a current-balance model following methods used to assess mass-dependent or constant contamination blanks in graphite targets (REFS).
+We used these data to develop a current-balance model following methods used to assess mass-dependent or constant contamination blanks in graphite targets \[@shah_walter_ultra-small_2015; \@sun_14_2020].
 The constant contamination blank was also assessed directly by measuring pure helium in vials using the same method.
 Results from both tests are shown in Figure 3a.
-The inverse of the sample current is used to linearize the data, allowing assessment with a simple linear model [@shah_walter_ultra-small_2015].
+The inverse of the sample current is used to linearize the data, allowing assessment with a simple linear model.
 Fits from modern and dead dilution tests should trend towards the fraction modern of the blank, crossing at the inverse of blank current.
@@
 This constant contamination blank is consistent with the fact that hybrid cathodes produce a minimum current with or without gas flow to the source.
-Given that this minimum current ranges from 100 nA to 400 nA, this current likely makes up a large fraction of the method blank.
-
-$$Fm_s = Fm_m + \frac{I_b(Fm_m-Fm_b)}{I_m-I_b}$$
-
-### Current independent blank
+Given that this blank cathode current is a significant fraction of measured sample current, this current likely makes up a large fraction of the method blank.
 
@@
 If the fraction modern of the blank is 0.8 modern as estimated above, radiocarbon-free samples will show the most effect of a constant contamination blank with current.
-Figure 3B shows the fraction modern of measured blanks compared to their mean ^12^C^-^ ion current.
-Fraction modern of measured blanks does decrease with increasing current, but the effect is small, less than F^14^C 0.01 over the range of currents measured.
+Figure 3b shows the fraction modern of measured blanks compared to their mean ^12^C^-^ ion current.
+Fraction modern of measured blanks does decrease with increasing current, but the decrease is small, less than F^14^C 0.01 over the range of currents measured.
 This change with current is roughly equal to the variability between blanks, so the error due to the use of a current-dependent blank based on samples measured over multiple wheels will likely be larger than a current independent blank correction based on blanks measured on each wheel.
-Constant current blank is likely the best model, but we simplify to a large blank correction since blanks vary by run, currents are relatively similar between blanks, unknowns and standards, and the simpler model works well within the error of the method.
-
-We use equation 1 for this correction.
+Constant current blank is likely the best model, but we simplify to a linear correction based on the measured ratio of process blanks and normalizing standards:
 
@@
 
-Where $R_{samp}$ is the ratio of the unknown sample, $R_{meas}$ is the measured ratio, $R_{blank}$ is the ratio of the blank, and $R_{std}$ is the ratio of the normalizing standard.
+where $R_{samp}$ is the ratio of the unknown sample, $R_{meas}$ is the measured ratio, $R_{blank}$ is the ratio of the blank, and $R_{std}$ is the ratio of the normalizing standard.
+This mass-independent blank correction works well for our data since blanks vary by run and currents are relatively similar between blanks, unknowns and standards.
 
@@
 of measurement time per sample.
-The simplicity of this approach...
-With lower cost and time per sample, the lower precision of the method may be balanced by the ability to measure more samples from a population, ie.
+The simplicity of of this approach reduces costs, improves reliablity, and reduces system maintenance.
+With lower cost and time per sample, the lower precision of the method may be balanced by the ability to measure more samples from a population, i.e.
 triplicate measurement or more core depths.
 This may lead to more accurate results compared to a more precise graphite measurement where reproducibility is not as well constrained.
-
-The accuracy of age-depth models is often driven by environmental factors and sparse data, rather than the precision of individual dates \[3\].
-The lower time and cost of the hybrid method allows analysis of more samples.
-This may produce more accurate age models, despite reduced precision of individual measurements.
 
@@
 
-This work is supported by NSF cooperative agreement. We are grateful to C. McIntyre, R. Hansman, G. Salazar, T. Ognibene, T. Broek.
+This work was supported by the NOSAMS NSF cooperative agreement 1755125.
+We are grateful to C.
+McIntyre, R.
+Hansman, G.
+Salazar, T.
+Ognibene, and T.
+Broek for discussions on gas handling and gas ion sourcery; and to J.
+Gunnard for improvements to the text.\
+
+# Tables
 

### 2022-04-01 doc/AMS-15/AMS-15_paper/tables/table_1_reference_materials.Rmd
Finish current and flow section. Sundry edits. Convert table to F14C.

@@
   arrange(fm_consensus) %>% 
-  mutate(across(c(fm_consensus, fm_corr_mean, fm_corr_sd, sig_fm_corr_mean), 
-         function(x) x * 100),
-         fm_corr_err = set_errors(fm_corr_mean, fm_corr_sd)) %>% 
+ # mutate(across(c(fm_consensus, fm_corr_mean, fm_corr_sd, sig_fm_corr_mean), 
+ #        function(x) x * 100),
+  mutate(fm_corr_err = set_errors(fm_corr_mean, fm_corr_sd)) %>% 
   select(Name, `Consensus` = fm_consensus,
@@
   tab_header(title = "Reference material precision and accuracy",
-             subtitle = "Values in pMC, Errors are SD of measurements") %>% 
+             subtitle = "Values in F14C, Errors are SD of measurements") %>% 
   fmt_number(4, 
@@
   fmt_number(1:3, 
-             scale_by = 100,
-             decimals = 2)
+             decimals = 4)
 

### 2022-04-01 doc/AMS-15/AMS-15_paper/AMS-15_paper.Rmd
Add information on core samples

@@
 Evacuation, acid addition, and heated reaction were performed by the automated gas handling system developed for the CFAMS GIS carbonate system [@Roberts2013].
+
+Reference and process blank materials were not pretreated or cleaned.
+Carbonate shell samples for the core comparison study were picked from a split core and cleaned by washing with distilled water and dried at 60 C.
+Shell samples were crushed and mixed, but not otherwise pretreated.
+Other?
+Sonication?
 
@@
 
-A suite of 15 carbonate macrofossils consisting of shell and coral fragments from a series of coastal marine cores REF were measured as graphite and as CO~2~ using the HGIS method to further test the method and serve to demonstrate a potential application.
-coral macrofossils from a subtropical coastal marine sediment core using our gas method and compared to the same samples measured using the traditional graphite method.
+A suite of 15 carbonate macrofossils consisting of shell and coral fragments from a series of coastal marine cores @moser2021 were measured as graphite and as CO~2~ using the HGIS method to further test the method and serve to demonstrate a potential application.
+Sediment cores were collected from Baie des Baradères, Haiti during a scientific cruise in May 2016 lead by Dr.
+Jeffrey Donnelly (WHOI).
+Coring was performed with a Rossfelder vibracorer from a 20-foot pontoon boat.
+The cores were collected from a coastal karst basin (CKB) at vertical depth \~11.7 m and coordinates 18.514° N, 73.629° W.
+Core HATC1 D1 is 145 cm long and was collected in a 3-inch poly tube as a control directly beside core HATC1 D2, which is a continuous 896 cm core collected in a 3-inch aluminum tube.
+Carbonate materials such as corals and bivalve shells were selected throughout the core for the age-depth model.
 
-To further test the method and a potential application, we measured 15 shell and coral macrofossils from a subtropical coastal marine sediment core using our gas method and compared to the same samples measured using the traditional graphite method.
+age range
 
@@
 The age-depth curves produced by the two datasets are very similar, with the largest differences being in the uncertainty of the individual dates.
+
+We applied a correction for the regional marine reservoir effect with a delta R of -216 and uncertainty 92 based on the oxcal model with 10 reservoir points.
+Samples with radiocarbon age \> 0 were calibrated using the Marine20 curve @heaton_marine20marine_2020.
+The NH1 curve from Hua et.
+al @hua_atmospheric_2013 was used for calibration of postmodern dates.
 

### 2022-04-03 doc/AMS-15/AMS-15_paper/AMS-15_paper.Rmd
Complete reference standards section. Lots of edits.

@@
 output:
+  html_document: default
+  pdf_document: default
   word_document: default
-  pdf_document: default
-  html_document: default
 bibliography: HGIS_paper.bib
@@
 
-The gas-capable 134-sample MCSNICS from National Electrostatics Corporation (NEC) [@Xu2007\] works on the same principle as hybrid sources developed by ETH Zürich \[@ruff_gas_2007; @Ruff2010] and High Voltage Engineering Europa (HVEE) [@BronkRamsey2004].
+The gas-capable 134-sample MCSNICS from National Electrostatics Corporation (NEC) [@Xu2007] works on the same principle as hybrid sources developed by ETH Zürich [@ruff_gas_2007; @Ruff2010] and High Voltage Engineering Europa (HVEE) [@BronkRamsey2004].
 The sample CO~2~ - helium mixture is introduced onto a titanium insert or frit in an aluminum target cathode using stainless tubing mounted to an arm with a probe that mates with the back of the cathode.
@@
 
-Reference and process blank materials were not pretreated or cleaned.
+Reference and process blank materials were not pretreated.
 Carbonate shell samples for the core comparison study were picked from a split core and cleaned by washing with distilled water and dried at 60 C.
 Shell samples were crushed and mixed, but not otherwise pretreated.
-Other?
-Sonication?
 
 Standards were prepared by evacuating vials and then filling with a modern standard CO~2~ using the automated gas handling system.
-Standards were adjusted to the same size as other samples by removing CO~2~ and replacing with helium using a syringe.
+Standards were adjusted to the same size as other samples by extracting an appropriate volume of CO~2~ from the full vial and replacing with helium using a syringe.
 Process blank samples were prepared from powdered C-1 or TIRI-F, acidified in vials using the method above.
@@
 
-Measurements were corrected for process and measurement blank using a linear correction that assumes that the corrected fraction modern of blank samples should be zero, samples with the fraction modern of the standard need no correction, and applies a proportional correction to samples of intermediate fraction modern \[@roberts_14c_2019\].
+Measurements were corrected for process and measurement blank using a linear correction that assumes that the corrected fraction modern of blank samples should be zero, samples with the fraction modern of the standard need no correction, and applies a proportional correction to samples of intermediate fraction modern $$@roberts_14c_2019$$.
 The correction uses measured values of standard gas and blank samples for correction of contaminant carbon and machine blank.
@@
 Peak current was produced at a CO~2~ flow rate of 7.5 μl min^-1^ CO~2~, but the source produced near optimal currents with flows from 7.5-25 μl min^-1^ CO~2~.
-This is comparable with other hybrid sources [@Xu2007], although improved target design and continued optimization of source parameters may improve performance \[@Salazar2013; \@Fahrni2013\].
+This is comparable with other hybrid sources [@Xu2007], although improved target design and continued optimization of source parameters may improve performance $$@Salazar2013; \@Fahrni2013$$.
 Typical currents for this study were between 4-9 μA and average current for all samples was `r curmean`.
@@
 Capillary clogging can be an issue, especially when sampling from vials containing acid.
-
-Precision of measurements of modern CO~2~ has been better than 1% (std error of mean of measurements of a target) for high current samples, with precision dropping off as current decreases below 6 μA or so.
-Accuracy is reasonable within the large errors on gas measurements.
-Uncorrected measurements of dead CO~2~ range from 1% to 5% of modern depending on sample current.
 
@@
 The most important feature of the current-flow relationship is the wide range of flows under which stable ion current is measured, which in turn produces relatively stable currents and ratios despite a changing carbon flow rate during measurement.
+Several studies have shown peak currents over a narrower range of flows @Fahrni2013; @Broek2022 at high helium dilution rates (95%), while Xu et.
+al @Xu2007 shows a somewhat broader peak with pure CO~2~.
+Broek et.
+al @Broek2022 shows peak broadening with decreasing dilution ratio.
+While lower rates of dilution decrease current slightly, the observed peak broadening allows stable measurement over the range of CO~2~ flows produced by the open split.
 
@@
 The HGIS method produces higher ratios for radiocarbon dead samples than the graphite method.
-Measurement of radiocarbon-free reference materials shows that the average normalized ratio for these samples is `r blankmean` ± `r blanksd` (N=`r nblank`).
-We investigated the sources of this method blank as a current-dependent blank and as a current-independent blank.
+Measurements of radiocarbon-free reference materials (IAEA C1 marble and TIRI F spar calcite) show that the average normalized ratio for these samples is `r blankmean` ± `r blanksd` (1σ uncertainty, N=`r nblank`).
+We investigated the sources of this method blank as a current-dependent blank and as a current-independent blank, and implemented a blank correction for the method.
 
@@
 For the range of currents in this study (3-9 μA), this baseline current represents a 1-6% contribution to the total measured sample current at the F^14^C of the target current baseline.
-This can be interpreted as a constant current dependent blank, analogous to a mass-dependent "constant contaminant" blank and may be modeled with the same tools [@haghipour_compound-specific_2019; @hanke_comprehensive_2017; @roberts_14c_2019; @sun_14_2020\].
+This can be interpreted as a constant current dependent blank, analogous to a mass-dependent "constant contaminant" blank and may be modeled with the same tools [@haghipour_compound-specific_2019; @hanke_comprehensive_2017; @roberts_14c_2019; @sun_14_2020].
 To assess blank contribution and ratio dependence on sample current, vials of live and dead standard gas were measured as the displacement helium flow diluted the samples, producing diminishing current over time.
-We used these data to develop a current-balance model following methods used to assess mass-dependent or constant contamination blanks in graphite targets \[@shah_walter_ultra-small_2015; \@sun_14_2020].
+We used these data to develop a current-balance model following methods used to assess mass-dependent or constant contamination blanks in graphite targets [@shah_walter_ultra-small_2015; \@sun_14_2020].
 The constant contamination blank was also assessed directly by measuring pure helium in vials using the same method.
@@
 
-Precision and accuracy of the method were assessed by measuring carbonate standards with consensus values from graphite measurements.
-Summary data for these samples are listed in Table 1, and measurement data for all samples is in supplementary information.
+```{r, message = FALSE}
+#library(gt)
+library(errors)
+
+options(digits = 3)
+options(scipen = 4)
+options(decimals = 4)
+options(errors.notation = "plus-minus")
+
+# Format data
+cons_df <- carb %>% 
+  filter(!(wheel %in% c("USAMS020521", "USAMS061121", "USAMS061821"))) %>% 
+  mutate(he12C = he12C * 1E6) %>% 
+  compare_consensus() %>% 
+  filter(Name %in% c("TIRI-I", "C-2", "NOSAMS2"))
+
+# Summary table
+cons_df %>% 
+  group_by(Name, fm_consensus) %>% 
+  summarize(across(c(fm_corr, sig_fm_corr, fm_diff, sigma),
+                   list(mean = mean, sd = sd)),
+            N = n()) %>% 
+  ungroup() %>% 
+  arrange(fm_consensus) %>% 
+  mutate (fm_corr_err = set_errors(fm_corr_mean, fm_corr_sd),
+          Name = case_when(Name == "TIRI-I" ~ "TIRI I",
+                           Name == "C-2" ~ "IAEA C2",
+                           Name == "NOSAMS2" ~ "NOSAMS 2")) %>% 
+  select(`Sample Name` = Name, 
+         `Reference F^14^C` = fm_consensus,
+         `HGIS F^14^C ± stdev` = fm_corr_err,
+         `Mean per-sample error` = sig_fm_corr_mean,
+         N) %>% 
+  knitr::kable(digits = 3,
+               caption = "Reference material precision and accuracy. Reference 
+               values are community consensus values for TIRI I and C2, and the 
+               NOSAMS internal graphite value for NOSAMS2. NOSAMS2 is a 
+               homogenized bivalve molusk reference. The mean of per-sample 
+               error estimates is included for comparison to the reproducibilty 
+               of the samples.")
+
+  # gt() %>% 
+  # tab_header(title = "Reference material precision and accuracy",
+  #            subtitle = "Values in pMC, Errors are SD of measurements") %>% 
+  # fmt_number(4, 
+  #            decimals = 3) %>% 
+  # cols_align("right", 3)
+
+# Consensus summary for inline data
+cons_sum <- cons_df %>% 
+  filter(Name %in% c("TIRI-I", "C-2", "NOSAMS2")) %>% 
+  ungroup() %>% 
+  summarize(across(c(sig_fm_corr, fm_diff),
+                   list(mean = mean, sd = sd)),
+            N = n()) 
+```
+
+Precision and accuracy of the method were assessed by measuring carbonate reference standards for comparison with consensus values from graphite measurements.
+We measured TIRI I, a roughly quarter-modern turbidite carbonate, IAEA C2, a half modern turbidite carbonate, and NOSAMS 2, a modern internal shell carbonate standard from a bivalve molusk REF.
+Summary data for these samples are given in Table 1, and measurement data for all samples is in supplementary information.
 Measurement, data reduction, and blank correction was performed as described above.
 
-Overall, reference materials run with the hybrid gas method were in good agreement with consensus values.
-Mean difference from consensus for all measurements was -0.04 pMC, and the standard deviation of the difference was 0.011 F^14^C.
-Mean propagated measurement error was 0.011 F^14^C for these samples.
-The hybrid method has no bias in aggregate.
-Method error is larger than graphite, but calculated per-sample error is similar to standard deviation of replicate measurements.
-A summary of accuracy and precision of these measurements is in Table 1.
+Reference materials run with the hybrid gas method were in good agreement with consensus values.
+Mean difference from consensus for all measurements was F^14^C `r format(cons_sum$fm_diff_mean, digits =1, nsmall = 4)` shows that the hybrid method has no bias in aggregate.
+The standard deviation of the difference in F^14^C was `r format(cons_sum$fm_diff_sd, nsmall = 4)`, which may be taken as an estimate of the 1σ precision of the method.
+The mean of propagated measurement errors for these samples was `r format(cons_sum$sig_fm_corr_mean, nsmall = 4)` F^14^C for these samples, which is in good agreement with the method precision above.
+The method error is about three times larger than the roughly 3 ‰ precision of graphite measurements, but is within the range of method precision of other gas sources REFS.
 
-The difference between reference samples and graphite consensus does not vary systematically by fraction modern, indicating that the method has no significant bias by fraction modern (Figure 4).
-
-The per-sample error agrees well with variability of replicates, and measurement of known-value standards shows that the method has no significant bias.
+The difference between reference samples and graphite consensus does not vary systematically by F^14^C, indicating that the method has no significant bias with radiocarbon content (Figure 4).
+Each reference standard type agrees with graphite consensus values within the standard deviation of samples measured, with slightly more measurements agreeing with graphite within 1σ uncertainties than would be expected statistically.
 
@@
 
-age range
-
-We found the HGIS and graphite measurements to be in good agreement, with no significant bias (Figure 5a).
+We found the HGIS and graphite measurements to be in good agreement, with no significant bias over the range in F^14^C of the core samples (0.4 - 1.15) (Figure 5a).
 Mean difference between hybrid and graphite measurements was -0.003 F^14^C and the standard deviation of this difference was 0.009 F^14^C.
-Mean per-sample error was 0.010 F^14^C for HGIS measurements, compared with 0.002 F^14^C for graphite.
+Mean per-sample 1σ uncertainty was 0.010 F^14^C for HGIS measurements, compared with 0.002 F^14^C for graphite.
+There is an apparent, but not significant positive relationship between the HGIS-graphite difference and measured value.
+The higher graphite values at lower F^14^C may indicate an excessive HGIS blank correction for these samples or an undercorrected method blank for graphite measurements.
 
@@
 We applied a correction for the regional marine reservoir effect with a delta R of -216 and uncertainty 92 based on the oxcal model with 10 reservoir points.
-Samples with radiocarbon age \> 0 were calibrated using the Marine20 curve @heaton_marine20marine_2020.
+Samples with pre-bomb radiocarbon ages were calibrated using the Marine20 curve @heaton_marine20marine_2020.
 The NH1 curve from Hua et.
-al @hua_atmospheric_2013 was used for calibration of postmodern dates.
+al @hua_atmospheric_2013 was used for calibration of postmodern dates, but these measurements were not included in the age model due to uncertainties in reservoir correction.
 
@@
 of measurement time per sample.
-The simplicity of of this approach reduces costs, improves reliablity, and reduces system maintenance.
+The simplicity of of this approach reduces costs, improves reliability, and reduces system maintenance.
 With lower cost and time per sample, the lower precision of the method may be balanced by the ability to measure more samples from a population, i.e.
@@
 Broek for discussions on gas handling and gas ion sourcery; and to J.
-Gunnard for improvements to the text.\
+Gunnard for improvements to the text.
 
@@
 Performance of the NOSAMS MCSNICS hybrid gas ion source.
-a.
-Modeled flows of helium and CO~2~ to the ion source during displacement of a pure CO~2~ vial sample with helium.
-b.
-Measured 12C- ion current during the same dilution experiment.
-c.
-Measured ion current for modeled CO~2~ flow.
+(a) Modeled flows of helium and CO~2~ to the ion source during displacement of a pure CO~2~ vial sample with helium.
+(b) Measured 12C- ion current during the same dilution experiment.
+(c) Measured ion current for modeled CO~2~ flow.
 The stability of the source over a range of flow produces stable current for up to 40 m of measurement time.
-
-Will probably skip this figure: Figure 3. Relationship of sample size to ^12^C^-^ ion current. There is no strong relationship between size and current for samples larger than 800 μg C.
 
@@
 Comparison of Haiti core samples measured as graphite and via HGIS.
-a.
-Difference between HGIS and graphite measurements of the same materials.
+(a) Difference between HGIS and graphite measurements of the same materials.
 Difference between HGIS and graphite measurements is shown on the vertical axis and plotted against the mean of the two measurements.
@@
 Graphite measurements are included to allow comparison of error estimates.
-b) comparison of age-depth models produced using HGIS and graphite measurements.
+(b) Comparison of age-depth models produced using HGIS and graphite measurements.
 

### 2022-04-04 doc/AMS-15/AMS-15_paper/figures/fig5_BChron_model.R
Use NH2 curve for bomb dates. Add reservoir ages. Don't use bomb dates for age model.

@@ -77,14 +77,14 @@ cr_no <- combined_results %>%
 #####
 
 # load curve Hua 2020 NH1 Curve
-bomb_NH1 <- IntCal::ccurve(postbomb = TRUE)
+bomb_NH2 <- IntCal::ccurve("nh2")
 
 # create Bchron cal curve
-createCalCurve("bomb_NH1", bomb_NH1[, 1], bomb_NH1[, 2], bomb_NH1[,3])
+createCalCurve("bomb_NH2", bomb_NH2[, 1], bomb_NH2[, 2], bomb_NH2[,3])
 
 # Copy to bchron install directory
 file.copy(
-  from = "bomb_NH1.rda",
+  from = "bomb_NH2.rda",
   to = system.file("data", package = "Bchron"),
   overwrite = TRUE
 )
@@ -94,42 +94,22 @@ file.copy(
 cal_sub <- cr_no %>% 
   mutate(id = paste(sample_name, "-", method)) %>% 
   select(sample_name, id, depth, rc_age, sig_rc_age, method) %>% 
-  mutate(curve = if_else(rc_age > 0, "marine20", "bomb_NH1")) %>% 
-  add_row(id = "Top", sample_name = "Top", depth = 0, rc_age = 0, sig_rc_age = 1, curve = "normal")
-
-cal_dates <- with(cal_sub,
-  BchronCalibrate(rc_age, sig_rc_age, positions = depth, ids = id, calCurves = curve))
-plot(cal_dates)
-
-cal_sub_hgis <- cal_sub %>% 
-  filter(method == "hgis" | is.na(method))
-
-cal_sub_gr <- cal_sub %>% 
-  filter(method == "graphite" | is.na(method)) 
+  mutate(curve = if_else(rc_age > 0, "marine20", "bomb_NH2"),
+         rc_age_corr = if_else(curve == "marine20", rc_age - 216, rc_age),
+         sig_rc_age_corr = if_else(curve == "marine20", sqrt(sig_rc_age^2 + 92^2), sig_rc_age)) %>% 
+  add_row(id = "Top", sample_name = "Top", depth = 0, 
+          rc_age = 0, sig_rc_age = 1, 
+          rc_age_corr = 0, sig_rc_age_corr = 1, 
+          curve = "normal")
 
 
 #####
 # Calibrate dates
 #####
 
-# Calibrate hgis ages
-cal_dates_hgis <- with(cal_sub_hgis,
[…]

### 2022-04-04 doc/AMS-15/AMS-15_paper/AMS-15_paper.Rmd
Sundry edits. Fix order of fig 2.

@@
 
-### Currents and flows
-
-One of the greatest contrasts of the open split approach with closed-source HGIS methods is changing CO~2~ flow and helium dilution ratio over the period of measurement due to the dilution of the sample by the displacement helium.
-We examined the effect of this variable carbon delivery and dilution by measuring ^12^C^-^ current evolution during helium displacement/dilution of vial CO~2~ (Figure 2b) and comparing this to the composition of the gas delivered to the source.
-We estimated gas composition (Figure 2a) using a simple mixing model:
+A contrast of the open split approach with closed-source HGIS methods is changing CO~2~ flow and helium dilution ratio over the period of measurement due to the dilution of the sample by the displacement helium.
+We examined the effect of this variable carbon delivery and dilution by measuring ^12^C^-^ current evolution during helium displacement/dilution of vial CO~2~ (Figure 2a) and comparing this to the composition of the gas delivered to the source.
+We estimated gas composition (Figure 2b) using a simple mixing model:
 
@@
 While lower rates of dilution decrease current slightly, the observed peak broadening allows stable measurement over the range of CO~2~ flows produced by the open split.
-
-### Sample size
 
@@
 The directly measured blank current (120 nA) agrees well with the model assessment, but has a higher fraction modern (F^14^C = 0.82).
-This is likely due to a combination of low measurement precision for the direct measurement and high sensitivity of the model to variability in the measured fraction modern and current in the depletion samples [@sun_14_2020].
+A combination of low measurement precision for the direct measurement and high sensitivity of the model to variability in the measured fraction modern and current in the depletion samples may explain this difference[@sun_14_2020].
 This constant contamination blank is consistent with the fact that hybrid cathodes produce a minimum current with or without gas flow to the source.
@@
 Core HATC1 D1 is 145 cm long and was collected in a 3-inch poly tube as a control directly beside core HATC1 D2, which is a continuous 896 cm core collected in a 3-inch aluminum tube.
-Carbonate materials such as corals and bivalve shells were selected throughout the core for the age-depth model.
 
-We found the HGIS and graphite measurements to be in good agreement, with no significant bias over the range in F^14^C of the core samples (0.4 - 1.15) (Figure 5a).
-Mean difference between hybrid and graphite measurements was -0.003 F^14^C and the standard deviation of this difference was 0.009 F^14^C.
-Mean per-sample 1σ uncertainty was 0.010 F^14^C for HGIS measurements, compared with 0.002 F^14^C for graphite.
-There is an apparent, but not significant positive relationship between the HGIS-graphite difference and measured value.
-The higher graphite values at lower F^14^C may indicate an excessive HGIS blank correction for these samples or an undercorrected method blank for graphite measurements.
-
+Carbonate materials such as corals and bivalve shells were selected throughout the core for the age-depth model, processed and measured as described above.
+Parallel graphite measurements were made following NOSAMS methods for acidification, graphitization, and measurement.
 One HGIS sample vial was noted to be at atmospheric pressure when the 16 mg sample should have produced only 0.5 atmosphere of CO~2~.
@@
 
-### Age model
+We found the HGIS and graphite measurements to be in good agreement, with no significant bias over the range in F^14^C of the core samples (0.4 to 1.15) (Figure 5a).
+Mean difference between hybrid and graphite measurements was -0.003 F^14^C and the standard deviation of this difference was 0.009 F^14^C.
+Mean per-sample 1σ uncertainty was 0.010 F^14^C for HGIS measurements, compared with 0.002 F^14^C for graphite.
+There is an apparent, but not significant positive relationship between the HGIS-graphite difference and measured value.
+The higher graphite values at lower F^14^C may indicate an overcorrection of HGIS blank for these samples.
+
+Mean and sd of difference of calibrated ages?
 
@@
 We used the Bchron Baesian radiochronology package [@haslett_simple_2008] to develop these models.
-The age-depth curves produced by the two datasets are very similar, with the largest differences being in the uncertainty of the individual dates.
-
 We applied a correction for the regional marine reservoir effect with a delta R of -216 and uncertainty 92 based on the oxcal model with 10 reservoir points.
 Samples with pre-bomb radiocarbon ages were calibrated using the Marine20 curve @heaton_marine20marine_2020.
-The NH1 curve from Hua et.
+The NH zone 2 curve from Hua et.
 al @hua_atmospheric_2013 was used for calibration of postmodern dates, but these measurements were not included in the age model due to uncertainties in reservoir correction.
-
-Mean and sd of difference of calibrated ages?
+The age-depth curves produced by the two datasets are very similar, with the largest differences being in the uncertainty of the individual dates.
 

### 2022-04-04 doc/AMS-15/AMS-15_paper/figures/fig2_current_flow.R

@@ -18,6 +18,15 @@ data <- process_hgis_results(here("data/USAMS040121R.txt"),
   mutate(time = cumsum(corr_lt) / 60,
          co2flow = concCO2(time, r = 244) * 30) #30ul/min
 
+# current vs time subplot
+cur_time <- ggplot(data, aes(time, he12C)) +
+  geom_smooth(span = .4, se = FALSE, color = "#00b7bd") +
+  geom_point(size = 3, color = "#0069b1") +
+  xlim(0, 125) +
+  labs(title = "A",
+       #title = "Ion current",
+       x = "Time (min)",
+       y = "12C current (μA)")
 
 # flow vs time subplot
 flow_time <- data.frame(x = 0:125) %>% 
@@ -27,7 +36,7 @@ flow_time <- data.frame(x = 0:125) %>%
   stat_function(fun = function(x) 30 - (concCO2(x, r = 244) * 30),
                 aes(color = "Helium"), size = 1.5) +
   scale_color_manual("Gas", values = c("#00b7bd", "#b7bf10")) +
-  labs(title = "A",
+  labs(title = "B",
        #title = "Gas flows to source",
        #subtitle = "7mL vial, 244μL/min helium, 30μL/min to source",
        y = "Gas flow (μL/min)") +
@@ -35,17 +44,6 @@ flow_time <- data.frame(x = 0:125) %>%
         axis.text.x = element_blank(),
         legend.position = c(0.87, 0.65),
         legend.background = element_rect(fill = "white", color = "black"))
-
-# current vs time subplot
-cur_time <- ggplot(data, aes(time, he12C)) +
-  geom_smooth(span = .4, se = FALSE, color = "#00b7bd") +
-  geom_point(size = 3, color = "#0069b1") +
-  xlim(0, 125) +
-  labs(title = "B",
-       #title = "Ion current",
-       x = "Time (min)",
-       y = "12C current (μA)")
-
 
 # Current vs flow subplot
 
@@ -59,6 +57,7 @@ cur_flow <- ggplot(data, aes(co2flow, he12C)) +
        y = TeX(r'(^{12}C^- current (μA))'))
 
 # Build figure and save
-flow_time / cur_time / cur_flow
+cur_time / flow_time / cur_flow
 
 ggsave(here("doc/AMS-15/AMS-15_paper/figures/fig2_current_flow.svg"))
+ggsave(here("doc/AMS-15/AMS-15_paper/figures/fig2_current_flow.pdf"))

### 2022-04-04 doc/AMS-15/AMS-15_paper/tables/table_1_reference_materials.R

@@ -1,4 +1,5 @@
 # Table 1 showing performance of reference standards and performance summary
+# 
 
 
 library(tidyverse)
@@ -49,7 +50,7 @@ gt() %>%
   cols_align("right", 3)
 
 # Consensus summary
-cons_df %>% 
+cons_sum <- cons_df %>% 
   filter(Name %in% c("TIRI-I", "C-2", "NOSAMS2")) %>% 
   ungroup() %>% 
   summarize(across(c(sig_fm_corr, fm_diff),
@@ -59,10 +60,4 @@ cons_df %>%
          `Std. Dev. of consensus difference` = fm_diff_sd,
          `per-sample error` = sig_fm_corr_mean,
          N
-         ) %>%  
-  gt() %>% 
-  tab_header(title = "Reference material performance summary",
-             subtitle = "TIRI-I, C-2, and NOSAMS2. Values in pMC") %>% 
-  fmt_number(1:3, 
-             scale_by = 100,
-             decimals = 2)
+         ) 

### 2022-04-04 doc/AMS-15/AMS-15_paper/AMS-15_paper.Rmd
Add core data

@@
 
+```{r core data}
+core <- read_csv(here("data_analysed/haiti_combined.csv"))
+
+# Remove outlier: HGIS sample leaked
+cr_no <- core %>% 
+  filter(rec_num != 171996)
+
+dups <- cr_no %>% 
+  filter(method == "hgis") %>% 
+  group_by(rec_num) %>% 
+         filter(n() > 1) %>% 
+  ungroup()
+
+# pick single core run for fairness?
+core_summary <- cr_no %>% 
+  group_by(method) %>% 
+  #filter(method == "hgis") %>% 
+  summarize(across(c(sig_fm_corr, fm_diff),
+                   list(mean = mean, sd = sd)),
+            N = n())
+```
+
 A suite of 15 carbonate macrofossils consisting of shell and coral fragments from a series of coastal marine cores @moser2021 were measured as graphite and as CO~2~ using the HGIS method to further test the method and serve to demonstrate a potential application.
@@
 Mean per-sample 1σ uncertainty was 0.010 F^14^C for HGIS measurements, compared with 0.002 F^14^C for graphite.
+Three samples were analysed by HGIS in duplicate.
+Two of the three pairs agreed within 1σ and all three agreed within 2σ.
 There is an apparent, but not significant positive relationship between the HGIS-graphite difference and measured value.
@@
 The age-depth curves produced by the two datasets are very similar, with the largest differences being in the uncertainty of the individual dates.
-
-The accuracy of age-depth models is often driven by environmental factors and sparse data, rather than the precision of individual dates [@blaauw_double_2018] ONE MORE?.
-The lower time and cost of the hybrid method allows analysis of more samples.
-This may produce more accurate age models, despite reduced precision of individual measurements.
 
@@
 
+The accuracy of age-depth models is often driven by environmental factors and sparse data, rather than the precision of individual dates [@blaauw_double_2018] ONE MORE?.
+The lower time and cost of the hybrid method allows analysis of more samples.
+This may produce more accurate age models, despite reduced precision of individual measurements.
 Sample requirements, analysis time, precision, accuracy.

### 2022-04-04 doc/AMS-15/AMS-15_paper/figures/fig2_current_flow.R
Figure updates. Working bchron figure

@@ -16,7 +16,8 @@ data <- process_hgis_results(here("data/USAMS040121R.txt"),
                              as.Date("2021-04-02")) %>%
   filter(pos == 5) %>% 
   mutate(time = cumsum(corr_lt) / 60,
-         co2flow = concCO2(time, r = 244) * 30) #30ul/min
+         co2flow = concCO2(time, r = 244) * 30, #30ul/min
+         he12C = he12C * 1E6)
 
 # current vs time subplot
 cur_time <- ggplot(data, aes(time, he12C)) +
@@ -26,7 +27,9 @@ cur_time <- ggplot(data, aes(time, he12C)) +
   labs(title = "A",
        #title = "Ion current",
        x = "Time (min)",
-       y = "12C current (μA)")
+       y = "12C current (μA)") +
+  theme(axis.title.x = element_blank(),
+        axis.text.x = element_blank())
 
 # flow vs time subplot
 flow_time <- data.frame(x = 0:125) %>% 
@@ -39,10 +42,9 @@ flow_time <- data.frame(x = 0:125) %>%
   labs(title = "B",
        #title = "Gas flows to source",
        #subtitle = "7mL vial, 244μL/min helium, 30μL/min to source",
+       x = "Time (min)",
        y = "Gas flow (μL/min)") +
-  theme(axis.title.x = element_blank(),
-        axis.text.x = element_blank(),
-        legend.position = c(0.87, 0.65),
+  theme(legend.position = c(0.87, 0.50),
         legend.background = element_rect(fill = "white", color = "black"))
 
 # Current vs flow subplot
@@ -54,10 +56,12 @@ cur_flow <- ggplot(data, aes(co2flow, he12C)) +
        #title = bquote('Ion current is stable for a range of CO'[2]~'flow'),
        #subtitle = bquote('Current vs. CO'[2]~'flow during vial dillution'),
        x = bquote('CO'[2]~'Flow (μl/min)'),
-       y = TeX(r'(^{12}C^- current (μA))'))
+       y = "12C- current (μA)"
+       #y = TeX("^{12}C^- current (uA)")
+       )
 
 # Build figure and save
 cur_time / flow_time / cur_flow
 
-ggsave(here("doc/AMS-15/AMS-15_paper/figures/fig2_current_flow.svg"))
-ggsave(here("doc/AMS-15/AMS-15_paper/figures/fig2_current_flow.pdf"))
[…]

### 2022-04-04 doc/AMS-15/AMS-15_paper/figures/fig3_blanks.R

@@ -162,4 +162,5 @@ fm_curr <- blanks %>%
 # Build figure and save
 blank_model / fm_curr
 
-ggsave(here("doc/AMS-15/AMS-15_paper/figures/fig3_blanks.svg"))
+ggsave(here("doc/AMS-15/AMS-15_paper/figures/fig3_blanks.svg"), width = 7, height = 5)
+ggsave(here("doc/AMS-15/AMS-15_paper/figures/fig3_blanks.pdf"), width = 7, height = 5)

### 2022-04-04 doc/AMS-15/AMS-15_paper/figures/fig4_consensus_samples.R

@@ -41,5 +41,6 @@ ggplot(aes(fm_consensus, fm_diff, color = Name, shape = Name)) +
        y = "F14C difference (HGIS - expected)")
 
 # Save figure
-ggsave(here("doc/AMS-15/AMS-15_paper/figures/fig4_consensus_samples.svg"))
+ggsave(here("doc/AMS-15/AMS-15_paper/figures/fig4_consensus_samples.svg"), width = 7, height = 4)
+ggsave(here("doc/AMS-15/AMS-15_paper/figures/fig4_consensus_samples.pdf"), width = 7, height = 4)
 

### 2022-04-04 doc/AMS-15/AMS-15_paper/figures/fig5_BChron_model.R

@@ -98,8 +98,8 @@ cal_sub <- cr_no %>%
          rc_age_corr = if_else(curve == "marine20", rc_age - 216, rc_age),
          sig_rc_age_corr = if_else(curve == "marine20", sqrt(sig_rc_age^2 + 92^2), sig_rc_age)) %>% 
   add_row(id = "Top", sample_name = "Top", depth = 0, 
-          rc_age = 0, sig_rc_age = 1, 
-          rc_age_corr = 0, sig_rc_age_corr = 1, 
+          rc_age = -66, sig_rc_age = 1, 
+          rc_age_corr = -66, sig_rc_age_corr = 1, 
           curve = "normal")
 
 
@@ -108,7 +108,10 @@ cal_sub <- cr_no %>%
 #####
 
 cal_dates <- with(cal_sub,
-  BchronCalibrate(rc_age_corr, sig_rc_age_corr, positions = depth, ids = id, calCurves = curve))
+  BchronCalibrate(rc_age_corr, sig_rc_age_corr, 
+                  positions = depth, ids = id, 
+                  calCurves = curve,
+                  allowOutside = TRUE))
 plot(cal_dates)
 
 
@@ -162,7 +165,9 @@ summary(chron_hgis, "convergence")
 chron_gr  <- with(cal_sub_gr,
                   Bchronology(rc_age_corr, sig_rc_age_corr, positions = depth, 
                               ids = sample_name, calCurves = curve, 
-                              artificialThickness = 1.00))
+                              #artificialThickness = 1.00,
+                              allowOutside = TRUE,
+                              extractDate = -66))
 plot(chron_gr) +
   ggtitle("Graphite chronology")
 

### 2022-04-04 doc/AMS-15/AMS-15_paper/figures/fig5_core_carbonates.R

@@ -115,11 +115,29 @@ age_depth <- cr_no %>%
   theme(legend.position = c(0.85, 0.30),
         legend.background = element_rect(fill = "white", color = "black")) 
 
+# Bchron age models
+
+# Read chronologies
+
+chron_hgis <- read_rds(here("data_analysed/hgis_chronology.rds"))
+chron_gr <- read_rds(here("data_analysed/graphite_chronology.rds"))
+
+hgis_bchron <- plot(chron_hgis) +
+  xlim(8000, -500) +
+  ggtitle("B")
+
+gr_bchron <- plot(chron_gr) +
+  xlim(8000, -500) +
+  ggtitle("C") +
+  theme(axis.title.y = element_blank(),
+        axis.text.y = element_blank())
 # Combine plots and save
 
-core_compare / age_depth
+core_compare / (hgis_bchron | gr_bchron)
 
-ggsave(here("doc/AMS-15/AMS-15_paper/figures/fig5_core_carbonates.svg"))
+# Save figure
+ggsave(here("doc/AMS-15/AMS-15_paper/figures/fig5_core_carbonates.svg"), width = 7, height = 7)
+ggsave(here("doc/AMS-15/AMS-15_paper/figures/fig5_core_carbonates.pdf"), width = 7, height = 7)
 
 # data for core carbonates section
 

### 2022-04-04 doc/AMS-15/AMS-15_paper/AMS-15_paper.Rmd
Readthrough edits. Fix references.

@@
 output:
-  html_document: default
   pdf_document: default
   word_document: default
+  html_document: default
 bibliography: HGIS_paper.bib
@@
 Samples larger than 8 mg carbonate can be measured.
-Results from measurements of consensus standards (TIRI-I, IAEA C2, and an internal modern shell standard), and samples from a marine core (F^14^C = 0.4 - 1.15) show that the method agrees well with traditional AMS measurement of the same samples as graphite, and that the method error is less than F^14^C 0.015.
+Results from measurements of consensus standards (TIRI-I, IAEA C2, and an internal modern shell standard), and samples from a marine core (F^14^C = 0.4 - 1.15) show that the method agrees well with traditional AMS measurement of the same samples as graphite, and that the 1σ uncertainty is about 1 %.
 We discuss advantages and disadvantages of continuous flow sample introduction, and the effect of reduced precision on calibrated age-depth models produced using gas-source data.
@@
 
-Most interfaces to gas-accepting sputter ion sources are a closed inlet design [@Xu2007; @Uhl2007; @ruff_gas_2007; \@ruff\_-line_2010; @Fahrni2013; @wacker_versatile_2013; @gottschalk_radiocarbon_2018].
+Most interfaces to gas-accepting sputter ion sources are a closed inlet design [@Xu2007; @Uhl2007; @ruff_gas_2007; @ruff_-line_2010; @Fahrni2013; @wacker_versatile_2013; @gottschalk_radiocarbon_2018].
 A mixture of sample and helium is introduced to the source through a capillary, and flow is controlled by the pressure in the inlet volume and the length and diameter of the capillary.
@@
 
-An open inlet has the advantage of simplicity; flow of CO~2~ to the source is preset and there are no separate sample transfer and dilution steps.
+An open inlet has the advantage of simplicity; flow of gas to the source is preset by capillary parameters and there are no separate sample transfer and dilution steps.
 This may translate to greater throughput by eliminating the time required for sample transfer, dilution, and setting of delivery pressure.
@@
 In gas-source AMS, Roberts et al.
-@Roberts2013 use an open split to introduce sample gas to a ECR ion source and Salazar and Szidat @salazar_low-cost_2019 designed a continuous flow system that maintains relatively stable pCO~2~ during acquisition by controlling pressure and venting excess gas to atmosphere.
+@Roberts2013 used an open split to introduce sample gas to a ECR ion source and Salazar and Szidat @salazar_low-cost_2019 designed a continuous flow system that maintains relatively stable pCO~2~ during acquisition by controlling pressure and venting excess gas to atmosphere.
 
@@
 The split is connected to the top of the inner needle, and the source delivery capillary is inserted into the split.
-The split has fairly low volume (\< 1 mL).
-The system was operated manually during development, but may be easily mounted on the arm of a gasbench-style autosampler to enable automated operation.
+The split has fairly low volume (< 1 mL).
+The system was operated manually during development, but may be mounted on the arm of a gasbench-style autosampler to enable automated operation.
 
@@
 
-Measurements were corrected for process and measurement blank using a linear correction that assumes that the corrected fraction modern of blank samples should be zero, samples with the fraction modern of the standard need no correction, and applies a proportional correction to samples of intermediate fraction modern $$@roberts_14c_2019$$.
+Measurements were corrected for process and measurement blank using a linear correction that assumes that the corrected fraction modern of blank samples should be zero, samples with the fraction modern of the standard need no correction, and applies a proportional correction to samples of intermediate fraction modern [@donahue_isotope-ratio_1990; @roberts_14c_2019].
 The correction uses measured values of standard gas and blank samples for correction of contaminant carbon and machine blank.
@@
 carbcur <- carb %>%  
-  map_dfr(2) %>% 
   compare_consensus() %>% 
@@
 Peak current was produced at a CO~2~ flow rate of 7.5 μl min^-1^ CO~2~, but the source produced near optimal currents with flows from 7.5-25 μl min^-1^ CO~2~.
-This is comparable with other hybrid sources [@Xu2007], although improved target design and continued optimization of source parameters may improve performance $$@Salazar2013; \@Fahrni2013$$.
-Typical currents for this study were between 4-9 μA and average current for all samples was `r curmean`.
+This is comparable with other hybrid sources [@Xu2007], although improved target design and continued optimization of source parameters may improve performance [@Salazar2013; @Fahrni2013].
+Typical currents for this study were between 4-9 μA and average current for all samples was `r curmean` ± `r cursd`.
 Several studies have shown best currents at lower CO~2~ flow rates [@Xu2007; @Ruff2010].
-While efficiency of the source is best at very low flow rates (\<1.5 μl min^-1^), measuring samples with flow set for optimal current produces more precise results due to higher count rates.
 
@@
 Factors affecting source current may include capillary condition, source and gas delivery arm alignment, slight variability in target geometry, source condition, and performance of the ionizer and cesium.
-Current also varies by target for a given set of gas and source conditions.
-Capillary clogging can be an issue, especially when sampling from vials containing acid.
+Capillary clogging caused occasional flow issues, especially when sampling from vials containing acid.
 
@@
 The most important feature of the current-flow relationship is the wide range of flows under which stable ion current is measured, which in turn produces relatively stable currents and ratios despite a changing carbon flow rate during measurement.
-Several studies have shown peak currents over a narrower range of flows @Fahrni2013; @Broek2022 at high helium dilution rates (95%), while Xu et.
-al @Xu2007 shows a somewhat broader peak with pure CO~2~.
+Several studies have measured peak currents over a narrower range of flows [@Fahrni2013; @broek_optimization_nodate] at high helium dilution rates (> 10:1), while Xu et.
+al [@Xu2007] shows a somewhat broader peak with pure CO~2~.
 Broek et.
-al @Broek2022 shows peak broadening with decreasing dilution ratio.
-While lower rates of dilution decrease current slightly, the observed peak broadening allows stable measurement over the range of CO~2~ flows produced by the open split.
+al [@broek_optimization_nodate] found peak broadening with decreasing dilution ratio.
+While lower rates of dilution used in our HGIS method may decrease current slightly, the observed peak broadening allows stable measurement over the range of CO~2~ flows produced by the open split.
 
@@
 HGIS cathodes produce measurable ion current, even with no gas flow.
-This current is initially \> 10 μA and drops exponentially, reaching a baseline of \~100 nA after 5-10 minutes.
+This current is initially \> 10 μA and drops exponentially, reaching a baseline of ~100 nA after 5-10 minutes.
 For the range of currents in this study (3-9 μA), this baseline current represents a 1-6% contribution to the total measured sample current at the F^14^C of the target current baseline.
@@
 To assess blank contribution and ratio dependence on sample current, vials of live and dead standard gas were measured as the displacement helium flow diluted the samples, producing diminishing current over time.
-We used these data to develop a current-balance model following methods used to assess mass-dependent or constant contamination blanks in graphite targets [@shah_walter_ultra-small_2015; \@sun_14_2020].
+We used these data to develop a current-balance model following methods used to assess mass-dependent or constant contamination blanks in graphite targets [@shah_walter_ultra-small_2015; @sun_14_2020].
 The constant contamination blank was also assessed directly by measuring pure helium in vials using the same method.
@@
 # Summary table
-cons_df %>% 
-  group_by(Name, fm_consensus) %>% 
-  summarize(across(c(fm_corr, sig_fm_corr, fm_diff, sigma),
-                   list(mean = mean, sd = sd)),
-            N = n()) %>% 
-  ungroup() %>% 
-  arrange(fm_consensus) %>% 
-  mutate (fm_corr_err = set_errors(fm_corr_mean, fm_corr_sd),
-          Name = case_when(Name == "TIRI-I" ~ "TIRI I",
-                           Name == "C-2" ~ "IAEA C2",
-                           Name == "NOSAMS2" ~ "NOSAMS 2")) %>% 
-  select(`Sample Name` = Name, 
-         `Reference F^14^C` = fm_consensus,
-         `HGIS F^14^C ± stdev` = fm_corr_err,
-         `Mean per-sample error` = sig_fm_corr_mean,
-         N) %>% 
-  knitr::kable(digits = 3,
-               caption = "Reference material precision and accuracy. Reference 
-               values are community consensus values for TIRI I and C2, and the 
-               NOSAMS internal graphite value for NOSAMS2. NOSAMS2 is a 
-               homogenized bivalve molusk reference. The mean of per-sample 
-               error estimates is included for comparison to the reproducibilty 
-               of the samples.")
+# cons_df %>% 
+#   group_by(Name, fm_consensus) %>% 
+#   summarize(across(c(fm_corr, sig_fm_corr, fm_diff, sigma),
+#                    list(mean = mean, sd = sd)),
+#             N = n()) %>% 
+#   ungroup() %>% 
+#   arrange(fm_consensus) %>% 
+#   mutate (fm_corr_err = set_errors(fm_corr_mean, fm_corr_sd),
+#           Name = case_when(Name == "TIRI-I" ~ "TIRI I",
+#                            Name == "C-2" ~ "IAEA C2",
+#                            Name == "NOSAMS2" ~ "NOSAMS 2")) %>% 
+#   select(`Sample Name` = Name, 
+#          `Reference F^14^C` = fm_consensus,
+#          `HGIS F^14^C ± stdev` = fm_corr_err,
+#          `Mean per-sample error` = sig_fm_corr_mean,
+#          N) %>% 
+#   knitr::kable(digits = 3,
+#                caption = "Reference material precision and accuracy. Reference 
+#                values are community consensus values for TIRI I and C2, and the 
+#                NOSAMS internal graphite value for NOSAMS2. NOSAMS2 is a 
+#                homogenized bivalve molusk reference. The mean of per-sample 
+#                error estimates is included for comparison to the reproducibilty 
+#                of the samples.")
 
@@
 Reference materials run with the hybrid gas method were in good agreement with consensus values.
-Mean difference from consensus for all measurements was F^14^C `r format(cons_sum$fm_diff_mean, digits =1, nsmall = 4)` shows that the hybrid method has no bias in aggregate.
+Mean difference from consensus for all measurements was F^14^C `r format(cons_sum$fm_diff_mean, digits = 1, nsmall = 4)` shows that the hybrid method has no bias in aggregate.
 The standard deviation of the difference in F^14^C was `r format(cons_sum$fm_diff_sd, nsmall = 4)`, which may be taken as an estimate of the 1σ precision of the method.
 The mean of propagated measurement errors for these samples was `r format(cons_sum$sig_fm_corr_mean, nsmall = 4)` F^14^C for these samples, which is in good agreement with the method precision above.
-The method error is about three times larger than the roughly 3 ‰ precision of graphite measurements, but is within the range of method precision of other gas sources REFS.
+The method error is about three times larger than the roughly 3 ‰ precision of graphite measurements, but is within the range of method precision of other gas sources.
 
@@
 
-```{r core data}
+```{r core data, message = FALSE}
 core <- read_csv(here("data_analysed/haiti_combined.csv"))
@@
 # Remove outlier: HGIS sample leaked
+# Remove HATC duplicates
 cr_no <- core %>% 
-  filter(rec_num != 171996)
+  filter(rec_num != 171996,
+         !str_starts(sample_name, "HATC"))
 
-dups <- cr_no %>% 
+# Data for duplicate discussion
+dups <- core %>% 
   filter(method == "hgis") %>% 
@@
 
-A suite of 15 carbonate macrofossils consisting of shell and coral fragments from a series of coastal marine cores @moser2021 were measured as graphite and as CO~2~ using the HGIS method to further test the method and serve to demonstrate a potential application.
+A suite of 15 carbonate macrofossils consisting of shell and coral fragments from a series of coastal marine cores @moser_multi-proxy_2021 were measured as graphite and as CO~2~ using the HGIS method to further test the method and serve to demonstrate a potential application.
 Sediment cores were collected from Baie des Baradères, Haiti during a scientific cruise in May 2016 lead by Dr.
@@
 Coring was performed with a Rossfelder vibracorer from a 20-foot pontoon boat.
-The cores were collected from a coastal karst basin (CKB) at vertical depth \~11.7 m and coordinates 18.514° N, 73.629° W.
+The cores were collected from a coastal karst basin (CKB) at vertical depth ~11.7 m and coordinates 18.514° N, 73.629° W.
 Core HATC1 D1 is 145 cm long and was collected in a 3-inch poly tube as a control directly beside core HATC1 D2, which is a continuous 896 cm core collected in a 3-inch aluminum tube.
@@
 
-Mean and sd of difference of calibrated ages?
+```{r}
+# Agreement within uncertainty of paired ages?
+# Mean and sd of difference of calibrated ages?
+```
 
-Age models were developed for both the traditional graphite measurements of the core carbonates and the HGIS measurements (Figure 5b).
+Age models were developed for both the HGIS measurements (Figure 5b) and the traditional graphite measurements (Figure 5c) of the core carbonates.
 We used the Bchron Baesian radiochronology package [@haslett_simple_2008] to develop these models.
@@
 
-Age model quality changes with lower precision.
-
 # Discussion
@@
 We developed a method for the analysis of carbonates using an open-split interface to a hybrid sputter gas ion source.
-This system produces results that agree with graphite measurements to better than 1% from 8-20 mg of carbonate.
+This system produces results that agree with graphite measurements within the approximately 1% method error from 8-20 mg of carbonate.
 Sample preparation takes about 10 min.
@@
 The simplicity of of this approach reduces costs, improves reliability, and reduces system maintenance.
+
 With lower cost and time per sample, the lower precision of the method may be balanced by the ability to measure more samples from a population, i.e.
@@
 
-The accuracy of age-depth models is often driven by environmental factors and sparse data, rather than the precision of individual dates [@blaauw_double_2018] ONE MORE?.
+The accuracy of age-depth models is often driven by environmental factors and sparse data, rather than the precision of individual dates [@blaauw_double_2018].
 The lower time and cost of the hybrid method allows analysis of more samples.
 This may produce more accurate age models, despite reduced precision of individual measurements.
-Sample requirements, analysis time, precision, accuracy.
-
-ideas for further development of method
 
@@
 
+```{r message = FALSE}
+cons_df %>% 
+  group_by(Name, fm_consensus) %>% 
+  summarize(across(c(fm_corr, sig_fm_corr, fm_diff, sigma),
+                   list(mean = mean, sd = sd)),
+            N = n()) %>% 
+  ungroup() %>% 
+  arrange(fm_consensus) %>% 
+  mutate (fm_corr_err = set_errors(fm_corr_mean, fm_corr_sd),
+          Name = case_when(Name == "TIRI-I" ~ "TIRI I",
+                           Name == "C-2" ~ "IAEA C2",
+                           Name == "NOSAMS2" ~ "NOSAMS 2")) %>% 
+  select(`Sample Name` = Name, 
+         `Reference F^14^C` = fm_consensus,
+         `HGIS F^14^C ± stdev` = fm_corr_err,
+         `Mean per-sample error` = sig_fm_corr_mean,
+         N) %>% 
+  knitr::kable(digits = 3,
+               caption = "Table 1. Reference material precision and accuracy. Reference 
+               values are community consensus values for TIRI I and C2, and the 
+               NOSAMS internal graphite value for NOSAMS2. NOSAMS2 is a 
+               homogenized bivalve molusk reference. The mean of per-sample 
+               error estimates is included for comparison to the reproducibilty 
+               of the samples.")
+
+```
+
 # Figures
 
-![Figure 1.
-Inlet layout.
-Helium displacement gas is delivered to vial through a double needle, mixed sample and helium flows to an open split, and a glass capillary delivers gas to the ion source.](images/HGIS_open_split_drawing.svg)
+![Figure 1](images/HGIS_open_split_drawing.svg)
 
@@
 Performance of the NOSAMS MCSNICS hybrid gas ion source.
-(a) Modeled flows of helium and CO~2~ to the ion source during displacement of a pure CO~2~ vial sample with helium.
-(b) Measured 12C- ion current during the same dilution experiment.
+(a) Measured ^12^C^-^ ion current during displacement of a pure CO~2~ vial sample with helium.
+(b) Modeled flows of helium and CO~2~ to the ion source during the same dilution experiment.
 (c) Measured ion current for modeled CO~2~ flow.
-The stability of the source over a range of flow produces stable current for up to 40 m of measurement time.
+The stability of the source over a range of flow produces stable current for up to 40 min of measurement time.
 
@@
 (b) Fraction modern of large, radiocarbon dead samples compared to sample current.
-Blank rises with decreasing current, as expected, but may be treated as a single blank with a simple correction for full-size samples producing greater than 3 μA ^12^C^-^ current.
+Horizontal and vertical bars represent measurement uncertainty in current and ratio, respectively.
 
@@
 Expected values are spread to avoid overlapping points.
-Error bars represent one sigma sample errors.
+Error bars represent one sigma sample uncertainty.
 The black line is the mean difference from expected value for all measurements, and the blue lines are one standard deviation of all measurements.
@@
 Graphite measurements are included to allow comparison of error estimates.
-(b) Comparison of age-depth models produced using HGIS and graphite measurements.
+The bottom row of plots show a comparison of age-depth models produced using HGIS (b) and graphite measurements (c).
 

### 2022-04-04 doc/AMS-15/AMS-15_paper/figures/fig5_BChron_model.R
Don't use duplicate measurements in age model.

@@ -140,17 +140,20 @@ cal_sub_nb <- cal_sub %>%
 
 # Separate hgis and graphite
 cal_sub_hgis <- cal_sub_nb %>% 
-  filter(method == "hgis" | is.na(method))
+  filter(method == "hgis" | is.na(method),
+         !(str_starts(sample_name, "HATC"))) %>% # Remove duplicates
+  arrange(depth)
 
 cal_sub_gr <- cal_sub_nb %>% 
-  filter(method == "graphite" | is.na(method)) 
+  filter(method == "graphite" | is.na(method)) %>% 
+  arrange(depth)
 
 # Make hgis chronology
 # Some of the MCMC settings are causing a crash
 chron_hgis  <- with(cal_sub_hgis,
                     Bchronology(rc_age_corr, sig_rc_age_corr, positions = depth, 
                                 ids = sample_name, calCurves = curve,
-                                artificialThickness = 1,
+                                #artificialThickness = 1,
                                 allowOutside = TRUE
                                 #iterations = 15000,
                                 #burn = 2000,
@@ -158,6 +161,7 @@ chron_hgis  <- with(cal_sub_hgis,
                                 #extractDate = -66
                                 ))
 plot(chron_hgis) +
+  xlim(8000, -500) +
   ggtitle("HGIS chronology")
 summary(chron_hgis, "convergence")
 
@@ -165,11 +169,16 @@ summary(chron_hgis, "convergence")
 chron_gr  <- with(cal_sub_gr,
                   Bchronology(rc_age_corr, sig_rc_age_corr, positions = depth, 
                               ids = sample_name, calCurves = curve, 
-                              #artificialThickness = 1.00,
+                              #artificialThickness = 10.00,
                               allowOutside = TRUE,
                               extractDate = -66))
 plot(chron_gr) +
+  xlim(8000, -500) +
   ggtitle("Graphite chronology")
 
 #predict(chron_gr, seq(0, 900, by = 50) )
 #class(chron_gr)
+
+# Save chronologies
+write_rds(chron_hgis, here("data_analysed/hgis_chronology.rds"))
+write_rds(chron_gr, here("data_analysed/graphite_chronology.rds"))

### 2022-04-04 doc/AMS-15/AMS-15_paper/figures/fig5_core_carbonates.R

@@ -57,9 +57,11 @@ combined_results <- left_join(combined_results, depth, by = "rec_num") %>%
          sig_rc_age = rc_age - -8033 * log(fm_corr + sig_fm_corr))
 write_csv(combined_results, here("data_analysed/haiti_combined.csv"))
 
-# No outliers
-cr_no <- combined_results %>% 
-  filter(abs(fm_diff_mean) < 0.02)
+# Remove outlier: HGIS sample leaked
+# Remove HATC duplicates
+cr_no <- core %>% 
+  filter(rec_num != 171996,
+         !str_starts(sample_name, "HATC"))
 
 mean_diff <- cr_no %>% 
   filter(method == "hgis") %>% 
@@ -143,21 +145,21 @@ ggsave(here("doc/AMS-15/AMS-15_paper/figures/fig5_core_carbonates.pdf"), width =
 
 # mean 
 
-cr_no %>% 
-  group_by(method) %>% 
-  #filter(method == "hgis") %>% 
-  summarize(across(c(sig_fm_corr, fm_diff),
-                   list(mean = mean, sd = sd)),
-            N = n()) %>% 
-  select(`Consensus difference` = fm_diff_mean,
-         `Std. Dev. of consensus difference` = fm_diff_sd,
-         `per-sample error` = sig_fm_corr_mean,
-         N
-         ) %>%  
-  gt() %>% 
-  tab_header(title = "Haiti carbonates performance summary",
-             subtitle = "TIRI-I, C-2, and NOSAMS2. Values in pMC") %>% 
-  fmt_number(1:3, 
-             scale_by = 100,
-             decimals = 2)
-
+# cr_no %>% 
+#   group_by(method) %>% 
+#   #filter(method == "hgis") %>% 
+#   summarize(across(c(sig_fm_corr, fm_diff),
+#                    list(mean = mean, sd = sd)),
+#             N = n()) %>% 
+#   select(`Consensus difference` = fm_diff_mean,
+#          `Std. Dev. of consensus difference` = fm_diff_sd,
+#          `per-sample error` = sig_fm_corr_mean,
+#          N
+#          ) %>%  
+#   gt() %>% 
+#   tab_header(title = "Haiti carbonates performance summary",
+#              subtitle = "TIRI-I, C-2, and NOSAMS2. Values in pMC") %>% 
+#   fmt_number(1:3, 
+#              scale_by = 100,
+#              decimals = 2)
+# 

### 2022-04-04 doc/AMS-15/AMS-15_paper/figures/fig5_core_carbonates.R
Fix data and axis titles.

@@ -59,7 +59,7 @@ write_csv(combined_results, here("data_analysed/haiti_combined.csv"))
 
 # Remove outlier: HGIS sample leaked
 # Remove HATC duplicates
-cr_no <- core %>% 
+cr_no <- combined_results %>% 
   filter(rec_num != 171996,
          !str_starts(sample_name, "HATC"))
 
@@ -126,11 +126,15 @@ chron_gr <- read_rds(here("data_analysed/graphite_chronology.rds"))
 
 hgis_bchron <- plot(chron_hgis) +
   xlim(8000, -500) +
-  ggtitle("B")
+  labs(title = "B",
+       y = "Years BP",
+       x = "Core depth (cm)")
 
 gr_bchron <- plot(chron_gr) +
   xlim(8000, -500) +
-  ggtitle("C") +
+  labs(title = "C",
+       y = "Years BP",
+       x = "Core depth (cm)") +
   theme(axis.title.y = element_blank(),
         axis.text.y = element_blank())
 # Combine plots and save

### 2022-04-05 doc/AMS-15/AMS-15_paper/AMS-15_paper.Rmd
typo

@@
 Precision and accuracy of the method were assessed by measuring carbonate reference standards for comparison with consensus values from graphite measurements.
-We measured TIRI I, a roughly quarter-modern turbidite carbonate, IAEA C2, a half modern turbidite carbonate, and NOSAMS 2, a modern internal shell carbonate standard from a bivalve molusk REF.
+We measured TIRI I, a roughly quarter-modern turbidite carbonate, IAEA C2, a half modern turbidite carbonate, and NOSAMS 2, a modern internal shell carbonate standard from a bivalve mollusk REF.
 Summary data for these samples are given in Table 1, and measurement data for all samples is in supplementary information.

### 2022-04-05 doc/AMS-15/AMS-15_paper/AMS_HGIS_els/AMS_HGIS_els.Rmd
add elsevier rticles format

--- 2022-04-05 doc/AMS-15/AMS-15_paper/AMS-15_paper.Rmd
@@
 author:
-- Brett Longworth
-- Josh Burton
-- Simon Pendleton
-- Sydney Moser
-- Mark Roberts
-- Mark Kurz
+  - name: Brett Longworth
+    email: blongworth@whoi.edu
+    affiliation: Woods Hole Oceanographic Institution
+    correspondingauthor: true
+  - name: Josh Burton
+    email: jburton@whoi.edu
+    affiliation: Woods Hole Oceanographic Institution
+  - name: Simon Pendleton
+    email: simon.pendleton@plymouth.edu
+    affiliation: Plymouth State University
+  - name: Sydney Moser
+    email: smose007@uottawa.ca
+    affiliation: University of Ottawa
+  - name: Mark Roberts
+    email: mroberts@whoi.edu
+    affiliation: Woods Hole Oceanographic Institution
+  - name: Mark Kurz
+    email: mkurz@whoi.edu
+    affiliation: Woods Hole Oceanographic Institution
+address:
+  - code: Woods Hole Oceanographic Institution
+    address: Department of Geology and Geophysics, 360 Woods Hole Rd., Woods Hole, MA, 02540
+  - code: Plymouth State University
+    address: Department, Street, City, State, Zip
+  - code: University of Ottawa
+    address: Department, Street, City, State, Zip
+footnote:
+abstract: |
+  We have developed a method for measuring radiocarbon in carbonate minerals as CO~2~ gas via a NEC MCSNICS hybrid sputter gas ion source (HGIS).
+The method uses helium as a carrier gas to displace CO~2~ from sample vials to an open split, where a glass capillary samples the mixture for delivery directly to the HGIS.
+This method skips the gas transfer and quantification steps used in a closed inlet HGIS system, simplifying sample measurement.
+Samples larger than 8 mg carbonate can be measured.
+Results from measurements of consensus standards (TIRI-I, IAEA C2, and an internal modern shell standard), and samples from a marine core (F^14^C = 0.4 - 1.15) show that the method agrees well with traditional AMS measurement of the same samples as graphite, and that the 1$\sigma$ uncertainty is about 1 %.
+We discuss advantages and disadvantages of continuous flow sample introduction, and the effect of reduced precision on calibrated age-depth models produced using gas-source data.
+keywords: 
+  - radiocarbon
+  - carbonate
+  - hybrid gas ion source
+journal: "Nuclear Instruments and Methods in Physics Research Section B"
+date: "`r Sys.Date()`"
+classoption: preprint, 3p, authoryear
+bibliography: HGIS_paper.bib
+linenumbers: true
+numbersections: true
+# Use a CSL with `citation_package = "default"`
+# csl: https://www.zotero.org/styles/elsevier-harvard
+csl: elsevier-without-titles.csl
 output: 
-  pdf_document: default
-  word_document: default
-  html_document: default
-bibliography: HGIS_paper.bib
-csl: elsevier-without-titles.csl
+  rticles::elsevier_article:
+    keep_tex: true
+    citation_package: default
 ---
@@
 ```
-
-# Abstract
-
-We have developed a method for measuring radiocarbon in carbonate minerals as CO~2~ gas via a NEC MCSNICS hybrid sputter gas ion source (HGIS).
-The method uses helium as a carrier gas to displace CO~2~ from sample vials to an open split, where a glass capillary samples the mixture for delivery directly to the HGIS.
-This method skips the gas transfer and quantification steps used in a closed inlet HGIS system, simplifying sample measurement.
-Samples larger than 8 mg carbonate can be measured.
-Results from measurements of consensus standards (TIRI-I, IAEA C2, and an internal modern shell standard), and samples from a marine core (F^14^C = 0.4 - 1.15) show that the method agrees well with traditional AMS measurement of the same samples as graphite, and that the 1σ uncertainty is about 1 %.
-We discuss advantages and disadvantages of continuous flow sample introduction, and the effect of reduced precision on calibrated age-depth models produced using gas-source data.
 
@@
 
-The open split interface (Figure 1) supplies a helium - CO~2~ mixture to a split where a capillary delivers gas to the source and the rest is vented.
+The open split interface (Figure \ref{fig1}) supplies a helium - CO~2~ mixture to a split where a capillary delivers gas to the source and the rest is vented.
 Flowing helium is used to displace sample CO~2~ from the sample vial to the delivery capillary in the split.
@@
 
-The helium displacement flow rate was set to 250 μl min^-1^ for initial tests, and this was reduced to 100 μl min^-1^ for later work, including most reference standard comparisons and the core comparison.
-The flow rate from the open split into the source was defined by 150 cm of 50 μm fused silica capillary, split into two sections by a zero-volume bulkhead connector at the source high-voltage cage.
+The helium displacement flow rate was set to 250 $\mu$l min^-1^ for initial tests, and this was reduced to 100 $\mu$l min^-1^ for later work, including most reference standard comparisons and the core comparison.
+The flow rate from the open split into the source was defined by 150 cm of 50 $\mu$m fused silica capillary, split into two sections by a zero-volume bulkhead connector at the source high-voltage cage.
 The section inside the cage was jacketed in 1/16 in.
 PFTE tubing to help protect against spark damage.
-The total flow rate of gas to the source varied due to the viscosity difference between helium and CO~2~ and helium, and was estimated to be between 30 μl min^-1^ for pure helium and 40 μl min^-1^ for pure CO~2~ using the Hagen-Poiseuille equation.
+The total flow rate of gas to the source varied due to the viscosity difference between helium and CO~2~ and helium, and was estimated to be between 30 $\mu$l min^-1^ for pure helium and 40 $\mu$l min^-1^ for pure CO~2~ using the Hagen-Poiseuille equation.
+
+```{r fig1, fig.width = 5, fig.height = 5, fig.align='center', out.width="50%", fig.cap = "\\label{fig1}Schematic of the open split inlet.
+Two valco valves provide the ability to admit helium and live or dead tank standards.
+The double needle is housed in a VICI cross and allows displacement helium to flow into the sample vial and the sample-helium mixture to flow to the open split, where a capillary delivers the mixture to the ion source.", echo = FALSE}
+knitr::include_graphics("../figures/fig1_HGIS_open_split_drawing.pdf")
+```
 
@@
 
-curmean <- mean(currents$c12ua) %>% format(digits = 2) %>% paste("μA")
+curmean <- mean(currents$c12ua) %>% format(digits = 2)
 cursd <- sd(currents$c12ua)
@@
 We used tank CO~2~ connected through a regulator and glass capillary to the ion source to measure source performance over a range of CO~2~ flows.
-Best currents for the source was 10 μA ^12^C^-^, or roughly 1/10 of graphite target current.
-Peak current was produced at a CO~2~ flow rate of 7.5 μl min^-1^ CO~2~, but the source produced near optimal currents with flows from 7.5-25 μl min^-1^ CO~2~.
+Best currents for the source was 10 $\mu$A ^12^C^-^, or roughly 1/10 of graphite target current.
+Peak current was produced at a CO~2~ flow rate of 7.5 $\mu$l min^-1^ CO~2~, but the source produced near optimal currents with flows from 7.5-25 $\mu$l min^-1^ CO~2~.
 This is comparable with other hybrid sources [@Xu2007], although improved target design and continued optimization of source parameters may improve performance [@Salazar2013; @Fahrni2013].
-Typical currents for this study were between 4-9 μA and average current for all samples was `r curmean` ± `r cursd`.
+Typical currents for this study were between 4-9 $\mu$A and average current for all samples was `r curmean` $\mu$A ± `r cursd`.
 Several studies have shown best currents at lower CO~2~ flow rates [@Xu2007; @Ruff2010].
+
+```{r fig2, fig.width = 5, fig.height = 5, fig.align='center', out.width="50%", fig.cap = "\\label{fig2}Performance of the NOSAMS MCSNICS hybrid gas ion source.
+(a) Measured $^{12}C^-$ ion current during displacement of a pure $CO_2$ vial sample with helium.
+(b) Modeled flows of helium and $CO_2$ to the ion source during the same dilution experiment.
+(c) Measured ion current for modeled $CO_2$ flow.
+The stability of the source over a range of flow produces stable current for up to 40 min of measurement time.", echo = FALSE}
+knitr::include_graphics("../figures/fig2_current_flow.pdf")
+```
 
@@
 A contrast of the open split approach with closed-source HGIS methods is changing CO~2~ flow and helium dilution ratio over the period of measurement due to the dilution of the sample by the displacement helium.
-We examined the effect of this variable carbon delivery and dilution by measuring ^12^C^-^ current evolution during helium displacement/dilution of vial CO~2~ (Figure 2a) and comparing this to the composition of the gas delivered to the source.
-We estimated gas composition (Figure 2b) using a simple mixing model:
+We examined the effect of this variable carbon delivery and dilution by measuring ^12^C^-^ current evolution during helium displacement/dilution of vial CO~2~ (Figure \ref{fig2}a) and comparing this to the composition of the gas delivered to the source.
+We estimated gas composition (Figure \ref{fig2}b) using a simple mixing model:
 
@@
 The HGIS method produces higher ratios for radiocarbon dead samples than the graphite method.
-Measurements of radiocarbon-free reference materials (IAEA C1 marble and TIRI F spar calcite) show that the average normalized ratio for these samples is `r blankmean` ± `r blanksd` (1σ uncertainty, N=`r nblank`).
+Measurements of radiocarbon-free reference materials (IAEA C1 marble and TIRI F spar calcite) show that the average normalized ratio for these samples is `r blankmean` ± `r blanksd` (1$\sigma$ uncertainty, N=`r nblank`).
 We investigated the sources of this method blank as a current-dependent blank and as a current-independent blank, and implemented a blank correction for the method.
@@
 HGIS cathodes produce measurable ion current, even with no gas flow.
-This current is initially \> 10 μA and drops exponentially, reaching a baseline of ~100 nA after 5-10 minutes.
-For the range of currents in this study (3-9 μA), this baseline current represents a 1-6% contribution to the total measured sample current at the F^14^C of the target current baseline.
+This current is initially \> 10 $\mu$A and drops exponentially, reaching a baseline of ~100 nA after 5-10 minutes.
+For the range of currents in this study (3-9 $\mu$A), this baseline current represents a 1-6% contribution to the total measured sample current at the F^14^C of the target current baseline.
 This can be interpreted as a constant current dependent blank, analogous to a mass-dependent "constant contaminant" blank and may be modeled with the same tools [@haghipour_compound-specific_2019; @hanke_comprehensive_2017; @roberts_14c_2019; @sun_14_2020].
@@
 Based on this assessment, we estimate the ratio of the blank as having F^14^C of 0.43 with a current of 110 nA.
+
+```{r fig3, fig.width = 5, fig.height = 5, fig.align='center', out.width="50%", fig.cap = "\\label{fig3}Assessment of blanks.
+(a) Current dependent blank model and data.
+$F^{14}C$ and current of measurements from dilution tests of live and dead gas are used to generate linear fits which should cross at the $F^{14}C$ and current of the constant current blank.
+Direct determination of $F^{14}C$ and current of the blank are also shown.
+(b) Fraction modern of large, radiocarbon dead samples compared to sample current.
+Horizontal and vertical bars represent measurement uncertainty in current and ratio, respectively.", echo = FALSE}
+knitr::include_graphics("../figures/fig3_blanks.pdf")
+```
 
@@
 
-# Summary table
-# cons_df %>% 
-#   group_by(Name, fm_consensus) %>% 
-#   summarize(across(c(fm_corr, sig_fm_corr, fm_diff, sigma),
-#                    list(mean = mean, sd = sd)),
-#             N = n()) %>% 
-#   ungroup() %>% 
-#   arrange(fm_consensus) %>% 
-#   mutate (fm_corr_err = set_errors(fm_corr_mean, fm_corr_sd),
-#           Name = case_when(Name == "TIRI-I" ~ "TIRI I",
-#                            Name == "C-2" ~ "IAEA C2",
-#                            Name == "NOSAMS2" ~ "NOSAMS 2")) %>% 
-#   select(`Sample Name` = Name, 
-#          `Reference F^14^C` = fm_consensus,
-#          `HGIS F^14^C ± stdev` = fm_corr_err,
-#          `Mean per-sample error` = sig_fm_corr_mean,
-#          N) %>% 
-#   knitr::kable(digits = 3,
-#                caption = "Reference material precision and accuracy. Reference 
-#                values are community consensus values for TIRI I and C2, and the 
-#                NOSAMS internal graphite value for NOSAMS2. NOSAMS2 is a 
-#                homogenized bivalve molusk reference. The mean of per-sample 
-#                error estimates is included for comparison to the reproducibilty 
-#                of the samples.")
-
-  # gt() %>% 
-  # tab_header(title = "Reference material precision and accuracy",
-  #            subtitle = "Values in pMC, Errors are SD of measurements") %>% 
-  # fmt_number(4, 
-  #            decimals = 3) %>% 
-  # cols_align("right", 3)
-
 # Consensus summary for inline data
@@
 Mean difference from consensus for all measurements was F^14^C `r format(cons_sum$fm_diff_mean, digits = 1, nsmall = 4)` shows that the hybrid method has no bias in aggregate.
-The standard deviation of the difference in F^14^C was `r format(cons_sum$fm_diff_sd, nsmall = 4)`, which may be taken as an estimate of the 1σ precision of the method.
+The standard deviation of the difference in F^14^C was `r format(cons_sum$fm_diff_sd, nsmall = 4)`, which may be taken as an estimate of the 1$\sigma$ precision of the method.
 The mean of propagated measurement errors for these samples was `r format(cons_sum$sig_fm_corr_mean, nsmall = 4)` F^14^C for these samples, which is in good agreement with the method precision above.
@@
 
+```{r tab1, message = FALSE}
+# Summary table
+cons_df %>% 
+  group_by(Name, fm_consensus) %>% 
+  summarize(across(c(fm_corr, sig_fm_corr, fm_diff, sigma),
+                   list(mean = mean, sd = sd)),
+            N = n()) %>% 
+  ungroup() %>% 
+  arrange(fm_consensus) %>% 
+  mutate (fm_corr_err = set_errors(fm_corr_mean, fm_corr_sd),
+          Name = case_when(Name == "TIRI-I" ~ "TIRI I",
+                           Name == "C-2" ~ "IAEA C2",
+                           Name == "NOSAMS2" ~ "NOSAMS 2")) %>% 
+  select(`Sample Name` = Name, 
+         `Reference F^14^C` = fm_consensus,
+         `HGIS F^14^C ± stdev` = fm_corr_err,
+         `Mean per-sample error` = sig_fm_corr_mean,
+         N) %>% 
+  knitr::kable(digits = 3,
+               caption = "\\label{tab1}Reference material precision and accuracy. Reference 
+               values are community consensus values for TIRI I and C2, and the 
+               NOSAMS internal graphite value for NOSAMS2. NOSAMS2 is a 
+               homogenized bivalve molusk reference. The mean of per-sample 
+               error estimates is included for comparison to the reproducibilty 
+               of the samples.")
+
+```
+
 The difference between reference samples and graphite consensus does not vary systematically by F^14^C, indicating that the method has no significant bias with radiocarbon content (Figure 4).
-Each reference standard type agrees with graphite consensus values within the standard deviation of samples measured, with slightly more measurements agreeing with graphite within 1σ uncertainties than would be expected statistically.
+Each reference standard type agrees with graphite consensus values within the standard deviation of samples measured, with slightly more measurements agreeing with graphite within 1$\sigma$ uncertainties than would be expected statistically.
+
+```{r fig4, fig.width = 5, fig.height = 5, fig.align='center', out.width="50%", fig.cap = "\\label{fig4}Agreement of standards with graphite consensus values.
+Difference of HGIS measurements from consensus for individual samples are plotted against their expected value.
+Expected values are spread to avoid overlapping points.
+Error bars represent one sigma sample uncertainty.
+The black line is the mean difference from expected value for all measurements, and the blue lines are one standard deviation of all measurements.", echo = FALSE}
+knitr::include_graphics("../figures/fig4_consensus_samples.pdf")
+```
 
@@
 Mean difference between hybrid and graphite measurements was -0.003 F^14^C and the standard deviation of this difference was 0.009 F^14^C.
-Mean per-sample 1σ uncertainty was 0.010 F^14^C for HGIS measurements, compared with 0.002 F^14^C for graphite.
+Mean per-sample 1$\sigma$ uncertainty was 0.010 F^14^C for HGIS measurements, compared with 0.002 F^14^C for graphite.
 Three samples were analysed by HGIS in duplicate.
-Two of the three pairs agreed within 1σ and all three agreed within 2σ.
+Two of the three pairs agreed within 1$\sigma$ and all three agreed within 2$\sigma$.
 There is an apparent, but not significant positive relationship between the HGIS-graphite difference and measured value.
 The higher graphite values at lower F^14^C may indicate an overcorrection of HGIS blank for these samples.
+
+```{r fig5, fig.width = 5, fig.height = 5, fig.align='center', out.width="50%", fig.cap = "\\label{fig5}Comparison of Haiti core samples measured as graphite and via HGIS.
+(a) Difference between HGIS and graphite measurements of the same materials.
+Difference between HGIS and graphite measurements is shown on the vertical axis and plotted against the mean of the two measurements.
+Samples are offset on the horizontal axis for clarity.
+Error bars are measurement error for each sample.
+Graphite measurements are included to allow comparison of error estimates.
+The bottom row of plots show a comparison of age-depth models produced using HGIS (b) and graphite measurements (c).", echo = FALSE}
+knitr::include_graphics("../figures/fig5_core_carbonates.pdf")
+```
 
@@
 
-# Tables
-
-```{r message = FALSE}
-cons_df %>% 
-  group_by(Name, fm_consensus) %>% 
-  summarize(across(c(fm_corr, sig_fm_corr, fm_diff, sigma),
-                   list(mean = mean, sd = sd)),
-            N = n()) %>% 
-  ungroup() %>% 
-  arrange(fm_consensus) %>% 
-  mutate (fm_corr_err = set_errors(fm_corr_mean, fm_corr_sd),
-          Name = case_when(Name == "TIRI-I" ~ "TIRI I",
-                           Name == "C-2" ~ "IAEA C2",
-                           Name == "NOSAMS2" ~ "NOSAMS 2")) %>% 
-  select(`Sample Name` = Name, 
-         `Reference F^14^C` = fm_consensus,
-         `HGIS F^14^C ± stdev` = fm_corr_err,
-         `Mean per-sample error` = sig_fm_corr_mean,
-         N) %>% 
-  knitr::kable(digits = 3,
-               caption = "Table 1. Reference material precision and accuracy. Reference 
-               values are community consensus values for TIRI I and C2, and the 
-               NOSAMS internal graphite value for NOSAMS2. NOSAMS2 is a 
-               homogenized bivalve molusk reference. The mean of per-sample 
-               error estimates is included for comparison to the reproducibilty 
-               of the samples.")
-
-```
-
-# Figures
-
-![Figure 1](images/HGIS_open_split_drawing.svg)
-
-![Figure 2](figures/fig2_current_flow.svg)
-
-![Figure 3](figures/fig3_blanks.svg)
-
-![Figure 4](figures/fig4_consensus_samples.svg)
-
-![Figure 5](figures/fig5_core_carbonates.svg)
-
-# Figure captions
-
-Figure 1.
-Schematic of the open split inlet.
-Two valco valves provide the ability to admit helium and live or dead tank standards.
-The double needle is housed in a VICI cross and allows displacement helium to flow into the sample vial and the sample-helium mixture to flow to the open split, where a capillary delivers the mixture to the ion source.
-
-Figure 2.
-Performance of the NOSAMS MCSNICS hybrid gas ion source.
-(a) Measured ^12^C^-^ ion current during displacement of a pure CO~2~ vial sample with helium.
-(b) Modeled flows of helium and CO~2~ to the ion source during the same dilution experiment.
-(c) Measured ion current for modeled CO~2~ flow.
-The stability of the source over a range of flow produces stable current for up to 40 min of measurement time.
-
-Figure 3.
-Assessment of blanks.
-(a) Current dependent blank model and data.
-F^14^C and current of measurements from dilution tests of live and dead gas are used to generate linear fits which should cross at the F^14^C and current of the constant current blank.
-Direct determination of F^14^C and current of the blank are also shown.
-(b) Fraction modern of large, radiocarbon dead samples compared to sample current.
-Horizontal and vertical bars represent measurement uncertainty in current and ratio, respectively.
-
-Figure 4.
-Agreement of standards with graphite consensus values.
-Difference of HGIS measurements from consensus for individual samples are plotted against their expected value.
-Expected values are spread to avoid overlapping points.
-Error bars represent one sigma sample uncertainty.
-The black line is the mean difference from expected value for all measurements, and the blue lines are one standard deviation of all measurements.
-
-Figure 5.
-Comparison of Haiti core samples measured as graphite and via HGIS.
-(a) Difference between HGIS and graphite measurements of the same materials.
-Difference between HGIS and graphite measurements is shown on the vertical axis and plotted against the mean of the two measurements.
-Samples are offset on the horizontal axis for clarity.
-Error bars are measurement error for each sample.
-Graphite measurements are included to allow comparison of error estimates.
-The bottom row of plots show a comparison of age-depth models produced using HGIS (b) and graphite measurements (c).
-
-# References
+# References {-}

### 2022-04-07 doc/AMS-15/AMS-15_paper/AMS_HGIS_els/AMS_HGIS_els.Rmd
Update refs, minor edits.

@@
 ```{r fig1, fig.width = 5, fig.height = 5, fig.align='center', out.width="50%", fig.cap = "\\label{fig1}Schematic of the open split inlet.
-Two valco valves provide the ability to admit helium and live or dead tank standards.
+Two valco 2-way valves provide the ability to admit helium and live or dead tank standards.
 The double needle is housed in a VICI cross and allows displacement helium to flow into the sample vial and the sample-helium mixture to flow to the open split, where a capillary delivers the mixture to the ion source.", echo = FALSE}
@@
 Sealed vials were evacuated, then 1 ml 85% phosphoric acid was added.
-The samples were allowed to react at 60 C for 4-6 h.
+The samples were allowed to react at 60 C for 6 h.
 Sample pressure was adjusted to 1 atm by allowing the sample to equilibrate with a helium-filled syringe connected through the vial septum.
@@
 First, each 90 s block was corrected for fractionation using the AMS ^13/12^C measurement.
-The raw ^14/12^C ratio divided by the ^13/12^C ratio squared gives the corrected ^14/12^C ratio.
 Blocks that are more than 1.5 times the interquartile range (IQR) below the first quartile of all measurements of a sample or more than 1.5 IQR above the third quartile are considered outliers and were dropped at this point.
-Remaining runs of a sample were averaged and per-sample error was taken as the greater of counting error or the standard error of the measurements.
+Remaining runs of a sample were averaged and sample uncertainty was taken as the greater of counting error or the standard error of the measurements.
 The sample ratio was then normalized using the mean of gas standards measured during the run.
@@
 curmean <- mean(currents$c12ua) %>% format(digits = 2)
-cursd <- sd(currents$c12ua)
+cursd <- sd(currents$c12ua) %>% format(digits = 2)
 ```
@@
 This is comparable with other hybrid sources [@Xu2007], although improved target design and continued optimization of source parameters may improve performance [@Salazar2013; @Fahrni2013].
-Typical currents for this study were between 4-9 $\mu$A and average current for all samples was `r curmean` $\mu$A ± `r cursd`.
+Typical currents for this study were between 4-9 $\mu$A and average current for all samples was `r curmean` ± `r cursd` $\mu$A.
 Several studies have shown best currents at lower CO~2~ flow rates [@Xu2007; @Ruff2010].
@@
 blankmean <- mean(blanks$norm_ratio) %>% format(digits = 3)
-blanksd <-  sd(blanks$norm_ratio) %>% format(digits = 3)
+blanksd <-  sd(blanks$norm_ratio) %>% format(digits = 2)
 
@@
 
-The HGIS method produces higher ratios for radiocarbon dead samples than the graphite method.
-Measurements of radiocarbon-free reference materials (IAEA C1 marble and TIRI F spar calcite) show that the average normalized ratio for these samples is `r blankmean` ± `r blanksd` (1$\sigma$ uncertainty, N=`r nblank`).
+The HGIS method produced higher ratios for radiocarbon dead samples than the graphite method.
+The average normalized ratio for radiocarbon-free reference materials IAEA C1 marble and TIRI F spar calcite was `r blankmean` ± `r blanksd` (1$\sigma$ uncertainty, N=`r nblank`).
 We investigated the sources of this method blank as a current-dependent blank and as a current-independent blank, and implemented a blank correction for the method.
@@
 HGIS cathodes produce measurable ion current, even with no gas flow.
-This current is initially \> 10 $\mu$A and drops exponentially, reaching a baseline of ~100 nA after 5-10 minutes.
+This current was initially \> 10 $\mu$A and dropped exponentially, reaching a baseline of ~100 nA after 5-10 minutes.
 For the range of currents in this study (3-9 $\mu$A), this baseline current represents a 1-6% contribution to the total measured sample current at the F^14^C of the target current baseline.
@@
 Fits from modern and dead dilution tests should trend towards the fraction modern of the blank, crossing at the inverse of blank current.
-Based on this assessment, we estimate the ratio of the blank as having F^14^C of 0.43 with a current of 110 nA.
+Based on this assessment, we estimate the F^14^C of the blank as 0.43 with a current of 110 nA.
 
@@
 
-The directly measured blank current (120 nA) agrees well with the model assessment, but has a higher fraction modern (F^14^C = 0.82).
-A combination of low measurement precision for the direct measurement and high sensitivity of the model to variability in the measured fraction modern and current in the depletion samples may explain this difference[@sun_14_2020].
+The directly measured blank current (120 nA) agrees well with the model assessment, but has higher F^14^C (0.82).
+A combination of low measurement precision for the direct measurement and high sensitivity of the model to variability in the measured fraction modern and current in the depletion samples may explain this difference [@sun_14_2020].
 This constant contamination blank is consistent with the fact that hybrid cathodes produce a minimum current with or without gas flow to the source.
@@
                values are community consensus values for TIRI I and C2, and the 
-               NOSAMS internal graphite value for NOSAMS2. NOSAMS2 is a 
-               homogenized bivalve molusk reference. The mean of per-sample 
+               NOSAMS internal graphite value for NOSAMS2. The mean of per-sample 
                error estimates is included for comparison to the reproducibilty 
@@
 A suite of 15 carbonate macrofossils consisting of shell and coral fragments from a series of coastal marine cores @moser_multi-proxy_2021 were measured as graphite and as CO~2~ using the HGIS method to further test the method and serve to demonstrate a potential application.
-Sediment cores were collected from Baie des Baradères, Haiti during a scientific cruise in May 2016 lead by Dr.
-Jeffrey Donnelly (WHOI).
+Sediment cores were collected from Baie des Baradères, Haiti in May 2016.
 Coring was performed with a Rossfelder vibracorer from a 20-foot pontoon boat.
-The cores were collected from a coastal karst basin (CKB) at vertical depth ~11.7 m and coordinates 18.514° N, 73.629° W.
+The cores were collected from a coastal karst basin at vertical depth ~11.7 m and coordinates 18.514° N, 73.629° W.
 Core HATC1 D1 is 145 cm long and was collected in a 3-inch poly tube as a control directly beside core HATC1 D2, which is a continuous 896 cm core collected in a 3-inch aluminum tube.
@@
 We used the Bchron Baesian radiochronology package [@haslett_simple_2008] to develop these models.
-We applied a correction for the regional marine reservoir effect with a delta R of -216 and uncertainty 92 based on the oxcal model with 10 reservoir points.
+We applied a correction for the regional marine reservoir effect with a $\Delta$ R of -216 ± 92 years based on the Oxcal model [@ramsey_bayesian_2009] with 10 reservoir points.
 Samples with pre-bomb radiocarbon ages were calibrated using the Marine20 curve @heaton_marine20marine_2020.

### 2022-04-10 doc/AMS-15/AMS-15_paper/AMS_HGIS_els/AMS_HGIS_els.Rmd
Fix references, start on edits.

@@
 Analysis of CO~2~ directly via a gas-accepting AMS ion source eliminates the cost and effort of graphitization and returns results more rapidly.
-The NOSAMS USAMS system at the National Oceanographic AMS lab (NOSAMS) is based on a 3MV Tandetron accelerator [@Longworth2015] combined with a NEC 134-sample gas-capable MCSNICS source [@Xu2007] .
+The NOSAMS USAMS system at the National Oceanographic AMS lab (NOSAMS) is based on a 3MV Tandetron accelerator [@Longworth2015] combined with a National Electrostatics Corporation (NEC) 134-sample gas-capable MCSNICS source [@Xu2007] .
 We present here a system for semi-automated analysis of carbonate minerals by acid hydrolysis in vials and introduction to the ion source in a flow of helium via an open split.
@@
 
-```{r fig1, fig.width = 5, fig.height = 5, fig.align='center', out.width="50%", fig.cap = "\\label{fig1}Schematic of the open split inlet.
+```{r fig1, fig.width = 5, fig.height = 5, fig.align='center', out.width="80%", fig.cap = "\\label{fig1}Schematic of the open split inlet.
 Two valco 2-way valves provide the ability to admit helium and live or dead tank standards.
 The double needle is housed in a VICI cross and allows displacement helium to flow into the sample vial and the sample-helium mixture to flow to the open split, where a capillary delivers the mixture to the ion source.", echo = FALSE}
-knitr::include_graphics("../figures/fig1_HGIS_open_split_drawing.pdf")
+knitr::include_graphics(here("doc/AMS-15/AMS-15_paper/figures/fig1_HGIS_open_split_drawing.pdf"))
 ```
@@
 Sample pressure was adjusted to 1 atm by allowing the sample to equilibrate with a helium-filled syringe connected through the vial septum.
-Evacuation, acid addition, and heated reaction were performed by the automated gas handling system developed for the CFAMS GIS carbonate system [@Roberts2013].
+Evacuation, acid addition, and heated reaction were performed by the automated gas handling system developed by Roberts et al.
+[@Roberts2013].
 
@@
 The delivery needle was then transferred to the sample vial, which started sample delivery to the split and source.
-Currents were allowed to stabilize for 30 s , at which point the data acquisition was started.
+Currents were allowed to stabilize for 30 s, at which point the data acquisition was started.
 Data was acquired in 10 blocks of 90 s each, for 15 min total acquisition time.
@@
 First, each 90 s block was corrected for fractionation using the AMS ^13/12^C measurement.
-Blocks that are more than 1.5 times the interquartile range (IQR) below the first quartile of all measurements of a sample or more than 1.5 IQR above the third quartile are considered outliers and were dropped at this point.
+Blocks that are more than 1.5 times the interquartile range (IQR) below the first quartile of all measurements of a sample or more than 1.5 IQR above the third quartile are considered outliers and were not further used in the analysis.
 Remaining runs of a sample were averaged and sample uncertainty was taken as the greater of counting error or the standard error of the measurements.
@@
 
-# Results
+# Results and Discussion
 
@@
 We used tank CO~2~ connected through a regulator and glass capillary to the ion source to measure source performance over a range of CO~2~ flows.
-Best currents for the source was 10 $\mu$A ^12^C^-^, or roughly 1/10 of graphite target current.
+Best currents for the source was 10 $\mu$A ^12^C^-^, or roughly 10% of graphite target current.
 Peak current was produced at a CO~2~ flow rate of 7.5 $\mu$l min^-1^ CO~2~, but the source produced near optimal currents with flows from 7.5-25 $\mu$l min^-1^ CO~2~.
@@
 The stability of the source over a range of flow produces stable current for up to 40 min of measurement time.", echo = FALSE}
-knitr::include_graphics("../figures/fig2_current_flow.pdf")
+knitr::include_graphics(here("doc/AMS-15/AMS-15_paper/figures/fig2_current_flow.pdf"))
 ```
@@
 Horizontal and vertical bars represent measurement uncertainty in current and ratio, respectively.", echo = FALSE}
-knitr::include_graphics("../figures/fig3_blanks.pdf")
+knitr::include_graphics(here("doc/AMS-15/AMS-15_paper/figures/fig3_blanks.pdf"))
 ```
@@
 The black line is the mean difference from expected value for all measurements, and the blue lines are one standard deviation of all measurements.", echo = FALSE}
-knitr::include_graphics("../figures/fig4_consensus_samples.pdf")
+knitr::include_graphics(here("doc/AMS-15/AMS-15_paper/figures/fig4_consensus_samples.pdf"))
 ```
@@
 
-```{r fig5, fig.width = 5, fig.height = 5, fig.align='center', out.width="50%", fig.cap = "\\label{fig5}Comparison of Haiti core samples measured as graphite and via HGIS.
+```{r fig5, fig.width = 5, fig.height = 5, fig.align='center', out.width="80%", fig.cap = "\\label{fig5}Comparison of Haiti core samples measured as graphite and via HGIS.
 (a) Difference between HGIS and graphite measurements of the same materials.
@@
 The bottom row of plots show a comparison of age-depth models produced using HGIS (b) and graphite measurements (c).", echo = FALSE}
-knitr::include_graphics("../figures/fig5_core_carbonates.pdf")
+knitr::include_graphics(here("doc/AMS-15/AMS-15_paper/figures/fig5_core_carbonates.pdf"))
 ```
@@
 
-# Discussion
+# Conclusions
 

### 2022-04-11 doc/AMS-15/AMS-15_paper/AMS_HGIS_els/AMS_HGIS_els.Rmd
Add MR edits

@@
   - code: Plymouth State University
-    address: Department, Street, City, State, Zip
+    address: Environmental Science and Policy, 17 High St., Plymouth, MA, 03264
   - code: University of Ottawa
-    address: Department, Street, City, State, Zip
-footnote:
+    address: Geography, Environment and Geomatics, 550 Cumberland St., Ottawa, ON, Canada, K1N 6N5
 abstract: |
@@
 classoption: preprint, 3p, authoryear
-bibliography: HGIS_paper.bib
-linenumbers: true
+bibliography: ../HGIS_paper.bib
+linenumbers: false
 numbersections: true
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
 ```
@@
 We designed a open inlet where a capillary samples gas for delivery to the ion source in a flowing stream that exhausts to the atmosphere, often called an open split.
-This defines the delivery pressure as 1 ATM, and therefore flow is controlled by the capillary parameters and is not easily adjustable during measurement.
-The flow of CO~2~ to the source may be varied by controlling the ratio of CO~2~ to helium in the sample stream.
-Changing the ratio of CO~2~ to helium changes the partial pressure of CO~2~ and therefore the carbon flow rate to the source.
+This defines the delivery pressure as 1 ATM.
+CO~2~ flow to the source can be controlled by changing the length or diameter of the capillary, or by changing the ratio of CO~2~ to He in the sample stream.
 
@@
 Our open-split interface is constructed of simple, off-the-shelf parts.
-The core of the design is a dual needle and split housed in a VICI cross fitting.
-Displacement helium flows in through one side of the cross and into the sample vial through 1/16 in.
+The core of the design is a dual needle and split housed in a Valco tee fitting.
+Displacement helium flows in through the sidearm of the tee and into the sample vial through 1/16 in.
 stainless steel tubing, which forms the outer flow path of the concentric dual needle.
 
-The mixture of sample CO~2~ and displacement helium from the vial flows up a 21 gauge (0.819 mm) needle inserted through the cross and stainless tubing which forms the inner flow path of the dual needle.
+The mixture of sample CO~2~ and displacement helium from the vial flows up a 21 gauge (0.819 mm) needle inserted through the tee and stainless tubing which forms the inner flow path of the dual needle.
 The open split is a VICI Luer adapter (part ZBUMLPK) with 5cm of 1/16 in.
@@
 The split has fairly low volume (< 1 mL).
-The system was operated manually during development, but may be mounted on the arm of a gasbench-style autosampler to enable automated operation.
+The system was operated manually during development, but may be mounted on the arm of a vial-sampling autosampler to enable automated operation.
 
@@
 The helium displacement flow rate was set to 250 $\mu$l min^-1^ for initial tests, and this was reduced to 100 $\mu$l min^-1^ for later work, including most reference standard comparisons and the core comparison.
-The flow rate from the open split into the source was defined by 150 cm of 50 $\mu$m fused silica capillary, split into two sections by a zero-volume bulkhead connector at the source high-voltage cage.
+The flow rate from the open split into the source was defined by 150 cm of 50 $\mu$m diameter fused silica capillary, split into two sections by a zero-volume bulkhead connector at the source high-voltage cage.
 The section inside the cage was jacketed in 1/16 in.
 PFTE tubing to help protect against spark damage.
-The total flow rate of gas to the source varied due to the viscosity difference between helium and CO~2~ and helium, and was estimated to be between 30 $\mu$l min^-1^ for pure helium and 40 $\mu$l min^-1^ for pure CO~2~ using the Hagen-Poiseuille equation.
+The total flow rate of gas to the source varied between 30 $\mu$l min^-1^ for pure helium and 40 $\mu$l min^-1^ for pure CO~2~ due to the viscosity difference between helium and CO~2~, estimated using the Hagen-Poiseuille equation.
 
@@
 Two valco 2-way valves provide the ability to admit helium and live or dead tank standards.
-The double needle is housed in a VICI cross and allows displacement helium to flow into the sample vial and the sample-helium mixture to flow to the open split, where a capillary delivers the mixture to the ion source.", echo = FALSE}
+The double needle is housed in a tee fitting and allows displacement helium to flow into the sample vial and the sample-helium mixture to flow to the open split, where a capillary delivers the mixture to the ion source.", echo = FALSE}
 knitr::include_graphics(here("doc/AMS-15/AMS-15_paper/figures/fig1_HGIS_open_split_drawing.pdf"))
@@
 
-Carbonate samples were weighed into 7 ml septa-sealed Vacutainer vials (Labco 768W).
+Carbonate samples were weighed into 7 ml septa-sealed glass vials (Labco 768W).
 Sealed vials were evacuated, then 1 ml 85% phosphoric acid was added.
@@
 
-Standards were prepared by evacuating vials and then filling with a modern standard CO~2~ using the automated gas handling system.
+Normalizing standards were prepared by evacuating vials and then filling with a modern standard CO~2~ using the automated gas handling system.
 Standards were adjusted to the same size as other samples by extracting an appropriate volume of CO~2~ from the full vial and replacing with helium using a syringe.
-Process blank samples were prepared from powdered C-1 or TIRI-F, acidified in vials using the method above.
+Process blank samples were prepared from powdered IAEA C1 or TIRI F, acidified in vials using the method above.
 
@@
 Samples were measured on a modified 3MV Tandetron AMS system [@Longworth2015].
-The displacement helium was allowed to flow constantly during analysis, and the sample needle was inserted into a helium-filled vial between samples.
 Measurement of a sample proceeded as follows: a clean sputter target cathode was rotated into position and the target was pre-cleaned by sputtering for 5 minutes or until cathode ^12^C^-^ current fell below 100 nA.
-The delivery needle was then transferred to the sample vial, which started sample delivery to the split and source.
+The delivery needle was then transferred to the sample vial, which started gas delivery to the split and source.
 Currents were allowed to stabilize for 30 s, at which point the data acquisition was started.
 Data was acquired in 10 blocks of 90 s each, for 15 min total acquisition time.
+The displacement helium was allowed to flow constantly during cathode cleaning and analysis, and the sample needle was inserted into a helium-filled vial between samples.
 Normalizing standards, blanks, and secondary standards were measured identically to unknowns.
-Total time for a measurement was 20-25 min including target pre-cleaning and sample changes.
+Total time for a measurement was 20-25 min including target pre-cleaning and changing samples.
 
@@
 
-Measurements were corrected for process and measurement blank using a linear correction that assumes that the corrected fraction modern of blank samples should be zero, samples with the fraction modern of the standard need no correction, and applies a proportional correction to samples of intermediate fraction modern [@donahue_isotope-ratio_1990; @roberts_14c_2019].
+Measurements were corrected for process and measurement blank using a linear correction that assumes that the corrected ratio of blank samples should be zero, samples with the same ratio as the standard need no correction, and samples of intermediate ratio need a proportional correction [@donahue_isotope-ratio_1990; @roberts_14c_2019].
 The correction uses measured values of standard gas and blank samples for correction of contaminant carbon and machine blank.
@@
 This correction does not account for the effect of a blank with constant mass or current on samples of varying mass or current.
-For all measurements used to assess the performance of the method and for the Haiti core comparison, sample, standard, and blank masses were kept constant.
+For all measurements used to assess the precision and accuracy of the method, sample, standard, and blank masses were kept constant.
 Sample current did vary, but we believe the linear correction described above produces the most accurate results within this range.
@@
 
-```{r fig2, fig.width = 5, fig.height = 5, fig.align='center', out.width="50%", fig.cap = "\\label{fig2}Performance of the NOSAMS MCSNICS hybrid gas ion source.
+```{r fig2, fig.width = 5, fig.height = 5, fig.align='center', out.width="80%", fig.cap = "\\label{fig2}Performance of the NOSAMS MCSNICS hybrid gas ion source.
 (a) Measured $^{12}C^-$ ion current during displacement of a pure $CO_2$ vial sample with helium.
@@
 
-$$C(t) = e^{-\frac{r}{V}t}$$
+\begin{equation}
+\label{eq1}
+C(t) = e^{-\frac{r}{V}t}
+\end{equation}
 
@@
 Figure 2c shows the resulting relationship of sample ion current to CO~2~ flow rate.
-The most important feature of the current-flow relationship is the wide range of flows under which stable ion current is measured, which in turn produces relatively stable currents and ratios despite a changing carbon flow rate during measurement.
+The most important feature of the current-flow relationship is the wide range of CO~2~ flow rates over which stable ion current is measured, which in turn produces relatively stable currents and ratios despite a changing carbon flow rate during measurement.
 Several studies have measured peak currents over a narrower range of flows [@Fahrni2013; @broek_optimization_nodate] at high helium dilution rates (> 10:1), while Xu et.
@@
 Broek et.
-al [@broek_optimization_nodate] found peak broadening with decreasing dilution ratio.
+al [@broek_optimization_nodate] found that the range of flow rates with peak currents increased with decreasing dilution ratio.
 While lower rates of dilution used in our HGIS method may decrease current slightly, the observed peak broadening allows stable measurement over the range of CO~2~ flows produced by the open split.
 
-Work with currents and flows (above and Figure 2) predicts that optimal currents should be produced until less than 2 ml CO~2~ (equivalent to 0.8 mg C) remains in the sample vial.
+Work with currents and flows (above and Figure 2) shows that stable currents are produced until less than 2 ml CO~2~ (0.8 mg C) remains in the sample vial.
 Tests of a range of sample sizes support this estimate and show that no reduction in measured current is seen in samples larger than 1 mg C which corresponds to 8 mg CaCO~3~ sample material.
 To ensure ample sample for optimal measurements, we chose 15 mg as a target sample mass.
-To ensure that samples started at or slightly above atmospheric pressure, we injected clean helium with a syringe to equilibrate pressure in the vials when the reaction with acid was complete.
+To ensure that samples started at or slightly above atmospheric pressure, we injected helium with a syringe to equilibrate pressure in the vials when the reaction with acid was complete.
 
@@
 The HGIS method produced higher ratios for radiocarbon dead samples than the graphite method.
-The average normalized ratio for radiocarbon-free reference materials IAEA C1 marble and TIRI F spar calcite was `r blankmean` ± `r blanksd` (1$\sigma$ uncertainty, N=`r nblank`).
+The average uncorrected $F^{14}C$ for radiocarbon-free reference materials IAEA C1 marble and TIRI F spar calcite was `r blankmean` ± `r blanksd` (1$\sigma$ uncertainty, N=`r nblank`).
 We investigated the sources of this method blank as a current-dependent blank and as a current-independent blank, and implemented a blank correction for the method.
@@
 
-```{r fig3, fig.width = 5, fig.height = 5, fig.align='center', out.width="50%", fig.cap = "\\label{fig3}Assessment of blanks.
+```{r fig3, fig.width = 5, fig.height = 5, fig.align='center', out.width="80%", fig.cap = "\\label{fig3}Assessment of blanks.
 (a) Current dependent blank model and data.
@@
 
-$$ R_{samp} = R_{meas} - \frac{R_{blank}(R_{std}-R_{meas})}{R_{std}} $$
+\begin{equation}
+\label{eq2}
+R_{samp} = R_{meas} - \frac{R_{blank}(R_{std}-R_{meas})}{R_{std}}
+\end{equation}
 
@@
 
-## Reference materials
+## Standard reference materials
 
@@
 Precision and accuracy of the method were assessed by measuring carbonate reference standards for comparison with consensus values from graphite measurements.
-We measured TIRI I, a roughly quarter-modern turbidite carbonate, IAEA C2, a half modern turbidite carbonate, and NOSAMS 2, a modern internal shell carbonate standard from a bivalve mollusk REF.
+We measured TIRI I, a roughly quarter-modern turbidite carbonate, IAEA C2, a half modern turbidite carbonate, and NOSAMS 2, a modern internal shell carbonate standard from a bivalve mollusk.
 Summary data for these samples are given in Table 1, and measurement data for all samples is in supplementary information.
@@
 
-```{r fig4, fig.width = 5, fig.height = 5, fig.align='center', out.width="50%", fig.cap = "\\label{fig4}Agreement of standards with graphite consensus values.
+```{r fig4, fig.width = 5, fig.height = 5, fig.align='center', out.width="80%", fig.cap = "\\label{fig4}Agreement of standards with graphite consensus values.
 Difference of HGIS measurements from consensus for individual samples are plotted against their expected value.

### 2022-04-11 doc/AMS-15/AMS-15_paper/AMS_HGIS_els/AMS_HGIS_els.Rmd
Edits to figures and text. Use xelatex to fix compile issues.

@@
     email: simon.pendleton@plymouth.edu
-    affiliation: Plymouth State University
+    affiliation:  Woods Hole Oceanographic Institution,Plymouth State University
   - name: Sydney Moser
@@
 # csl: https://www.zotero.org/styles/elsevier-harvard
-csl: elsevier-without-titles.csl
+csl: elsevier-with-titles.csl
 output: 
   rticles::elsevier_article:
+    latex_engine: xelatex
     keep_tex: true
@@
 A primary disadvantage of the method is its cost, which often limits the number of samples analyzed.
-This forces researchers to carefully choose samples to be run and limits repeat measurements for error determination.
+This forces researchers to carefully choose which samples to analyze and limits repeat measurements for error determination.
 One driver of the cost of AMS analysis is the time and effort required for sample preparation, including reduction of the sample to graphite.
@@
 A disadvantage of the open split is that the flow of CO~2~ to the source varies with sample size (CO~2~:He) and changes over the course of a measurement due to dilution of the sample by displacement helium.
-Our method compensates for this by limiting CO~2~ flow to a range that produces relatively constant C- ion current and stable ratios.
+Our method compensates for this by limiting CO~2~ flow to a range that produces relatively constant ^12^C^-^ ion current and stable ratios.
 In gas-source AMS, Roberts et al.
@@
 The delivery needle was then transferred to the sample vial, which started gas delivery to the split and source.
-Currents were allowed to stabilize for 30 s, at which point the data acquisition was started.
+Ion current was allowed to stabilize for 30 s, at which point the data acquisition was started.
 Data was acquired in 10 blocks of 90 s each, for 15 min total acquisition time.
@@
 Blocks that are more than 1.5 times the interquartile range (IQR) below the first quartile of all measurements of a sample or more than 1.5 IQR above the third quartile are considered outliers and were not further used in the analysis.
-Remaining runs of a sample were averaged and sample uncertainty was taken as the greater of counting error or the standard error of the measurements.
-The sample ratio was then normalized using the mean of gas standards measured during the run.
+Remaining blocks of a sample were averaged and sample uncertainty was taken as the greater of counting error or the standard error of the measurements.
+The sample ratio was then normalized using the mean of gas standards measured during the measurement session.
 The measurement error of the standards was taken as the greater of the standard error of all standards or the errors of the standards added in quadrature.
@@
 The errors in the blanks and standards were propagated with errors of the samples to produce the final error estimate for the sample.
-This correction does not account for the effect of a blank with constant mass or current on samples of varying mass or current.
+This correction does not account for the effect of a blank with constant mass or ion current on samples of varying mass or current.
 For all measurements used to assess the precision and accuracy of the method, sample, standard, and blank masses were kept constant.
@@
 We used tank CO~2~ connected through a regulator and glass capillary to the ion source to measure source performance over a range of CO~2~ flows.
-Best currents for the source was 10 $\mu$A ^12^C^-^, or roughly 10% of graphite target current.
+Best ^12^C^-^ ion current for the source was 10 $\mu$A, or roughly 10% of graphite target current.
 Peak current was produced at a CO~2~ flow rate of 7.5 $\mu$l min^-1^ CO~2~, but the source produced near optimal currents with flows from 7.5-25 $\mu$l min^-1^ CO~2~.
@@
 
-Source currents are variable by day and by target under similar source, inlet, and sample conditions.
+Source ion currents are variable by day and by target under similar source, inlet, and sample conditions.
 Factors affecting source current may include capillary condition, source and gas delivery arm alignment, slight variability in target geometry, source condition, and performance of the ionizer and cesium.
@@
 (a) Current dependent blank model and data.
-$F^{14}C$ and current of measurements from dilution tests of live and dead gas are used to generate linear fits which should cross at the $F^{14}C$ and current of the constant current blank.
-Direct determination of $F^{14}C$ and current of the blank are also shown.
-(b) Fraction modern of large, radiocarbon dead samples compared to sample current.
+$F^{14}C$ and current of measurements from dilution tests of modern and dead tank standards are used to generate linear fits which should cross at the $F^{14}C$ and current of the constant current blank.
+Direct determination of $F^{14}C$ and current of the blank using clean targets and helium without CO~2~ are also shown.
+(b) Fraction modern of large IAEA C1 and TIRI-F carbonate blank samples compared to sample current.
 Horizontal and vertical bars represent measurement uncertainty in current and ratio, respectively.", echo = FALSE}
@@
 where $R_{samp}$ is the ratio of the unknown sample, $R_{meas}$ is the measured ratio, $R_{blank}$ is the ratio of the blank, and $R_{std}$ is the ratio of the normalizing standard.
-This mass-independent blank correction works well for our data since blanks vary by run and currents are relatively similar between blanks, unknowns and standards.
+This mass-independent blank correction works well for our data since blanks vary by run and ion currents are relatively similar between blanks, unknowns and standards.
 
@@
 
-Reference materials run with the hybrid gas method were in good agreement with consensus values.
+Reference materials analyzed with the hybrid gas method were in good agreement with consensus values.
 Mean difference from consensus for all measurements was F^14^C `r format(cons_sum$fm_diff_mean, digits = 1, nsmall = 4)` shows that the hybrid method has no bias in aggregate.
@@
 Expected values are spread to avoid overlapping points.
-Error bars represent one sigma sample uncertainty.
+Error bars represent 1 $\\sigma$ sample uncertainty.
 The black line is the mean difference from expected value for all measurements, and the blue lines are one standard deviation of all measurements.", echo = FALSE}
@@
 Carbonate materials such as corals and bivalve shells were selected throughout the core for the age-depth model, processed and measured as described above.
-Parallel graphite measurements were made following NOSAMS methods for acidification, graphitization, and measurement.
+Parallel graphite measurements were made following standard NOSAMS methods.
 One HGIS sample vial was noted to be at atmospheric pressure when the 16 mg sample should have produced only 0.5 atmosphere of CO~2~.
@@
 Two of the three pairs agreed within 1$\sigma$ and all three agreed within 2$\sigma$.
-There is an apparent, but not significant positive relationship between the HGIS-graphite difference and measured value.
-The higher graphite values at lower F^14^C may indicate an overcorrection of HGIS blank for these samples.
 
@@
 Samples are offset on the horizontal axis for clarity.
-Error bars are measurement error for each sample.
-Graphite measurements are included to allow comparison of error estimates.
-The bottom row of plots show a comparison of age-depth models produced using HGIS (b) and graphite measurements (c).", echo = FALSE}
+Error bars are 1 $\\sigma$ measurement uncertainty for each sample.
+Graphite measurements are included to allow comparison of uncertainty.
+The bottom row compares age-depth models produced using HGIS (b) and graphite measurements (c).
+Grey subplots show probability distribution of calibrated ages.", echo = FALSE}
 knitr::include_graphics(here("doc/AMS-15/AMS-15_paper/figures/fig5_core_carbonates.pdf"))
@@
 of measurement time per sample.
-The simplicity of of this approach reduces costs, improves reliability, and reduces system maintenance.
+The simplicity of of this approach reduces sample cost by eliminating graphitization, and may improve reliability and reduce maintenance compared to more complex sample inlet designs.
 
@@
 This may lead to more accurate results compared to a more precise graphite measurement where reproducibility is not as well constrained.
-
 The accuracy of age-depth models is often driven by environmental factors and sparse data, rather than the precision of individual dates [@blaauw_double_2018].
@@
 This work was supported by the NOSAMS NSF cooperative agreement 1755125.
-We are grateful to C.
-McIntyre, R.
-Hansman, G.
-Salazar, T.
-Ognibene, and T.
-Broek for discussions on gas handling and gas ion sourcery; and to J.
-Gunnard for improvements to the text.
+We are grateful to K.
+von Reden, A.
+Gagnon, K.
+Gospodinova, and J.
+Hlavenka for initial work and discussions on the NOSAMS HGIS; C.
+McIntyre, T.
+Broek, G.
+Salazar, and T.
+Ognibene for discussions on gas handling and gas ion sourcery; J.
+Gunnard for improvements to the text; and the staff of NOSAMS for assistance with the HGIS and preparation of graphite samples.
 

### 2022-04-11 doc/AMS-15/AMS-15_paper/figures/fig3_blanks.R

@@ -78,8 +78,8 @@ ds <- sum_hgis(data) %>%
   filter((str_detect(Sample.Name, "LiveGas") & Cur > 1) |
          (str_detect(Sample.Name, "DeadGas") & Cur > 1) |
           str_detect(Sample.Name, "Blank")) %>% 
-  mutate(Gas = case_when(str_detect(Sample.Name, "LiveGas") ~ "LiveGas",
-                         str_detect(Sample.Name, "DeadGas") ~ "DeadGas",
+  mutate(Gas = case_when(str_detect(Sample.Name, "LiveGas") ~ "Modern",
+                         str_detect(Sample.Name, "DeadGas") ~ "Dead",
                          TRUE ~ "Blank"),
          Cur_inv = 1/Cur)
 
@@ -94,8 +94,8 @@ fits <- ds %>%
 blankfit <- fits %>% 
   select(Gas, term, estimate) %>% 
   pivot_wider(names_from = c(Gas, term), values_from = estimate) %>% 
-  mutate(inv_m_blank = -(`DeadGas_(Intercept)` - `LiveGas_(Intercept)`)/(DeadGas_Cur_inv - LiveGas_Cur_inv),
-         Fm_blank = `DeadGas_(Intercept)` + DeadGas_Cur_inv * inv_m_blank,
+  mutate(inv_m_blank = -(`Dead_(Intercept)` - `Modern_(Intercept)`)/(Dead_Cur_inv - Modern_Cur_inv),
+         Fm_blank = `Dead_(Intercept)` + Dead_Cur_inv * inv_m_blank,
          m_blank = 1/inv_m_blank) 
 
 # Modeled Fm and mass of blank
@@ -111,12 +111,12 @@ blank_meas <- ds %>%
 
 # Make plot of fm vs current with model fits
 
-blank_model <- ggplot(ds, aes(Cur_inv, mean, color = Gas)) +
-  geom_abline(slope = blankfit$LiveGas_Cur_inv, 
-              intercept = blankfit$`LiveGas_(Intercept)`,
+blank_model <- ggplot(ds, aes(Cur_inv, mean, color = Gas, shape = Gas)) +
+  geom_abline(slope = blankfit$Modern_Cur_inv, 
+              intercept = blankfit$`Modern_(Intercept)`,
               color = "#0069b1") +
-  geom_abline(slope = blankfit$DeadGas_Cur_inv, 
[…]

### 2022-04-11 doc/AMS-15/AMS-15_paper/figures/fig5_core_carbonates.R

@@ -124,13 +124,13 @@ age_depth <- cr_no %>%
 chron_hgis <- read_rds(here("data_analysed/hgis_chronology.rds"))
 chron_gr <- read_rds(here("data_analysed/graphite_chronology.rds"))
 
-hgis_bchron <- plot(chron_hgis) +
+hgis_bchron <- plot(chron_hgis, dateLabels = FALSE) +
   xlim(8000, -500) +
   labs(title = "B",
        y = "Years BP",
        x = "Core depth (cm)")
 
-gr_bchron <- plot(chron_gr) +
+gr_bchron <- plot(chron_gr, dateLabels = FALSE) +
   xlim(8000, -500) +
   labs(title = "C",
        y = "Years BP",

### 2022-04-12 doc/AMS-15/AMS-15_paper/figures/fig5_core_carbonates.R
Add combined age models and model difference.

@@ -91,31 +91,31 @@ core_compare <- ggplot(cr_no) +
        x = "Mean of measurements (F14C)",
        y = "HGIS - graphite (F14C)") +
   theme(legend.position = "none")
-  # theme(legend.position = c(0.85, 0.15),
-  #       legend.direction = "horizontal",
-  #       legend.background = element_rect(fill = "white", color = "black")) 
+  # theme(legend.position = c(0.25, 0.85),
+  #       legend.direction = "horizontal") #,
+  #       #legend.background = element_rect(fill = "white", color = "black")) 
 
 
 # Simple age depth plot
 # If this is used, should use calibrated ages
 # or at least bomb correction
 
-age_depth <- cr_no %>% 
-  ggplot(aes(depth, rc_age, color = method, shape = method)) +
-  geom_pointrange(aes(ymin = rc_age - sig_rc_age, 
-                      ymax = rc_age + sig_rc_age),
-                  size = 0.5) +
-  geom_smooth(alpha = .2) +
-  coord_flip() +
-  scale_y_reverse() +
-  scale_x_reverse() +
-  scale_color_manual(values = c("#00b7bd", "#b7bf10")) +
-  labs(title = "B",
-       #subtitle = "Age vs. depth for graphite and HGIS",
-       y = "Radiocarbon years (BP)",
-       x = "Core depth (cm)") +
-  theme(legend.position = c(0.85, 0.30),
-        legend.background = element_rect(fill = "white", color = "black")) 
+# age_depth <- cr_no %>% 
+#   ggplot(aes(depth, rc_age, color = method, shape = method)) +
+#   geom_pointrange(aes(ymin = rc_age - sig_rc_age, 
+#                       ymax = rc_age + sig_rc_age),
+#                   size = 0.5) +
+#   geom_smooth(alpha = .2) +
+#   coord_flip() +
+#   scale_y_reverse() +
+#   scale_x_reverse() +
+#   scale_color_manual(values = c("#00b7bd", "#b7bf10")) +
+#   labs(title = "B",
+#        #subtitle = "Age vs. depth for graphite and HGIS",
+#        y = "Radiocarbon years (BP)",
+#        x = "Core depth (cm)") +
+#   theme(legend.position = c(0.85, 0.30),
+#         legend.background = element_rect(fill = "white", color = "black")) 
 
 # Bchron age models
 
@@ -124,46 +124,141 @@ age_depth <- cr_no %>%
[…]

### 2022-04-12 doc/AMS-15/AMS-15_paper/AMS_HGIS_els/AMS_HGIS_els.Rmd
Add MK edits.

@@
 ---
-title: "Radiocarbon measurement of CO~2~ from carbonate minerals using a gas ion source with an open split interface"
+title: "Radiocarbon measurement of CO~2~ from carbonate minerals using a hybrid gas ion source with an open split interface"
 author:
@@
 Radiocarbon Accelerator Mass Spectrometry (AMS) is a sensitive method for determining the age, provenance, and composition of carbon-containing materials.
-A primary disadvantage of the method is its cost, which often limits the number of samples analyzed.
-This forces researchers to carefully choose which samples to analyze and limits repeat measurements for error determination.
+A primary disadvantage of the method is cost, which can limit the number of samples analyzed.
+This forces researchers to carefully select samples for analysis and may limit repeat measurements for error determination.
 One driver of the cost of AMS analysis is the time and effort required for sample preparation, including reduction of the sample to graphite.
@@
 Pressure is adjusted by changing either the inlet volume or the amount of gas in the inlet.
-We designed a open inlet where a capillary samples gas for delivery to the ion source in a flowing stream that exhausts to the atmosphere, often called an open split.
+We designed a open inlet where a capillary samples gas for delivery to the ion source in a flowing stream that exhausts to the atmosphere, referred to here as an open split.
 This defines the delivery pressure as 1 ATM.
@@
 The sample CO~2~ - helium mixture is introduced onto a titanium insert or frit in an aluminum target cathode using stainless tubing mounted to an arm with a probe that mates with the back of the cathode.
-Sample gas is exposed to a beam of charged cesium atoms, which causes dissociation of CO~2~ and ionization of molecular and atomic species.
+Sample gas is exposed to a beam of charged cesium atoms as it flows over the titanium insert, which causes dissociation of CO~2~ and ionization of molecular and atomic species.
 Negative ions are extracted from the target, carbon species are selected, and interfering masses filtered out by the USAMS Tandetron beamline.
@@
 
-The open split interface (Figure \ref{fig1}) supplies a helium - CO~2~ mixture to a split where a capillary delivers gas to the source and the rest is vented.
+The open split interface (Figure \ref{fig1}) supplies a helium - CO~2~ mixture to a split where a capillary transfers gas to the ion source and the gas not taken by the capillary is vented.
 Flowing helium is used to displace sample CO~2~ from the sample vial to the delivery capillary in the split.
-Our open-split interface is constructed of simple, off-the-shelf parts.
+The open-split interface is constructed of simple, off-the-shelf parts.
 The core of the design is a dual needle and split housed in a Valco tee fitting.
@@
 The split has fairly low volume (< 1 mL).
-The system was operated manually during development, but may be mounted on the arm of a vial-sampling autosampler to enable automated operation.
+The system was operated manually during development, but could be mounted on the arm of a vial-sampling autosampler to enable automated operation.
 
@@
 Shell samples were crushed and mixed, but not otherwise pretreated.
+The homogenized carbonates were sub-sampled for measurement by HGIS (15mg) and as graphite (10mg).
 
@@
 
-Samples were measured on a modified 3MV Tandetron AMS system [@Longworth2015].
+The HGIS measurements were carried out on a modified 3MV Tandetron AMS system [@Longworth2015].
 Measurement of a sample proceeded as follows: a clean sputter target cathode was rotated into position and the target was pre-cleaned by sputtering for 5 minutes or until cathode ^12^C^-^ current fell below 100 nA.
@@
 Total time for a measurement was 20-25 min including target pre-cleaning and changing samples.
+Graphite samples were prepared at NOSAMS and measured on the CFAMS system using standard procedures [@gagnon_nosams_2000; @roberts_high-performance_2010].
 
@@
 We used tank CO~2~ connected through a regulator and glass capillary to the ion source to measure source performance over a range of CO~2~ flows.
-Best ^12^C^-^ ion current for the source was 10 $\mu$A, or roughly 10% of graphite target current.
+Best ^12^C^-^ ion current for the source was 10 $\mu$A, or roughly 10% of typical graphite target current.
 Peak current was produced at a CO~2~ flow rate of 7.5 $\mu$l min^-1^ CO~2~, but the source produced near optimal currents with flows from 7.5-25 $\mu$l min^-1^ CO~2~.
-This is comparable with other hybrid sources [@Xu2007], although improved target design and continued optimization of source parameters may improve performance [@Salazar2013; @Fahrni2013].
+This is comparable with other hybrid sources [@Xu2007], although improved target design and continued optimization of source parameters may improve performance [@Salazar2013; @Fahrni2013; @broek_optimization_nodate].
 Typical currents for this study were between 4-9 $\mu$A and average current for all samples was `r curmean` ± `r cursd` $\mu$A.
@@
 ```{r fig2, fig.width = 5, fig.height = 5, fig.align='center', out.width="80%", fig.cap = "\\label{fig2}Performance of the NOSAMS MCSNICS hybrid gas ion source.
-(a) Measured $^{12}C^-$ ion current during displacement of a pure $CO_2$ vial sample with helium.
+(a) Measured $^{12}C^-$ ion current during displacement of modern tank $CO_2$ from a sample vial with helium.
 (b) Modeled flows of helium and $CO_2$ to the ion source during the same dilution experiment.
-(c) Measured ion current for modeled $CO_2$ flow.
+(c) Measured ion current for the modeled $CO_2$ flow from b.
 The stability of the source over a range of flow produces stable current for up to 40 min of measurement time.", echo = FALSE}
@@
 Source ion currents are variable by day and by target under similar source, inlet, and sample conditions.
-Factors affecting source current may include capillary condition, source and gas delivery arm alignment, slight variability in target geometry, source condition, and performance of the ionizer and cesium.
+Factors affecting source current may include capillary condition, source and gas delivery arm alignment, slight variability in target geometry, source condition, and performance of the ionizer and cesium reservoir.
 Capillary clogging caused occasional flow issues, especially when sampling from vials containing acid.
@@
 The most important feature of the current-flow relationship is the wide range of CO~2~ flow rates over which stable ion current is measured, which in turn produces relatively stable currents and ratios despite a changing carbon flow rate during measurement.
-Several studies have measured peak currents over a narrower range of flows [@Fahrni2013; @broek_optimization_nodate] at high helium dilution rates (> 10:1), while Xu et.
+Several studies have measured peak currents over a narrower range of flows [@Fahrni2013; @broek_optimization_nodate] at high helium dilution rates (> 90% helium), while Xu et.
 al [@Xu2007] shows a somewhat broader peak with pure CO~2~.
@@
 al [@broek_optimization_nodate] found that the range of flow rates with peak currents increased with decreasing dilution ratio.
-While lower rates of dilution used in our HGIS method may decrease current slightly, the observed peak broadening allows stable measurement over the range of CO~2~ flows produced by the open split.
+While lower rates of dilution used in our HGIS method may decrease current slightly, the observed relationship between ion current and flow rate allows stable measurement over the range of CO~2~ flows produced by the open split.
 
@@
 Tests of a range of sample sizes support this estimate and show that no reduction in measured current is seen in samples larger than 1 mg C which corresponds to 8 mg CaCO~3~ sample material.
-To ensure ample sample for optimal measurements, we chose 15 mg as a target sample mass.
+To ensure sufficient sample for optimal measurements, we chose 15 mg as a target sample mass.
 To ensure that samples started at or slightly above atmospheric pressure, we injected helium with a syringe to equilibrate pressure in the vials when the reaction with acid was complete.
@@
 
-The HGIS method produced higher ratios for radiocarbon dead samples than the graphite method.
-The average uncorrected $F^{14}C$ for radiocarbon-free reference materials IAEA C1 marble and TIRI F spar calcite was `r blankmean` ± `r blanksd` (1$\sigma$ uncertainty, N=`r nblank`).
+The HGIS method produced higher $^{14/12}C$ ratios for radiocarbon dead samples than the graphite method.
+The average uncorrected $F^{14}C$ for radiocarbon-free reference materials IAEA C1 marble and TIRI F spar calcite was `r blankmean` ± `r blanksd` (1$\sigma$ uncertainty, N=`r nblank`), compared to XXX for graphite.
 We investigated the sources of this method blank as a current-dependent blank and as a current-independent blank, and implemented a blank correction for the method.
@@
 $F^{14}C$ and current of measurements from dilution tests of modern and dead tank standards are used to generate linear fits which should cross at the $F^{14}C$ and current of the constant current blank.
-Direct determination of $F^{14}C$ and current of the blank using clean targets and helium without CO~2~ are also shown.
+Direct determination of $F^{14}C$ and current of the blank using clean targets and helium without $CO_2$ are also shown.
 (b) Fraction modern of large IAEA C1 and TIRI-F carbonate blank samples compared to sample current.
@@
 Fraction modern of measured blanks does decrease with increasing current, but the decrease is small, less than F^14^C 0.01 over the range of currents measured.
-This change with current is roughly equal to the variability between blanks, so the error due to the use of a current-dependent blank based on samples measured over multiple wheels will likely be larger than a current independent blank correction based on blanks measured on each wheel.
+This change with current is roughly equal to the variability between blanks, so the error due to the use of a current-dependent blank based on samples measured over multiple measurement sessions will likely be larger than a current independent blank correction based on blanks measured during each session.
 Constant current blank is likely the best model, but we simplify to a linear correction based on the measured ratio of process blanks and normalizing standards:
@@
 The mean of propagated measurement errors for these samples was `r format(cons_sum$sig_fm_corr_mean, nsmall = 4)` F^14^C for these samples, which is in good agreement with the method precision above.
-The method error is about three times larger than the roughly 3 ‰ precision of graphite measurements, but is within the range of method precision of other gas sources.
+The method error is about three times larger than the 2--3 ‰ precision of graphite measurements, but is within the range of method precision of other gas sources.
 
@@
 Difference between HGIS and graphite measurements is shown on the vertical axis and plotted against the mean of the two measurements.
-Samples are offset on the horizontal axis for clarity.
 Error bars are 1 $\\sigma$ measurement uncertainty for each sample.
 Graphite measurements are included to allow comparison of uncertainty.
-The bottom row compares age-depth models produced using HGIS (b) and graphite measurements (c).
-Grey subplots show probability distribution of calibrated ages.", echo = FALSE}
+Age-depth models produced using HGIS (b) and graphite measurements (c).
+Grey subplots show probability distribution of calibrated ages.
+(d) Combined age-depth prediction and 95\\% confidence intervals for HGIS and graphte.
+(e) Offset of HGIS and graphite values from the mean of the models with confidence intervals.", echo = FALSE}
 knitr::include_graphics(here("doc/AMS-15/AMS-15_paper/figures/fig5_core_carbonates.pdf"))
@@
 al @hua_atmospheric_2013 was used for calibration of postmodern dates, but these measurements were not included in the age model due to uncertainties in reservoir correction.
-The age-depth curves produced by the two datasets are very similar, with the largest differences being in the uncertainty of the individual dates.
+The age-depth curves produced by the two datasets agree within the uncertainty of the chronologies, with the largest differences being in the uncertainty of the individual dates.
 
@@
 of measurement time per sample.
-The simplicity of of this approach reduces sample cost by eliminating graphitization, and may improve reliability and reduce maintenance compared to more complex sample inlet designs.
+The simplicity of of this approach can potentially reduce sample cost by eliminating graphitization, and may improve reliability and reduce maintenance compared to more complex sample inlet designs.
 
@@
 This work was supported by the NOSAMS NSF cooperative agreement 1755125.
+S.
+Pendleton was supported by a WHOI/NOSAMS Postdoctoral fellowship.
 We are grateful to K.
@@
 Ognibene for discussions on gas handling and gas ion sourcery; J.
-Gunnard for improvements to the text; and the staff of NOSAMS for assistance with the HGIS and preparation of graphite samples.
+Gunnard and T.
+Broek for improvements to the text; and the staff of NOSAMS for assistance with the HGIS and preparation of graphite samples.
 

### 2022-04-13 doc/AMS-15/AMS-15_paper/AMS_HGIS_els/AMS_HGIS_els.Rmd
Format figure axes, add supplementary table + data, lots of minor edits, improve references.

@@
 author:
-  - name: Brett Longworth
+  - name: Brett E. Longworth
     email: blongworth@whoi.edu
@@
     correspondingauthor: true
-  - name: Josh Burton
+  - name: Josh R. Burton
     email: jburton@whoi.edu
@@
     affiliation:  Woods Hole Oceanographic Institution,Plymouth State University
-  - name: Sydney Moser
+  - name: Sydney D. Moser
     email: smose007@uottawa.ca
     affiliation: University of Ottawa
-  - name: Mark Roberts
+  - name: Mark L. Roberts
     email: mroberts@whoi.edu
     affiliation: Woods Hole Oceanographic Institution
-  - name: Mark Kurz
+  - name: Mark D. Kurz
     email: mkurz@whoi.edu
@@
 Samples larger than 8 mg carbonate can be measured.
-Results from measurements of consensus standards (TIRI-I, IAEA C2, and an internal modern shell standard), and samples from a marine core (F^14^C = 0.4 - 1.15) show that the method agrees well with traditional AMS measurement of the same samples as graphite, and that the 1$\sigma$ uncertainty is about 1 %.
+Results from measurements of consensus standards (TIRI I, IAEA C2, and an internal modern shell standard), and samples from a marine core (F^14^C = 0.4--1.15) show that the method agrees well with traditional AMS measurement of the same samples as graphite, and that the 1$\sigma$ uncertainty is about 1 %.
 We discuss advantages and disadvantages of continuous flow sample introduction, and the effect of reduced precision on calibrated age-depth models produced using gas-source data.
@@
 library(here)
-#library(errors)
+library(errors)
 
-#options(errors.notation = "plus-minus")
+options(errors.notation = "plus-minus")
 
@@
 Analysis of CO~2~ directly via a gas-accepting AMS ion source eliminates the cost and effort of graphitization and returns results more rapidly.
-The NOSAMS USAMS system at the National Oceanographic AMS lab (NOSAMS) is based on a 3MV Tandetron accelerator [@Longworth2015] combined with a National Electrostatics Corporation (NEC) 134-sample gas-capable MCSNICS source [@Xu2007] .
+The NOSAMS USAMS system at the National Oceanographic AMS lab (NOSAMS) is based on a 3MV Tandetron accelerator [@Longworth2015] combined with a National Electrostatics Corporation (NEC) 134-sample gas-capable MCSNICS source.
+The hybrid gas capability of this source was developed as a joint project between NEC and the Scottish Universities Environmental Research Centre (SUERC) [@Xu2007], and further improved by work at Lawrence Livermore National Laboratory (LLNL) [@Salazar2013].
 We present here a system for semi-automated analysis of carbonate minerals by acid hydrolysis in vials and introduction to the ion source in a flow of helium via an open split.
@@
 
-The open split interface (Figure \ref{fig1}) supplies a helium - CO~2~ mixture to a split where a capillary transfers gas to the ion source and the gas not taken by the capillary is vented.
+The open split interface (Figure \ref{fig1}) supplies a CO~2~ - helium mixture to a split where a capillary transfers gas to the ion source and the gas not taken by the capillary is vented.
 Flowing helium is used to displace sample CO~2~ from the sample vial to the delivery capillary in the split.
@@
 Normalizing standards, blanks, and secondary standards were measured identically to unknowns.
+One target is used per gas sample.
 Total time for a measurement was 20-25 min including target pre-cleaning and changing samples.
@@
 Best ^12^C^-^ ion current for the source was 10 $\mu$A, or roughly 10% of typical graphite target current.
-Peak current was produced at a CO~2~ flow rate of 7.5 $\mu$l min^-1^ CO~2~, but the source produced near optimal currents with flows from 7.5-25 $\mu$l min^-1^ CO~2~.
+Peak current was produced at a CO~2~ flow rate of 7.5 $\mu$l min^-1^ CO~2~, but the source produced near optimal currents with flows from 7.5--25 $\mu$l min^-1^ CO~2~.
 This is comparable with other hybrid sources [@Xu2007], although improved target design and continued optimization of source parameters may improve performance [@Salazar2013; @Fahrni2013; @broek_optimization_nodate].
@@
          sample_type == "B")
-blankmean <- mean(blanks$norm_ratio) %>% format(digits = 3)
-blanksd <-  sd(blanks$norm_ratio) %>% format(digits = 2)
+blankmean <- mean(blanks$norm_ratio) %>% format(digits = 2)
+blanksd <-  sd(blanks$norm_ratio) %>% format(digits = 1)
 
@@
 The HGIS method produced higher $^{14/12}C$ ratios for radiocarbon dead samples than the graphite method.
-The average uncorrected $F^{14}C$ for radiocarbon-free reference materials IAEA C1 marble and TIRI F spar calcite was `r blankmean` ± `r blanksd` (1$\sigma$ uncertainty, N=`r nblank`), compared to XXX for graphite.
+The average uncorrected $F^{14}C$ for radiocarbon-free reference materials IAEA C1 marble and TIRI F spar calcite was `r blankmean` ± `r blanksd` (1$\sigma$ uncertainty, N=`r nblank`), compared to 0.003 ± 0.001 for graphite.
 We investigated the sources of this method blank as a current-dependent blank and as a current-independent blank, and implemented a blank correction for the method.
@@
 This current was initially \> 10 $\mu$A and dropped exponentially, reaching a baseline of ~100 nA after 5-10 minutes.
-For the range of currents in this study (3-9 $\mu$A), this baseline current represents a 1-6% contribution to the total measured sample current at the F^14^C of the target current baseline.
+For the range of currents in this study (3--9 $\mu$A), this baseline current represents a 1--6% contribution to the total measured sample current at the F^14^C of the target current baseline.
 This can be interpreted as a constant current dependent blank, analogous to a mass-dependent "constant contaminant" blank and may be modeled with the same tools [@haghipour_compound-specific_2019; @hanke_comprehensive_2017; @roberts_14c_2019; @sun_14_2020].
@@
 Direct determination of $F^{14}C$ and current of the blank using clean targets and helium without $CO_2$ are also shown.
-(b) Fraction modern of large IAEA C1 and TIRI-F carbonate blank samples compared to sample current.
+(b) Fraction modern of large IAEA C1 and TIRI F carbonate blank samples compared to sample current.
 Horizontal and vertical bars represent measurement uncertainty in current and ratio, respectively.", echo = FALSE}
@@
 
-options(digits = 3)
-options(scipen = 4)
+#options(digits = 3)
+options(scipen = 99)
 options(decimals = 4)
@@
 Mean difference from consensus for all measurements was F^14^C `r format(cons_sum$fm_diff_mean, digits = 1, nsmall = 4)` shows that the hybrid method has no bias in aggregate.
-The standard deviation of the difference in F^14^C was `r format(cons_sum$fm_diff_sd, nsmall = 4)`, which may be taken as an estimate of the 1$\sigma$ precision of the method.
-The mean of propagated measurement errors for these samples was `r format(cons_sum$sig_fm_corr_mean, nsmall = 4)` F^14^C for these samples, which is in good agreement with the method precision above.
+The standard deviation of the difference in F^14^C was `r format(cons_sum$fm_diff_sd, digits = 2, nsmall = 4)`, which may be taken as an estimate of the 1$\sigma$ precision of the method.
+The mean of propagated measurement errors for these samples was `r format(cons_sum$sig_fm_corr_mean, digits = 2, nsmall = 4)` F^14^C for these samples, which is in good agreement with the method precision above.
 The method error is about three times larger than the 2--3 ‰ precision of graphite measurements, but is within the range of method precision of other gas sources.
@@
          N) %>% 
-  knitr::kable(digits = 3,
+  knitr::kable(digits = c(4, 4, 4, 3, 2),
                caption = "\\label{tab1}Reference material precision and accuracy. Reference 
@@
 Expected values are spread to avoid overlapping points.
-Error bars represent 1 $\\sigma$ sample uncertainty.
+Error bars represent 1$\\sigma$ sample uncertainty.
 The black line is the mean difference from expected value for all measurements, and the blue lines are one standard deviation of all measurements.", echo = FALSE}
@@
 Difference between HGIS and graphite measurements is shown on the vertical axis and plotted against the mean of the two measurements.
-Error bars are 1 $\\sigma$ measurement uncertainty for each sample.
+Error bars are 1$\\sigma$ measurement uncertainty for each sample.
 Graphite measurements are included to allow comparison of uncertainty.
-Age-depth models produced using HGIS (b) and graphite measurements (c).
+Bayesian age-depth models produced using HGIS (b) and graphite measurements (c).
 Grey subplots show probability distribution of calibrated ages.
@@
 Age models were developed for both the HGIS measurements (Figure 5b) and the traditional graphite measurements (Figure 5c) of the core carbonates.
-We used the Bchron Baesian radiochronology package [@haslett_simple_2008] to develop these models.
-We applied a correction for the regional marine reservoir effect with a $\Delta$ R of -216 ± 92 years based on the Oxcal model [@ramsey_bayesian_2009] with 10 reservoir points.
+We used the Bchron Bayesian radiocarbon chronology package [@haslett_simple_2008] to develop these models.
+We applied a correction for the regional marine reservoir effect with a $\Delta$ R of -216 ± 92 years based on the Oxcal model [@bronk_ramsey_bayesian_2009] with 10 reservoir points.
 Samples with pre-bomb radiocarbon ages were calibrated using the Marine20 curve @heaton_marine20marine_2020.
@@
 al @hua_atmospheric_2013 was used for calibration of postmodern dates, but these measurements were not included in the age model due to uncertainties in reservoir correction.
-The age-depth curves produced by the two datasets agree within the uncertainty of the chronologies, with the largest differences being in the uncertainty of the individual dates.
+The age-depth curves produced by the two datasets agree within the uncertainty of the chronologies (Figure 5d & 5e), with the largest differences being in the uncertainty of the individual dates.
 
@@
 We developed a method for the analysis of carbonates using an open-split interface to a hybrid sputter gas ion source.
-This system produces results that agree with graphite measurements within the approximately 1% method error from 8-20 mg of carbonate.
+This system produces results that agree with graphite measurements within the approximately 1% method error from 8--20 mg of carbonate.
 Sample preparation takes about 10 min.

### 2022-04-13 doc/AMS-15/AMS-15_paper/figures/fig2_current_flow.R

@@ -27,7 +27,7 @@ cur_time <- ggplot(data, aes(time, he12C)) +
   labs(title = "A",
        #title = "Ion current",
        x = "Time (min)",
-       y = "12C current (μA)") +
+       y = TeX(r"( $^{12}C^-$ current (μA))")) +
   theme(axis.title.x = element_blank(),
         axis.text.x = element_blank())
 
@@ -43,7 +43,7 @@ flow_time <- data.frame(x = 0:125) %>%
        #title = "Gas flows to source",
        #subtitle = "7mL vial, 244μL/min helium, 30μL/min to source",
        x = "Time (min)",
-       y = "Gas flow (μL/min)") +
+       y = TeX(r"(Gas flow (μl $min^{-1}$))")) +
   theme(legend.position = c(0.87, 0.50),
         legend.background = element_rect(fill = "white", color = "black"))
 
@@ -53,12 +53,8 @@ cur_flow <- ggplot(data, aes(co2flow, he12C)) +
   geom_smooth(span = .4, se = FALSE, color = "#00b7bd") +
   geom_point(size = 3, color = "#0069b1") +
   labs(title = "C",
-       #title = bquote('Ion current is stable for a range of CO'[2]~'flow'),
-       #subtitle = bquote('Current vs. CO'[2]~'flow during vial dillution'),
-       x = bquote('CO'[2]~'Flow (μl/min)'),
-       y = "12C- current (μA)"
-       #y = TeX("^{12}C^- current (uA)")
-       )
+       x = TeX(r"($CO^2$ flow (μl $min^{-1}$))"),
+       y = TeX(r"( $^{12}C^-$ current (μA))")) 
 
 # Build figure and save
 cur_time / flow_time / cur_flow

### 2022-04-13 doc/AMS-15/AMS-15_paper/figures/fig3_blanks.R

@@ -123,8 +123,8 @@ blank_model <- ggplot(ds, aes(Cur_inv, mean, color = Gas, shape = Gas)) +
   xlim(0, 10) +
   labs(title = "A",
        #subtitle = "Fits cross at Fm and current of blank",
-       x = "Inverse 12C Current (μA-1)",
-       y = "F14C") +
+       x = TeX(r"(Inverse $^{12}C^-$ current ($μA^{-1}$))"), 
+       y = TeX(r"($F^{14}C$)")) +
   theme(#legend.position = "none",
         legend.background = element_rect(fill = "white", color = "black")) 
 
@@ -153,8 +153,8 @@ fm_curr <- blanks %>%
   geom_errorbarh(aes(xmin = he12C - sig_he12C, xmax = he12C + sig_he12C)) +
   geom_point(color = "#00a9e0", size = 3) +
   labs(title = "B",
-       x = "12C Current (μA)",
-       y = "F14C") +
+       x = TeX(r"( $^{12}C^-$ current (μA))"), 
+       y = TeX(r"($F^{14}C$)")) +
   theme(legend.position = c(0.82, 0.76),
         legend.background = element_rect(fill = "white", color = "black"))
 

### 2022-04-13 doc/AMS-15/AMS-15_paper/figures/fig4_consensus_samples.R

@@ -1,4 +1,4 @@
-# Figure 5 bland altman plot of carbonate consensus standards run as
+# Figure 4 bland altman plot of carbonate consensus standards run as
 # hgis compared to accepted value
 
 library(tidyverse)
@@ -15,7 +15,8 @@ cons_df <- results_df %>%
   filter(!(wheel %in% c("USAMS020521", "USAMS061121", "USAMS061821"))) %>% 
   mutate(he12C = he12C * 1E6) %>% 
   compare_consensus() %>% 
-  filter(!str_detect(Name, "Gas"))
+  filter(!str_detect(Name, "Gas")) %>% 
+  arrange(Name)
 
 mean_diff <- mean(cons_df$fm_diff)
 sd_diff <- sd(cons_df$fm_diff)
@@ -37,8 +38,8 @@ ggplot(aes(fm_consensus, fm_diff, color = Name, shape = Name)) +
   scale_y_continuous(breaks = seq(-0.03,0.03,0.01)) +
   labs(#title = "Agreement of HGIS with graphite measurements",
        #subtitle = "Blank corrected HGIS measurements of consensus carbonate standards",
-       x = "F14C expected (points jittered for clarity)", 
-       y = "F14C difference (HGIS - expected)")
+       x = TeX(r"($F^{14}C$ expected (points jittered for clarity))"), 
+       y = TeX(r"($F^{14}C$ difference (HGIS - expected))")) 
 
 # Save figure
 ggsave(here("doc/AMS-15/AMS-15_paper/figures/fig4_consensus_samples.svg"), width = 7, height = 4)

### 2022-04-13 doc/AMS-15/AMS-15_paper/figures/fig5_core_carbonates.R

@@ -88,8 +88,8 @@ core_compare <- ggplot(cr_no) +
   scale_color_manual(values = c("#00b7bd", "#b7bf10")) +
   labs(title = "A",
        #subtitle = "Sediment core macrofossils measured via HGIS and graphite",
-       x = "Mean of measurements (F14C)",
-       y = "HGIS - graphite (F14C)") +
+       x = TeX(r"(Mean of measurements ($F^{14}C$))"),
+       y = TeX(r"(HGIS - graphite ($F^{14}C$))")) +
   theme(legend.position = "none")
   # theme(legend.position = c(0.25, 0.85),
   #       legend.direction = "horizontal") #,

### 2022-04-13 doc/AMS-15/AMS-15_paper/tables/table_s1_reference.R

@@ -1,0 +1,41 @@
+# Table S1 Reference standards and blanks
+
+library(tidyverse)
+library(here)
+library(hgis)
+library(errors)
+
+options(errors.notation = "plus-minus")
+options(scipen = 999)
+
+theme_set(theme_classic())
+
+# Read data. Don't use first carbonate test, smaller samples, or gas standards.
+results_df <- read_csv(here("data_analysed/carb_all_results.csv"))
+
+cons_df <- results_df %>% 
+  filter(!(wheel %in% c("USAMS020521", "USAMS061121", "USAMS061821"))) %>% 
+  mutate(he12C = he12C * 1E6) %>% 
+  compare_consensus() %>% 
+  filter(!str_detect(Name, "Gas")) %>% 
+  arrange(fm_consensus) %>% 
+  mutate (fm_corr_err = set_errors(fm_corr, sig_fm_corr),
+          Name = case_when(Name == "TIRI-F" ~ "TIRI F",
+                           Name == "TIRI-I" ~ "TIRI I",
+                           Name == "C-2" ~ "IAEA C2",
+                           Name == "C-2" ~ "IAEA C1",
+                           Name == "NOSAMS2" ~ "NOSAMS 2"))
+
+cons_df %>% 
+  select(Name, fm_consensus, fm_corr, sig_fm_corr, current = he12C) %>% 
+  write_csv(here("data_analysed/reference_standards.csv"))
+
+cons_df %>% 
+  select(`Sample Name` = Name, 
+         `Reference F^14^C` = fm_consensus,
+         `HGIS F^14^C ± 1$\\sigma$` = fm_corr_err,
+         `^12^C^-^ current ($\\mu$A)` = he12C) %>% 
+  knitr::kable(digits = c(3, 4, 4, 1),
+               caption = "\\label{tabS1}Measurements of reference standard materials. Reference 
+               values are community consensus values for TIRI I and C2, and the 
+               NOSAMS internal graphite value for NOSAMS2.")

### 2022-04-14 doc/AMS-15/AMS-15_paper/AMS_HGIS_els/AMS_HGIS_els.Rmd
Another round of edits

@@
     affiliation: Woods Hole Oceanographic Institution
-  - name: Simon Pendleton
+  - name: Simon L. Pendleton
     email: simon.pendleton@plymouth.edu
@@
 The split has fairly low volume (< 1 mL).
-The system was operated manually during development, but could be mounted on the arm of a vial-sampling autosampler to enable automated operation.
+The system was operated manually during development but could be mounted on the arm of a vial-sampling autosampler to enable automated operation.
 
@@
 ```{r fig1, fig.width = 5, fig.height = 5, fig.align='center', out.width="80%", fig.cap = "\\label{fig1}Schematic of the open split inlet.
-Two valco 2-way valves provide the ability to admit helium and live or dead tank standards.
 The double needle is housed in a tee fitting and allows displacement helium to flow into the sample vial and the sample-helium mixture to flow to the open split, where a capillary delivers the mixture to the ion source.", echo = FALSE}
@@
 One target is used per gas sample.
-Total time for a measurement was 20-25 min including target pre-cleaning and changing samples.
+Total time for a measurement was 20--25 min including target pre-cleaning and changing samples.
 Graphite samples were prepared at NOSAMS and measured on the CFAMS system using standard procedures [@gagnon_nosams_2000; @roberts_high-performance_2010].
@@
 
-Measurements were corrected for process and measurement blank using a linear correction that assumes that the corrected ratio of blank samples should be zero, samples with the same ratio as the standard need no correction, and samples of intermediate ratio need a proportional correction [@donahue_isotope-ratio_1990; @roberts_14c_2019].
+Measurements were corrected for process and measurement blank using a linear correction that assumes that the corrected ratio of blank samples should be zero, samples with the same ratio as the standard need no correction, and samples of intermediate ratio need a proportional correction (Equation \ref{eq2}) [@donahue_isotope-ratio_1990; @roberts_14c_2019].
 The correction uses measured values of standard gas and blank samples for correction of contaminant carbon and machine blank.
 The errors in the blanks and standards were propagated with errors of the samples to produce the final error estimate for the sample.
-This correction does not account for the effect of a blank with constant mass or ion current on samples of varying mass or current.
 For all measurements used to assess the precision and accuracy of the method, sample, standard, and blank masses were kept constant.
-Sample current did vary, but we believe the linear correction described above produces the most accurate results within this range.
-We discuss this constant contaminant blank in the results below.
 Data were reduced using a package written in R [@r_core_team_r_2022] which is published on github (<https://github.com/blongworth/hgis>).
@@
 This is comparable with other hybrid sources [@Xu2007], although improved target design and continued optimization of source parameters may improve performance [@Salazar2013; @Fahrni2013; @broek_optimization_nodate].
-Typical currents for this study were between 4-9 $\mu$A and average current for all samples was `r curmean` ± `r cursd` $\mu$A.
+Typical currents for this study were 4--9 $\mu$A and average current for all samples was `r curmean` ± `r cursd` $\mu$A.
 Several studies have shown best currents at lower CO~2~ flow rates [@Xu2007; @Ruff2010].
@@
 The most important feature of the current-flow relationship is the wide range of CO~2~ flow rates over which stable ion current is measured, which in turn produces relatively stable currents and ratios despite a changing carbon flow rate during measurement.
-Several studies have measured peak currents over a narrower range of flows [@Fahrni2013; @broek_optimization_nodate] at high helium dilution rates (> 90% helium), while Xu et.
-al [@Xu2007] shows a somewhat broader peak with pure CO~2~.
-Broek et.
-al [@broek_optimization_nodate] found that the range of flow rates with peak currents increased with decreasing dilution ratio.
+Several studies have measured peak currents over a narrower range of flows [@Fahrni2013; @broek_optimization_nodate] at high helium dilution rates (> 90% helium), while Xu et al.
+[@Xu2007] shows a somewhat wider flow range producing high currents with pure CO~2~.
+Broek et al.
+[@broek_optimization_nodate] found that the range of flow rates with peak currents increased with decreasing dilution ratio.
 While lower rates of dilution used in our HGIS method may decrease current slightly, the observed relationship between ion current and flow rate allows stable measurement over the range of CO~2~ flows produced by the open split.
@@
 The HGIS method produced higher $^{14/12}C$ ratios for radiocarbon dead samples than the graphite method.
-The average uncorrected $F^{14}C$ for radiocarbon-free reference materials IAEA C1 marble and TIRI F spar calcite was `r blankmean` ± `r blanksd` (1$\sigma$ uncertainty, N=`r nblank`), compared to 0.003 ± 0.001 for graphite.
+The average uncorrected $F^{14}C$ for radiocarbon-free reference materials IAEA C1 marble and TIRI F spar calcite was `r blankmean` ± `r blanksd` (1$\sigma$ uncertainty, N=`r nblank`), compared to 0.003 ± 0.001 for these materials run as graphite on the CFAMS system.
 We investigated the sources of this method blank as a current-dependent blank and as a current-independent blank, and implemented a blank correction for the method.
@@
 Precision and accuracy of the method were assessed by measuring carbonate reference standards for comparison with consensus values from graphite measurements.
-We measured TIRI I, a roughly quarter-modern turbidite carbonate, IAEA C2, a half modern turbidite carbonate, and NOSAMS 2, a modern internal shell carbonate standard from a bivalve mollusk.
+We measured TIRI I, a roughly quarter-modern turbidite carbonate, IAEA C2, a half modern travertine, and NOSAMS 2, a modern bivalve mollusk internal standard.
 Summary data for these samples are given in Table 1, and measurement data for all samples is in supplementary information.
@@
 Samples with pre-bomb radiocarbon ages were calibrated using the Marine20 curve @heaton_marine20marine_2020.
-The NH zone 2 curve from Hua et.
-al @hua_atmospheric_2013 was used for calibration of postmodern dates, but these measurements were not included in the age model due to uncertainties in reservoir correction.
+The NH zone 2 curve from Hua et al.
+@hua_atmospheric_2013 was used for calibration of postmodern dates, but these measurements were not included in the age model due to uncertainties in reservoir correction.
 The age-depth curves produced by the two datasets agree within the uncertainty of the chronologies (Figure 5d & 5e), with the largest differences being in the uncertainty of the individual dates.
@@
 This system produces results that agree with graphite measurements within the approximately 1% method error from 8--20 mg of carbonate.
-Sample preparation takes about 10 min.
-of person time and 6 h of automated processing and reaction time, and 20 min.
-of measurement time per sample.
+Sample preparation takes about 10 min of person time and 6 h of automated processing and reaction time, and 20 min of measurement time per sample.
 The simplicity of of this approach can potentially reduce sample cost by eliminating graphitization, and may improve reliability and reduce maintenance compared to more complex sample inlet designs.
@@
 This may lead to more accurate results compared to a more precise graphite measurement where reproducibility is not as well constrained.
-The accuracy of age-depth models is often driven by environmental factors and sparse data, rather than the precision of individual dates [@blaauw_double_2018].
+The accuracy of age-depth models is often limited by environmental factors and sparse data, rather than the precision of individual dates [@blaauw_double_2018].
 The lower time and cost of the hybrid method allows analysis of more samples.

### 2022-04-14 doc/AMS-15/AMS-15_paper/AMS_HGIS_els/AMS_HGIS_els.Rmd
Add supplementary tables and current size figure. Many small edits.

@@
   - code: Plymouth State University
-    address: Environmental Science and Policy, 17 High St., Plymouth, MA, 03264
+    address: Environmental Science and Policy, 17 High St., Plymouth, NH, 03264
   - code: University of Ottawa
@@
 Samples larger than 8 mg carbonate can be measured.
-Results from measurements of consensus standards (TIRI I, IAEA C2, and an internal modern shell standard), and samples from a marine core (F^14^C = 0.4--1.15) show that the method agrees well with traditional AMS measurement of the same samples as graphite, and that the 1$\sigma$ uncertainty is about 1 %.
+Results from measurements of consensus standards (TIRI I, IAEA C2, and an internal modern shell standard), and samples from a marine core (F^14^C = 0.4--1.15) show that the method agrees well with traditional AMS measurement of the same samples as graphite, and that the 1$\sigma$ uncertainty is about 1%.
 We discuss advantages and disadvantages of continuous flow sample introduction, and the effect of reduced precision on calibrated age-depth models produced using gas-source data.
@@
 The mixture of sample CO~2~ and displacement helium from the vial flows up a 21 gauge (0.819 mm) needle inserted through the tee and stainless tubing which forms the inner flow path of the dual needle.
-The open split is a VICI Luer adapter (part ZBUMLPK) with 5cm of 1/16 in.
+The open split is a VICI Luer adapter (part ZBUMLPK) with 5\ cm of 1/16 in.
 PFTE tubing attached.
 The split is connected to the top of the inner needle, and the source delivery capillary is inserted into the split.
-The split has fairly low volume (< 1 mL).
+The split has fairly low volume (<\ 1\ mL).
 The system was operated manually during development but could be mounted on the arm of a vial-sampling autosampler to enable automated operation.
@@
 
-The helium displacement flow rate was set to 250 $\mu$l min^-1^ for initial tests, and this was reduced to 100 $\mu$l min^-1^ for later work, including most reference standard comparisons and the core comparison.
+The helium displacement flow rate was set to 250 $\mu$l min^-1^ for initial tests, and this was reduced to 100\ $\mu$l\ min^-1^ for later work, including most reference standard comparisons and the core comparison.
 The flow rate from the open split into the source was defined by 150 cm of 50 $\mu$m diameter fused silica capillary, split into two sections by a zero-volume bulkhead connector at the source high-voltage cage.
@@
 
-Where $C(t)$ is the fraction CO~2~, $r$ is the rate of helium displacement flow, and $V$ is the initial volume of the CO~2~-filled vial.
+Where $C(t)$ is the fraction CO~2~ at time $t$, $r$ is the rate of helium displacement flow, and $V$ is the initial volume of the CO~2~-filled vial.
 Figure 2c shows the resulting relationship of sample ion current to CO~2~ flow rate.
@@
 To ensure sufficient sample for optimal measurements, we chose 15 mg as a target sample mass.
-To ensure that samples started at or slightly above atmospheric pressure, we injected helium with a syringe to equilibrate pressure in the vials when the reaction with acid was complete.
+We allowed samples to equilibrate to 1 atm pressure using a syringe filled with helium when the reaction with acid was complete.
 
@@
 HGIS cathodes produce measurable ion current, even with no gas flow.
-This current was initially \> 10 $\mu$A and dropped exponentially, reaching a baseline of ~100 nA after 5-10 minutes.
+This current was initially \>\ 10\ $\mu$A and dropped exponentially, reaching a baseline of ~100 nA after 5-10 minutes.
 For the range of currents in this study (3--9 $\mu$A), this baseline current represents a 1--6% contribution to the total measured sample current at the F^14^C of the target current baseline.
@@
 Parallel graphite measurements were made following standard NOSAMS methods.
-One HGIS sample vial was noted to be at atmospheric pressure when the 16 mg sample should have produced only 0.5 atmosphere of CO~2~.
+One HGIS sample vial was noted to be at atmospheric pressure when the 16\ mg sample should have produced only 0.5 atmosphere of CO~2~.
 The HGIS measurement for this sample was 0.053 F^14^C higher than the measurement as graphite.
@@
 triplicate measurement or more core depths.
-This may lead to more accurate results compared to a more precise graphite measurement where reproducibility is not as well constrained.
+This may lead to more accurate results compared to a more precise graphite measurement where reproducibility is not well constrained by multiple measurements.
 The accuracy of age-depth models is often limited by environmental factors and sparse data, rather than the precision of individual dates [@blaauw_double_2018].
@@
 
+# Supplementary Materials
+
+```{r tabS1, child=here("doc/AMS-15/AMS-15_paper/tables/table_s1_reference.Rmd"), message = FALSE}
+```
+
+```{r tabS2, child=here("doc/AMS-15/AMS-15_paper/tables/table_s2_haiti_core.Rmd"), message = FALSE}
+```
+
 # References {-}

### 2022-04-14 doc/AMS-15/AMS-15_paper/tables/table_s1_reference.Rmd
Add supplementary tables and current size figure. Many small edits.

```{r}
library(tidyverse)
library(here)
library(hgis)
library(errors)

options(errors.notation = "plus-minus")
options(scipen = 999)

theme_set(theme_classic())

# Read data. Don't use first carbonate test, smaller samples, or gas standards.
results_df <- read_csv(here("data_analysed/carb_all_results.csv"))

cons_df <- results_df %>% 
  filter(!(wheel %in% c("USAMS020521", "USAMS061121", "USAMS061821"))) %>% 
  mutate(he12C = he12C * 1E6) %>% 
  compare_consensus() %>% 
  filter(!str_detect(Name, "Gas")) %>% 
  arrange(fm_consensus) %>% 
  mutate (fm_corr_err = set_errors(fm_corr, sig_fm_corr),
          Name = case_when(Name == "TIRI-F" ~ "TIRI F",
                           Name == "TIRI-I" ~ "TIRI I",
                           Name == "C-2" ~ "IAEA C2",
                           Name == "C-1" ~ "IAEA C1",
                           Name == "NOSAMS2" ~ "NOSAMS 2"))

cons_df %>% 
  select(Name, fm_consensus, fm_corr, sig_fm_corr, current = he12C) %>% 
  write_csv(here("data_analysed/reference_standards.csv"))

cons_df %>% 
  select(`Sample Name` = Name, 
         `Reference F^14^C` = fm_consensus,
         `HGIS F^14^C ± 1$\\sigma$` = fm_corr_err,
         `^12^C^-^ current ($\\mu$A)` = he12C) %>% 
  knitr::kable(digits = c(3, 4, 4, 1),
               caption = "\\label{tabS1}Measurements of reference standard materials. Reference 
               values are community consensus values for TIRI I and C2, and the 
               NOSAMS internal graphite value for NOSAMS2.")
```

### 2022-04-14 doc/AMS-15/AMS-15_paper/tables/table_s2_haiti_core.Rmd
Add supplementary tables and current size figure. Many small edits.

```{r}
library(tidyverse)
library(here)
library(hgis)
library(errors)

options(errors.notation = "plus-minus")
options(scipen = 999)

# Read data. Don't use first carbonate test, smaller samples, or gas standards.
core <- read_csv(here("data_analysed/haiti_combined.csv")) %>% 
  select(sample_name, depth, method, fm_corr, sig_fm_corr)

dups <- core %>% 
  filter(str_starts(sample_name, "HATC")) %>% 
  select(sample_name, depth, fm_corr_hgis = fm_corr, sig_fm_corr_hgis = sig_fm_corr) %>% 
  mutate(fm_corr_graphite = NA, sig_fm_corr_graphite = NA)

core_long <- core %>% 
  filter(!str_starts(sample_name, "HATC")) %>% 
  pivot_wider(names_from = method, values_from = c(fm_corr, sig_fm_corr)) %>% 
  bind_rows(dups) %>% 
  arrange(depth)

core_long %>% 
  write_csv(here("data_analysed/haiti_summary.csv"))

core_long %>% 
  mutate (hgis = set_errors(fm_corr_hgis, sig_fm_corr_hgis),
          graphite = set_errors(fm_corr_graphite, sig_fm_corr_graphite)) %>% 
  select(`Sample Name` = sample_name, 
         `Core Depth` = depth,
         `HGIS F^14^C ± 1$\\sigma$` = hgis,
         `graphite F^14^C ± 1$\\sigma$` = graphite) %>% 
  knitr::kable(digits = c(3, 4, 4, 4),
               caption = "\\label{tabS2}HGIS and graphite measurements of Haiti core carbonates.
                          HGIS replicates are listed without a corresponding graphite value.")
```

### 2022-04-14 doc/AMS-15/AMS-15_paper/figures/fig5_core_carbonates.R

@@ -5,10 +5,9 @@ library(tidyverse)
 library(here)
 library(hgis)
 library(amstools)
-library(scales)
-library(gt)
 library(Bchron)
 library(patchwork)
+library(latex2exp)
 
 theme_set(theme_classic())
 
@@ -248,6 +247,7 @@ chron_diff <- ggplot(chron_comb, aes(Age, pos_diff, color = method,
   scale_color_manual(values = c("#00b7bd", "#b7bf10")) +
   scale_fill_manual(values = c("#00b7bd", "#b7bf10")) +
   scale_x_reverse() +
+  scale_y_reverse() +
   xlim(8000, -500) +
   labs(title = "E",
        x = "Years BP",

### 2022-04-14 doc/AMS-15/AMS-15_paper/figures/figs1_size_current.R

@@ -1,0 +1,24 @@
+# Relationship of sample size to ion current
+
+library(tidyverse)
+library(hgis)
+library(here)
+
+# Read reduced data as produced by analyse_carbonates.R
+df <- readRDS(here("data_analysed/carb_all.rds")) %>% 
+  map_dfr(2) %>% 
+  compare_consensus() %>% 
+  mutate(c12ua = he12C * 1E6)
+samples <- read_csv(here("data/carb_data.csv")) %>% 
+  select(wheel, pos, mass, he_added)
+df <- left_join(df, samples, by = c("wheel", "pos")) %>% 
+  filter(wheel != "USAMS020521")
+
+
+# Basic plot of current vs mass
+df %>% 
+ggplot(aes(mass, c12ua, color = wheel)) +
+  geom_point()
+
+ggsave(here("doc/AMS-15/AMS-15_paper/figures/figs1_size_current.svg"), width = 4, height = 4)
+ggsave(here("doc/AMS-15/AMS-15_paper/figures/figs1_size_current.pdf"), width = 4, height = 4)

### 2022-04-18 doc/AMS-15/AMS-15_paper/figures/fig3_blanks.R
Final bits before submission

@@ -138,8 +138,8 @@ blank_model <- ggplot(ds, aes(Cur_inv, mean, color = Gas, shape = Gas)) +
 # Get data
 blanks <- readRDS(here("data_analysed/carb_all.rds")) %>% 
   map_dfr(2) %>% 
-  filter(rec_num %in% c(2138, 83028),
-         he12C > 2E-6,
+  filter(!(wheel %in% c("USAMS020521", "USAMS061121", "USAMS061821")),
+         rec_num %in% c(2138, 83028),
          sample_type == "B") %>% 
   mutate(he12C = he12C * 1E6,
          sig_he12C = sig_he12C * 1E6)

### 2022-04-18 doc/AMS-15/AMS-15_paper/figures/fig4_consensus_samples.R

@@ -12,11 +12,17 @@ theme_set(theme_classic())
 results_df <- read_csv(here("data_analysed/carb_all_results.csv"))
 
 cons_df <- results_df %>% 
-  filter(!(wheel %in% c("USAMS020521", "USAMS061121", "USAMS061821"))) %>% 
-  mutate(he12C = he12C * 1E6) %>% 
   compare_consensus() %>% 
-  filter(!str_detect(Name, "Gas")) %>% 
-  arrange(Name)
+  filter(!(wheel %in% c("USAMS020521", "USAMS061121", "USAMS061821")),
+         !str_detect(Name, "Gas")) %>% 
+  mutate(he12C = he12C * 1E6,
+         Name = case_when(Name == "TIRI-F" ~ "TIRI F",
+                          Name == "TIRI-I" ~ "TIRI I",
+                          Name == "C-2" ~ "IAEA C2",
+                          Name == "C-1" ~ "IAEA C1",
+                          Name == "NOSAMS2" ~ "NOSAMS 2"),
+         Name = factor(Name, levels = c("IAEA C1", "TIRI F", "TIRI I", 
+                                        "IAEA C2", "NOSAMS 2")))
 
 mean_diff <- mean(cons_df$fm_diff)
 sd_diff <- sd(cons_df$fm_diff)
@@ -28,12 +34,15 @@ ggplot(aes(fm_consensus, fm_diff, color = Name, shape = Name)) +
   geom_hline(yintercept = mean_diff, color = "blue") +
   geom_hline(yintercept = mean_diff + sd_diff, color = "lightblue") +
   geom_hline(yintercept = mean_diff - sd_diff, color = "lightblue") +
-  geom_smooth(method = "lm",  se = FALSE) + 
+  #geom_smooth(method = "lm",  se = FALSE) + 
+  #geom_boxplot(aes(fm_consensus, fm_diff, group = fm_consensus), width = .3) +
   geom_pointrange(aes(ymin = fm_diff - sig_fm_corr, 
                       ymax = fm_diff + sig_fm_corr), 
                   position = position_dodge2(width = 0.1), 
-                  ) + #size = 0.5) + 
+                  size = .6,
+                  fatten = 5) + 
   scale_color_brewer(palette = "Dark2") +
+  scale_shape_manual(values = c(2, 6, 5, 0, 1)) +
[…]

### 2022-04-18 doc/AMS-15/AMS-15_paper/AMS_HGIS_els/AMS_HGIS_els.Rmd
Final bits before submission
Unnumber extra sections
Modifications for submission

@@
     correspondingauthor: true
-  - name: Josh R. Burton
+  - name: Joshua R. Burton
     email: jburton@whoi.edu
@@
 date: "`r Sys.Date()`"
-classoption: preprint, 3p, authoryear
+classoption: 12pt, review, 3p, authoryear
 bibliography: ../HGIS_paper.bib
@@
     citation_package: default
+header-includes: 
+ - \usepackage[section]{placeins}
 ---
@@
 ```
+
+\newpage
 
@@
 The total flow rate of gas to the source varied between 30 $\mu$l min^-1^ for pure helium and 40 $\mu$l min^-1^ for pure CO~2~ due to the viscosity difference between helium and CO~2~, estimated using the Hagen-Poiseuille equation.
-
-```{r fig1, fig.width = 5, fig.height = 5, fig.align='center', out.width="80%", fig.cap = "\\label{fig1}Schematic of the open split inlet.
-The double needle is housed in a tee fitting and allows displacement helium to flow into the sample vial and the sample-helium mixture to flow to the open split, where a capillary delivers the mixture to the ion source.", echo = FALSE}
-knitr::include_graphics(here("doc/AMS-15/AMS-15_paper/figures/fig1_HGIS_open_split_drawing.pdf"))
-```
 
@@
 
-```{r fig2, fig.width = 5, fig.height = 5, fig.align='center', out.width="80%", fig.cap = "\\label{fig2}Performance of the NOSAMS MCSNICS hybrid gas ion source.
-(a) Measured $^{12}C^-$ ion current during displacement of modern tank $CO_2$ from a sample vial with helium.
-(b) Modeled flows of helium and $CO_2$ to the ion source during the same dilution experiment.
-(c) Measured ion current for the modeled $CO_2$ flow from b.
-The stability of the source over a range of flow produces stable current for up to 40 min of measurement time.", echo = FALSE}
-knitr::include_graphics(here("doc/AMS-15/AMS-15_paper/figures/fig2_current_flow.pdf"))
-```
-
 Source ion currents are variable by day and by target under similar source, inlet, and sample conditions.
@@
 
-```{r fig3, fig.width = 5, fig.height = 5, fig.align='center', out.width="80%", fig.cap = "\\label{fig3}Assessment of blanks.
-(a) Current dependent blank model and data.
-$F^{14}C$ and current of measurements from dilution tests of modern and dead tank standards are used to generate linear fits which should cross at the $F^{14}C$ and current of the constant current blank.
-Direct determination of $F^{14}C$ and current of the blank using clean targets and helium without $CO_2$ are also shown.
-(b) Fraction modern of large IAEA C1 and TIRI F carbonate blank samples compared to sample current.
-Horizontal and vertical bars represent measurement uncertainty in current and ratio, respectively.", echo = FALSE}
-knitr::include_graphics(here("doc/AMS-15/AMS-15_paper/figures/fig3_blanks.pdf"))
-```
-
 The directly measured blank current (120 nA) agrees well with the model assessment, but has higher F^14^C (0.82).
@@
 
-```{r tab1, message = FALSE}
-# Summary table
-cons_df %>% 
-  group_by(Name, fm_consensus) %>% 
-  summarize(across(c(fm_corr, sig_fm_corr, fm_diff, sigma),
-                   list(mean = mean, sd = sd)),
-            N = n()) %>% 
-  ungroup() %>% 
-  arrange(fm_consensus) %>% 
-  mutate (fm_corr_err = set_errors(fm_corr_mean, fm_corr_sd),
-          Name = case_when(Name == "TIRI-I" ~ "TIRI I",
-                           Name == "C-2" ~ "IAEA C2",
-                           Name == "NOSAMS2" ~ "NOSAMS 2")) %>% 
-  select(`Sample Name` = Name, 
-         `Reference F^14^C` = fm_consensus,
-         `HGIS F^14^C ± stdev` = fm_corr_err,
-         `Mean per-sample error` = sig_fm_corr_mean,
-         N) %>% 
-  knitr::kable(digits = c(4, 4, 4, 3, 2),
-               caption = "\\label{tab1}Reference material precision and accuracy. Reference 
-               values are community consensus values for TIRI I and C2, and the 
-               NOSAMS internal graphite value for NOSAMS2. The mean of per-sample 
-               error estimates is included for comparison to the reproducibilty 
-               of the samples.")
-
-```
-
 The difference between reference samples and graphite consensus does not vary systematically by F^14^C, indicating that the method has no significant bias with radiocarbon content (Figure 4).
 Each reference standard type agrees with graphite consensus values within the standard deviation of samples measured, with slightly more measurements agreeing with graphite within 1$\sigma$ uncertainties than would be expected statistically.
-
-```{r fig4, fig.width = 5, fig.height = 5, fig.align='center', out.width="80%", fig.cap = "\\label{fig4}Agreement of standards with graphite consensus values.
-Difference of HGIS measurements from consensus for individual samples are plotted against their expected value.
-Expected values are spread to avoid overlapping points.
-Error bars represent 1$\\sigma$ sample uncertainty.
-The black line is the mean difference from expected value for all measurements, and the blue lines are one standard deviation of all measurements.", echo = FALSE}
-knitr::include_graphics(here("doc/AMS-15/AMS-15_paper/figures/fig4_consensus_samples.pdf"))
-```
 
@@
 
-```{r fig5, fig.width = 5, fig.height = 5, fig.align='center', out.width="80%", fig.cap = "\\label{fig5}Comparison of Haiti core samples measured as graphite and via HGIS.
-(a) Difference between HGIS and graphite measurements of the same materials.
-Difference between HGIS and graphite measurements is shown on the vertical axis and plotted against the mean of the two measurements.
-Error bars are 1$\\sigma$ measurement uncertainty for each sample.
-Graphite measurements are included to allow comparison of uncertainty.
-Bayesian age-depth models produced using HGIS (b) and graphite measurements (c).
-Grey subplots show probability distribution of calibrated ages.
-(d) Combined age-depth prediction and 95\\% confidence intervals for HGIS and graphte.
-(e) Offset of HGIS and graphite values from the mean of the models with confidence intervals.", echo = FALSE}
-knitr::include_graphics(here("doc/AMS-15/AMS-15_paper/figures/fig5_core_carbonates.pdf"))
-```
-
 ```{r}
@@
 
-# Acknowledgements
+\section*{Acknowledgements}
 
@@
 
-# Supplementary Materials
+\newpage
+
+\section*{Figures}
+
+```{r fig1, fig.width = 5, fig.height = 5, fig.align='center', out.width="60%", fig.cap = "\\label{fig1}Schematic of the open split inlet.
+The double needle is housed in a tee fitting and allows displacement helium to flow into the sample vial and the sample-helium mixture to flow to the open split, where a capillary delivers the mixture to the ion source.", echo = FALSE}
+knitr::include_graphics(here("doc/AMS-15/AMS-15_paper/figures/fig1_HGIS_open_split_drawing.pdf"))
+```
+
+```{r fig2, fig.width = 5, fig.height = 5, fig.align='center', out.width="80%", fig.cap = "\\label{fig2}Performance of the NOSAMS MCSNICS hybrid gas ion source.
+(a) Measured $^{12}C^-$ ion current during displacement of modern tank $CO_2$ from a sample vial with helium.
+(b) Modeled flows of helium and $CO_2$ to the ion source during the same dilution experiment.
+(c) Measured ion current for the modeled $CO_2$ flow from b.
+The stability of the source over a range of flow produces stable current for up to 40 min of measurement time.", echo = FALSE}
+knitr::include_graphics(here("doc/AMS-15/AMS-15_paper/figures/fig2_current_flow.pdf"))
+```
+
+```{r fig3, fig.width = 5, fig.height = 5, fig.align='center', out.width="80%", fig.cap = "\\label{fig3}Assessment of blanks.
+(a) Current dependent blank model and data.
+$F^{14}C$ and current of measurements from dilution tests of modern and dead tank standards are used to generate linear fits which should cross at the $F^{14}C$ and current of the constant current blank.
+Direct determination of $F^{14}C$ and current of the blank using clean targets and helium without $CO_2$ are also shown.
+(b) Fraction modern of large IAEA C1 and TIRI F carbonate blank samples compared to sample current.
+Horizontal and vertical bars represent measurement uncertainty in current and ratio, respectively.", echo = FALSE}
+knitr::include_graphics(here("doc/AMS-15/AMS-15_paper/figures/fig3_blanks.pdf"))
+```
+
+```{r fig4, fig.width = 5, fig.height = 5, fig.align='center', out.width="80%", fig.cap = "\\label{fig4}Agreement of standards with graphite consensus values.
+Difference of HGIS measurements from consensus for individual samples are plotted against their expected value.
+Expected values are spread to avoid overlapping points.
+Error bars represent 1$\\sigma$ sample uncertainty.
+The black line is the mean difference from expected value for all measurements, and the blue lines are one standard deviation of all measurements.", echo = FALSE}
+knitr::include_graphics(here("doc/AMS-15/AMS-15_paper/figures/fig4_consensus_samples.pdf"))
+```
+
+```{r fig5, fig.width = 5, fig.height = 5, fig.align='center', out.width="80%", fig.cap = "\\label{fig5}Comparison of Haiti core samples measured as graphite and via HGIS.
+(a) Difference between HGIS and graphite measurements of the same materials.
+Difference between HGIS and graphite measurements is shown on the vertical axis and plotted against the mean of the two measurements.
+Error bars are 1$\\sigma$ measurement uncertainty for each sample.
+Graphite measurements are included to allow comparison of uncertainty.
+Bayesian age-depth models produced using HGIS (b) and graphite measurements (c).
+Grey subplots show probability distribution of calibrated ages.
+(d) Combined age-depth prediction and 95\\% confidence intervals for HGIS and graphte.
+(e) Offset of HGIS and graphite values from the mean of the models with confidence intervals.", echo = FALSE}
+knitr::include_graphics(here("doc/AMS-15/AMS-15_paper/figures/fig5_core_carbonates.pdf"))
+```
+
+\newpage
+
+\section*{Tables}
+
+```{r tab1, message = FALSE}
+# Summary table
+cons_df %>% 
+  group_by(Name, fm_consensus) %>% 
+  summarize(across(c(fm_corr, sig_fm_corr, fm_diff, sigma),
+                   list(mean = mean, sd = sd)),
+            N = n()) %>% 
+  ungroup() %>% 
+  arrange(fm_consensus) %>% 
+  mutate (fm_corr_err = set_errors(fm_corr_mean, fm_corr_sd),
+          Name = case_when(Name == "TIRI-I" ~ "TIRI I",
+                           Name == "C-2" ~ "IAEA C2",
+                           Name == "NOSAMS2" ~ "NOSAMS 2")) %>% 
+  select(`Sample Name` = Name, 
+         `Reference F^14^C` = fm_consensus,
+         `HGIS F^14^C ± stdev` = fm_corr_err,
+         `Mean per-sample error` = sig_fm_corr_mean,
+         N) %>% 
+  knitr::kable(digits = c(4, 4, 4, 3, 2),
+               caption = "\\label{tab1}Reference material precision and accuracy. Reference 
+               values are community consensus values for TIRI I and C2, and the 
+               NOSAMS internal graphite value for NOSAMS2. The mean of per-sample 
+               error estimates is included for comparison to the reproducibilty 
+               of the samples.")
+
+```
+
+\newpage
+
+\section*{Supplementary Materials}
 
@@
 
+\newpage
+
 # References {-}

### 2022-08-04 doc/AMS-15/AMS-15_paper/response_to_reviewers_HGIS.txt
add response to reviewers

Review reply letter for HGIS

We thank the two reviewers for their time and attention to improving our work. The second reviewer had no specific questions or suggestions, and recommended publication as is. The first reviewer had several suggestions on comparison with previous methods and the provenance of our suite of marine core shell carbonates. We address these concerns in the attached revised manuscript and interleaved with their comments below.

    More figures to describe each experimental steps should be included. For examples, photos and more comprehensive figures of the new system is useful rather than the simple figure currently depicted in the figure 1. What is the difference between previous methods with current apparatus (eg., Roberts et al., 2013 Radiocarbon)?

​We have added more detail to figure 1. Our apparatus for delivery of the helium CO2 mixture is very simple and is complete as described by figure 1 and in the text. We can add photographs if neccessary but don't feel they add anything to the clarity of the method. We have referenced previous methods in the text and have discussed differences where relevant. We have added and edited text to make these sections clearer.

    Figures comparing previous methods/papers reference material measurements with the present "new" methods should be shown to indicate advantage of the current method.

We have not seen a methods paper with these kinds of comparisons. We reference all relevant previous papers, which include detailed discussion of those methods, including data on measurements. We believe these kinds of comparisons belong in a review paper, rather than a short methods paper.

    Figures to show performance of AMS measurements has to be included such as the one similar to the figure 3 in Roberts et al., 2013.

​We have added a plot of ratio over time to our ion current figure (fig 3).

    Flowcharts is useful to describe steps of experiments compare to Roberts et al., 2013.

​We believe this kind of comparison is beyond the scope of a methods paper. We reference Roberts et al 2013 as another method that uses an open split interface. We have clarified the similarities and differences between these two methods in the text.

    At the end of the introduction, sentence to describe the purpose of this study should be added.
    In the conclusion section, authors are advised to present future perspectives to improve accuracy should be stated.

​I'll look at this and add as necessary.

    I think it is useful to show the comparison between graphitized carbon results with the current method using sediment cores obtained from Haiti. However, if they describe details of sampling, geological information such as columnar sections, photos of sediments, and so forth should be put in the section 3.4. I think that the detailed sampling protocol are needed so can be amended.

    If they employ new local reservoir age, they should provide further information for 10 samples. Otherwise references should be properly cited.

We believe these details are beyond the scope of this methods paper. We make clear that the same samples are used for the comparison between HGIS and graphite measurements, and therefore their provenance is not the subject of this work. Details about the core setting and local reservoir age do not affect our conclusions. Details of the core, geologic context, and additional data are published in S. Moser's thesis (referenced in the text and available online), and these details will be published in a paper in prep.
​

### 2022-08-04 doc/AMS-15/AMS-15_paper/AMS_HGIS_els/AMS_HGIS_els.Rmd
address reviews, renumber figures

@@
 The system was operated manually during development but could be mounted on the arm of a vial-sampling autosampler to enable automated operation.
+The dual needle and open split design are nearly identical to those used by Roberts et al.
+@Roberts2013, with the difference that helium, rather than water, is used for sample displacement.
 
@@
 A contrast of the open split approach with closed-source HGIS methods is changing CO~2~ flow and helium dilution ratio over the period of measurement due to the dilution of the sample by the displacement helium.
-We examined the effect of this variable carbon delivery and dilution by measuring ^12^C^-^ current evolution during helium displacement/dilution of vial CO~2~ (Figure \ref{fig2}a) and comparing this to the composition of the gas delivered to the source.
-We estimated gas composition (Figure \ref{fig2}b) using a simple mixing model:
+We examined the effect of this variable carbon delivery and dilution by measuring ^12^C^-^ current (Figure \ref{fig2}a) and 14/12 C ratio (Figure \ref{fig2}b) evolution during helium displacement/dilution of vial CO~2~ and comparing this to the composition of the gas delivered to the source.
+We estimated gas composition (Figure \ref{fig2}c) using a simple mixing model:
 
@@
 Where $C(t)$ is the fraction CO~2~ at time $t$, $r$ is the rate of helium displacement flow, and $V$ is the initial volume of the CO~2~-filled vial.
-Figure 2c shows the resulting relationship of sample ion current to CO~2~ flow rate.
+Figure 3 shows the resulting relationship of sample ion current to CO~2~ flow rate.
 The most important feature of the current-flow relationship is the wide range of CO~2~ flow rates over which stable ion current is measured, which in turn produces relatively stable currents and ratios despite a changing carbon flow rate during measurement.
@@
 The constant contamination blank was also assessed directly by measuring pure helium in vials using the same method.
-Results from both tests are shown in Figure 3a.
+Results from both tests are shown in Figure 4a.
 The inverse of the sample current is used to linearize the data, allowing assessment with a simple linear model.
@@
 If the fraction modern of the blank is 0.8 modern as estimated above, radiocarbon-free samples will show the most effect of a constant contamination blank with current.
-Figure 3b shows the fraction modern of measured blanks compared to their mean ^12^C^-^ ion current.
+Figure 4b shows the fraction modern of measured blanks compared to their mean ^12^C^-^ ion current.
 Fraction modern of measured blanks does decrease with increasing current, but the decrease is small, less than F^14^C 0.01 over the range of currents measured.
@@
 
-The difference between reference samples and graphite consensus does not vary systematically by F^14^C, indicating that the method has no significant bias with radiocarbon content (Figure 4).
+The difference between reference samples and graphite consensus does not vary systematically by F^14^C, indicating that the method has no significant bias with radiocarbon content (Figure 5).
 Each reference standard type agrees with graphite consensus values within the standard deviation of samples measured, with slightly more measurements agreeing with graphite within 1$\sigma$ uncertainties than would be expected statistically.
@@
 
-We found the HGIS and graphite measurements to be in good agreement, with no significant bias over the range in F^14^C of the core samples (0.4 to 1.15) (Figure 5a).
+We found the HGIS and graphite measurements to be in good agreement, with no significant bias over the range in F^14^C of the core samples (0.4 to 1.15) (Figure 6a).
 Mean difference between hybrid and graphite measurements was -0.003 F^14^C and the standard deviation of this difference was 0.009 F^14^C.
@@
 
-Age models were developed for both the HGIS measurements (Figure 5b) and the traditional graphite measurements (Figure 5c) of the core carbonates.
+Age models were developed for both the HGIS measurements (Figure 6b) and the traditional graphite measurements (Figure 6c) of the core carbonates.
 We used the Bchron Bayesian radiocarbon chronology package [@haslett_simple_2008] to develop these models.
@@
 @hua_atmospheric_2013 was used for calibration of postmodern dates, but these measurements were not included in the age model due to uncertainties in reservoir correction.
-The age-depth curves produced by the two datasets agree within the uncertainty of the chronologies (Figure 5d & 5e), with the largest differences being in the uncertainty of the individual dates.
+The age-depth curves produced by the two datasets agree within the uncertainty of the chronologies (Figure 6d & 6e), with the largest differences being in the uncertainty of the individual dates.
 
@@
 (a) Measured $^{12}C^-$ ion current during displacement of modern tank $CO_2$ from a sample vial with helium.
-(b) Modeled flows of helium and $CO_2$ to the ion source during the same dilution experiment.
-(c) Measured ion current for the modeled $CO_2$ flow from b.
-The stability of the source over a range of flow produces stable current for up to 40 min of measurement time.", echo = FALSE}
+(c) Modeled flows of helium and $CO_2$ to the ion source during the same dilution experiment.", echo = FALSE}
 knitr::include_graphics(here("doc/AMS-15/AMS-15_paper/figures/fig2_current_flow.pdf"))
@@
 
-```{r fig3, fig.width = 5, fig.height = 5, fig.align='center', out.width="80%", fig.cap = "\\label{fig3}Assessment of blanks.
+```{r fig3, fig.width = 5, fig.height = 5, fig.align='center', out.width="80%", fig.cap = "\\label{fig3}Measured ion current for the modeled $CO_2$ flow from b.
+The stability of the source over a range of flow produces stable current for up to 40 min of measurement time.", echo = FALSE}
+knitr::include_graphics(here("doc/AMS-15/AMS-15_paper/figures/fig3_current_flow.pdf"))
+```
+
+```{r fig4, fig.width = 5, fig.height = 5, fig.align='center', out.width="80%", fig.cap = "\\label{fig4}Assessment of blanks.
 (a) Current dependent blank model and data.
@@
 
-```{r fig4, fig.width = 5, fig.height = 5, fig.align='center', out.width="80%", fig.cap = "\\label{fig4}Agreement of standards with graphite consensus values.
+```{r fig5, fig.width = 5, fig.height = 5, fig.align='center', out.width="80%", fig.cap = "\\label{fig5}Agreement of standards with graphite consensus values.
 Difference of HGIS measurements from consensus for individual samples are plotted against their expected value.
@@
 
-```{r fig5, fig.width = 5, fig.height = 5, fig.align='center', out.width="80%", fig.cap = "\\label{fig5}Comparison of Haiti core samples measured as graphite and via HGIS.
+```{r fig6, fig.width = 5, fig.height = 5, fig.align='center', out.width="80%", fig.cap = "\\label{fig6}Comparison of Haiti core samples measured as graphite and via HGIS.
 (a) Difference between HGIS and graphite measurements of the same materials.

### 2022-08-04 doc/AMS-15/AMS-15_paper/figures/fig2_current_flow.R
Add ratio plot, break out current flow figure

@@ -31,6 +31,18 @@ cur_time <- ggplot(data, aes(time, he12C)) +
   theme(axis.title.x = element_blank(),
         axis.text.x = element_blank())
 
+# ratio vs time subplot
+cur_rat <- ggplot(data, aes(time, he14_12)) +
+  geom_smooth(span = .4, se = FALSE, color = "#00b7bd") +
+  geom_point(size = 3, color = "#0069b1") +
+  xlim(0, 125) +
+  labs(title = "B",
+       #title = "Ion current",
+       x = "Time (min)",
+       y = TeX(r"( $^\frac{14}{12}C^-$ ratio )")) +
+  theme(axis.title.x = element_blank(),
+        axis.text.x = element_blank())
+
 # flow vs time subplot
 flow_time <- data.frame(x = 0:125) %>% 
   ggplot(aes(x)) +
@@ -39,7 +51,7 @@ flow_time <- data.frame(x = 0:125) %>%
   stat_function(fun = function(x) 30 - (concCO2(x, r = 244) * 30),
                 aes(color = "Helium"), size = 1.5) +
   scale_color_manual("Gas", values = c("#00b7bd", "#b7bf10")) +
-  labs(title = "B",
+  labs(title = "C",
        #title = "Gas flows to source",
        #subtitle = "7mL vial, 244μL/min helium, 30μL/min to source",
        x = "Time (min)",
@@ -48,16 +60,20 @@ flow_time <- data.frame(x = 0:125) %>%
         legend.background = element_rect(fill = "white", color = "black"))
 
 # Current vs flow subplot
-
 cur_flow <- ggplot(data, aes(co2flow, he12C)) +
   geom_smooth(span = .4, se = FALSE, color = "#00b7bd") +
   geom_point(size = 3, color = "#0069b1") +
-  labs(title = "C",
+  labs(title = "D",
        x = TeX(r"($CO^2$ flow (μl $min^{-1}$))"),
        y = TeX(r"( $^{12}C^-$ current (μA))")) 
 
 # Build figure and save
-cur_time / flow_time / cur_flow
+cur_time / cur_rat / flow_time
 
 ggsave(here("doc/AMS-15/AMS-15_paper/figures/fig2_current_flow.svg"), width = 7, height = 7)
 ggsave(here("doc/AMS-15/AMS-15_paper/figures/fig2_current_flow.pdf"), width = 7, height = 7)
+
+cur_flow
+
[…]

### 2022-08-09 doc/AMS-15/AMS-15_paper/AMS_HGIS_els/AMS_HGIS_els.Rmd
More edits for review revisions

@@
 The hybrid gas capability of this source was developed as a joint project between NEC and the Scottish Universities Environmental Research Centre (SUERC) [@Xu2007], and further improved by work at Lawrence Livermore National Laboratory (LLNL) [@Salazar2013].
-We present here a system for semi-automated analysis of carbonate minerals by acid hydrolysis in vials and introduction to the ion source in a flow of helium via an open split.
+We present here a system for semi-automated analysis of carbonate minerals by acid hydrolysis in vials and introduction to the ion source in a flow of helium via an open split, and results for measurements of reference materials and shell macrofossils from a marine sediment core.
 
@@
 The system was operated manually during development but could be mounted on the arm of a vial-sampling autosampler to enable automated operation.
-The dual needle and open split design are nearly identical to those used by Roberts et al.
-@Roberts2013, with the difference that helium, rather than water, is used for sample displacement.
+The dual needle and open split design are similar to those used by Roberts et al.
+@Roberts2013, with the differences that helium, rather than water, is used for sample displacement, and displacement helium flows into the vial through the outer/upper flow path.
 
@@
 The total flow rate of gas to the source varied between 30 $\mu$l min^-1^ for pure helium and 40 $\mu$l min^-1^ for pure CO~2~ due to the viscosity difference between helium and CO~2~, estimated using the Hagen-Poiseuille equation.
+The HGIS itself is as described by Xu et al.
+[Xu2007].
 
@@
 Sealed vials were evacuated, then 1 ml 85% phosphoric acid was added.
-The samples were allowed to react at 60 C for 6 h.
+The samples were allowed to react at 60 C for 1 h to overnight.
 Sample pressure was adjusted to 1 atm by allowing the sample to equilibrate with a helium-filled syringe connected through the vial septum.
@@
 ```{r}
-# Fix to use current form of processed data
 blanks <- carb %>% 
@@
 blanksd <-  sd(blanks$norm_ratio) %>% format(digits = 1)
-
-# errors should produce an "all in one" for this, but it doesn't work with inline code
-# meanblank <- as.errors(mean(blanks$norm_ratio), sd(blanks$norm_ratio))
 nblank <- length(blanks$norm_ratio)
@@
 
-```{r}
-# Agreement within uncertainty of paired ages?
-# Mean and sd of difference of calibrated ages?
-```
-
 Age models were developed for both the HGIS measurements (Figure 6b) and the traditional graphite measurements (Figure 6c) of the core carbonates.
@@
 This system produces results that agree with graphite measurements within the approximately 1% method error from 8--20 mg of carbonate.
-Sample preparation takes about 10 min of person time and 6 h of automated processing and reaction time, and 20 min of measurement time per sample.
+Sample preparation takes about 10 min of person time and 1 h of automated processing and reaction time, and 20 min of measurement time per sample.
 The simplicity of of this approach can potentially reduce sample cost by eliminating graphitization, and may improve reliability and reduce maintenance compared to more complex sample inlet designs.
+Further development should focus on improvement of precision and accuracy by increasing currents and current stability and reducing the magnitude and uncertainty of the blank.
 
@@
 (a) Measured $^{12}C^-$ ion current during displacement of modern tank $CO_2$ from a sample vial with helium.
+(b) Uncorrected $^{14/12}C$ ratio measured during the same test.
 (c) Modeled flows of helium and $CO_2$ to the ion source during the same dilution experiment.", echo = FALSE}

### 2022-08-09 doc/AMS-15/AMS-15_paper/figures/fig2_current_flow.R

@@ -39,7 +39,7 @@ cur_rat <- ggplot(data, aes(time, he14_12)) +
   labs(title = "B",
        #title = "Ion current",
        x = "Time (min)",
-       y = TeX(r"( $^\frac{14}{12}C^-$ ratio )")) +
+       y = TeX(r"( $^{14/12}C$ ratio )")) +
   theme(axis.title.x = element_blank(),
         axis.text.x = element_blank())
 
